### 2024-03-19 review.qmd
First commit of Quarto document

---
title: "Measuring contrast processing in the visual system using SSVEP (and VEP?)"
format: html
---

## Baker, Wade: Review: Vis Neurosci

Figures:

*1: a) A typical SSVEP setup: a screen, a grating, a flicker rate: b)
The responses from a single electrode (time domain) c) Frequency
domain.*

```{python}
import matplotlib.pyplot as plt
import numpy as np
import scipy as scp

inputFrequencies_Hz=np.array([5,7])

duration_S=1

digitisationRate_Hz=1000

support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)

inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))

for i, freq in enumerate(inputFrequencies_Hz):
    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)

# Create a figure with two subplots
fig, axs = plt.subplots(len(inputFrequencies_Hz), 1, figsize=(8, 6))

# Plot each input modulation in a separate subplot
for i, freq in enumerate(inputFrequencies_Hz):
    axs[i].plot(support_S, inputModulations[:, i])
    axs[i].set_xlabel('Time (s)')
    axs[i].set_ylabel('Amplitude')
    axs[i].set_title('Frequency: {} Hz'.format(freq))

# Adjust the spacing between subplots
plt.tight_layout()

```

The two inputs are represented here as amplitude modulations across time.

If the two inputs are simply added together, the representation of the resulting signal in the Fourier domain is just the linear sum of the two independent signals:

```{python}

freqCutoff_Hz=30
combinedSignal=np.sum(inputModulations,axis=1)
FT_combinedSignal=np.fft.fft(combinedSignal)/len(combinedSignal)
fig,axs=plt.subplots(1,2,figsize=(8,6))
axs[0].plot(support_S,combinedSignal)
axs[0].set_xlabel('Time (s)')
axs[0].set_ylabel('Amplitude (a.u.)')
axs[0].set_title('Input signals with linear combination')

# Create the frequency support for the FFT plot
frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

# Plot the absolute values of the FFT in the second axis as a bar plot
axs[1].bar(frequencySupport, np.abs(FT_combinedSignal[frequencySupport]))
axs[1].set_xlabel('Frequency (Hz)')
axs[1].set_ylabel('Magnitude (a.u.)')
axs[1].set_title('FFT of Combined Signal')

plt.tight_layout()
plt.show()

```

*2: Sweeps and CRFs. Measurements of contrast sensitivity, extrapolating
the sweep to zero response to get t'hold. Infants and adults?*

*3: Measurements of modulation. Figures adapted from other papers:
Attention to space, attention to features, adaptation (?).
masking/surround suppression\\*

*4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*

*5: Future directions (decoding in frequency domain?, animals? BCI?)*

[]{#anchor-1}ABSTRACT:

[]{#anchor-2}History

[]{#anchor-3}1. The basics of SSVEP and contrast sensitivity including a
history of both fields

SSVEP (Steady-State Visual Evoked Potential): A continuous electrical
response evoked in the brain by visual stimuli flickering at a constant
frequency (Regan, 1966).

Contrast Sensitivity: The ability to detect differences in luminance
between an object and its background (Campbell & Green, 1965).

Regan, D. (1966). Some characteristics of average steady-state and
transient responses evoked by modulated light. Electroencephalography
and clinical neurophysiology, 20(3), 238-248.

Campbell, F. W., & Green, D. G. (1965). Optical and retinal factors
affecting visual resolution. J Physiology, 181, 576-593.

Check Regan paper for earlier (e.g. EEG refs). Norcia review will be
helpful!

(From Tyler / Levi / Apkarian paper):

6\. Regan, D.: Rapid methods for refracting the eye and assessing the
visual acuity in amblyopia using steady-state visual evoked potentials.
In Desmedt, J.E., editor: Visual Evoked Potentials in Man: New
Developments, Oxford, 1977, Clarendon Press, pp. 418-426.

7\. Fricker, S.J.: Narrow-band filter techniques for the detection and
measurement of evoked responses, Electroencephalogr. Clin. Neurophysiol.
14:411, 1962.

8\. Van der Tweel, L. H., Sem-Jacobsen, C.W., Kamp, A., Van Leeuwen,
W.S., and Verings, F.T.H.: Objective determination of response to
modulated light, Acta Physiol. Pharmacol. Neerl. 7:528, 1958.

9\. Regan, D.: Latencies of evoked potentials to flicker and to pattern
speedily estimated by simultaneous stimulation method,
Electroencephalogr. Clin. Neurophysiol. 40:654, 1976.

10\. Tyler, C.W., Apkarian, P., and Nakayama, K.: Multiple spatial
frequency tuning of electrical responses from the human\....

\[Around here perhaps a section about what gain control is, mentioning
other methods as well including psychophysics, MRI, electrophysiology,
and other EEG markers including ERPs and evoked gamma band oscillations.
Maybe outline the Heeger gain control model and its cousins.\]

\[Sure. But in fact gain control is only a small part of this story -
especially in the early days. They were, I think, more interested in
using SSVEP to measure absolute contrast sensitivity and to get lower
bounds on things like infant visual development. Gain control might be
better as a separate section later.\]

[]{#anchor-4}2. Spatial, temporal frequency and contrast sensitivity
measurements

Basically using SSVEP to measure a cortical output amplitude for any
given input contrast. You can vary parameters like SF, TF, position,
color and of course contrast. Early on people realised that you can
'sweep' the stimulus to get a CRF. You broadly get a line in log
contrast space if you do that (Tyler) - then you can extrapolate that
line down to zero response to estimate the threshold. That doesn't
\>quite\< work but it's pretty close.

You can also use this for 'difficult' populations like babies. One
interesting story was about how SSVEP become a replacement for
preferential looking (which was the other way of looking at infant
visual development). See e.g. Davida Teller. SSVEPs allowed people to
make objective measurements of contrast sensitivity development and
deduce that the visual system was more mature (e.g. more functional) in
infancy than previously expected. Also measures of colour sensitivity.
Tinyeyes is based off those measurements. Other people: Tyler, Norcia,
Gunilla H-P, many of the people at SKERI in the 1980s and 90s. Norcia
86,88,90 - mentioned in Regan's nice autobiography :
https://www.google.co.uk/books/edition/Seeing_Spatial_Form/HqaKlR_GpP0C?hl=en&gbpv=1&dq=Analysis+of+EEG+responses+in+man,+evoked+by+sine+watre+modulated+light.+Thesis&pg=PA215&printsec=frontcover

In parallel of course, people were using frequency tagging to do single
unit work - the 1F vs 2F simple/complex cell classification scheme was
all about this (Lennie and others).

[]{#anchor-5}3. SSVEP in functional localization

Techniques like fMRI have been combined with SSVEP to achieve more
precise spatial localization (Di Russo et al., 2007). See however the
Ales cruciform paper.

Di Russo, F., Pitzalis, S., Spitoni, G., Aprile, T., Patria, F., Stella,
A., \... & Hillyard, S. A. (2007). Identification of the neural sources
of the pattern-reversal VEP. Neuroimage, 34(1), 177-189.

Ales and Norcia:\... (showed that people's intuition about V1 upper /
lower v.f reversing polarity wasn't really correct). This feels a little
outside our scope though\... We have also combined SSVEP with source
imaging techniques to probe responses in different cortical locations
(again, Ales papers, some early stuff from Stan K? Appelbaum, Wade,
Norcia figure/ground\...) and then a host of later work from that lab
and others.

[]{#anchor-6}4. SSVEP and contrast gain control including adaptation,
masking, and attention

\[The story goes: Up to the lte 90s people were primarily interested in
measuring contrast sensitivity - the shape of the response function was
assumed to be basically log-linear - and they fit it with straight lines
to extrapolate back to zero response. But then (once Heeger's 1992 paper
had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley &
Victor, 1981)](https://www.zotero.org/google-docs/?zAmtgj)), people
started thinking about gain control - Candy and Norcia in about 1999,
Porciatti, probably a load of Tyler papers that I don't even know
about\...\]. Probably look in p

And then people worked out that if you can use SSVEP to measure contrast
responses, you can also use it to measure things that modulate contrast
responses. These include adaptation, masking, suppression, attention
(feature and space), clinical things.

[]{#anchor-7}Adaptation:

Continuous exposure to high-contrast patterns reduces contrast
sensitivity, which can be measured using SSVEP (Ross et al., 1989).
Others? Baker recent gain control paper is perhaps worth mentioning here
as a 'confound' of sorts. Engel 2018 [*(Vergeer et al.,
2018)*](https://www.zotero.org/google-docs/?kdURiS). This paper is
interesting [*(Rideaux et al.,
2023)*](https://www.zotero.org/google-docs/?J2EO3d) and \>sort\< of
SSVEP.

[]{#anchor-8}Masking:

High contrast masks can suppress the visibility of low contrast
patterns, which has implications in SSVEP amplitude (Haynes et al.,
2003).

Attention: Directing attention can enhance contrast sensitivity, as
shown in studies using SSVEP (Müller et al., 2006). Also Tsai
(dynamics), Baker / Wade (several), Winawer? I think JW has a nice
dynamic model of normalization with some MEG data. Busse et al cat/human
comparison. Candy and Norcia 2001 JNS [(Candy et al.,
2001)](https://www.zotero.org/google-docs/?7hhs5Z)

Ross, J., Speed, H. D., & Morgan, M. J. (1989). The effects of
adaptation and masking on incremental thresholds for contrast. Vision
research, 29(2), 205-215.

Haynes, J. D., Roth, G., Stadler, M., & Heinze, H. J. (2003).
Neuromagnetic correlates of perceived contrast in primary visual cortex.
Journal of Neurophysiology, 89(6), 2655-2666.

Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
Teder-Sälejärvi, W. A., & Hillyard, S. A. (2006). Effects of spatial
selective attention on the steady-state visual evoked potential in the
20--28 Hz range. Cognitive Brain Research, 24(1), 1-13.

5\. Clinical implications

[]{#anchor-9}Clinical applications (this is a whole section)

Not sure exactly how SSVEP used in clinic. mfVEP?

, e.g., in monitoring visual impairments, tracking neuronal diseases, or
neurofeedback (Norcia et al., 2015).

Tsai epilepsy [(Tsai et al.,
2011)](https://www.zotero.org/google-docs/?AdhH3f). Other photosensitive
epilepsy: Porciatti [(Porciatti et al.,
2000)](https://www.zotero.org/google-docs/?UTBQbV)

Citation:

Migraine (Regan et al). Autism?

Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., &
Rossion, B. (2015). The steady-state visual evoked potential in vision
research: A review. Journal of vision, 15(6), 4-4.

Animal work: Flies (Eliott, West, Himmelberg, Ales, Norcia): Mice/rats
(probably many - can't think off the top of my head - we were working
with a mouse EEG person at UCSF about 12 years ago\...), Monkeys
(Kiorpes?),

[]{#anchor-10}6. Future directions

Use as a readout of modulations. TMS? FUS? Marmite B12 / Fluoxetine /
amblyopia in Rats,

GABA Huang

There is a lot of SSVEP interest these days because of BCIs. I think
it's pretty weak but there is \>so much\< of it that it might be worth
mentioning\...

Advanced signal processing techniques and machine learning can be
integrated to improve SSVEP-based systems (Zhu et al., 2010).

Exploring new clinical and diagnostic applications, understanding
neurological diseases, and developing novel therapeutic interventions.

Citation:

Zhu, D., Bieger, J., Molina, G. G., & Aarts, R. M. (2010). A survey of
stimulation methods used in SSVEP-based BCIs. Computational intelligence
and neuroscience, 2010.

### 2024-03-19 review.qmd
Adding biblio
Adding contents of review_files folder.Perhaps ignore?

@@
 format: html
+bibliography: WadeBaker_Review.bib
 ---
@@
 *4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*
-
+[@porciatti_lack_2000][@tsai_contrast_2011]
 *5: Future directions (decoding in frequency domain?, animals? BCI?)*

### 2024-03-19 review.qmd
Started writing!

@@
 
+Neurons in the visual areas of the brain are primarily responsive to differences between light and dark in an image.
+This property, referred to as contrast, sets the fundamental limits of our visual abilities, which remain steady over a remarkably wide range of environmental light levels.
+The human response to contrast can be studied using many different techniques.
+Early work used psychophysical methods to measure contrast sensitivity (Campbell & Green, 1965), defined as the inverse of the lowest contrast that can be reliably detected.
+But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG.
+Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.
+
+The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency (Regan, 1966).
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal.
+The flickering stimulus entrains neural responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
+Two common variants involve sinusoidal on-off flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
+On-off flicker produces a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on.
+Counterphase flicker does not produce a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
+Variants involving square wave flicker can additionally produce more complex spectral components.
+The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system (Regan & Regan, 1988).
+SSVEP signals can also be elicited by periodic changes of stimulus properties other than contrast, such as chromaticity, motion, stereo depth, and facial identity or expression (see Norcia et al., 2015, for an overview); however our focus here is on the contrast response.
+
 Figures:
@@
 ```{python}
+
 import matplotlib.pyplot as plt

### 2024-03-19 review.qmd
Fix issue with ampersands

@@
 The human response to contrast can be studied using many different techniques.
-Early work used psychophysical methods to measure contrast sensitivity (Campbell & Green, 1965), defined as the inverse of the lowest contrast that can be reliably detected.
+Early work used psychophysical methods to measure contrast sensitivity (Campbell \& Green, 1965), defined as the inverse of the lowest contrast that can be reliably detected.
 But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG.
@@
 Variants involving square wave flicker can additionally produce more complex spectral components.
-The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system (Regan & Regan, 1988).
+The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system (Regan \& Regan, 1988).
 SSVEP signals can also be elicited by periodic changes of stimulus properties other than contrast, such as chromaticity, motion, stereo depth, and facial identity or expression (see Norcia et al., 2015, for an overview); however our focus here is on the contrast response.
@@
 Contrast Sensitivity: The ability to detect differences in luminance
-between an object and its background (Campbell & Green, 1965).
+between an object and its background (Campbell \& Green, 1965).
 
@@
 
-Campbell, F. W., & Green, D. G. (1965). Optical and retinal factors
+Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
 affecting visual resolution. J Physiology, 181, 576-593.
@@
 86,88,90 - mentioned in Regan's nice autobiography :
-https://www.google.co.uk/books/edition/Seeing_Spatial_Form/HqaKlR_GpP0C?hl=en&gbpv=1&dq=Analysis+of+EEG+responses+in+man,+evoked+by+sine+watre+modulated+light.+Thesis&pg=PA215&printsec=frontcover
 
@@
 Di Russo, F., Pitzalis, S., Spitoni, G., Aprile, T., Patria, F., Stella,
-A., \... & Hillyard, S. A. (2007). Identification of the neural sources
+A., \... \& Hillyard, S. A. (2007). Identification of the neural sources
 of the pattern-reversal VEP. Neuroimage, 34(1), 177-189.
@@
 to extrapolate back to zero response. But then (once Heeger's 1992 paper
-had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley &
+had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley \&
 Victor, 1981)](https://www.zotero.org/google-docs/?zAmtgj)), people
@@
 
-Ross, J., Speed, H. D., & Morgan, M. J. (1989). The effects of
+Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
 adaptation and masking on incremental thresholds for contrast. Vision
@@
 
-Haynes, J. D., Roth, G., Stadler, M., & Heinze, H. J. (2003).
+Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
 Neuromagnetic correlates of perceived contrast in primary visual cortex.
@@
 Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
-Teder-Sälejärvi, W. A., & Hillyard, S. A. (2006). Effects of spatial
+Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
 selective attention on the steady-state visual evoked potential in the
@@
 
-Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., &
+Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
 Rossion, B. (2015). The steady-state visual evoked potential in vision
@@
 
-Zhu, D., Bieger, J., Molina, G. G., & Aarts, R. M. (2010). A survey of
+Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
 stimulation methods used in SSVEP-based BCIs. Computational intelligence

### 2024-03-19 review.qmd
Testing references
Regularised cite keys in .bib file

@@
 The human response to contrast can be studied using many different techniques.
-Early work used psychophysical methods to measure contrast sensitivity (Campbell \& Green, 1965), defined as the inverse of the lowest contrast that can be reliably detected.
+Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965], defined as the inverse of the lowest contrast that can be reliably detected.
 But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG.
@@
 
-The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency (Regan, 1966).
+The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
 For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal.
@@
 Variants involving square wave flicker can additionally produce more complex spectral components.
-The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system (Regan \& Regan, 1988).
-SSVEP signals can also be elicited by periodic changes of stimulus properties other than contrast, such as chromaticity, motion, stereo depth, and facial identity or expression (see Norcia et al., 2015, for an overview); however our focus here is on the contrast response.
+The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system [@Regan1988].
+SSVEP signals can also be elicited by periodic changes of stimulus properties other than contrast, such as chromaticity, motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
 

### 2024-03-20 review.qmd
Writing

@@
 SSVEP signals can also be elicited by periodic changes of stimulus properties other than contrast, such as chromaticity, motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
+
+In early work on infants, the ‘sweep VEP’ paradigm was developed to measure SSVEP activity in response to a stimulus that changes its properties throughout a trial, for example gradually increasing in contrast or spatial frequency.
+Other studies keep the stimulus properties constant for the entire trial, which avoids serial effects and makes the analysis more straightforward.
+
+Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction.
+By parametrically varying the peak stimulus contrast, a monotonically increasing contrast response function can be measured.
+This corresponds closely to similar functions measured in single unit studies, or those measuring local field potentials invasively in the cortex.
+However SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
+An early use of the contrast response function was to provide an automated estimate of contrast sensitivity, without requiring behavioural responses from the participant.
+This was achieved by measuring the contrast response function, and extrapolating back along the function to estimate its intercept with the x-axis, which was shown to correspond with psychophysically measured detection thresholds (REFS).
+This general approach revealed much about the development of visual abilities in infants (Braddick), as well as deficits in neurological disease.
+However, psychophysical measurements remain the gold standard, and do not always correspond closely with SSVEP estimates in individuals with normal adult visual systems.
 

### 2024-03-20 review.qmd
Added R code , anstract

--- 2024-03-19 review.qmd
@@
 
-Neurons in the visual areas of the brain are primarily responsive to differences between light and dark in an image.
-This property, referred to as contrast, sets the fundamental limits of our visual abilities, which remain steady over a remarkably wide range of environmental light levels.
-The human response to contrast can be studied using many different techniques.
-Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965], defined as the inverse of the lowest contrast that can be reliably detected.
-But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG.
-Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.
+## ABSTRACT:
 
-The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal.
-The flickering stimulus entrains neural responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
-Two common variants involve sinusoidal on-off flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
-On-off flicker produces a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on.
-Counterphase flicker does not produce a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
-Variants involving square wave flicker can additionally produce more complex spectral components.
-The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system [@Regan1988].
-SSVEP signals can also be elicited by periodic changes of stimulus properties other than contrast, such as chromaticity, motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
+Cone photoreceptor contrast modulations are the currency of the early visual system.
+Measuring the way that sensitivity to contrast depends on factors such as spatial and temporal frequency, age, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
+One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
+SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high SNR and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
+Here we describe how SSVEPs have been used to study visual contrast over the past 70 years[@dawson_summation_1954].
+Because our thinking about SSVEPs is best described in simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.
+This document can therefore be used both as a review of the use of SSVEP in measuring human contrast processing, and as an interactive learning aid.
 
@@
 ```{python}
-
 import matplotlib.pyplot as plt
@@
 
-inputFrequencies_Hz=np.array([5,7])
+inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
 
-duration_S=1
+duration_S=1 # All variables in this code have an _[units] where possible
 
@@
 plt.tight_layout()
+plt.show()
+```
 
+It is relatively easy to write the same thing in R.
+This is a 'machine translation' using Anthropic's Claude3 LLM.
+You need to modify your R environment to include the reticulate package (and probably some others...
+rmarkdown for one?)
+
+Note that once you have this working, you can pass data back and fore between R and python!
+
+```{r}
+library(ggplot2)
+
+inputFrequencies_Hz <- c(5, 7)
+duration_S <- 1
+digitisationRate_Hz <- 1000
+
+support_S <- seq(0, duration_S, length.out = duration_S * digitisationRate_Hz)
+
+inputModulations <- matrix(0, nrow = length(support_S), ncol = length(inputFrequencies_Hz))
+for (i in seq_along(inputFrequencies_Hz)) {
+  freq <- inputFrequencies_Hz[i]
+  inputModulations[, i] <- sin(2 * pi * freq * support_S)
+}
+
+# Create a data frame for plotting
+data <- data.frame(
+  Time = rep(support_S, length(inputFrequencies_Hz)),
+  Amplitude = c(inputModulations),
+  Frequency = factor(rep(inputFrequencies_Hz, each = length(support_S)))
+)
+
+# Create the plot using ggplot2
+plot <- ggplot(data, aes(x = Time, y = Amplitude)) +
+  geom_line() +
+  facet_wrap(~ Frequency, ncol = 1) +
+  scale_x_continuous(name = "Time (s)", limits = c(0, duration_S)) +
+  scale_y_continuous(name = "Amplitude") +
+  ggtitle("Input Modulations") +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+print(plot)
 ```
@@
 
-[]{#anchor-1}ABSTRACT:
-
 []{#anchor-2}History
@@
 Contrast Sensitivity: The ability to detect differences in luminance
-between an object and its background (Campbell \& Green, 1965).
+between an object and its background (Campbell & Green, 1965).
 
@@
 
-Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
+Campbell, F. W., & Green, D. G. (1965). Optical and retinal factors
 affecting visual resolution. J Physiology, 181, 576-593.
@@
 86,88,90 - mentioned in Regan's nice autobiography :
+https://www.google.co.uk/books/edition/Seeing_Spatial_Form/HqaKlR_GpP0C?hl=en&gbpv=1&dq=Analysis+of+EEG+responses+in+man,+evoked+by+sine+watre+modulated+light.+Thesis&pg=PA215&printsec=frontcover
 
@@
 Di Russo, F., Pitzalis, S., Spitoni, G., Aprile, T., Patria, F., Stella,
-A., \... \& Hillyard, S. A. (2007). Identification of the neural sources
+A., \... & Hillyard, S. A. (2007). Identification of the neural sources
 of the pattern-reversal VEP. Neuroimage, 34(1), 177-189.
@@
 to extrapolate back to zero response. But then (once Heeger's 1992 paper
-had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley \&
+had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley &
 Victor, 1981)](https://www.zotero.org/google-docs/?zAmtgj)), people
@@
 
-Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
+Ross, J., Speed, H. D., & Morgan, M. J. (1989). The effects of
 adaptation and masking on incremental thresholds for contrast. Vision
@@
 
-Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
+Haynes, J. D., Roth, G., Stadler, M., & Heinze, H. J. (2003).
 Neuromagnetic correlates of perceived contrast in primary visual cortex.
@@
 Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
-Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
+Teder-Sälejärvi, W. A., & Hillyard, S. A. (2006). Effects of spatial
 selective attention on the steady-state visual evoked potential in the
@@
 
-Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
+Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., &
 Rossion, B. (2015). The steady-state visual evoked potential in vision
@@
 
-Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
+Zhu, D., Bieger, J., Molina, G. G., & Aarts, R. M. (2010). A survey of
 stimulation methods used in SSVEP-based BCIs. Computational intelligence

### 2024-03-20 review.qmd

--- 2024-03-20 review.qmd
@@
 
+## ABSTRACT:
+
+Cone photoreceptor contrast modulations are the currency of the early visual system.
+Measuring the way that sensitivity to contrast depends on factors such as spatial and temporal frequency, age, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
+One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
+SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high SNR and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
+Here we describe how SSVEPs have been used to study visual contrast over the past 70 years[@dawson_summation_1954].
+Because our thinking about SSVEPs is best described in simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.
+This document can therefore be used both as a review of the use of SSVEP in measuring human contrast processing, and as an interactive learning aid.
+
+## INTRODUCTION:
 Neurons in the visual areas of the brain are primarily responsive to differences between light and dark in an image.
@@
 *5: Future directions (decoding in frequency domain?, animals? BCI?)*
-
-[]{#anchor-1}ABSTRACT:
 

### 2024-03-20 review.qmd
Minor tinkering

@@
 
-*1: a) A typical SSVEP setup: a screen, a grating, a flicker rate: b)
-The responses from a single electrode (time domain) c) Frequency
-domain.*
-
 ```{python}
 
+#| label: plotsinewaves
+#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain"
+#| fig-cap-location: bottom
+#| 
 import matplotlib.pyplot as plt
@@
 ```{r}
+
 library(ggplot2)
@@
 print(plot)
+
 ```
@@
 ```{python}
+
+#| label: plotspectra
+#| fig-cap: "a"
+#| fig-cap-location: bottom
 
@@
 *4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*
-[@porciatti_lack_2000][@tsai_contrast_2011]
+[@Porciatti2000; @Tsai2011]
 *5: Future directions (decoding in frequency domain?, animals? BCI?)*

### 2024-03-20 review.qmd

--- 2024-03-20 review.qmd
@@
 SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high SNR and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
-Here we describe how SSVEPs have been used to study visual contrast over the past 70 years[@dawson_summation_1954].
+Here we describe how SSVEPs have been used to study visual contrast over the past 70 years[@Dawson1954].
 Because our thinking about SSVEPs is best described in simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.
@@
 *4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*
-[@porciatti_lack_2000][@tsai_contrast_2011]
+[@Porciatti2000][@tsai_contrast_2011]
 *5: Future directions (decoding in frequency domain?, animals? BCI?)*

### 2024-03-27 review.qmd
Added ,some motivation

--- 2024-03-20 review.qmd
@@
 
+### Why measure responses to contrast?
+Contrast is one of the most fundamental pieces of information that the eye transmits to the brain.
+It can be defined as the change in cone photoreceptor activity over time ('spatial contrast') or space ('spatial contrast').
+Cone photoreceptors contribute to both chromatic and achromatic contrast and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well.
+[McKeefry, Baseler, DiRusso...].
+
+Contrast is relatively simple to define: typically, contrast is specificed as the percentage deviation of uniform stimulus from the background.
+So, for example, a disk of 100 units, $I_{\mathrm{stim}}$ of cone activation surrounded by a 'background' of 50 units of activation $I_{\mathrm{background}}$ has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
+Where patterns are more complex (for example, the sine-wave gratings or Gabors common in vision science), the Michaelson definition of contrast is specified by the maximum and minimum excursions from the mean: $${\displaystyle {\frac {I_{\mathrm {stimmax} }-I_{\mathrm {stimmin} }}{I_{\mathrm {stimmax} }+I_{\mathrm {stimmin} }}},} $$.
+These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance [LenniePokornySmith]) and also to definitions based on cone excitations [DKL, MacBoynton] which are more common in work on chromatic processing.
+
+Although its definition is simple, the computations that underlie contrast processing have been the subject of intense research for many decades.
+The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
+This normalization, achieved through a computation called 'contrast gain control' [Heeger 92, Heeger and Carandini 2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
+As an example, a grating seen on a low-contrast background tpyically appears more intense than the same grating when superimposed on a high contrast background.
+A signifcant part of research into contrast processing is concerned with how these normalization mechanisms depend on colour, orientation, eye of origin, spatial and temporal frequency, age, and the presence of neurological disorders.
+The SSVEP has proven to be invaliable in this research because it provides an objective readout of contrast representation at different parts of the visual system and allows us to 'tag' the probe and background separately.
+
+### The development of contrast processing
 In early work on infants, the ‘sweep VEP’ paradigm was developed to measure SSVEP activity in response to a stimulus that changes its properties throughout a trial, for example gradually increasing in contrast or spatial frequency.
@@
 Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction.
-By parametrically varying the peak stimulus contrast, a monotonically increasing contrast response function can be measured.
+By varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency or a low multiple thereof.
 This corresponds closely to similar functions measured in single unit studies, or those measuring local field potentials invasively in the cortex.
@@
 This general approach revealed much about the development of visual abilities in infants (Braddick), as well as deficits in neurological disease.
-However, psychophysical measurements remain the gold standard, and do not always correspond closely with SSVEP estimates in individuals with normal adult visual systems.
+In well-motivated adults, psychophysical measurements remain the gold standard.
+However, it is hard to obtain reliable psychophysical data from infants, children and adults with neuological disorders.
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses.
 

### 2024-03-27 review.qmd
Figure to show response compression

@@
 ### The development of contrast processing
-In early work on infants, the ‘sweep VEP’ paradigm was developed to measure SSVEP activity in response to a stimulus that changes its properties throughout a trial, for example gradually increasing in contrast or spatial frequency.
-Other studies keep the stimulus properties constant for the entire trial, which avoids serial effects and makes the analysis more straightforward.
-
 Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction.
@@
 In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses.
+
+```{python}
+
+```
+
+In early work on infants, the ‘sweep VEP’ paradigm was developed to measure SSVEP activity in response to a stimulus that changes its properties throughout a trial, for example gradually increasing in contrast or spatial frequency.
+Other studies keep the stimulus properties constant for the entire trial, which avoids serial effects and makes the analysis more straightforward.
 
@@
 
-It is relatively easy to write the same thing in R.
-This is a 'machine translation' using Anthropic's Claude3 LLM.
-You need to modify your R environment to include the reticulate package (and probably some others...
-rmarkdown for one?)
-
-Note that once you have this working, you can pass data back and fore between R and python!
-
-```{r}
-
-library(ggplot2)
-
-inputFrequencies_Hz <- c(5, 7)
-duration_S <- 1
-digitisationRate_Hz <- 1000
-
-support_S <- seq(0, duration_S, length.out = duration_S * digitisationRate_Hz)
-
-inputModulations <- matrix(0, nrow = length(support_S), ncol = length(inputFrequencies_Hz))
-for (i in seq_along(inputFrequencies_Hz)) {
-  freq <- inputFrequencies_Hz[i]
-  inputModulations[, i] <- sin(2 * pi * freq * support_S)
-}
-
-# Create a data frame for plotting
-data <- data.frame(
-  Time = rep(support_S, length(inputFrequencies_Hz)),
-  Amplitude = c(inputModulations),
-  Frequency = factor(rep(inputFrequencies_Hz, each = length(support_S)))
-)
-
-# Create the plot using ggplot2
-plot <- ggplot(data, aes(x = Time, y = Amplitude)) +
-  geom_line() +
-  facet_wrap(~ Frequency, ncol = 1) +
-  scale_x_continuous(name = "Time (s)", limits = c(0, duration_S)) +
-  scale_y_continuous(name = "Amplitude") +
-  ggtitle("Input Modulations") +
-  theme_bw() +
-  theme(plot.title = element_text(hjust = 0.5))
-
-print(plot)
-
-```
-
 The two inputs are represented here as amplitude modulations across time.
@@
 #| fig-cap-location: bottom
+
+#| label: plotsinewaves
+#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain"
+#| fig-cap-location: bottom
+#| 
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy as scp
+
+inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
+
+duration_S=1 # All variables in this code have an _[units] where possible
+
+digitisationRate_Hz=1000
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))
+for i, freq in enumerate(inputFrequencies_Hz):
+    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)
 
@@
 
+plt.tight_layout()
+plt.show()
+
+```
+
+```{python}
+
+#| label: ContrastResponseFunction
+#| fig-cap: "a"
+#| fig-cap-location: bottom
+
+#| label: plotsinewaves
+#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain"
+#| fig-cap-location: bottom
+#| 
+import matplotlib.pyplot as plt
+from matplotlib import colormaps
+import numpy as np
+import scipy as scp
+
+def simpleTransducerFunction(inputContrast, c50, RMax):
+ # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
+ # RMax defines the reponse gain of the transducer (how big the biggest output is). 
+ # c50 defines the conrast sensitivity. Smaller values are more sensitive.
+ # The exponent can also vary - here we fix it to 2
+  expnt=2.5
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
+  return output
+
+# Define some constants
+inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
+contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal=[]
+
+fig, axs = plt.subplots(1,3, figsize=(8, 3))
+
+initialSignal = np.zeros(len(contrastRange))
+transducedSignal = np.zeros(len(contrastRange))
+for contrastIndex,thisContrast in enumerate(contrastRange):
+  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
+
+  axs[0].plot(support_S, inputModulation)
+
+  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+
+  responseAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
+  initialSignal[contrastIndex]=responseAmplitude[0]
+  transducedSignal[contrastIndex]=simpleTransducerFunction(responseAmplitude[0], 0.05, 10)
+
+axs[1].plot(contrastRange,initialSignal)
+axs[1].set_xlabel('Contrast')
+axs[1].set_ylabel('Amplitude (a.u.)')
+axs[1].set_title('Linear computation stage')
+axs[1].grid()
+
+axs[2].plot(contrastRange,transducedSignal)
+axs[2].set_xlabel('Contrast')
+axs[2].set_ylabel('Amplitude (a.u.)')
+axs[2].set_title('After nonlinear transducer')
+
+axs[2].grid()
 plt.tight_layout()

### 2024-03-27 review.qmd
Turned off code echo
Single word

@@
 bibliography: WadeBaker_Review.bib
+execute:
+  echo: false
 ---
@@
 
-Cone photoreceptor contrast modulations are the currency of the early visual system.
+Contrast is the currency of the early visual system.
 Measuring the way that sensitivity to contrast depends on factors such as spatial and temporal frequency, age, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.

### 2024-03-27 review.qmd
Added sweep VEP figure

@@
 Contrast is the currency of the early visual system.
-Measuring the way that sensitivity to contrast depends on factors such as spatial and temporal frequency, age, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
+Measuring the way that the computations underlying contrast processing depend on factors such as spatial and temporal frequency, age, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
 One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
@@
 ## INTRODUCTION:
-Neurons in the visual areas of the brain are primarily responsive to differences between light and dark in an image.
+Neurons in the visual areas of the brain are primarily responsive to changes in cone photoreceptor activations across time and space.
 This property, referred to as contrast, sets the fundamental limits of our visual abilities, which remain steady over a remarkably wide range of environmental light levels.
@@
 Cone photoreceptors contribute to both chromatic and achromatic contrast and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well.
-[McKeefry, Baseler, DiRusso...].
+[McKeefry, Sutter, Baseler, DiRusso...].
 
@@
 So, for example, a disk of 100 units, $I_{\mathrm{stim}}$ of cone activation surrounded by a 'background' of 50 units of activation $I_{\mathrm{background}}$ has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
-Where patterns are more complex (for example, the sine-wave gratings or Gabors common in vision science), the Michaelson definition of contrast is specified by the maximum and minimum excursions from the mean: $${\displaystyle {\frac {I_{\mathrm {stimmax} }-I_{\mathrm {stimmin} }}{I_{\mathrm {stimmax} }+I_{\mathrm {stimmin} }}},} $$.
+Where patterns are more complex (for example, the sine-wave gratings or Gabor patches common in vision science), the Michaelson definition of contrast is specified by the maximum and minimum excursions from the mean: $${{\frac {I_{\mathrm {stimmax} }-I_{\mathrm {stimmin} }}{I_{\mathrm {stimmax} }+I_{\mathrm {stimmin} }}},} $$.
 These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance [LenniePokornySmith]) and also to definitions based on cone excitations [DKL, MacBoynton] which are more common in work on chromatic processing.
 
-Although its definition is simple, the computations that underlie contrast processing have been the subject of intense research for many decades.
+Although its definition as a stimulus is simple, the computations that underlie contrast processing in the brain have been the subject of intense research for many decades.
 The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
 This normalization, achieved through a computation called 'contrast gain control' [Heeger 92, Heeger and Carandini 2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
-As an example, a grating seen on a low-contrast background tpyically appears more intense than the same grating when superimposed on a high contrast background.
+As an example, a grating placed at the centre of a low-contrast background tpyically appears more intense than the same grating when superimposed on a high contrast background.
+```{python}
+import numpy as np
+import matplotlib.pyplot as plt
+
+def create_grating(size, frequency, phase, contrast):
+    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
+    grating = np.sin(2 * np.pi * frequency * np.sqrt(x**2 + y**2) + phase)
+    grating = (grating + 1) / 2  # Normalize to [0, 1]
+    grating = grating * contrast + (1 - contrast) / 2  # Adjust contrast
+    return grating
+
+def create_circular_mask(size):
+    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
+    mask = np.sqrt(x**2 + y**2) <= 0.3
+    return mask
+
+# Set parameters
+size = 256
+frequency = 10
+phase = 0
+contrast_center = 0.5
+contrast_surround_low = 0.1
+contrast_surround_high = 0.9
+
+# Create gratings and circular masks
+grating_center = create_grating(size, frequency, phase, contrast_center)
+grating_surround_low = create_grating(size, frequency, phase, contrast_surround_low)
+grating_surround_high = create_grating(size, frequency, phase, contrast_surround_high)
+mask = create_circular_mask(size)
+
+# Combine gratings and masks
+image_low_contrast = grating_surround_low.copy()
+image_low_contrast[mask] = grating_center[mask]
+image_high_contrast = grating_surround_high.copy()
+image_high_contrast[mask] = grating_center[mask]
+
+# Create subplots
+fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(7, 3))
+
+# Plot low-contrast background
+ax1.imshow(image_low_contrast, cmap='gray', vmin=0, vmax=1)
+ax1.axis('off')
+ax1.set_title('Low-Contrast Background', )
+
+# Plot high-contrast background
+ax2.imshow(image_high_contrast, cmap='gray', vmin=0, vmax=1)
+ax2.axis('off')
+ax2.set_title('High-Contrast Background')
+
+plt.tight_layout()
+plt.show()
+
+```
+
  A signifcant part of research into contrast processing is concerned with how these normalization mechanisms depend on colour, orientation, eye of origin, spatial and temporal frequency, age, and the presence of neurological disorders.
@@
 However SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
+
 An early use of the contrast response function was to provide an automated estimate of contrast sensitivity, without requiring behavioural responses from the participant.
-This was achieved by measuring the contrast response function, and extrapolating back along the function to estimate its intercept with the x-axis, which was shown to correspond with psychophysically measured detection thresholds (REFS).
+Because SSVEP response at detection threshold are very small, this was achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function to estimate its intercept with the x-axis, which was shown to correspond with psychophysically measured detection thresholds (REFS).
 This general approach revealed much about the development of visual abilities in infants (Braddick), as well as deficits in neurological disease.
+
 In well-motivated adults, psychophysical measurements remain the gold standard.
 However, it is hard to obtain reliable psychophysical data from infants, children and adults with neuological disorders.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses.
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick].
+
+Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changes its contrast (or some other property) throughout a trial - for example gradually increasing in contrast or spatial frequency.
+To avoid hysteresis effects, the sweep was sometimes conducted both up and down in the same experiment [Tyler, Norcia etc].
 
@@
 
+#| label: SweepVEP Deteection threshold estimate
+#| fig-cap: "a"
+#| fig-cap-location: bottom
+
+import matplotlib.pyplot as plt
+from matplotlib import colormaps
+import numpy as np
+import scipy as scp
+from sklearn.linear_model import LinearRegression
+
+def simpleTransducerFunction(inputContrast, c50, RMax):
+ # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
+ # RMax defines the reponse gain of the transducer (how big the biggest output is). 
+ # c50 defines the conrast sensitivity. Smaller values are more sensitive.
+ # The exponent can also vary - here we fix it to 2
+  expnt=2
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
+  return output
+
+# Define some constants
+nLevels=8
+internalNoise=1
+inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
+contrastRange=np.logspace(1,1.9,nLevels)/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal=[]
+
+fig, axs = plt.subplots(1,1, figsize=(4, 3))
+
+initialSignal = np.zeros(len(contrastRange))
+transducedSignal = np.zeros(len(contrastRange))
+for contrastIndex,thisContrast in enumerate(contrastRange):
+  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast + np.random.rand(len(support_S))*internalNoise
+
+  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+
+  responseAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
+  initialSignal[contrastIndex]=responseAmplitude[0]
+  transducedSignal[contrastIndex]=simpleTransducerFunction(responseAmplitude[0], 0.05, 10)
+
+# Now fit a linear regression to the first 4 contrast levels (before it saturates) and find out where this crosses the Y-axis
+X = np.log10(contrastRange[:4]).reshape(-1, 1)
+y = transducedSignal[:4]
+reg = LinearRegression().fit(X, y)
+x_intercept = 10 ** (-reg.intercept_ / reg.coef_[0])
+
+axs.set_title(f'Simulated sweep VEP at {nLevels} different contrasts', fontsize=10)
+axs.semilogx(contrastRange, transducedSignal, 'o')
+axs.semilogx(contrastRange[:4], reg.predict(X), 'r--', label='Linear Fit')
+axs.semilogx([x_intercept], [0], 'ro', label='Threshold')
+axs.set_xlabel('Contrast')
+axs.set_ylabel('Amplitude (a.u.)')
+axs.set_xlim(0.01, 1)
+axs.set_ylim(0, 10)
+axs.legend()
+axs.grid()
+plt.tight_layout()
+plt.show()
+
 ```
-
-In early work on infants, the ‘sweep VEP’ paradigm was developed to measure SSVEP activity in response to a stimulus that changes its properties throughout a trial, for example gradually increasing in contrast or spatial frequency.
-Other studies keep the stimulus properties constant for the entire trial, which avoids serial effects and makes the analysis more straightforward.
 
@@
   transducedSignal[contrastIndex]=simpleTransducerFunction(responseAmplitude[0], 0.05, 10)
+axs[0].set_title('Sine-wave responses at\n different contrast levels', fontsize=10)
 
@@
 axs[1].set_ylabel('Amplitude (a.u.)')
-axs[1].set_title('Linear computation stage')
+axs[1].set_title('Linear computation stage', fontsize=10)
 axs[1].grid()
@@
 axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('After nonlinear transducer')
+axs[2].set_title('After nonlinear transducer', fontsize=10)
 

### 2024-03-29 review.qmd
tidied
Some edits on square / sine / harmonics

@@
 Here we describe how SSVEPs have been used to study visual contrast over the past 70 years[@Dawson1954].
-Because our thinking about SSVEPs is best described in simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.
+Because our thinking about SSVEPs is well-described by simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.
 This document can therefore be used both as a review of the use of SSVEP in measuring human contrast processing, and as an interactive learning aid.
@@
 Two common variants involve sinusoidal on-off flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
-On-off flicker produces a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on.
-Counterphase flicker does not produce a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
-Variants involving square wave flicker can additionally produce more complex spectral components.
+On-off flicker can drive independent populations of on- and off-cells in the retina once per cycle and can therefore produce a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on.
+Counterphase flicker contains two transients per cycle and therefore does not produce a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
+Because square-waves are spectrally broad-band, square wave flicker tends to produce more complex spectral harmonics than sine-wave flicker.
+
 The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system [@Regan1988].
-SSVEP signals can also be elicited by periodic changes of stimulus properties other than contrast, such as chromaticity, motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
+SSVEP signals can also be elicited by periodic changes of stimulus properties other than achromatic and chromatic contrast, such as, motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
 
@@
 It can be defined as the change in cone photoreceptor activity over time ('spatial contrast') or space ('spatial contrast').
-Cone photoreceptors contribute to both chromatic and achromatic contrast and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well.
+Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well.
 [McKeefry, Sutter, Baseler, DiRusso...].
 
-Contrast is relatively simple to define: typically, contrast is specificed as the percentage deviation of uniform stimulus from the background.
+Contrast is relatively simple to define: typically, contrast is specified as the percentage deviation of uniform stimulus from the background.
 So, for example, a disk of 100 units, $I_{\mathrm{stim}}$ of cone activation surrounded by a 'background' of 50 units of activation $I_{\mathrm{background}}$ has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
@@
 support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
-transducedSignal=[]
 

### 2024-03-29 review.qmd
Adding stuff about sweep

@@
 The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal.
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di Russo, Hillebrand:].
 The flickering stimulus entrains neural responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
@@
 
- A signifcant part of research into contrast processing is concerned with how these normalization mechanisms depend on colour, orientation, eye of origin, spatial and temporal frequency, age, and the presence of neurological disorders.
+ A signifcant part of research into contrast processing is concerned with how these normalization mechanisms depend on colour, orientation, eye of origin, spatial and temporal frequency, age, and the presence of neurological disorders [refs: Chen and Foley, Petrov etc Carandini, Binocular normalization / suppression (Baker/Meese but also e.g.
+Heeger), Tadin, Winawer, Wade, Porciatti, Tsai...].
 The SSVEP has proven to be invaliable in this research because it provides an objective readout of contrast representation at different parts of the visual system and allows us to 'tag' the probe and background separately.
@@
 Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction.
-By varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency or a low multiple thereof.
-This corresponds closely to similar functions measured in single unit studies, or those measuring local field potentials invasively in the cortex.
-However SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
+For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency or a low multiple thereof.
+This corresponds closely to similar functions measured in single unit studies, or those measuring local field potentials invasively in the cortex[@ShapleyVictor etc].
+However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
 
 An early use of the contrast response function was to provide an automated estimate of contrast sensitivity, without requiring behavioural responses from the participant.
-Because SSVEP response at detection threshold are very small, this was achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function to estimate its intercept with the x-axis, which was shown to correspond with psychophysically measured detection thresholds (REFS).
-This general approach revealed much about the development of visual abilities in infants (Braddick), as well as deficits in neurological disease.
+Because SSVEP response at detection threshold are very small, this was achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function to estimate its intercept with the x-axis, which was shown to correspond with psychophysically measured detection thresholds [(REFS)].
+This general approach revealed much about the development of visual abilities in infants [@braddick], as well as deficits in neurological disease.
 
-In well-motivated adults, psychophysical measurements remain the gold standard.
+In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
 However, it is hard to obtain reliable psychophysical data from infants, children and adults with neuological disorders.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick].
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick, Porciatti, Tsai...].
 
-Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changes its contrast (or some other property) throughout a trial - for example gradually increasing in contrast or spatial frequency.
-To avoid hysteresis effects, the sweep was sometimes conducted both up and down in the same experiment [Tyler, Norcia etc].
+Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, or some other property, throughout a trial - for example gradually increasing in spatial frequency [Tyler, Nakayama etc 1979].
+To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, Norcia etc: 1985, 1990 etc].
+
+In general, SSVEP measurements of infant vision have revealed that the development of spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception was faster than had been supposed previously based on behavioural measures [@DobsonTeller78, @NorciaTyler85, @NorciaGerhard 15]
 
@@
 ```
+
+As well as measuing contrast thresholds for a stimulus with a fixed spatiotemporal frequency, this technique can also be used to assess spatial frequency thresholds [ REFS - Almoqbel etc.].
+Because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in young subjects [Ridder2004 Diva vs Enfant, Hoffman, Bach and Farmer].
 

### 2024-03-29 review.qmd
Fixing sweep stuff more

@@
 
-### The development of contrast processing
 Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction.
@@
 
-An early use of the contrast response function was to provide an automated estimate of contrast sensitivity, without requiring behavioural responses from the participant.
-Because SSVEP response at detection threshold are very small, this was achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function to estimate its intercept with the x-axis, which was shown to correspond with psychophysically measured detection thresholds [(REFS)].
+### The development of contrast processing
+
+An early use of the contrast response function was to provide an automated estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
+
+In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
+However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick].
+
+Because SSVEP response at detection threshold are very small, this was achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis.
+This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@DaleNorciaTyler86].
+
+Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [Tyler, Nakayama etc 1979].
+To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, Norcia etc: 1985, 1990 etc].
 This general approach revealed much about the development of visual abilities in infants [@braddick], as well as deficits in neurological disease.
 
-In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
-However, it is hard to obtain reliable psychophysical data from infants, children and adults with neuological disorders.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick, Porciatti, Tsai...].
-
-Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, or some other property, throughout a trial - for example gradually increasing in spatial frequency [Tyler, Nakayama etc 1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, Norcia etc: 1985, 1990 etc].
-
-In general, SSVEP measurements of infant vision have revealed that the development of spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception was faster than had been supposed previously based on behavioural measures [@DobsonTeller78, @NorciaTyler85, @NorciaGerhard 15]
+In general, SSVEP measurements of infant vision have revealed that the development of spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception is faster than had been supposed previously based on behavioural readouts [@DobsonTeller78, @NorciaTyler85, @NorciaGerhard 15].
+At least some of this difference is likely due to the high SNR of the SSVEP technique.
 
@@
 
-As well as measuing contrast thresholds for a stimulus with a fixed spatiotemporal frequency, this technique can also be used to assess spatial frequency thresholds [ REFS - Almoqbel etc.].
-Because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in young subjects [Ridder2004 Diva vs Enfant, Hoffman, Bach and Farmer].
+The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
+In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [Ridder2004 Diva vs Enfant, Hoffman, Bach and Farmer].
+
+# Contrast processing - gain control
+Previous figures showed how a single-input SSVEP might generate a single peak in the FT response which, in turn, can be used as a read-out of neuronal activity.
+Here, we show that adding a second contrast component can generate intermodulation and that the two components interact.
+The nature of the interaction is determined by the precise computations happening as the contrast signal moves from retina to cortex.
 

### 2024-03-29 review.qmd
Zotero integration

@@
 To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, Norcia etc: 1985, 1990 etc].
-This general approach revealed much about the development of visual abilities in infants [@braddick], as well as deficits in neurological disease.
-
-In general, SSVEP measurements of infant vision have revealed that the development of spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception is faster than had been supposed previously based on behavioural readouts [@DobsonTeller78, @NorciaTyler85, @NorciaGerhard 15].
+This general approach revealed much about the development of visual abilities in infants [@braddick, @atkinson].
+In general, SSVEP measurements of infant vision have revealed that the development of spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception is faster than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ].
 At least some of this difference is likely due to the high SNR of the SSVEP technique.

### 2024-04-01 review.qmd
No message

@@
 # Contrast processing - gain control
-Previous figures showed how a single-input SSVEP might generate a single peak in the FT response which, in turn, can be used as a read-out of neuronal activity.
+Previous figures showed how a single-input SSVEP might generate a single set of peaks in the frequency domain response which, in turn, can be used as a read-out of neuronal activity.
+
 Here, we show that adding a second contrast component can generate intermodulation and that the two components interact.

### 2024-04-01 review.qmd
Adding gain control words

--- 2024-03-29 review.qmd
@@
 
-### The development of contrast processing
+### Measuring the development of contrast processing
 
-An early use of the contrast response function was to provide an automated estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
+An early use of the contrast response function was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
 
@@
 However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick].
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
 
-Because SSVEP response at detection threshold are very small, this was achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis.
+Because SSVEP response at detection threshold are very small, estimating this threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis.
 This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@DaleNorciaTyler86].
 
-Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [Tyler, Nakayama etc 1979].
+Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler, Nakayama etc 1979].
 To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, Norcia etc: 1985, 1990 etc].
-This general approach revealed much about the development of visual abilities in infants [@braddick, @atkinson].
-In general, SSVEP measurements of infant vision have revealed that the development of spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception is faster than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ].
-At least some of this difference is likely due to the high SNR of the SSVEP technique.
+
+This general approach has revealed much about the development of visual abilities in infants [@braddick, @atkinson].
+In general, SSVEP measurements of infant vision have revealed that spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+At least some of this difference is likely due to the relatively objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking which require careful measurement of the infant's gaze direction.
+One open question is how more complex contrast computations develop - and in particular how the human brain develop the tightly-tuned 'gain control' mechanisms that allow it to detect and discriminate contrast across very different visual scenes.
 
@@
 The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
-In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [Ridder2004 Diva vs Enfant, Hoffman, Bach and Farmer].
+In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004 Diva vs Enfant, @Hoffman, @Bach and @Farmer].
 
-# Contrast processing - gain control
-Previous figures showed how a single-input SSVEP might generate a single peak in the FT response which, in turn, can be used as a read-out of neuronal activity.
+# Contrast processing - linear and nonlinear
+
+Neurons have a limited dynamic range yet they can transmit information about  visual stimuli that span many order of magnitude.
+In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear,sigmoidal contrast vs response (CRF) transducer and different neurons exhibit peak sensitivity (the maximum slope of the function) at different contrast levels [@Early matteo stuff, Busse refs].
+Individual neurons at multiple stages of the visual hierarchy also change their sensitivity depending on the average spatiotemporal contrast energy of their environment.
+This normalisation process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed respnses of the local neuronal population.[@Heeger 92, @Carandini and Heeger 2011 @Busse 2008, @Baker and Wade]
+
+Previous figures showed how a single-input SSVEP might generate a single set of peaks in the frequency domain response which, in turn, can be used as a read-out of neuronal activity.
+
 Here, we show that adding a second contrast component can generate intermodulation and that the two components interact.

### 2024-04-02 review.qmd
Adding / fixing two-input descriptions - not done yet.

@@
 In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004 Diva vs Enfant, @Hoffman, @Bach and @Farmer].
+The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
+In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffman; @Bach; @Farmer].
 
@@
 Neurons have a limited dynamic range yet they can transmit information about  visual stimuli that span many order of magnitude.
-In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear,sigmoidal contrast vs response (CRF) transducer and different neurons exhibit peak sensitivity (the maximum slope of the function) at different contrast levels [@Early matteo stuff, Busse refs].
+In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear,sigmoidal contrast vs response (CRF) transducer and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Early matteo stuff, Busse refs].
+A neuronal population will therefor span a sensitivity range greater than any individual member,
+Neurons have a limited dynamic range yet they can transmit information about visual stimuli that span many orders of magnitude.
+In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear, sigmoidal contrast vs response (CRF) transducer and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@EarlyMatteoStuff; @BusseRefs].
+A neuronal population will therefore span a sensitivity range greater than any individual member.
+
  Individual neurons at multiple stages of the visual hierarchy also change their sensitivity depending on the average spatiotemporal contrast energy of their environment.
-This normalisation process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed respnses of the local neuronal population.[@Heeger 92, @Carandini and Heeger 2011 @Busse 2008, @Baker and Wade]
+This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed respnses of the local neuronal population.[@Heeger 92, @Carandini and Heeger 2011 @Busse 2008, @Baker and Wade]
+Individual neurons at multiple stages of the visual hierarchy also change their sensitivity depending on the average spatiotemporal contrast energy of their environment.
+This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed responses of the local neuronal population [@Heeger1992; @CarandiniHeeger2011; @Busse2008; @BakerWade].
 
@@
 
-Here, we show that adding a second contrast component can generate intermodulation and that the two components interact.
-The nature of the interaction is determined by the precise computations happening as the contrast signal moves from retina to cortex.
+Here, we show the effect of adding a second contrast component: in general, the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
+The nature of the interaction --- and therefore the pattern of intermodulation terms ---  is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan]
+
+Here, we show the effect of adding a second contrast component: in general, the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
+The nature of the interaction --- and therefore the pattern of intermodulation terms --- is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan].
+First, we implement a simple non-linear transducer with a single input to illustrate the typical signoidal shape of a contrast response function.
+(Figure xx)
+
+Next, we show that adding a second component which contributes to the gain pool of the first, results in a complex pattern of intermodulation terms.
 
@@
 
-If the two inputs are simply added together, the representation of the resulting signal in the Fourier domain is just the linear sum of the two independent signals:
+If the two inputs are simply added together, the representation of the resulting signal in the Fourier domain is just the linear sum of the two independent signals.
+However, it is important to take physiology into account.
+Contrast is represented as a combination of rectified signals from 'on and 'off'-sensitive cells (broadly, cells sensitive to positive or negative changes in spatiotemporal luminace).
+To model the population response to a contrast-reversing luminance sine wave, we therefore introduce an initial full-wave rectification stage that simulates the outputs of two equal populations of on- and off-polarity detectors.
+Once this is done, we see that the FT of the resulting signal contains power only at the second harmonics and above.
+It is possible to generate first-hamonic responses in contrast-response SSVEPs but only when the inputs are modulated in an 'on/off' manner rather than contrast reversed.
 
@@
 combinedSignal=np.sum(inputModulations,axis=1)
+combinedRectifiedSignal=np.sum(np.abs(inputModulations), axis=1)
 FT_combinedSignal=np.fft.fft(combinedSignal)/len(combinedSignal)
-fig,axs=plt.subplots(1,2,figsize=(8,6))
+FT_combinedRectifiedSignal=np.fft.fft(combinedRectifiedSignal)/len(combinedSignal)
+fig,axs=plt.subplots(1,4,figsize=(8,3))
 axs[0].plot(support_S,combinedSignal)
@@
 axs[0].set_ylabel('Amplitude (a.u.)')
-axs[0].set_title('Input signals with linear combination')
+axs[0].set_title('Input signals:\nlinear combination')
 
@@
 axs[1].set_ylabel('Magnitude (a.u.)')
-axs[1].set_title('FFT of Combined Signal')
+axs[1].set_title('FFT\n linear combination')
+# Plot the absolute values of the FFT in the second axis as a bar plot
+
+axs[2].plot(support_S,combinedRectifiedSignal)
+axs[2].set_xlabel('Time (s)')
+axs[2].set_ylabel('Amplitude (a.u.)')
+axs[2].set_title('Sum rectified inputs')
+
+axs[3].bar(frequencySupport, np.abs(FT_combinedRectifiedSignal[frequencySupport]))
+axs[3].set_xlabel('Frequency (Hz)')
+axs[3].set_ylabel('Magnitude (a.u.)')
+axs[3].set_title('FFT of\n sum rectified inputs')
 
@@
 
+Althoug
 ```{python}
@@
 
-#| label: plotsinewaves
-#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain"
-#| fig-cap-location: bottom
-#| 
 import matplotlib.pyplot as plt
@@
  # The exponent can also vary - here we fix it to 2
-  expnt=2.5
+  expnt=2
+
+  # Contrast cannot be negative. To model the responses of a set of on and off 
+  # contrast detectors we perform full-wave rectification of the signal
+  inputContrast=np.abs(inputContrast)
   output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
@@
 freqCutoff_Hz=30
+c50=.10
+RMax=10
 
@@
 
-  responseAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
-  initialSignal[contrastIndex]=responseAmplitude[0]
-  transducedSignal[contrastIndex]=simpleTransducerFunction(responseAmplitude[0], 0.05, 10)
+  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
+  initialSignal[contrastIndex]=inputAmplitude[0]
+
+   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
+   # using the transducer functiona above.
+  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
+  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
+  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequencies_Hz*2])
+  transducedSignal[contrastIndex]=transducedAmplitude[0]
+
 axs[0].set_title('Sine-wave responses at\n different contrast levels', fontsize=10)
@@
 axs[2].set_title('After nonlinear transducer', fontsize=10)
+
+axs[2].grid()
+plt.tight_layout()
+plt.show()
+
+```
+
+Notice the slight roll-off in the amplitude of the second harmonic at high input contrasts.
+This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
+Power at other harmonics increases and the total power generated by the input is monotonic.
+This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler]
+
+Here, we include a second component that contributes to the gain pool of the first. At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT.
+
+```{python}
+#| label: MaskedContrastResponseFunction
+#| fig-cap: "a"
+#| fig-cap-location: bottom
+
+import matplotlib.pyplot as plt
+from matplotlib import colormaps
+import numpy as np
+import scipy as scp
+
+def simpleTransducerFunction(inputContrast, maskContrast, c50, RMax):
+ # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
+ # RMax defines the reponse gain of the transducer (how big the biggest output is). 
+ # c50 defines the conrast sensitivity. Smaller values are more sensitive.
+ # The exponent can also vary - here we fix it to 2
+  expnt=2
+  inputContrast=np.abs(inputContrast)
+  maskContrast=np.abs(maskContrast)
+
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
+  return output
+
+# Define some constants
+inputFrequency_Hz=np.array([5,7]) # We simulate a system with two inputs at a single contrast
+inputContrasts=np.array([.2 ,.2])
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+c50=.10
+RMax=10
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal=[]
+
+fig, axs = plt.subplots(1,3, figsize=(8, 3))
+
+inputModulation = np.sin(2 * np.pi * inputFrequency_Hz[0] * support_S) * inputContrasts[0]
+maskModulation =  np.sin(2 * np.pi * inputFrequency_Hz[1] * support_S) * inputContrasts[1]
+
+axs[0].plot(support_S, inputModulation,'k')
+axs[0].plot(support_S, maskModulation,'r')
+axs[0].set_title('Input modulations', fontsize=10)
+axs[0].grid()
+
+combinedSignal=simpleTransducerFunction(inputModulation, maskModulation, c50,RMax)
+FT_combinedSignal=np.abs(np.fft.fft(combinedSignal)/ len(combinedSignal))
+axs[1].plot(support_S, combinedSignal,'k')
+axs[1].set_title('Combined signal after transducer', fontsize=10)
+axs[1].grid()
+
+axs[2].bar(np.arange(1,19), FT_combinedSignal[1:19])
 

### 2024-04-02 review.qmd
Tidying

@@
 However, it is important to take physiology into account.
-Contrast is represented as a combination of rectified signals from 'on and 'off'-sensitive cells (broadly, cells sensitive to positive or negative changes in spatiotemporal luminace).
-To model the population response to a contrast-reversing luminance sine wave, we therefore introduce an initial full-wave rectification stage that simulates the outputs of two equal populations of on- and off-polarity detectors.
+If we ignore cells sensitive to contrast transients, contrast is represented as a combination of rectified signals from 'on and 'off'-sensitive cells (broadly, cells that code positive or negative changes in spatiotemporal luminance).
+To model the population response of these cells to a contrast-reversing luminance sine wave, we therefore introduce an initial full-wave rectification stage that simulates the outputs of two, equal populations of on- and off-polarity detectors.
 Once this is done, we see that the FT of the resulting signal contains power only at the second harmonics and above.
@@
 #| label: plotspectra
-#| fig-cap: "a"
+#| fig-cap: "Temporal frequency representations of linar combinations of inputs."
 #| fig-cap-location: bottom
@@
 
-Althoug
+Even without considering a spatial component, the early visual system is far more complex than the model here suggests.
+For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code termporal changes in contrast.
+These cells (@ Movshon?
+Levitt?
+Earlier...
+Kuffler?
+H+W?
+Alpern?) will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
+Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position.
+The SSVEP response to a contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
+
+One of the simplest nonlinearities is the response function that describes a cell's response to different levels of contrast.
+In Figure xxx this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity.
+${R{\mathrm{max}}}\frac{C{\mathrm{in}}^n}{C{\mathrm{in}}^n+ C{\mathrm{50}}^n}$ where ${R{\mathrm{max}}}$ describes the maximum response level, $C{\mathrm{50}}$ is the 'semi-saturation constant' (the point at which the reponse is at half-maximum) and n controls the steepness of the curve.
+
 ```{python}
@@
 #| label: ContrastResponseFunction
-#| fig-cap: "a"
+#| fig-cap: "Non-linear contrast response function"
 #| fig-cap-location: bottom
@@
 
-Notice the slight roll-off in the amplitude of the second harmonic at high input contrasts.
+The hyperbolic ration function is monotinc but notice the slight roll-off in the amplitude of the second harmonic at high input contrasts.
 This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
@@
 ```{python}
-#| label: MaskedContrastResponseFunction
+#| label: Intermodulation
 #| fig-cap: "a"

### 2024-04-02 review.qmd
Rearranged

@@
 The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
-This normalization, achieved through a computation called 'contrast gain control' [Heeger 92, Heeger and Carandini 2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
+This normalization, achieved through a computation called 'contrast gain control' [Nachmas and Sansbury?
+Foley?
+Heeger 92, Heeger and Carandini 2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
 As an example, a grating placed at the centre of a low-contrast background tpyically appears more intense than the same grating when superimposed on a high contrast background.
@@
 
-### Measuring the development of contrast processing
-
-An early use of the contrast response function was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
-
-In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
-However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
-
-Because SSVEP response at detection threshold are very small, estimating this threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis.
-This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@DaleNorciaTyler86].
-
-Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler, Nakayama etc 1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, Norcia etc: 1985, 1990 etc].
-
-This general approach has revealed much about the development of visual abilities in infants [@braddick, @atkinson].
-In general, SSVEP measurements of infant vision have revealed that spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
-At least some of this difference is likely due to the relatively objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking which require careful measurement of the infant's gaze direction.
-One open question is how more complex contrast computations develop - and in particular how the human brain develop the tightly-tuned 'gain control' mechanisms that allow it to detect and discriminate contrast across very different visual scenes.
-
-```{python}
-
-#| label: SweepVEP Deteection threshold estimate
-#| fig-cap: "a"
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-from matplotlib import colormaps
-import numpy as np
-import scipy as scp
-from sklearn.linear_model import LinearRegression
-
-def simpleTransducerFunction(inputContrast, c50, RMax):
- # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
- # RMax defines the reponse gain of the transducer (how big the biggest output is). 
- # c50 defines the conrast sensitivity. Smaller values are more sensitive.
- # The exponent can also vary - here we fix it to 2
-  expnt=2
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
-  return output
-
-# Define some constants
-nLevels=8
-internalNoise=1
-inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
-contrastRange=np.logspace(1,1.9,nLevels)/100 # Equally spaced log contrast between 1 and 10^1.5
-duration_S=1 # All variables in this code have an _[units] where possible
-digitisationRate_Hz=1000
-freqCutoff_Hz=30
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-fig, axs = plt.subplots(1,1, figsize=(4, 3))
-
-initialSignal = np.zeros(len(contrastRange))
-transducedSignal = np.zeros(len(contrastRange))
-for contrastIndex,thisContrast in enumerate(contrastRange):
-  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast + np.random.rand(len(support_S))*internalNoise
-
-  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
-
-  responseAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
-  initialSignal[contrastIndex]=responseAmplitude[0]
-  transducedSignal[contrastIndex]=simpleTransducerFunction(responseAmplitude[0], 0.05, 10)
-
-# Now fit a linear regression to the first 4 contrast levels (before it saturates) and find out where this crosses the Y-axis
-X = np.log10(contrastRange[:4]).reshape(-1, 1)
-y = transducedSignal[:4]
-reg = LinearRegression().fit(X, y)
-x_intercept = 10 ** (-reg.intercept_ / reg.coef_[0])
-
-axs.set_title(f'Simulated sweep VEP at {nLevels} different contrasts', fontsize=10)
-axs.semilogx(contrastRange, transducedSignal, 'o')
-axs.semilogx(contrastRange[:4], reg.predict(X), 'r--', label='Linear Fit')
-axs.semilogx([x_intercept], [0], 'ro', label='Threshold')
-axs.set_xlabel('Contrast')
-axs.set_ylabel('Amplitude (a.u.)')
-axs.set_xlim(0.01, 1)
-axs.set_ylim(0, 10)
-axs.legend()
-axs.grid()
-plt.tight_layout()
-plt.show()
-
-```
-
-The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
-In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004 Diva vs Enfant, @Hoffman, @Bach and @Farmer].
-The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
-In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffman; @Bach; @Farmer].
+To understand the utility of the contrast SSVEP, it is helpful to identify the cascade of processing stages in the early visual system that give rise to it.
+In the following section we illustrate how a typical SSVEP signal measured over early visual cortex might contain information about a large number of early visual computations.
 
@@
 
-The hyperbolic ration function is monotinc but notice the slight roll-off in the amplitude of the second harmonic at high input contrasts.
+The hyperbolic ration function is monotonic but notice the slight roll-off in the amplitude of the second harmonic at high input contrasts.
 This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
@@
 ```
+
+### Measuring the development of contrast processing
+
+An early use of the SSVEP was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
+
+In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
+However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
+
+Because SSVEP response at detection threshold are very small, estimating this threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis.
+This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@DaleNorciaTyler86].
+
+Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler, Nakayama etc 1979].
+To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, Norcia etc: 1985, 1990 etc].
+
+This general approach has revealed much about the development of visual abilities in infants [@braddick, @atkinson].
+In general, SSVEP measurements of infant vision have revealed that spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+At least some of this difference is likely due to the relatively objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking which require careful measurement of the infant's gaze direction.
+One open question is how more complex contrast computations develop - and in particular how the human brain develop the tightly-tuned 'gain control' mechanisms that allow it to detect and discriminate contrast across very different visual scenes.
+
+```{python}
+
+#| label: SweepVEP Deteection threshold estimate
+#| fig-cap: "a"
+#| fig-cap-location: bottom
+
+import matplotlib.pyplot as plt
+from matplotlib import colormaps
+import numpy as np
+import scipy as scp
+from sklearn.linear_model import LinearRegression
+
+def simpleTransducerFunction(inputContrast, c50, RMax):
+ # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
+ # RMax defines the reponse gain of the transducer (how big the biggest output is). 
+ # c50 defines the conrast sensitivity. Smaller values are more sensitive.
+ # The exponent can also vary - here we fix it to 2
+  expnt=2
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
+  return output
+
+# Define some constants
+nLevels=8
+internalNoise=1
+inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
+contrastRange=np.logspace(1,1.9,nLevels)/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+fig, axs = plt.subplots(1,1, figsize=(4, 3))
+
+initialSignal = np.zeros(len(contrastRange))
+transducedSignal = np.zeros(len(contrastRange))
+for contrastIndex,thisContrast in enumerate(contrastRange):
+  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast + np.random.rand(len(support_S))*internalNoise
+
+  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+
+  responseAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
+  initialSignal[contrastIndex]=responseAmplitude[0]
+  transducedSignal[contrastIndex]=simpleTransducerFunction(responseAmplitude[0], 0.05, 10)
+
+# Now fit a linear regression to the first 4 contrast levels (before it saturates) and find out where this crosses the Y-axis
+X = np.log10(contrastRange[:4]).reshape(-1, 1)
+y = transducedSignal[:4]
+reg = LinearRegression().fit(X, y)
+x_intercept = 10 ** (-reg.intercept_ / reg.coef_[0])
+
+axs.set_title(f'Simulated sweep VEP at {nLevels} different contrasts', fontsize=10)
+axs.semilogx(contrastRange, transducedSignal, 'o')
+axs.semilogx(contrastRange[:4], reg.predict(X), 'r--', label='Linear Fit')
+axs.semilogx([x_intercept], [0], 'ro', label='Threshold')
+axs.set_xlabel('Contrast')
+axs.set_ylabel('Amplitude (a.u.)')
+axs.set_xlim(0.01, 1)
+axs.set_ylim(0, 10)
+axs.legend()
+axs.grid()
+plt.tight_layout()
+plt.show()
+
+```
+
+The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
+In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004 Diva vs Enfant, @Hoffman, @Bach and @Farmer].
+The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
+In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffman; @Bach; @Farmer].
 

### 2024-04-02 review.qmd
More on infants

@@
 Heeger 92, Heeger and Carandini 2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
-As an example, a grating placed at the centre of a low-contrast background tpyically appears more intense than the same grating when superimposed on a high contrast background.
+As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background.
 ```{python}
+
+#| label: Perceptual gain control
+#| fig-cap: The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region"
+#| fig-cap-location: bottom
+#|
 import numpy as np
@@
 Neurons have a limited dynamic range yet they can transmit information about  visual stimuli that span many order of magnitude.
-In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear,sigmoidal contrast vs response (CRF) transducer and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Early matteo stuff, Busse refs].
-A neuronal population will therefor span a sensitivity range greater than any individual member,
+In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear,sigmoidal contrast vs response (CRF) transducer (@Albrecht and Hamilton1982, @Tolhurst et al1981) and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Early matteo stuff, Busse refs].
+A neuronal population will therefor span a sensitivity range greater than any individual member.
 Neurons have a limited dynamic range yet they can transmit information about visual stimuli that span many orders of magnitude.
@@
 #| label: plotspectra
-#| fig-cap: "Temporal frequency representations of linar combinations of inputs."
+#| fig-cap: "Temporal frequency representations of combining two inputs at frequencies F1=5Hz and F2=7Hz combination.
+A pure sine-wave input would generate responses only at F1 and F2.
+However, the rectification that accompanies early visual processing converts this contrast reversal into a more complex frequency-domain representation with power predominantly at the second harmonics 2F1 and 2F2"
 #| fig-cap-location: bottom
@@
 #| label: ContrastResponseFunction
-#| fig-cap: "Non-linear contrast response function"
+#| fig-cap: "Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
+Note the 'supersaturation' at high contrast levels."
 #| fig-cap-location: bottom
@@
 
-The hyperbolic ration function is monotonic but notice the slight roll-off in the amplitude of the second harmonic at high input contrasts.
+The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the 2F component contains slight roll-off  at high input contrasts.
 This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
 Power at other harmonics increases and the total power generated by the input is monotonic.
-This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler]
+This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler].
 
-Here, we include a second component that contributes to the gain pool of the first. At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT.
+To illustrate the effect of contrast gain control[@Heeger92], we include a second component that contributes to the gain pool of the first.
+At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT.
 
@@
 #| label: Intermodulation
-#| fig-cap: "a"
+#| fig-cap: "Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies"
 #| fig-cap-location: bottom
@@
 However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [Norcia, Tyler, Braddick] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [@Norcia, Tyler, Braddick] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
 
@@
 This general approach has revealed much about the development of visual abilities in infants [@braddick, @atkinson].
-In general, SSVEP measurements of infant vision have revealed that spatial acuity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months while spatial acuity as measured by SSVEP reaches adult levels more slowly but near-adult levels are recorded around 1 year [[@NorciaTyler1985, @HamerVisRed1989]] compared to  around 6-7 years with behavioural measures [@Ellemberg vis res 1999, @Atkinson].
 At least some of this difference is likely due to the relatively objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking which require careful measurement of the infant's gaze direction.
-One open question is how more complex contrast computations develop - and in particular how the human brain develop the tightly-tuned 'gain control' mechanisms that allow it to detect and discriminate contrast across very different visual scenes.
+The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section.[@Candy 2001, @PeiNorcia 2017].
+Although contrast gain control is measurable in infants as young as six weeks old (@Candy2001, @MorroneBurr1986, @Skoczenski & Norcia, 1998), development appears to be slower with adult levels being reached in approximately 11 years.
+[@PeiNorcia].
 
@@
 #| label: SweepVEP Deteection threshold estimate
-#| fig-cap: "a"
+#| fig-cap: "Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts"
 #| fig-cap-location: bottom

### 2024-04-02 review.qmd
Adding DeVries

@@
 This general approach has revealed much about the development of visual abilities in infants [@braddick, @atkinson].
-In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months while spatial acuity as measured by SSVEP reaches adult levels more slowly but near-adult levels are recorded around 1 year [[@NorciaTyler1985, @HamerVisRed1989]] compared to  around 6-7 years with behavioural measures [@Ellemberg vis res 1999, @Atkinson].
-At least some of this difference is likely due to the relatively objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking which require careful measurement of the infant's gaze direction.
+In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months while spatial acuity as measured by SSVEP reaches adult levels more slowly but near-adult levels are recorded around 1 year [@NorciaTyler1985, @HamerVisRed1989] compared to  around 6-7 years with behavioural measures [@Ellemberg vis res 1999, @Atkinson].
+At least some of this difference is likely due to the relatively objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking which require careful measurement of the infant's gaze direction although it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@de1982maturation].
+
 The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section.[@Candy 2001, @PeiNorcia 2017].

### 2024-04-02 review.qmd
More on fitting

@@
 
+The complexity of even a simple simulation of the frequency-domain signal arising from non-linear interactions is intriguing.
+Presumably, the signal measured from early visual cortex contains is the result of a cascade of nonlinear retinal and cortical operations up to that point and therefore contains a 'signature' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
+In principle, that information could be recovered from the SSVEP signal.
+
+This possibility was recognised in the early days of the technique [@reganandRegan1988].
+Although characterising the complete set of computations along entire processing pathway is problematic, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
+...
+MORE HERE - Although this idea was suggested as early as 1988 (probably earlier) not many people seem to have done it much.
+Candy and Norcia (in infants) 2001, Busse et al2008, Tsai/Norcia/Wade 2010s, Then lots of Baker stuff including the CerebCort paper.
+: Focusing only on SSVEP: [@NorciaCandy2001, @Regan1990?
+]
+
 ### Measuring the development of contrast processing
@@
 
-The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
-In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004 Diva vs Enfant, @Hoffman, @Bach and @Farmer].
+The sweep VEP (technically, the sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
+Because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004 Diva vs Enfant, @Hoffman, @Bach and @Farmer].
 The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
 In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffman; @Bach; @Farmer].
+
+##
 

### 2024-04-03 review.qmd
Switched to pdf output

@@
 title: "Measuring contrast processing in the visual system using SSVEP (and VEP?)"
-format: html
+format: pdf
 bibliography: WadeBaker_Review.bib

### 2024-04-03 review.qmd
Changed citation style

@@
 bibliography: WadeBaker_Review.bib
+csl: visual-neuroscience.csl
 execute:
@@
 Measuring the way that the computations underlying contrast processing depend on factors such as spatial and temporal frequency, age, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
-One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
-SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high SNR and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
-Here we describe how SSVEPs have been used to study visual contrast over the past 70 years[@Dawson1954].
+One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged' (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
+SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high SNR, and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
+Here we describe how SSVEPs have been used to study visual contrast over the past 70 years [@Dawson1954].
 Because our thinking about SSVEPs is well-described by simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.
@@
 
-## INTRODUCTION:
+## INTRODUCTION
+
 Neurons in the visual areas of the brain are primarily responsive to changes in cone photoreceptor activations across time and space.
@@
 The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di Russo, Hillebrand:].
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2007] (Hillebrand?).
 The flickering stimulus entrains neural responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
@@
 The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system [@Regan1988].
-SSVEP signals can also be elicited by periodic changes of stimulus properties other than achromatic and chromatic contrast, such as, motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
+SSVEP signals can also be elicited by periodic changes of stimulus properties other than achromatic and chromatic contrast, such as motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
 
 ### Why measure responses to contrast?
+
 Contrast is one of the most fundamental pieces of information that the eye transmits to the brain.
-It can be defined as the change in cone photoreceptor activity over time ('spatial contrast') or space ('spatial contrast').
+It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
 Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well.
@@
 
-Contrast is relatively simple to define: typically, contrast is specified as the percentage deviation of uniform stimulus from the background.
-So, for example, a disk of 100 units, $I_{\mathrm{stim}}$ of cone activation surrounded by a 'background' of 50 units of activation $I_{\mathrm{background}}$ has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
-Where patterns are more complex (for example, the sine-wave gratings or Gabor patches common in vision science), the Michaelson definition of contrast is specified by the maximum and minimum excursions from the mean: $${{\frac {I_{\mathrm {stimmax} }-I_{\mathrm {stimmin} }}{I_{\mathrm {stimmax} }+I_{\mathrm {stimmin} }}},} $$.
-These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance [LenniePokornySmith]) and also to definitions based on cone excitations [DKL, MacBoynton] which are more common in work on chromatic processing.
+Contrast is relatively simple to define: typically, contrast is specified as the percentage deviation of a uniform stimulus from the background.
+So, for example, a disk of 100 units of cone activation ($I_{\mathrm{stim}}$) surrounded by a 'background' of 50 units of activation ($I_{\mathrm{background}}$) has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
+Where patterns are more complex (for example, the sine-wave gratings or Gabor patches common in vision science), the @Michelson1927 definition of contrast is specified by the maximum and minimum excursions from the mean: $${{\frac {I_{\mathrm {stimmax} }-I_{\mathrm {stimmin} }}{I_{\mathrm {stimmax} }+I_{\mathrm {stimmin} }}}.} $$ These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance [LenniePokornySmith]) and also to definitions based on cone excitations [@Derrington1984, @MacLeod1979] which are more common in work on chromatic processing.
 
@@
 The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
-This normalization, achieved through a computation called 'contrast gain control' [Nachmas and Sansbury?
-Foley?
-Heeger 92, Heeger and Carandini 2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
+This normalization, achieved through a computation called 'contrast gain control' [@Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
 As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background.
+
 ```{python}
-
 #| label: Perceptual gain control
@@
 #| fig-cap-location: bottom
-#|
+
 import numpy as np
@@
 This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed respnses of the local neuronal population.[@Heeger 92, @Carandini and Heeger 2011 @Busse 2008, @Baker and Wade]
+
 Individual neurons at multiple stages of the visual hierarchy also change their sensitivity depending on the average spatiotemporal contrast energy of their environment.
@@
 The nature of the interaction --- and therefore the pattern of intermodulation terms --- is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan].
+
 First, we implement a simple non-linear transducer with a single input to illustrate the typical signoidal shape of a contrast response function.
@@
 ```{python}
-
 #| label: plotsinewaves
@@
 #| fig-cap-location: bottom
-#| 
+
 import matplotlib.pyplot as plt
@@
 plt.show()
+
 ```
@@
 ```{python}
-
 #| label: plotspectra
@@
 ```{python}
-
 #| label: ContrastResponseFunction
@@
 Candy and Norcia (in infants) 2001, Busse et al2008, Tsai/Norcia/Wade 2010s, Then lots of Baker stuff including the CerebCort paper.
-: Focusing only on SSVEP: [@NorciaCandy2001, @Regan1990?
-]
+: Focusing only on SSVEP: [@NorciaCandy2001, @Regan1990]
 
@@
 Because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004 Diva vs Enfant, @Hoffman, @Bach and @Farmer].
+
 The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
@@
 and neuroscience, 2010.
+
+# References

### 2024-04-03 review.qmd
Fixed some references, removed duplicated paragraphs

@@
 
-A signifcant part of research into contrast processing is concerned with how these normalization mechanisms depend on colour, orientation, eye of origin, spatial and temporal frequency, age, and the presence of neurological disorders [refs: Chen and Foley, Petrov etc Carandini, Binocular normalization / suppression (Baker/Meese but also e.g.
-Heeger), Tadin, Winawer, Wade, Porciatti, Tsai...].
-The SSVEP has proven to be invaliable in this research because it provides an objective readout of contrast representation at different parts of the visual system and allows us to 'tag' the probe and background separately.
+A significant part of research into contrast processing is concerned with how these normalization mechanisms depend on colour [@Chen2000], orientation [@Foley1994], eye of origin [@Legge1979; @Baker2007], spatial and temporal frequency [@Meese2009], location [@Petrov2005; @Tadin2003; @Polat1993], age [@Betts2005], and the presence of neurological disorders [@Porciatti2000; @Tsai2011].
+The SSVEP has proven to be invaliable in this research because it provides an objective readout of contrast representation at different parts of the visual system and allows us to 'tag' the probe and background at separate frequencies.
 
@@
 For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency or a low multiple thereof.
-This corresponds closely to similar functions measured in single unit studies, or those measuring local field potentials invasively in the cortex[@ShapleyVictor etc].
+This corresponds closely to similar functions measured in single unit studies, or those measuring local field potentials invasively in the cortex [@Shapley1980].
 However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
@@
 
-Neurons have a limited dynamic range yet they can transmit information about  visual stimuli that span many order of magnitude.
-In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear,sigmoidal contrast vs response (CRF) transducer (@Albrecht and Hamilton1982, @Tolhurst et al1981) and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Early matteo stuff, Busse refs].
+Neurons have a limited dynamic range yet they can transmit information about visual stimuli that span many orders of magnitude.
+In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear, sigmoidal contrast vs response (CRF) transducer [@Albrecht1982; @Tolhurst1981] and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Carandini1994; @Carandini1998; @Busse2009].
 A neuronal population will therefor span a sensitivity range greater than any individual member.
-Neurons have a limited dynamic range yet they can transmit information about visual stimuli that span many orders of magnitude.
-In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear, sigmoidal contrast vs response (CRF) transducer and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@EarlyMatteoStuff; @BusseRefs].
-A neuronal population will therefore span a sensitivity range greater than any individual member.
 
 Individual neurons at multiple stages of the visual hierarchy also change their sensitivity depending on the average spatiotemporal contrast energy of their environment.
-This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed respnses of the local neuronal population.[@Heeger 92, @Carandini and Heeger 2011 @Busse 2008, @Baker and Wade]
-
-Individual neurons at multiple stages of the visual hierarchy also change their sensitivity depending on the average spatiotemporal contrast energy of their environment.
-This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed responses of the local neuronal population [@Heeger1992; @CarandiniHeeger2011; @Busse2008; @BakerWade].
+This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed responses of the local neuronal population [@Heeger1992; @Carandini2011; @Busse2009; @Baker2017].
 
@@
 The nature of the interaction --- and therefore the pattern of intermodulation terms ---  is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan]
-
-Here, we show the effect of adding a second contrast component: in general, the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
-The nature of the interaction --- and therefore the pattern of intermodulation terms --- is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan].
 
@@
 However, it is important to take physiology into account.
-If we ignore cells sensitive to contrast transients, contrast is represented as a combination of rectified signals from 'on and 'off'-sensitive cells (broadly, cells that code positive or negative changes in spatiotemporal luminance).
+If we ignore cells sensitive to contrast transients, contrast is represented as a combination of rectified signals from 'on' and 'off'-sensitive cells (broadly, cells that code positive or negative changes in spatiotemporal luminance).
 To model the population response of these cells to a contrast-reversing luminance sine wave, we therefore introduce an initial full-wave rectification stage that simulates the outputs of two, equal populations of on- and off-polarity detectors.
-Once this is done, we see that the FT of the resulting signal contains power only at the second harmonics and above.
+Once this is done, we see that the Fourier spectrum of the resulting signal contains power only at the second harmonics and above.
 It is possible to generate first-hamonic responses in contrast-response SSVEPs but only when the inputs are modulated in an 'on/off' manner rather than contrast reversed.
@@
 Even without considering a spatial component, the early visual system is far more complex than the model here suggests.
-For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code termporal changes in contrast.
+For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code temporal changes in contrast.
 These cells (@ Movshon?
@@
 One of the simplest nonlinearities is the response function that describes a cell's response to different levels of contrast.
-In Figure xxx this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity.
-${R{\mathrm{max}}}\frac{C{\mathrm{in}}^n}{C{\mathrm{in}}^n+ C{\mathrm{50}}^n}$ where ${R{\mathrm{max}}}$ describes the maximum response level, $C{\mathrm{50}}$ is the 'semi-saturation constant' (the point at which the reponse is at half-maximum) and n controls the steepness of the curve.
+In Figure xxx this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity:
+
+\begin{equation}
+R_{max} = \frac{C_{in}^n}{C_{50}^n + C_{in}^n},
+\end{equation}
+
+where ${R_{max}}$ describes the maximum response level, $C_{in}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the reponse is at half-maximum) and *n* controls the steepness of the curve.
 
@@
 
-The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the 2F component contains slight roll-off  at high input contrasts.
+The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the 2F component contains a slight roll-off at high input contrasts.
 This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
 Power at other harmonics increases and the total power generated by the input is monotonic.
-This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler].
+This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler1985; @Peirce2007].
 
-To illustrate the effect of contrast gain control[@Heeger92], we include a second component that contributes to the gain pool of the first.
+To illustrate the effect of contrast gain control [@Heeger1992], we include a second component that contributes to the gain pool of the first.
 At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT.
@@
 
-This possibility was recognised in the early days of the technique [@reganandRegan1988].
+This possibility was recognised in the early days of the technique [@Regan1988].
 Although characterising the complete set of computations along entire processing pathway is problematic, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.

### 2024-04-04 review.qmd
Fixed some references

@@
 ---
-title: "Measuring contrast processing in the visual system using SSVEP (and VEP?)"
+title: "Measuring contrast processing in the visual system using SSVEP"
 format: pdf
@@
 
-## Baker, Wade: Review: Vis Neurosci
+## Wade, A.R. & Baker, D.H.: Review for Vis Neurosci
 
@@
 The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2007] (Hillebrand?).
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2007] (@Hillebrand?).
 The flickering stimulus entrains neural responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
@@
 It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
-Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well.
-[McKeefry, Sutter, Baseler, DiRusso...].
+Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@McKeefry1996; @Baseler1997; @Di-Russo2001].
 
@@
 So, for example, a disk of 100 units of cone activation ($I_{\mathrm{stim}}$) surrounded by a 'background' of 50 units of activation ($I_{\mathrm{background}}$) has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
-Where patterns are more complex (for example, the sine-wave gratings or Gabor patches common in vision science), the @Michelson1927 definition of contrast is specified by the maximum and minimum excursions from the mean: $${{\frac {I_{\mathrm {stimmax} }-I_{\mathrm {stimmin} }}{I_{\mathrm {stimmax} }+I_{\mathrm {stimmin} }}}.} $$ These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance [LenniePokornySmith]) and also to definitions based on cone excitations [@Derrington1984, @MacLeod1979] which are more common in work on chromatic processing.
+Where patterns are more complex (for example, the sine-wave gratings or Gabor patches common in vision science), the @Michelson1927 definition of contrast is specified by the maximum and minimum excursions from the mean:
+
+\begin{equation}
+\frac {I_{\mathrm{stimmax} }-I_{\mathrm{stimmin} }}{I_{\mathrm{stimmax} }+I_{\mathrm{stimmin} }}.
+\end{equation}
+
+These contrast definitions are appropriate both to photometric measures of stimulus contrast [for example, luminance; @Lennie1993] and also to definitions based on cone excitations [@Derrington1984, @MacLeod1979] which are more common in work on chromatic processing.
 
@@
 Here, we show the effect of adding a second contrast component: in general, the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
-The nature of the interaction --- and therefore the pattern of intermodulation terms ---  is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan]
+The nature of the interaction --- and therefore the pattern of intermodulation terms ---  is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan1988]
 
@@
 \begin{equation}
-R_{max} = \frac{C_{in}^n}{C_{50}^n + C_{in}^n},
+R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},
 \end{equation}
 
-where ${R_{max}}$ describes the maximum response level, $C_{in}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the reponse is at half-maximum) and *n* controls the steepness of the curve.
+where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the reponse is at half-maximum) and *n* controls the steepness of the curve.
 
@@
 MORE HERE - Although this idea was suggested as early as 1988 (probably earlier) not many people seem to have done it much.
-Candy and Norcia (in infants) 2001, Busse et al2008, Tsai/Norcia/Wade 2010s, Then lots of Baker stuff including the CerebCort paper.
-: Focusing only on SSVEP: [@NorciaCandy2001, @Regan1990]
+@Candy2001; @Busse2009; @Tsai2012; @Baker2017 Focusing only on SSVEP: [@NorciaCandy2001, @Regan1990]
 
@@
 However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [@Norcia, Tyler, Braddick] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [@Norcia1990; Tyler, Braddick] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
 
 Because SSVEP response at detection threshold are very small, estimating this threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis.
-This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@DaleNorciaTyler86].
+This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
 
 Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler, Nakayama etc 1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, Norcia etc: 1985, 1990 etc].
+To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, @Norcia1985a; @Norcia1990].
 
-This general approach has revealed much about the development of visual abilities in infants [@braddick, @atkinson].
-In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@dobsonVisualAcuityHuman1978; @norciaSpatialFrequencySweep1985, ] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months while spatial acuity as measured by SSVEP reaches adult levels more slowly but near-adult levels are recorded around 1 year [@NorciaTyler1985, @HamerVisRed1989] compared to  around 6-7 years with behavioural measures [@Ellemberg vis res 1999, @Atkinson].
-At least some of this difference is likely due to the relatively objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking which require careful measurement of the infant's gaze direction although it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@de1982maturation].
+This general approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
+In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around 1 year [@Norcia1985b, @HamerVisRed1989] compared to  around 6-7 years with behavioural measures [@Ellemberg1999; @Atkinson].
+At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
+However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
 
-The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section.[@Candy 2001, @PeiNorcia 2017].
-Although contrast gain control is measurable in infants as young as six weeks old (@Candy2001, @MorroneBurr1986, @Skoczenski & Norcia, 1998), development appears to be slower with adult levels being reached in approximately 11 years.
-[@PeiNorcia].
+The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section.[@Candy2001; @Pei2017].
+Although contrast gain control is measurable in infants as young as six weeks old [@Candy2001; @Morrone1986; @Skoczenski1998], its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
 
@@
 ```
-
-The sweep VEP (technically, the sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
-Because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004 Diva vs Enfant, @Hoffman, @Bach and @Farmer].
 

### 2024-04-04 review.qmd
Fixed more references

@@
 The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2007] (@Hillebrand?).
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2007] @Hillebrand.
 The flickering stimulus entrains neural responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
@@
 In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear, sigmoidal contrast vs response (CRF) transducer [@Albrecht1982; @Tolhurst1981] and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Carandini1994; @Carandini1998; @Busse2009].
-A neuronal population will therefor span a sensitivity range greater than any individual member.
+A neuronal population will therefore span a sensitivity range greater than any individual member.
 
@@
 For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code temporal changes in contrast.
-These cells (@ Movshon?
-Levitt?
-Earlier...
-Kuffler?
-H+W?
-Alpern?) will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
+These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
 Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position.
@@
 However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [@Norcia1990; Tyler, Braddick] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [@Norcia1990; @Tyler1979; @Braddick1986] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
 
@@
 
-Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler, Nakayama etc 1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [Tyler, @Norcia1985a; @Norcia1990].
+Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
+To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990].
 
@@
 In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
-Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around 1 year [@Norcia1985b, @HamerVisRed1989] compared to  around 6-7 years with behavioural measures [@Ellemberg1999; @Atkinson].
+Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around 1 year [@Norcia1985b; @Hamer1989] compared to  around 6-7 years with behavioural measures [@Ellemberg1999; @Atkinson].
 At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
@@
 The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
-In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffman; @Bach; @Farmer].
+In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffman2017; @Bach2008; @Bach2020].
 
@@
 *4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*
-[@Porciatti2000; @Tsai2011]
+
 *5: Future directions (decoding in frequency domain?, animals? BCI?)*

### 2024-04-04 review.qmd
Minor updates

@@
 
-These contrast definitions are appropriate both to photometric measures of stimulus contrast [for example, luminance; @Lennie1993] and also to definitions based on cone excitations [@Derrington1984, @MacLeod1979] which are more common in work on chromatic processing.
+These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance; @Lennie1993) and also to definitions based on cone excitations [@Derrington1984, @MacLeod1979] which are more common in work on chromatic processing.
 
@@
 In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
-Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around 1 year [@Norcia1985b; @Hamer1989] compared to  around 6-7 years with behavioural measures [@Ellemberg1999; @Atkinson].
+Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around 1 year [@Norcia1985b; @Hamer1989] compared to around 6-7 years with behavioural measures [@Ellemberg1999; @Atkinson1983].
 At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
@@
 The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
-In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffman2017; @Bach2008; @Bach2020].
+In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].
 

### 2024-04-06 review.qmd
Implemented figure referencing

@@
 Because our thinking about SSVEPs is well-described by simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.
-This document can therefore be used both as a review of the use of SSVEP in measuring human contrast processing, and as an interactive learning aid.
+This paper can therefore be used both as a review of the use of SSVEP in measuring human contrast processing, and as an interactive learning aid.
 
@@
 For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2007] @Hillebrand.
-The flickering stimulus entrains neural responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
+The flickering stimulus entrains neural population responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
 Two common variants involve sinusoidal on-off flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
@@
 
-These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance; @Lennie1993) and also to definitions based on cone excitations [@Derrington1984, @MacLeod1979] which are more common in work on chromatic processing.
+These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance; @Lennie1993) and also to definitions based on cone excitations [@Derrington1984; @MacLeod1979] which are more common in work on chromatic processing.
 
-Although its definition as a stimulus is simple, the computations that underlie contrast processing in the brain have been the subject of intense research for many decades.
+Although its mathematical definition is straightforward, the computations that underlie contrast processing in the brain have been the subject of intense research for many decades.
 The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
 This normalization, achieved through a computation called 'contrast gain control' [@Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
-As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background.
+As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround).
 
 ```{python}
-#| label: Perceptual gain control
+#| label: fig-centresurround
 #| fig-cap: The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region"
@@
 
-A significant part of research into contrast processing is concerned with how these normalization mechanisms depend on colour [@Chen2000], orientation [@Foley1994], eye of origin [@Legge1979; @Baker2007], spatial and temporal frequency [@Meese2009], location [@Petrov2005; @Tadin2003; @Polat1993], age [@Betts2005], and the presence of neurological disorders [@Porciatti2000; @Tsai2011].
-The SSVEP has proven to be invaliable in this research because it provides an objective readout of contrast representation at different parts of the visual system and allows us to 'tag' the probe and background at separate frequencies.
+A significant body of research into contrast processing is concerned with how these normalization mechanisms depend on colour [@Chen2000], orientation [@Foley1994], eye of origin [@Legge1979; @Baker2007], spatial and temporal frequency [@Meese2009], location [@Petrov2005; @Tadin2003; @Polat1993], age [@Betts2005], and the presence of neurological disorders [@Porciatti2000; @Tsai2011].
+The SSVEP has proven to be invaliable in this research because it provides an objective readout of contrast representation at different stages of the visual system, and allows us to 'tag' the probe and background at separate frequencies.
 
@@
 For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency or a low multiple thereof.
-This corresponds closely to similar functions measured in single unit studies, or those measuring local field potentials invasively in the cortex [@Shapley1980].
+This corresponds closely to similar functions reported by invasive studies measuring single unit activity or local field potentials in the cortex [@Shapley1980; @Morrone1982].
 However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
@@
 
-Neurons have a limited dynamic range yet they can transmit information about visual stimuli that span many orders of magnitude.
+Neurons have a limited dynamic range, yet they can transmit information about visual stimuli that span many orders of magnitude.
 In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear, sigmoidal contrast vs response (CRF) transducer [@Albrecht1982; @Tolhurst1981] and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Carandini1994; @Carandini1998; @Busse2009].
@@
 Previous figures showed how a single-input SSVEP might generate a single set of peaks in the frequency domain response which, in turn, can be used as a read-out of neuronal activity.
-
-Here, we show the effect of adding a second contrast component: in general, the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
+In @fig-plotspectra, we show the effect of adding a second contrast component: the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
 The nature of the interaction --- and therefore the pattern of intermodulation terms ---  is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan1988]
 
-First, we implement a simple non-linear transducer with a single input to illustrate the typical signoidal shape of a contrast response function.
-(Figure xx)
+First, we implement a simple non-linear transducer with a single input to illustrate the typical sigmoidal shape of a contrast response function [@fig-ContrastResponseFunction].
 
-Next, we show that adding a second component which contributes to the gain pool of the first, results in a complex pattern of intermodulation terms.
-
-Figures:
+Next, we show that adding a second component which contributes to the gain pool of the first, results in a complex pattern of intermodulation terms [@fig-plotspectra].
 
 ```{python}
-#| label: plotsinewaves
+#| label: fig-plotsinewaves
 #| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain"
@@
 
-If the two inputs are simply added together, the representation of the resulting signal in the Fourier domain is just the linear sum of the two independent signals.
+If the two inputs are simply added together, the representation of the resulting signal in the Fourier domain is just the linear sum of the two independent signals (first two panels of @fig-plotspectra).
 However, it is important to take physiology into account.
@@
 To model the population response of these cells to a contrast-reversing luminance sine wave, we therefore introduce an initial full-wave rectification stage that simulates the outputs of two, equal populations of on- and off-polarity detectors.
-Once this is done, we see that the Fourier spectrum of the resulting signal contains power only at the second harmonics and above.
-It is possible to generate first-hamonic responses in contrast-response SSVEPs but only when the inputs are modulated in an 'on/off' manner rather than contrast reversed.
+Once this is done, we see that the Fourier spectrum of the resulting signal contains power only at the second harmonics and above (final two panels of @fig-plotspectra).
+It is possible to generate first-hamonic responses in contrast-response SSVEPs, but only when the inputs are modulated in an 'on/off' manner rather than contrast reversed.
 
 ```{python}
-#| label: plotspectra
+#| label: fig-plotspectra
 #| fig-cap: "Temporal frequency representations of combining two inputs at frequencies F1=5Hz and F2=7Hz combination.
@@
 
-One of the simplest nonlinearities is the response function that describes a cell's response to different levels of contrast.
-In Figure xxx this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity:
+One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
+In @fig-ContrastResponseFunction this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity:
 
@@
 
-where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the reponse is at half-maximum) and *n* controls the steepness of the curve.
+where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the response is at half-maximum) and *n* controls the steepness of the curve.
 
 ```{python}
-#| label: ContrastResponseFunction
+#| label: fig-ContrastResponseFunction
 #| fig-cap: "Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
@@
 This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
-Power at other harmonics increases and the total power generated by the input is monotonic.
+Power at other harmonics increases, and the total power generated by the input is monotonic.
 This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler1985; @Peirce2007].
@@
 To illustrate the effect of contrast gain control [@Heeger1992], we include a second component that contributes to the gain pool of the first.
-At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT.
+At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT [@fig-Intermodulation].
 
 ```{python}
-#| label: Intermodulation
+#| label: fig-Intermodulation
 #| fig-cap: "Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies"
@@
 The complexity of even a simple simulation of the frequency-domain signal arising from non-linear interactions is intriguing.
-Presumably, the signal measured from early visual cortex contains is the result of a cascade of nonlinear retinal and cortical operations up to that point and therefore contains a 'signature' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
+Presumably, the signal measured from early visual cortex contains is the result of a cascade of nonlinear retinal and cortical operations up to that point.
+It therefore contains a 'signature' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
 In principle, that information could be recovered from the SSVEP signal.
@@
 An early use of the SSVEP was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
-
 In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
@@
 
-Because SSVEP response at detection threshold are very small, estimating this threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis.
+Because SSVEP responses at detection threshold are very small, estimating threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-SweepVEP).
 This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
-
-Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990].
-
-This general approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
-In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
-Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around 1 year [@Norcia1985b; @Hamer1989] compared to around 6-7 years with behavioural measures [@Ellemberg1999; @Atkinson1983].
-At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
-However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
-
-The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section.[@Candy2001; @Pei2017].
-Although contrast gain control is measurable in infants as young as six weeks old [@Candy2001; @Morrone1986; @Skoczenski1998], its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
 
@@
 
-#| label: SweepVEP Deteection threshold estimate
+#| label: fig-SweepVEP
 #| fig-cap: "Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts"
@@
 
+Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
+To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990].
 The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
 In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].
+
+This approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
+In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around 1 year [@Norcia1985b; @Hamer1989] compared to around 6-7 years with behavioural measures [@Ellemberg1999; @Atkinson1983].
+At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
+However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
+
+The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section [@Candy2001; @Pei2017].
+Although contrast gain control is measurable in infants as young as six weeks old [@Candy2001; @Morrone1986; @Skoczenski1998], its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
 

### 2024-04-06 review.qmd
I think figure references need to be lower case

@@
 
-First, we implement a simple non-linear transducer with a single input to illustrate the typical sigmoidal shape of a contrast response function [@fig-ContrastResponseFunction].
+First, we implement a simple non-linear transducer with a single input to illustrate the typical sigmoidal shape of a contrast response function [@fig-contrastresponsefunction].
 
@@
 
-One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
-In @fig-ContrastResponseFunction this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity:
-
-\begin{equation}
-R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},
-\end{equation}
-
-where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the response is at half-maximum) and *n* controls the steepness of the curve.
-
 ```{python}
-#| label: fig-ContrastResponseFunction
+#| label: fig-contrastresponsefunction
 #| fig-cap: "Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
@@
 
+One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
+In @fig-contrastresponsefunction this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity:
+
+\begin{equation}
+R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},
+\end{equation}
+
+where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the response is at half-maximum) and *n* controls the steepness of the curve.
+
 The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the 2F component contains a slight roll-off at high input contrasts.
@@
 To illustrate the effect of contrast gain control [@Heeger1992], we include a second component that contributes to the gain pool of the first.
-At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT [@fig-Intermodulation].
+At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT [@fig-intermodulation].
 
 ```{python}
-#| label: fig-Intermodulation
+#| label: fig-intermodulation
 #| fig-cap: "Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies"
@@
 
-Because SSVEP responses at detection threshold are very small, estimating threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-SweepVEP).
+Because SSVEP responses at detection threshold are very small, estimating threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
 This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
@@
 
-#| label: fig-SweepVEP
+#| label: fig-sweepvep
 #| fig-cap: "Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts"

### 2024-04-06 review.qmd
Started clinical section

@@
 title: "Measuring contrast processing in the visual system using SSVEP"
-format: pdf
+author:
+  - name: Alex R. Wade
+  - name: Daniel H. Baker
+format: 
+  pdf:
+    keep-tex: true
 bibliography: WadeBaker_Review.bib
@@
 ---
+<p style="text-align: center;">Department of Psychology and York Biomedical Research Institute
 
-## Wade, A.R. & Baker, D.H.: Review for Vis Neurosci
+University of York, UK</p>
 
-## ABSTRACT:
+## Review for Vis Neurosci
+
+# Abstract:
 
@@
 
-## INTRODUCTION
+# Introduction
 
@@
 The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2007] @Hillebrand.
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2007].
 The flickering stimulus entrains neural population responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
@@
 
-### Why measure responses to contrast?
+# Why measure responses to contrast?
 
@@
 MORE HERE - Although this idea was suggested as early as 1988 (probably earlier) not many people seem to have done it much.
-@Candy2001; @Busse2009; @Tsai2012; @Baker2017 Focusing only on SSVEP: [@NorciaCandy2001, @Regan1990]
+@Candy2001; @Busse2009; @Tsai2012; @Baker2017
 
-### Measuring the development of contrast processing
+# Measuring the development of contrast processing
 
@@
 
-##
+# Contrast processing in clinical conditions
+
+The SSVEP technique has also been used to study clinical conditions, such as diseases and developmental disorders.
+This can often be informative regarding the underlying mechanism that characterises the condition.
+Here we focus on four conditions, but there is potential to apply the method more broadly, either as a diagnostic technique, or to monitor disease severity and progression, or the efficacy of treatments.
+
+Epilepsy is a neurological condition in which patients experience seizures - episodes of uncontrolled neural activity that can cause unconsciousness, involuntary movements and convulsions, and atypical sensory experiences.
+@Porciatti2000 showed that individuals with photosensitive epilepsy generate larger steady-state signals in response to flickering visual stimuli, and their contrast response functions saturate less than those of healthy controls.
+This is consistent with the idea that epilepsy involves a cortical hyperexciteability that makes seizures more likely.
+It is also the case for individuals with idiopathic generalised epilepsy [@Tsai2011], a subtype of epilepsy that has a less obvious link to vision.
+The differences apply across the whole contrast-response function, and so resemble a response gain effect, which might be due to reduced inhibition from neighbouring neurons.
+Differences in SSVEP amplitudes have also been reported in individuals with migraine [@Shibata2008], a condition also associated with cortical hyperexciteability.
+
+Amblyopia
+
+Autism
+
+Parkinson's
+
+# Brain-computer interfaces
+
+# Conclusions
 

### 2024-04-06 review.qmd
Added amblyopia section

@@
 ---
-<p style="text-align: center;">Department of Psychology and York Biomedical Research Institute
-
-University of York, UK</p>
+Department of Psychology and York Biomedical Research Institute, University of York, UK
 
@@
 
-Amblyopia
+Amblyopia is a disorder of binocular vision, characterised by one eye contributing much less to perception than the other.
+This is often due to strabismus (squint) or anisometropia (difference in optical prescription between the eyes) during development.
+Contemporary accounts suggest that the amblyopic eye is suppressed by signals from the fellow eye.
+SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye across the contrast range [@Baker2015].
+There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.
+The steady-state approach may be more sensitive and objective than typical acuity measurements, and also has the potential to measure suppression between the eyes directly [e.g.
+@Zheng2019; @Du2023; @Hu2023].
 

### 2024-04-07 review.qmd
Added autism section

@@
 The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2007].
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2005; @Di-Russo2007].
 The flickering stimulus entrains neural population responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
@@
 Contemporary accounts suggest that the amblyopic eye is suppressed by signals from the fellow eye.
-SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye across the contrast range [@Baker2015].
+SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye across the contrast range [@Baker2015; @Lygo2021].
 There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.
@@
 
-Autism
+Autism is a condition often associated with differences in vision [@Simmons2009] and other senses [@MacLennan2022].
+@Pei2014 used a sweep-VEP method with counterphase flickering stimuli, and found weaker responses in autistic children at spatial frequencies around 8c/deg, compared with age-matched controls.
+This was subsequently replicated in a further pediatric sample by @Vilidaite2018, who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
+Interestingly this study replicated its key findings in a *drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.
 
 Parkinson's
+
+(maybe here we want a 4-panel figure plotting cartoon models of the four conditions - it would make for a good comparison, and we could potentially base it on data from the original studies)
 
@@
 Ales cruciform paper.
-
-Di Russo, F., Pitzalis, S., Spitoni, G., Aprile, T., Patria, F., Stella,
-A., \... \& Hillyard, S. A. (2007). Identification of the neural sources
-of the pattern-reversal VEP. Neuroimage, 34(1), 177-189.
 

### 2024-04-07 review.qmd
Started figure summarising clinical differences

@@
 
-First, we implement a simple non-linear transducer with a single input to illustrate the typical sigmoidal shape of a contrast response function [@fig-contrastresponsefunction].
+First, we implement a simple non-linear transducer with a single input to illustrate the typical sigmoidal shape of a contrast response function (@fig-contrastresponsefunction).
 
-Next, we show that adding a second component which contributes to the gain pool of the first, results in a complex pattern of intermodulation terms [@fig-plotspectra].
+Next, we show that adding a second component which contributes to the gain pool of the first, results in a complex pattern of intermodulation terms (@fig-plotspectra).
 
@@
 To illustrate the effect of contrast gain control [@Heeger1992], we include a second component that contributes to the gain pool of the first.
-At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT [@fig-intermodulation].
+At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT (@fig-intermodulation).
 
@@
 @Pei2014 used a sweep-VEP method with counterphase flickering stimuli, and found weaker responses in autistic children at spatial frequencies around 8c/deg, compared with age-matched controls.
-This was subsequently replicated in a further pediatric sample by @Vilidaite2018, who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
+This was subsequently replicated in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
 Interestingly this study replicated its key findings in a *drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.
@@
 
-(maybe here we want a 4-panel figure plotting cartoon models of the four conditions - it would make for a good comparison, and we could potentially base it on data from the original studies)
+```{r}
+
+pdf('Figures/clinicalplot.pdf',width=12,height=10)
+par(mfrow=c(2,2))
+
+finecontrastsdB <- seq(-18,36,1)
+finecontrasts <- 10^(finecontrastsdB/20)
+ticklocsx <- seq(-18,36,6)
+ticklabelsx <- c('1/8','1/4','1/2',1,2,4,8,16,32,64)
+
+# fits to data from Tsai et al. (2011)
+epilepsyparams <- matrix(0,nrow=2,ncol=4)
+epilepsyparams[1,] <- c(3.4,0.15,1.4,1)  # controls
+epilepsyparams[2,] <- c(5.3,0.23,1.4,1)  # patients
+
+respEPIcontrols <- epilepsyparams[1,1] * ((finecontrasts/100)^epilepsyparams[1,3]/(epilepsyparams[1,2]^epilepsyparams[1,3] + (finecontrasts/100)^epilepsyparams[1,3])) + epilepsyparams[1,4]
+
+respEPIpatients <- epilepsyparams[2,1] * ((finecontrasts/100)^epilepsyparams[2,3]/(epilepsyparams[2,2]^epilepsyparams[2,3] + (finecontrasts/100)^epilepsyparams[2,3])) + epilepsyparams[2,4]
+
+ticklocsy <- seq(0,6,2)
+ticklabelsy <- ticklocsy
+
+par(pty="s")  # make axis square
+plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,6))   
+axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
+mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
+mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
+title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
+title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+title(main='(a) Epilepsy',cex.main=2)
+
+lines(finecontrastsdB,respEPIpatients,lwd=4,col='blue')
+lines(finecontrastsdB,respEPIcontrols,lwd=4,lty=2,col='black')
+
+legend(-18,6,c('Epilepsy patients','Control participants'),lwd=4,lty=c(1,2),col=c('blue','black'),box.lwd=2)
+
+ticklocsy <- seq(0,6,2)
+ticklabelsy <- ticklocsy
+
+par(pty="s")  # make axis square
+plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,6))   
+axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
+mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
+mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
+title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
+title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+title(main='(b) Amblyopia',cex.main=2)
+
+# fits to data from Vilidaite et al (2018)
+autismparams <- matrix(0,nrow=2,ncol=4)
+autismparams[1,] <- c(0.6253,0.2701,125550,5.6199)  # ASD group
+autismparams[2,] <- c(1.0131,0.3369,3589900,6.8622) # control group
+
+respASD <- (autismparams[1,1]*(finecontrasts^autismparams[1,4])/(autismparams[1,3] + finecontrasts^autismparams[1,4])) + autismparams[1,2]
+respCON <- (autismparams[2,1]*(finecontrasts^autismparams[2,4])/(autismparams[2,3] + finecontrasts^autismparams[2,4])) + autismparams[2,2]
+
+ticklocsy <- seq(0,1.5,0.5)
+ticklabelsy <- ticklocsy
+
+par(pty="s")  # make axis square
+plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,1.5))   
+axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
+mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
+mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
+title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
+title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+title(main='(c) Autism',cex.main=2)
+
+lines(finecontrastsdB,respASD,lwd=4,col='darkgreen')
+lines(finecontrastsdB,respCON,lwd=4,lty=2,col='black')
+
+legend(-18,1.5,c('ASD children','Control children'),lwd=4,lty=c(1,2),col=c('darkgreen','black'),box.lwd=2)
+
+ticklocsy <- seq(0,6,2)
+ticklabelsy <- ticklocsy
+
+par(pty="s")  # make axis square
+plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,6))   
+axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
+mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
+mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
+title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
+title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+title(main="(d) Parkinson's",cex.main=2)
+
+dev.off()
+
+```
+
+![caption.](Figures/clinicalplot.pdf){#fig-plotclinical}
 

### 2024-04-07 review.qmd
Added python for clinical figure. Commented R (but you can uncomment it!)

@@
 Parkinson's
+```{python}
+import numpy as np
+import matplotlib.pyplot as plt
 
+# Parameters
+finecontrastsdB = np.arange(-18, 37, 1)
+finecontrasts = 10**(finecontrastsdB / 20)
+ticklocsx = np.arange(-18, 37, 6)
+ticklabelsx = ['1/8', '1/4', '1/2', 1, 2, 4, 8, 16, 32, 64]
+
+epilepsyparams = np.array([
+    [3.4, 0.15, 1.4, 1],  # controls
+    [5.3, 0.23, 1.4, 1]   # patients
+])
+
+# Calculate responses
+respEPIcontrols = epilepsyparams[0,0] * ((finecontrasts/100)**epilepsyparams[0,2] /
+                (epilepsyparams[0,1]**epilepsyparams[0,2] + (finecontrasts/100)**epilepsyparams[0,2])) + epilepsyparams[0,3]
+
+respEPIpatients = epilepsyparams[1,0] * ((finecontrasts/100)**epilepsyparams[1,2] /
+                (epilepsyparams[1,1]**epilepsyparams[1,2] + (finecontrasts/100)**epilepsyparams[1,2])) + epilepsyparams[1,3]
+
+plt.figure(figsize=(12, 10))
+plt.subplot(2, 2, 1)
+plt.plot(finecontrastsdB, respEPIpatients, lw=4, color='blue', label='Epilepsy patients')
+plt.plot(finecontrastsdB, respEPIcontrols, lw=4, color='black', linestyle='--', label='Control participants')
+plt.xticks(ticklocsx, ticklabelsx)
+plt.yticks(np.arange(0, 7, 2))
+plt.xlabel('Target contrast (%)')
+plt.ylabel('SSVEP response')
+plt.title('(a) Epilepsy')
+plt.legend()
+
+# Amblyopia parameters would follow a similar structure if they were provided.
+# Since they're not detailed in your R code, I'll move on to the autism plot.
+
+# Autism and control groups for the autism plot
+autismparams = np.array([
+    [0.6253, 0.2701, 125550, 5.6199],   # ASD group
+    [1.0131, 0.3369, 3589900, 6.8622]   # control group
+])
+
+respASD = (autismparams[0,0] * (finecontrasts**autismparams[0,3]) /
+           (autismparams[0,2] + finecontrasts**autismparams[0,3])) + autismparams[0,1]
+respCON = (autismparams[1,0] * (finecontrasts**autismparams[1,3]) /
+           (autismparams[1,2] + finecontrasts**autismparams[1,3])) + autismparams[1,1]
+
+plt.subplot(2, 2, 3)  # Assuming this is the third plot for Autism
+plt.plot(finecontrastsdB, respASD, lw=4, color='darkgreen', label='ASD children')
+plt.plot(finecontrastsdB, respCON, lw=4, color='black', linestyle='--', label='Control children')
+plt.xticks(ticklocsx, ticklabelsx)
+plt.yticks(np.arange(0, 1.6, 0.5))
+plt.xlabel('Target contrast (%)')
+plt.ylabel('SSVEP response')
+plt.title('(c) Autism')
+plt.legend()
+
+# Assuming Parkinson's parameters are similar to those given for epilepsy/autism,
+# but with specifics for Parkinson's. We'll use the epilepsy parameters as a placeholder.
+
+# Note: Replace epilepsyparams with Parkinson's specific parameters as necessary.
+respParkControls = epilepsyparams[0,0] * ((finecontrasts/100)**epilepsyparams[0,2] /
+                      (epilepsyparams[0,1]**epilepsyparams[0,2] + (finecontrasts/100)**epilepsyparams[0,2])) + epilepsyparams[0,3]
+
+respParkPatients = epilepsyparams[1,0] * ((finecontrasts/100)**epilepsyparams[1,2] /
+                      (epilepsyparams[1,1]**epilepsyparams[1,2] + (finecontrasts/100)**epilepsyparams[1,2])) + epilepsyparams[1,3]
+
+plt.subplot(2, 2, 4)  # Fourth plot for Parkinson's
+plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='Parkinson\'s patients')
+plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='Control participants')
+plt.xticks(ticklocsx, ticklabelsx)
+plt.yticks(np.arange(0, 7, 2))
+plt.xlabel('Target contrast (%)')
+plt.ylabel('SSVEP response')
+plt.title("(d) Parkinson's")
+plt.legend()
+# To save the figure
+plt.savefig('clinicalplot.pdf')
+```
+
+<!---
 ```{r}
@@
 title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+
 title(main='(b) Amblyopia',cex.main=2)
@@
 
-```
+``` commented out R because of the weird reticulate bug...--->
 

### 2024-04-07 review.tex
Added python for clinical figure. Commented R (but you can uncomment it!)

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Measuring contrast processing in the visual system using SSVEP},
  pdfauthor={Alex R. Wade; Daniel H. Baker},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Measuring contrast processing in the visual system using SSVEP}
\author{Alex R. Wade \and Daniel H. Baker}
\date{}

\begin{document}
\maketitle

Department of Psychology and York Biomedical Research Institute,
University of York, UK

\subsection{Review for Vis Neurosci}\label{review-for-vis-neurosci}

\section{Abstract:}\label{abstract}

Contrast is the currency of the early visual system. Measuring the way
that the computations underlying contrast processing depend on factors
such as spatial and temporal frequency, age, eccentricity, chromaticity
and the presence of other stimuli has been a focus of vision science for
at least 100 years. One of the most productive experimental approaches
in this field has been the use of the `steady-state visually-evoked
potential' (SSVEP): a technique where contrast modulating inputs are
`frequency tagged' (presented at well-defined frequencies and phases)
and the electrical signals that they generate in the brain are analyzed
in the temporal frequency domain. SSVEPs have several advantages over
conventional measures of visually-evoked responses: they have relatively
unambiguous ouput measures, a high SNR, and they allow us to analyze
interactions between stimulus components using a convenient mathematical
framework. Here we describe how SSVEPs have been used to study visual
contrast over the past 70 years (Dawson, 1954). Because our thinking
about SSVEPs is well-described by simple mathematical models, we embed
code that illustrates key steps in the modelling and analysis. This
paper can therefore be used both as a review of the use of SSVEP in
measuring human contrast processing, and as an interactive learning aid.

\section{Introduction}\label{introduction}

Neurons in the visual areas of the brain are primarily responsive to
changes in cone photoreceptor activations across time and space. This
property, referred to as contrast, sets the fundamental limits of our
visual abilities, which remain steady over a remarkably wide range of
environmental light levels. The human response to contrast can be
studied using many different techniques. Early work used psychophysical
methods to measure contrast sensitivity (Campbell \& Green, 1965),
defined as the inverse of the lowest contrast that can be reliably
detected. But neural responses can also be measured more directly using
techniques such as fMRI, MEG, and EEG. Here we will describe how an EEG
method known as the steady state visually evoked potential (SSVEP)
technique has contributed to our understanding of human contrast
processing in health, disease and throughout development.

The SSVEP is a continuous electrical response evoked in the brain by
visual stimuli flickering at a constant frequency (Regan, 1966). For
contrast-defined stimuli, such as sine-wave gratings, it is strongest at
the occipital pole, adjacent to the early visual areas that generate the
signal, although careful analysis of individual VEPs reveals multiple
generators throughout visual cortex (Di Russo \emph{et al.}, 2005,
2007). The flickering stimulus entrains neural population responses at
multiples of the stimulus frequency, so continuous EEG data are
typically analysed by taking the Fourier transform, and estimating the
amplitude at these frequencies. Two common variants involve sinusoidal
on-off flicker, where the stimulus alternates between a blank background
and the peak contrast, and sinusoidal counterphase flicker, where the
stimulus alternates in phase (i.e the black regions become white and the
white regions become black). On-off flicker can drive independent
populations of on- and off-cells in the retina once per cycle and can
therefore produce a response at the fundamental flicker frequency, known
as 1F, and its integer harmonics: 2F, 3F, 4F and so on. Counterphase
flicker contains two transients per cycle and therefore does not produce
a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
Because square-waves are spectrally broad-band, square wave flicker
tends to produce more complex spectral harmonics than sine-wave flicker.

The higher harmonics of the steady-state signal are generally thought to
reflect nonlinear processing in the visual system (Regan \& Regan,
1988). SSVEP signals can also be elicited by periodic changes of
stimulus properties other than achromatic and chromatic contrast, such
as motion, stereo depth, and facial identity or expression (see Norcia
\emph{et al.}, 2015, for an overview); however our focus here is on the
contrast response.

\section{Why measure responses to
contrast?}\label{why-measure-responses-to-contrast}

Contrast is one of the most fundamental pieces of information that the
eye transmits to the brain. It can be defined as the change in cone
photoreceptor activity over space (`spatial contrast') or time
(`temporal contrast'). Cone photoreceptors - which drive precortical
opponent pathways - contribute to both chromatic and achromatic contrast
and although most of the research we describe here focuses on achromatic
contrast, SSVEPs have proven to be an excellent measure of early
chromatic processing as well (McKeefry \emph{et al.}, 1996; Baseler \&
Sutter, 1997; Di Russo \emph{et al.}, 2001).

Contrast is relatively simple to define: typically, contrast is
specified as the percentage deviation of a uniform stimulus from the
background. So, for example, a disk of 100 units of cone activation
(\(I_{\mathrm{stim}}\)) surrounded by a `background' of 50 units of
activation (\(I_{\mathrm{background}}\)) has a contrast of
\(\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}\)
= 100\%. Where patterns are more complex (for example, the sine-wave
gratings or Gabor patches common in vision science), the Michelson
(1927) definition of contrast is specified by the maximum and minimum
excursions from the mean:

\begin{equation}
\frac {I_{\mathrm{stimmax} }-I_{\mathrm{stimmin} }}{I_{\mathrm{stimmax} }+I_{\mathrm{stimmin} }}.
\end{equation}

These contrast definitions are appropriate both to photometric measures
of stimulus contrast (for example, luminance; Lennie \emph{et al.}
(1993)) and also to definitions based on cone excitations (MacLeod \&
Boynton, 1979; Derrington \emph{et al.}, 1984) which are more common in
work on chromatic processing.

Although its mathematical definition is straightforward, the
computations that underlie contrast processing in the brain have been
the subject of intense research for many decades. The neural code for
contrast, even in the earliest parts of visual cortex, is not simply a
linear transform of the contrast at the retina - instead, contrast
signals undergo a cascade of nonlinear processing stages that, broadly,
attempt to normalise the output relative to the spatiotemporal
environment. This normalization, achieved through a computation called
`contrast gain control' (Heeger, 1992; Foley, 1994; Carandini \& Heeger,
2011) maximises the sensitivity of the visual system by making optimal
use of neuronal bandwidth. As an example, a grating placed at the centre
of a low-contrast background typically appears more intense than the
same grating when superimposed on a high contrast background (see
Figure~\ref{fig-centresurround}).

\begin{figure}

\centering{

\includegraphics{review_files/figure-pdf/fig-centresurround-output-1.pdf}

}

\caption{\label{fig-centresurround}The perceived contrast of a stimulus
depends on its context. A high contrast surround reduces the apparent
contrast of the central `probe' region''}

\end{figure}%

A significant body of research into contrast processing is concerned
with how these normalization mechanisms depend on colour (Chen \emph{et
al.}, 2000), orientation (Foley, 1994), eye of origin (Legge, 1979;
Baker \emph{et al.}, 2007), spatial and temporal frequency (Meese \&
Baker, 2009), location (Polat \& Sagi, 1993; Tadin \emph{et al.}, 2003;
Petrov \emph{et al.}, 2005), age (Betts \emph{et al.}, 2005), and the
presence of neurological disorders (Porciatti \emph{et al.}, 2000; Tsai
\emph{et al.}, 2011). The SSVEP has proven to be invaliable in this
research because it provides an objective readout of contrast
representation at different stages of the visual system, and allows us
to `tag' the probe and background at separate frequencies.

Because it provides a direct read-out of neural population activity, the
SSVEP signal can reveal key features of neural signal transduction. For
example, by varying the peak stimulus contrast parametrically, a
`contrast vs response' function can be measured - where the `response'
is typically defined as the amplitude of the SSVEP frequency component
at the stimulus frequency or a low multiple thereof. This corresponds
closely to similar functions reported by invasive studies measuring
single unit activity or local field potentials in the cortex (Shapley \&
Victor, 1980; Morrone \emph{et al.}, 1982). However the SSVEP has the
advantage that it is non-invasive, and so can be measured in awake,
behaving human participants.

To understand the utility of the contrast SSVEP, it is helpful to
identify the cascade of processing stages in the early visual system
that give rise to it. In the following section we illustrate how a
typical SSVEP signal measured over early visual cortex might contain
information about a large number of early visual computations.

\section{Contrast processing - linear and
nonlinear}\label{contrast-processing---linear-and-nonlinear}

Neurons have a limited dynamic range, yet they can transmit information
about visual stimuli that span many orders of magnitude. In the domain
of contrast, to some extent this is accomplished at a population level -
individual neurons typically implement a non-linear, sigmoidal contrast
vs response (CRF) transducer (Tolhurst \emph{et al.}, 1981; Albrecht \&
Hamilton, 1982) and different neurons exhibit peak sensitivity (defined
as the maximum slope of the function) at different contrast levels
(Carandini \& Heeger, 1994; Carandini \emph{et al.}, 1998; Busse
\emph{et al.}, 2009). A neuronal population will therefore span a
sensitivity range greater than any individual member.

Individual neurons at multiple stages of the visual hierarchy also
change their sensitivity depending on the average spatiotemporal
contrast energy of their environment. This ``normalisation'' process is
dynamic and nonlinear and is well-modeled by a hyperbolic ratio function
in which the response of each neuron is modulated by a local `gain pool'
composed of the summed responses of the local neuronal population
(Heeger, 1992; Busse \emph{et al.}, 2009; Carandini \& Heeger, 2011;
Baker \& Wade, 2017).

Previous figures showed how a single-input SSVEP might generate a single
set of peaks in the frequency domain response which, in turn, can be
used as a read-out of neuronal activity. In
Figure~\ref{fig-plotspectra}, we show the effect of adding a second
contrast component: the two components interact generating nonlinear
intermodulation terms at sums and differences of the input frequencies
(F1, F2). The nature of the interaction --- and therefore the pattern of
intermodulation terms --- is determined by the computations happening as
the contrast signal moves from retina to cortex (Regan \& Regan, 1988)

First, we implement a simple non-linear transducer with a single input
to illustrate the typical sigmoidal shape of a contrast response
function (Figure~\ref{fig-contrastresponsefunction}).

Next, we show that adding a second component which contributes to the
gain pool of the first, results in a complex pattern of intermodulation
terms (Figure~\ref{fig-plotspectra}).

\begin{figure}

\centering{

\includegraphics{review_files/figure-pdf/fig-plotsinewaves-output-1.pdf}

}

\caption{\label{fig-plotsinewaves}a) A typical SSVEP setup: a screen, a
grating, a flicker rate: b) The responses from a single electrode (time
domain) c) Frequency domain}

\end{figure}%

The two inputs are represented here as amplitude modulations across
time.

If the two inputs are simply added together, the representation of the
resulting signal in the Fourier domain is just the linear sum of the two
independent signals (first two panels of Figure~\ref{fig-plotspectra}).
However, it is important to take physiology into account. If we ignore
cells sensitive to contrast transients, contrast is represented as a
combination of rectified signals from `on' and `off'-sensitive cells
(broadly, cells that code positive or negative changes in spatiotemporal
luminance). To model the population response of these cells to a
contrast-reversing luminance sine wave, we therefore introduce an
initial full-wave rectification stage that simulates the outputs of two,
equal populations of on- and off-polarity detectors. Once this is done,
we see that the Fourier spectrum of the resulting signal contains power
only at the second harmonics and above (final two panels of
Figure~\ref{fig-plotspectra}). It is possible to generate first-hamonic
responses in contrast-response SSVEPs, but only when the inputs are
modulated in an `on/off' manner rather than contrast reversed.

\begin{figure}

\centering{

\includegraphics{review_files/figure-pdf/fig-plotspectra-output-1.pdf}

}

\caption{\label{fig-plotspectra}Temporal frequency representations of
combining two inputs at frequencies F1=5Hz and F2=7Hz combination. A
pure sine-wave input would generate responses only at F1 and F2.
However, the rectification that accompanies early visual processing
converts this contrast reversal into a more complex frequency-domain
representation with power predominantly at the second harmonics 2F1 and
2F2}

\end{figure}%

Even without considering a spatial component, the early visual system is
far more complex than the model here suggests. For example as well as
cells that code positive or negative contrast in a more or less
continuous manner, the retina also contains `transient' cells that code
temporal changes in contrast. These cells (Kuffler \emph{et al.}, 1957;
Alpern, 1971), and cells with similar properties in the LGN (Levitt
\emph{et al.}, 2001) and cortex (Hubel \& Wiesel, 1959; Movshon, 1975),
will introduce second harmonic components even when the stimulus itself
is modulated in an on-off fashion. Analogously, in the spatial domain,
so-called `simple cells' are sensitive to the polarity of a spatial
contrast modulation while `complex cells' respond to the presence of
patterned spatial contrast irrespective of its spatial position. The
SSVEP response to a contrast-reversing sine-wave grating therefore
contains information about nonlinear computations performed across a
range of retinal and cortical cell types.

\begin{figure}

\centering{

\includegraphics{review_files/figure-pdf/fig-contrastresponsefunction-output-1.pdf}

}

\caption{\label{fig-contrastresponsefunction}Non-linear contrast
response function resulting from the population average of two full-wave
rectified signals passing through a non-linear transducer. Note the
`supersaturation' at high contrast levels.}

\end{figure}%

One of the simplest nonlinearities is the function that describes a
cell's response to different levels of contrast. In
Figure~\ref{fig-contrastresponsefunction} this is modeled as described
above by a hyperbolic ratio function resulting in a saturating
non-linearity:

\begin{equation}
R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},
\end{equation}

where \({R_{\mathrm{max}}}\) describes the maximum response level,
\(C_{\mathrm{in}}\) is the input contrast, \(C_{50}\) is the
`semi-saturation constant' (the point at which the response is at
half-maximum) and \emph{n} controls the steepness of the curve.

The hyperbolic ratio function is monotonic but the CRF resulting from
measuring the amplitude of the 2F component contains a slight roll-off
at high input contrasts. This results from the distortion of the input
sine waves at high contrast due to a combination of the full-wave
rectification and saturating non-linearity. Power at other harmonics
increases, and the total power generated by the input is monotonic. This
roll-off is often seen in experimental data and has been referred to as
`supersaturation' (Tyler \& Apkarian, 1985; Peirce, 2007).

To illustrate the effect of contrast gain control (Heeger, 1992), we
include a second component that contributes to the gain pool of the
first. At a single pair of (matched) contrast levels, we see a complex
pattern of intermodulation terms in the FT
(Figure~\ref{fig-intermodulation}).

\begin{figure}

\centering{

\includegraphics{review_files/figure-pdf/fig-intermodulation-output-1.pdf}

}

\caption{\label{fig-intermodulation}Non-linear combination of two
independent contrast modulations at F1 and F2 results in
`intermodulation' terms that appear at low-order sums and differences of
those frequencies}

\end{figure}%

The complexity of even a simple simulation of the frequency-domain
signal arising from non-linear interactions is intriguing. Presumably,
the signal measured from early visual cortex contains is the result of a
cascade of nonlinear retinal and cortical operations up to that point.
It therefore contains a `signature' of the computational nature, order
and parameters of those operations - including the shape of the
transducer functions and the computations involved in signal
combination. In principle, that information could be recovered from the
SSVEP signal.

This possibility was recognised in the early days of the technique
(Regan \& Regan, 1988). Although characterising the complete set of
computations along entire processing pathway is problematic, careful
parametric variation of the input stimuli does allow us to fit models of
early visual gain control using SSVEP data. \ldots{} MORE HERE -
Although this idea was suggested as early as 1988 (probably earlier) not
many people seem to have done it much. Candy \emph{et al.} (2001); Busse
\emph{et al.} (2009); Tsai \emph{et al.} (2012); Baker \& Wade (2017)

\section{Measuring the development of contrast
processing}\label{measuring-the-development-of-contrast-processing}

An early use of the SSVEP was to provide an objective estimate of
spatial contrast sensitivity in infants, without requiring behavioural
responses. In well-motivated adults, psychophysical measurements of
contrast sensitivity remain the gold standard. However, it is difficult
and time consuming to obtain reliable psychophysical data from infants.
In these cases, SSVEP measurements represent a fast and efficient method
for measuring low-level visual responses (Tyler \emph{et al.}, 1979;
Braddick \emph{et al.}, 1986; Norcia \emph{et al.}, 1990) and the high
SNR of SSVEP means that infants need only look at the screen for short
periods of time.

Because SSVEP responses at detection threshold are very small,
estimating threshold is achieved by measuring the contrast response
function at relatively high levels, and extrapolating back along the
function (either contrast vs response measured at a constant spatial
frequency or spatial frequency vs response at a constant contrast level)
to estimate its intercept with the x-axis (see
Figure~\ref{fig-sweepvep}). This contrast level was shown to correspond
approximately with psychophysically measured detection thresholds
(Norcia \emph{et al.}, 1986).

\begin{figure}

\centering{

\includegraphics{review_files/figure-pdf/fig-sweepvep-output-1.pdf}

}

\caption{\label{fig-sweepvep}Sweep VEP simulation showing how a contrast
detection threshold can be estimated from sweep VEP data measured at
higher contrasts}

\end{figure}%

Measuring the SSVEP at multiple levels was made faster by the
development of the `sweep VEP' paradigm in which the stimulus changed
its contrast, spatial frequency or some other property, throughout a
trial (Tyler \emph{et al.}, 1979). To avoid hysteresis effects, the
sweep is sometimes conducted both up and down in the same experiment
Norcia \& Tyler (1985\emph{a}). The sweep VEP (really, a sweep `SSVEP')
technique is now commonly-used to obtain a rapid and objective
measurement of visual acuity. In particular, because of its relative
speed and simplicity, this technique has now become a standard for
conducting tests of visual acuity in very young subjects or where
behavioural tests are not appropriate (Ridder, 2004; Bach \emph{et al.},
2008; Hoffmann \emph{et al.}, 2017; Bach \& Farmer, 2020).

This approach has revealed much about the development of visual
abilities in infants (Harris \emph{et al.}, 1976; Atkinson \emph{et
al.}, 1979; Braddick \emph{et al.}, 1986). In general, SSVEP
measurements of infant vision have revealed that contrast sensitivity
for both achromatic and chromatic contrast as well as stereoscopic depth
perception develops earlier than had been supposed previously based on
behavioural readouts (Dobson \emph{et al.}, 1978; Norcia \& Tyler,
1985\emph{b}) with both chromatic and achromatic contrast detection
reaching near-adult levels by around six months. Spatial acuity as
measured by SSVEP reaches adult levels more slowly, but near-adult
levels are recorded around 1 year (Norcia \& Tyler, 1985\emph{b}; Hamer
\emph{et al.}, 1989) compared to around 6-7 years with behavioural
measures (Atkinson \& Braddick, 1983; Ellemberg \emph{et al.}, 1999). At
least some of this difference is likely due to the relative objectivity
and high SNR of the SSVEP technique compared to other methods such as
preferential looking, which require careful measurement of the infant's
gaze direction. However it should be noted that other groups have
reported electrophysiological correlates of visual acuity that more
closely match the behavioural measures (De Vries-Khoe \& Spekreijse,
1982).

The SSVEP technique has also been used to study the development of the
contrast gain control mechanisms described in the previous section
(Candy \emph{et al.}, 2001; Pei \emph{et al.}, 2017). Although contrast
gain control is measurable in infants as young as six weeks old (Morrone
\& Burr, 1986; Skoczenski \& Norcia, 1998; Candy \emph{et al.}, 2001),
its development appears to be slower, with adult levels being reached at
approximately 11 years (Pei \emph{et al.}, 2017).

\section{Contrast processing in clinical
conditions}\label{contrast-processing-in-clinical-conditions}

The SSVEP technique has also been used to study clinical conditions,
such as diseases and developmental disorders. This can often be
informative regarding the underlying mechanism that characterises the
condition. Here we focus on four conditions, but there is potential to
apply the method more broadly, either as a diagnostic technique, or to
monitor disease severity and progression, or the efficacy of treatments.

Epilepsy is a neurological condition in which patients experience
seizures - episodes of uncontrolled neural activity that can cause
unconsciousness, involuntary movements and convulsions, and atypical
sensory experiences. Porciatti \emph{et al.} (2000) showed that
individuals with photosensitive epilepsy generate larger steady-state
signals in response to flickering visual stimuli, and their contrast
response functions saturate less than those of healthy controls. This is
consistent with the idea that epilepsy involves a cortical
hyperexciteability that makes seizures more likely. It is also the case
for individuals with idiopathic generalised epilepsy (Tsai \emph{et
al.}, 2011), a subtype of epilepsy that has a less obvious link to
vision. The differences apply across the whole contrast-response
function, and so resemble a response gain effect, which might be due to
reduced inhibition from neighbouring neurons. Differences in SSVEP
amplitudes have also been reported in individuals with migraine (Shibata
\emph{et al.}, 2008), a condition also associated with cortical
hyperexciteability.

Amblyopia is a disorder of binocular vision, characterised by one eye
contributing much less to perception than the other. This is often due
to strabismus (squint) or anisometropia (difference in optical
prescription between the eyes) during development. Contemporary accounts
suggest that the amblyopic eye is suppressed by signals from the fellow
eye. SSVEPs provide a convenient and objective method to characterise
the difference in neural response to signals in each eye, and typically
show reduced responses to stimuli in the amblyopic eye across the
contrast range (Baker \emph{et al.}, 2015; Lygo \emph{et al.}, 2021).
There are currently many novel binocular treatments for amblyopia under
development, often involving virtual reality or stereo display systems
designed to encourage the two eyes to work together. The steady-state
approach may be more sensitive and objective than typical acuity
measurements, and also has the potential to measure suppression between
the eyes directly (e.g. Zheng \emph{et al.}, 2019; Hu \emph{et al.},
2023; Du \emph{et al.}, 2023).

Autism is a condition often associated with differences in vision
(Simmons \emph{et al.}, 2009) and other senses (MacLennan \emph{et al.},
2022). Pei \emph{et al.} (2014) used a sweep-VEP method with
counterphase flickering stimuli, and found weaker responses in autistic
children at spatial frequencies around 8c/deg, compared with age-matched
controls. This was subsequently replicated in a further pediatric sample
by Vilidaite \emph{et al.} (2018) (see Figure~\ref{fig-plotclinical}c),
who additionally found weaker responses in autistic adults at the second
harmonic (using on/off flicker). Interestingly this study replicated its
key findings in a \emph{drosophila} genetic model of autism (\emph{Nhe3}
mutations), illustrating the translational potential of the steady-state
approach, as well as identifying a possible biomarker for autism.

Parkinson's

\includegraphics{review_files/figure-pdf/cell-8-output-1.pdf}

\begin{figure}

\centering{

\includegraphics{Figures/clinicalplot.pdf}

}

\caption{\label{fig-plotclinical}caption.}

\end{figure}%

\section{Brain-computer interfaces}\label{brain-computer-interfaces}

\section{Conclusions}\label{conclusions}

\emph{2: Sweeps and CRFs. Measurements of contrast sensitivity,
extrapolating the sweep to zero response to get t'hold. Infants and
adults?}

\emph{3: Measurements of modulation. Figures adapted from other papers:
Attention to space, attention to features, adaptation (?).
masking/surround suppression\textbackslash{}}

\emph{4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)}

\emph{5: Future directions (decoding in frequency domain?, animals?
BCI?)}

\phantomsection\label{anchor-2}{}History

\phantomsection\label{anchor-3}{}1. The basics of SSVEP and contrast
sensitivity including a history of both fields

SSVEP (Steady-State Visual Evoked Potential): A continuous electrical
response evoked in the brain by visual stimuli flickering at a constant
frequency (Regan, 1966).

Contrast Sensitivity: The ability to detect differences in luminance
between an object and its background (Campbell \& Green, 1965).

Regan, D. (1966). Some characteristics of average steady-state and
transient responses evoked by modulated light. Electroencephalography
and clinical neurophysiology, 20(3), 238-248.

Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
affecting visual resolution. J Physiology, 181, 576-593.

Check Regan paper for earlier (e.g.~EEG refs). Norcia review will be
helpful!

(From Tyler / Levi / Apkarian paper):

6. Regan, D.: Rapid methods for refracting the eye and assessing the
visual acuity in amblyopia using steady-state visual evoked potentials.
In Desmedt, J.E., editor: Visual Evoked Potentials in Man: New
Developments, Oxford, 1977, Clarendon Press, pp.~418-426.

7. Fricker, S.J.: Narrow-band filter techniques for the detection and
measurement of evoked responses, Electroencephalogr. Clin. Neurophysiol.
14:411, 1962.

8. Van der Tweel, L. H., Sem-Jacobsen, C.W., Kamp, A., Van Leeuwen,
W.S., and Verings, F.T.H.: Objective determination of response to
modulated light, Acta Physiol. Pharmacol. Neerl. 7:528, 1958.

9. Regan, D.: Latencies of evoked potentials to flicker and to pattern
speedily estimated by simultaneous stimulation method,
Electroencephalogr. Clin. Neurophysiol. 40:654, 1976.

10. Tyler, C.W., Apkarian, P., and Nakayama, K.: Multiple spatial
frequency tuning of electrical responses from the human.\ldots{}

{[}Around here perhaps a section about what gain control is, mentioning
other methods as well including psychophysics, MRI, electrophysiology,
and other EEG markers including ERPs and evoked gamma band oscillations.
Maybe outline the Heeger gain control model and its cousins.{]}

{[}Sure. But in fact gain control is only a small part of this story -
especially in the early days. They were, I think, more interested in
using SSVEP to measure absolute contrast sensitivity and to get lower
bounds on things like infant visual development. Gain control might be
better as a separate section later.{]}

\phantomsection\label{anchor-4}{}2. Spatial, temporal frequency and
contrast sensitivity measurements

Basically using SSVEP to measure a cortical output amplitude for any
given input contrast. You can vary parameters like SF, TF, position,
color and of course contrast. Early on people realised that you can
`sweep' the stimulus to get a CRF. You broadly get a line in log
contrast space if you do that (Tyler) - then you can extrapolate that
line down to zero response to estimate the threshold. That doesn't
\textgreater quite\textless{} work but it's pretty close.

You can also use this for `difficult' populations like babies. One
interesting story was about how SSVEP become a replacement for
preferential looking (which was the other way of looking at infant
visual development). See e.g.~Davida Teller. SSVEPs allowed people to
make objective measurements of contrast sensitivity development and
deduce that the visual system was more mature (e.g.~more functional) in
infancy than previously expected. Also measures of colour sensitivity.
Tinyeyes is based off those measurements. Other people: Tyler, Norcia,
Gunilla H-P, many of the people at SKERI in the 1980s and 90s. Norcia
86,88,90 - mentioned in Regan's nice autobiography :

In parallel of course, people were using frequency tagging to do single
unit work - the 1F vs 2F simple/complex cell classification scheme was
all about this (Lennie and others).

\phantomsection\label{anchor-5}{}3. SSVEP in functional localization

Techniques like fMRI have been combined with SSVEP to achieve more
precise spatial localization (Di Russo et al., 2007). See however the
Ales cruciform paper.

Ales and Norcia:... (showed that people's intuition about V1 upper /
lower v.f reversing polarity wasn't really correct). This feels a little
outside our scope though... We have also combined SSVEP with source
imaging techniques to probe responses in different cortical locations
(again, Ales papers, some early stuff from Stan K? Appelbaum, Wade,
Norcia figure/ground...) and then a host of later work from that lab and
others.

\phantomsection\label{anchor-6}{}4. SSVEP and contrast gain control
including adaptation, masking, and attention

{[}The story goes: Up to the lte 90s people were primarily interested in
measuring contrast sensitivity - the shape of the response function was
assumed to be basically log-linear - and they fit it with straight lines
to extrapolate back to zero response. But then (once Heeger's 1992 paper
had sunk in - see also stuff like Shapley and Victor 1981
\href{https://www.zotero.org/google-docs/?zAmtgj}{(Shapley \& Victor,
1981)}), people started thinking about gain control - Candy and Norcia
in about 1999, Porciatti, probably a load of Tyler papers that I don't
even know about...{]}. Probably look in p

And then people worked out that if you can use SSVEP to measure contrast
responses, you can also use it to measure things that modulate contrast
responses. These include adaptation, masking, suppression, attention
(feature and space), clinical things.

\phantomsection\label{anchor-7}{}Adaptation:

Continuous exposure to high-contrast patterns reduces contrast
sensitivity, which can be measured using SSVEP (Ross et al., 1989).
Others? Baker recent gain control paper is perhaps worth mentioning here
as a `confound' of sorts. Engel 2018
\href{https://www.zotero.org/google-docs/?kdURiS}{\emph{(Vergeer et al.,
2018)}}. This paper is interesting
\href{https://www.zotero.org/google-docs/?J2EO3d}{\emph{(Rideaux et al.,
2023)}} and \textgreater sort\textless{} of SSVEP.

\phantomsection\label{anchor-8}{}Masking:

High contrast masks can suppress the visibility of low contrast
patterns, which has implications in SSVEP amplitude (Haynes et al.,
2003).

Attention: Directing attention can enhance contrast sensitivity, as
shown in studies using SSVEP (Müller et al., 2006). Also Tsai
(dynamics), Baker / Wade (several), Winawer? I think JW has a nice
dynamic model of normalization with some MEG data. Busse et al cat/human
comparison. Candy and Norcia 2001 JNS
\href{https://www.zotero.org/google-docs/?7hhs5Z}{(Candy et al., 2001)}

Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
adaptation and masking on incremental thresholds for contrast. Vision
research, 29(2), 205-215.

Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
Neuromagnetic correlates of perceived contrast in primary visual cortex.
Journal of Neurophysiology, 89(6), 2655-2666.

Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
selective attention on the steady-state visual evoked potential in the
20--28 Hz range. Cognitive Brain Research, 24(1), 1-13.

5. Clinical implications

\phantomsection\label{anchor-9}{}Clinical applications (this is a whole
section)

Not sure exactly how SSVEP used in clinic. mfVEP?

, e.g., in monitoring visual impairments, tracking neuronal diseases, or
neurofeedback (Norcia et al., 2015).

Tsai epilepsy \href{https://www.zotero.org/google-docs/?AdhH3f}{(Tsai et
al., 2011)}. Other photosensitive epilepsy: Porciatti
\href{https://www.zotero.org/google-docs/?UTBQbV}{(Porciatti et al.,
2000)}

Citation:

Migraine (Regan et al). Autism?

Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
Rossion, B. (2015). The steady-state visual evoked potential in vision
research: A review. Journal of vision, 15(6), 4-4.

Animal work: Flies (Eliott, West, Himmelberg, Ales, Norcia): Mice/rats
(probably many - can't think off the top of my head - we were working
with a mouse EEG person at UCSF about 12 years ago...), Monkeys
(Kiorpes?),

\phantomsection\label{anchor-10}{}6. Future directions

Use as a readout of modulations. TMS? FUS? Marmite B12 / Fluoxetine /
amblyopia in Rats,

GABA Huang

There is a lot of SSVEP interest these days because of BCIs. I think
it's pretty weak but there is \textgreater so much\textless{} of it that
it might be worth mentioning...

Advanced signal processing techniques and machine learning can be
integrated to improve SSVEP-based systems (Zhu et al., 2010).

Exploring new clinical and diagnostic applications, understanding
neurological diseases, and developing novel therapeutic interventions.

Citation:

Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
stimulation methods used in SSVEP-based BCIs. Computational intelligence
and neuroscience, 2010.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-Albrecht1982}
Albrecht DG \& Hamilton DB (1982).
\href{https://doi.org/10.1152/jn.1982.48.1.217}{Striate cortex of monkey
and cat: Contrast response function}. \emph{J Neurophysiol} \textbf{48,}
217--237.

\bibitem[\citeproctext]{ref-Alpern1971}
Alpern M (1971).
\href{https://doi.org/10.1113/jphysiol.1971.sp009580}{Rhodopsin kinetics
in the human eye}. \emph{J Physiol} \textbf{217,} 447--471.

\bibitem[\citeproctext]{ref-Atkinson1983}
Atkinson J \& Braddick O (1983).
\href{https://doi.org/10.1111/j.1755-3768.1983.tb03927.x}{Assessment of
visual acuity in infancy and early childhood}. \emph{Acta Ophthalmol
Suppl} \textbf{157,} 18--26.

\bibitem[\citeproctext]{ref-Atkinson1979}
Atkinson J, Braddick O \& French J (1979).
\href{https://www.ncbi.nlm.nih.gov/pubmed/761974}{Contrast sensitivity
of the human neonate measured by the visual evoked potential}.
\emph{Invest Ophthalmol Vis Sci} \textbf{18,} 210--213.

\bibitem[\citeproctext]{ref-Bach2020}
Bach M \& Farmer JD (2020).
\href{https://doi.org/10.1007/s10633-019-09726-2}{Evaluation of the
"freiburg acuity VEP" on commercial equipment}. \emph{Doc Ophthalmol}
\textbf{140,} 139--145.

\bibitem[\citeproctext]{ref-Bach2008}
Bach M, Maurer JP \& Wolf ME (2008).
\href{https://doi.org/10.1136/bjo.2007.130245}{Visual evoked
potential-based acuity assessment in normal vision, artificially
degraded vision, and in patients}. \emph{Br J Ophthalmol} \textbf{92,}
396--403.

\bibitem[\citeproctext]{ref-Baker2007}
Baker DH, Meese TS \& Summers RJ (2007).
\href{https://doi.org/10.1016/j.neuroscience.2007.01.030}{Psychophysical
evidence for two routes to suppression before binocular summation of
signals in human vision}. \emph{Neuroscience} \textbf{146,} 435--448.

\bibitem[\citeproctext]{ref-Baker2015}
Baker DH, Simard M, Saint-Amour D \& Hess RF (2015).
\href{https://doi.org/10.1167/iovs.14-15611}{Steady-state contrast
response functions provide a sensitive and objective index of amblyopic
deficits}. \emph{Invest Ophthalmol Vis Sci} \textbf{56,} 1208--1216.

\bibitem[\citeproctext]{ref-Baker2017}
Baker DH \& Wade AR (2017).
\href{https://doi.org/10.1093/cercor/bhw395}{Evidence for an {Optimal}
{Algorithm} {Underlying} {Signal} {Combination} in {Human} {Visual}
{Cortex}}. \emph{Cerebral Cortex (New York, NY: 1991)} \textbf{27,}
254--264.

\bibitem[\citeproctext]{ref-Baseler1997}
Baseler HA \& Sutter EE (1997).
\href{https://doi.org/10.1016/s0042-6989(96)00209-x}{M and p components
of the VEP and their visual field distribution}. \emph{Vision Res}
\textbf{37,} 675--690.

\bibitem[\citeproctext]{ref-Betts2005}
Betts LR, Taylor CP, Sekuler AB \& Bennett PJ (2005).
\href{https://doi.org/10.1016/j.neuron.2004.12.041}{Aging reduces
center-surround antagonism in visual motion processing}. \emph{Neuron}
\textbf{45,} 361--366.

\bibitem[\citeproctext]{ref-Braddick1986}
Braddick OJ, Wattam-Bell J \& Atkinson J (1986).
\href{https://doi.org/10.1038/320617a0}{Orientation-specific cortical
responses develop in early infancy}. \emph{Nature} \textbf{320,}
617--619.

\bibitem[\citeproctext]{ref-Busse2009}
Busse L, Wade AR \& Carandini M (2009).
\href{https://doi.org/10.1016/j.neuron.2009.11.004}{Representation of
concurrent stimuli by population activity in visual cortex}.
\emph{Neuron} \textbf{64,} 931--942.

\bibitem[\citeproctext]{ref-Campbell1965}
Campbell FW \& Green DG (1965).
\href{https://doi.org/10.1113/jphysiol.1965.sp007784}{Optical and
retinal factors affecting visual resolution}. \emph{J Physiol}
\textbf{181,} 576--593.

\bibitem[\citeproctext]{ref-Candy2001}
Candy TR, Skoczenski AM \& Norcia AM (2001).
\href{https://doi.org/10.1523/JNEUROSCI.21-12-04530.2001}{Normalization
models applied to orientation masking in the human infant}. \emph{The
Journal of Neuroscience: The Official Journal of the Society for
Neuroscience} \textbf{21,} 4530--4541.

\bibitem[\citeproctext]{ref-Carandini1994}
Carandini M \& Heeger DJ (1994).
\href{https://doi.org/10.1126/science.8191289}{Summation and division by
neurons in primate visual cortex}. \emph{Science} \textbf{264,}
1333--1336.

\bibitem[\citeproctext]{ref-Carandini2011}
Carandini M \& Heeger DJ (2011).
\href{https://doi.org/10.1038/nrn3136}{Normalization as a canonical
neural computation}. \emph{Nat Rev Neurosci} \textbf{13,} 51--62.

\bibitem[\citeproctext]{ref-Carandini1998}
Carandini M, Movshon JA \& Ferster D (1998).
\href{https://doi.org/10.1016/s0028-3908(98)00069-0}{Pattern adaptation
and cross-orientation interactions in the primary visual cortex}.
\emph{Neuropharmacology} \textbf{37,} 501--511.

\bibitem[\citeproctext]{ref-Chen2000}
Chen C, Foley JM \& Brainard DH (2000).
\href{https://doi.org/10.1016/s0042-6989(99)00227-8}{Detection of
chromoluminance patterns on chromoluminance pedestals i: Threshold
measurements}. \emph{Vision Res} \textbf{40,} 773--788.

\bibitem[\citeproctext]{ref-Dawson1954}
Dawson GD (1954). \href{https://doi.org/10.1016/0013-4694(54)90007-3}{A
summation technique for the detection of small evoked potentials}.
\emph{Electroencephalography and Clinical Neurophysiology} \textbf{6,}
65--84.

\bibitem[\citeproctext]{ref-De-Vries-Khoe1982}
De Vries-Khoe L \& Spekreijse H (1982). Maturation of luminance and
pattern EPs in man. \emph{Doc Ophthalmol Proc Ser} \textbf{31,}
461--475.

\bibitem[\citeproctext]{ref-Derrington1984}
Derrington AM, Krauskopf J \& Lennie P (1984).
\href{https://doi.org/10.1113/jphysiol.1984.sp015499}{Chromatic
mechanisms in lateral geniculate nucleus of macaque}. \emph{J Physiol}
\textbf{357,} 241--265.

\bibitem[\citeproctext]{ref-Di-Russo2007}
Di Russo F, Pitzalis S, Aprile T, Spitoni G, Patria F, Stella A,
Spinelli D \& Hillyard SA (2007).
\href{https://doi.org/10.1002/hbm.20276}{Spatiotemporal analysis of the
cortical sources of the steady-state visual evoked potential}. \emph{Hum
Brain Mapp} \textbf{28,} 323--334.

\bibitem[\citeproctext]{ref-Di-Russo2005}
Di Russo F, Pitzalis S, Spitoni G, Aprile T, Patria F, Spinelli D \&
Hillyard SA (2005).
\href{https://doi.org/10.1016/j.neuroimage.2004.09.029}{Identification
of the neural sources of the pattern-reversal VEP}. \emph{Neuroimage}
\textbf{24,} 874--886.

\bibitem[\citeproctext]{ref-Di-Russo2001}
Di Russo F, Spinelli D \& Morrone MC (2001).
\href{https://doi.org/10.1016/s0042-6989(01)00134-1}{Automatic gain
control contrast mechanisms are modulated by attention in humans:
Evidence from visual evoked potentials}. \emph{Vision Res} \textbf{41,}
2435--2447.

\bibitem[\citeproctext]{ref-Dobson1978}
Dobson V, Teller DY \& Belgum J (1978).
\href{https://doi.org/10.1016/0042-6989(78)90109-8}{Visual acuity in
human infants assessed with stationary stripes and phase-alternated
checkerboards}. \emph{Vision Res} \textbf{18,} 1233--1238.

\bibitem[\citeproctext]{ref-Du2023}
Du X, Liu L, Dong X \& Bao M (2023).
\href{https://doi.org/10.1111/nyas.14969}{Effects of altered-reality
training on interocular disinhibition in amblyopia}. \emph{Ann N Y Acad
Sci} \textbf{1522,} 126--138.

\bibitem[\citeproctext]{ref-Ellemberg1999}
Ellemberg D, Lewis TL, Liu CH \& Maurer D (1999).
\href{https://doi.org/10.1016/s0042-6989(98)00280-6}{Development of
spatial and temporal vision during childhood}. \emph{Vision Res}
\textbf{39,} 2325--2333.

\bibitem[\citeproctext]{ref-Foley1994}
Foley JM (1994). \href{https://doi.org/10.1364/josaa.11.001710}{Human
luminance pattern-vision mechanisms: Masking experiments require a new
model}. \emph{J Opt Soc Am A Opt Image Sci Vis} \textbf{11,} 1710--1719.

\bibitem[\citeproctext]{ref-Hamer1989}
Hamer RD, Norcia AM, Tyler CW \& Hsu-Winges C (1989).
\href{https://doi.org/10.1016/0042-6989(89)90004-7}{The development of
monocular and binocular VEP acuity}. \emph{Vision Res} \textbf{29,}
397--408.

\bibitem[\citeproctext]{ref-Harris1976}
Harris L, Atkinson J \& Braddick O (1976).
\href{https://doi.org/10.1038/264570a0}{Visual contrast sensitivity of a
6-month-old infant measured by the evoked potential}. \emph{Nature}
\textbf{264,} 570--571.

\bibitem[\citeproctext]{ref-Heeger1992}
Heeger DJ (1992).
\href{https://doi.org/10.1017/s0952523800009640}{Normalization of cell
responses in cat striate cortex}. \emph{Vis Neurosci} \textbf{9,}
181--197.

\bibitem[\citeproctext]{ref-Hoffmann2017}
Hoffmann MB, Brands J, Behrens-Baumann W \& Bach M (2017).
\href{https://doi.org/10.1007/s10633-017-9613-y}{VEP-based acuity
assessment in low vision}. \emph{Doc Ophthalmol} \textbf{135,} 209--218.

\bibitem[\citeproctext]{ref-Hu2023}
Hu J, Chen J, Ku Y \& Yu M (2023).
\href{https://doi.org/10.3389/fnins.2023.1280436}{Reduced interocular
suppression after inverse patching in anisometropic amblyopia}.
\emph{Front Neurosci} \textbf{17,} 1280436.

\bibitem[\citeproctext]{ref-Hubel1959}
Hubel DH \& Wiesel TN (1959).
\href{https://doi.org/10.1113/jphysiol.1959.sp006308}{Receptive fields
of single neurones in the cat's striate cortex}. \emph{J Physiol}
\textbf{148,} 574--591.

\bibitem[\citeproctext]{ref-Kuffler1957}
Kuffler SW, Fitzhugh R \& Barlow HB (1957).
\href{https://doi.org/10.1085/jgp.40.5.683}{Maintained activity in the
cat's retina in light and darkness}. \emph{J Gen Physiol} \textbf{40,}
683--702.

\bibitem[\citeproctext]{ref-Legge1979}
Legge GE (1979). \href{https://doi.org/10.1364/josa.69.000838}{Spatial
frequency masking in human vision: Binocular interactions}. \emph{J Opt
Soc Am} \textbf{69,} 838--847.

\bibitem[\citeproctext]{ref-Lennie1993}
Lennie P, Pokorny J \& Smith VC (1993).
\href{https://doi.org/10.1364/josaa.10.001283}{Luminance}. \emph{J Opt
Soc Am A} \textbf{10,} 1283--1293.

\bibitem[\citeproctext]{ref-Levitt2001}
Levitt JB, Schumer RA, Sherman SM, Spear PD \& Movshon JA (2001).
\href{https://doi.org/10.1152/jn.2001.85.5.2111}{Visual response
properties of neurons in the LGN of normally reared and visually
deprived macaque monkeys}. \emph{J Neurophysiol} \textbf{85,}
2111--2129.

\bibitem[\citeproctext]{ref-Lygo2021}
Lygo FA, Richard B, Wade AR, Morland AB \& Baker DH (2021).
\href{https://doi.org/10.1016/j.neuroimage.2021.117780}{Neural markers
of suppression in impaired binocular vision}. \emph{NeuroImage}
\textbf{230,} 117780.

\bibitem[\citeproctext]{ref-MacLennan2022}
MacLennan K, O'Brien S \& Tavassoli T (2022).
\href{https://doi.org/10.1007/s10803-021-05186-3}{In our own words: The
complex sensory experiences of autistic adults}. \emph{J Autism Dev
Disord} \textbf{52,} 3061--3075.

\bibitem[\citeproctext]{ref-MacLeod1979}
MacLeod DI \& Boynton RM (1979).
\href{https://doi.org/10.1364/josa.69.001183}{Chromaticity diagram
showing cone excitation by stimuli of equal luminance}. \emph{J Opt Soc
Am} \textbf{69,} 1183--1186.

\bibitem[\citeproctext]{ref-McKeefry1996}
McKeefry DJ, Russell MH, Murray IJ \& Kulikowski JJ (1996).
\href{https://doi.org/10.1017/s0952523800008543}{Amplitude and phase
variations of harmonic components in human achromatic and chromatic
visual evoked potentials}. \emph{Vis Neurosci} \textbf{13,} 639--653.

\bibitem[\citeproctext]{ref-Meese2009}
Meese TS \& Baker DH (2009).
\href{https://doi.org/10.1167/9.5.2}{Cross-orientation masking is speed
invariant between ocular pathways but speed dependent within them}.
\emph{J Vis} \textbf{9,} 2.1--15.

\bibitem[\citeproctext]{ref-Michelson1927}
Michelson A (1927). \emph{Studies in optics}. University of Chicago
Press.

\bibitem[\citeproctext]{ref-Morrone1986}
Morrone MC \& Burr DC (1986).
\href{https://doi.org/10.1038/321235a0}{Evidence for the existence and
development of visual inhibition in humans}. \emph{Nature} \textbf{321,}
235--237.

\bibitem[\citeproctext]{ref-Morrone1982}
Morrone MC, Burr DC \& Maffei L (1982).
\href{https://doi.org/10.1098/rspb.1982.0078}{Functional implications of
cross-orientation inhibition of cortical visual cells. I.
Neurophysiological evidence}. \emph{Proc R Soc Lond B Biol Sci}
\textbf{216,} 335--354.

\bibitem[\citeproctext]{ref-Movshon1975}
Movshon JA (1975).
\href{https://doi.org/10.1113/jphysiol.1975.sp011025}{The velocity
tuning of single units in cat striate cortex}. \emph{J Physiol}
\textbf{249,} 445--468.

\bibitem[\citeproctext]{ref-Norcia2015}
Norcia AM, Appelbaum LG, Ales JM, Cottereau BR \& Rossion B (2015).
\href{https://doi.org/10.1167/15.6.4}{The steady-state visual evoked
potential in vision research: {A} review}. \emph{Journal of vision}
\textbf{15,} 4--4.

\bibitem[\citeproctext]{ref-Norcia1985b}
Norcia AM \& Tyler CW (1985\emph{b}).
\href{https://doi.org/10.1016/0013-4694(85)91026-0}{Infant {VEP} acuity
measurements: Analysis of individual differences and measurement error}.
\emph{Electroencephalography and Clinical Neurophysiology} \textbf{61,}
359--369.

\bibitem[\citeproctext]{ref-Norcia1985a}
Norcia AM \& Tyler CW (1985\emph{a}).
\href{https://doi.org/10.1016/0042-6989(85)90217-2}{Spatial frequency
sweep {VEP}: Visual acuity during the first year of life}. \emph{Vision
Research} \textbf{25,} 1399--1408.

\bibitem[\citeproctext]{ref-Norcia1986}
Norcia AM, Tyler CW \& Allen D (1986).
\href{https://doi.org/10.1097/00006324-198601000-00003}{Electrophysiological
assessment of contrast sensitivity in human infants}. \emph{Am J Optom
Physiol Opt} \textbf{63,} 12--15.

\bibitem[\citeproctext]{ref-Norcia1990}
Norcia AM, Tyler CW \& Hamer RD (1990).
\href{https://doi.org/10.1016/0042-6989(90)90028-j}{Development of
contrast sensitivity in the human infant}. \emph{Vision Res}
\textbf{30,} 1475--1486.

\bibitem[\citeproctext]{ref-Pei2014}
Pei F, Baldassi S \& Norcia AM (2014). Electrophysiological measures of
low-level vision reveal spatial processing deficits and hemispheric
asymmetry in autism spectrum disorder. \emph{J Vis}; DOI:
\href{https://doi.org/10.1167/14.11.3}{10.1167/14.11.3}.

\bibitem[\citeproctext]{ref-Pei2017}
Pei F, Baldassi S, Tsai JJ, Gerhard HE \& Norcia AM (2017).
\href{https://doi.org/10.1016/j.visres.2016.03.010}{Development of
contrast normalization mechanisms during childhood and adolescence}.
\emph{Vision Res} \textbf{133,} 12--20.

\bibitem[\citeproctext]{ref-Peirce2007}
Peirce JW (2007). \href{https://doi.org/10.1167/7.6.13}{The potential
importance of saturating and supersaturating contrast response functions
in visual cortex}. \emph{J Vis} \textbf{7,} 13.

\bibitem[\citeproctext]{ref-Petrov2005}
Petrov Y, Carandini M \& McKee S (2005).
\href{https://doi.org/10.1523/JNEUROSCI.2871-05.2005}{Two distinct
mechanisms of suppression in human vision}. \emph{J Neurosci}
\textbf{25,} 8704--8707.

\bibitem[\citeproctext]{ref-Polat1993}
Polat U \& Sagi D (1993).
\href{https://doi.org/10.1016/0042-6989(93)90081-7}{Lateral interactions
between spatial channels: Suppression and facilitation revealed by
lateral masking experiments}. \emph{Vision Res} \textbf{33,} 993--999.

\bibitem[\citeproctext]{ref-Porciatti2000}
Porciatti V, Bonanni P, Fiorentini A \& Guerrini R (2000).
\href{https://doi.org/10.1038/72972}{Lack of cortical contrast gain
control in human photosensitive epilepsy}. \emph{Nature Neuroscience}
\textbf{3,} 259--263.

\bibitem[\citeproctext]{ref-Regan1966}
Regan D (1966). \href{https://doi.org/10.1016/0013-4694(66)90088-5}{Some
characteristics of average steady-state and transient responses evoked
by modulated light}. \emph{Electroencephalogr Clin Neurophysiol}
\textbf{20,} 238--248.

\bibitem[\citeproctext]{ref-Regan1988}
Regan MP \& Regan D (1988).
\href{https://doi.org/10.1016/S0022-5193(88)80323-0}{A frequency domain
technique for characterizing nonlinearities in biological systems}.
\emph{Journal of Theoretical Biology} \textbf{133,} 293--317.

\bibitem[\citeproctext]{ref-Ridder2004}
Ridder WH 3rd (2004).
\href{https://doi.org/10.1007/s10633-004-8053-7}{Methods of visual
acuity determination with the spatial frequency sweep visual evoked
potential}. \emph{Doc Ophthalmol} \textbf{109,} 239--247.

\bibitem[\citeproctext]{ref-Shapley1980}
Shapley RM \& Victor JD (1980).
\href{https://doi.org/10.1113/jphysiol.1980.sp013259}{The effect of
contrast on the non-linear response of the y cell}. \emph{J Physiol}
\textbf{302,} 535--547.

\bibitem[\citeproctext]{ref-Shibata2008}
Shibata K, Yamane K, Otuka K \& Iwata M (2008).
\href{https://doi.org/10.1016/j.jns.2008.04.004}{Abnormal visual
processing in migraine with aura: A study of steady-state visual evoked
potentials}. \emph{J Neurol Sci} \textbf{271,} 119--126.

\bibitem[\citeproctext]{ref-Simmons2009}
Simmons DR, Robertson AE, McKay LS, Toal E, McAleer P \& Pollick FE
(2009). \href{https://doi.org/10.1016/j.visres.2009.08.005}{Vision in
autism spectrum disorders}. \emph{Vision Res} \textbf{49,} 2705--2739.

\bibitem[\citeproctext]{ref-Skoczenski1998}
Skoczenski AM \& Norcia AM (1998).
\href{https://doi.org/10.1038/35630}{Neural noise limitations on infant
visual sensitivity}. \emph{Nature} \textbf{391,} 697--700.

\bibitem[\citeproctext]{ref-Tadin2003}
Tadin D, Lappin JS, Gilroy LA \& Blake R (2003).
\href{https://doi.org/10.1038/nature01800}{Perceptual consequences of
centre-surround antagonism in visual motion processing}. \emph{Nature}
\textbf{424,} 312--315.

\bibitem[\citeproctext]{ref-Tolhurst1981}
Tolhurst DJ, Movshon JA \& Thompson ID (1981).
\href{https://doi.org/10.1007/BF00238900}{The dependence of response
amplitude and variance of cat visual cortical neurones on stimulus
contrast}. \emph{Exp Brain Res} \textbf{41,} 414--419.

\bibitem[\citeproctext]{ref-Tsai2011}
Tsai JJ, Norcia AM, Ales JM \& Wade AR (2011).
\href{https://doi.org/10.1002/ana.22462}{Contrast gain control
abnormalities in idiopathic generalized epilepsy}. \emph{Annals of
Neurology} \textbf{70,} 574--582.

\bibitem[\citeproctext]{ref-Tsai2012}
Tsai JJ, Wade AR \& Norcia AM (2012).
\href{https://doi.org/10.1523/JNEUROSCI.4485-11.2012}{Dynamics of
normalization underlying masking in human visual cortex}. \emph{The
Journal of Neuroscience: The Official Journal of the Society for
Neuroscience} \textbf{32,} 2783--2789.

\bibitem[\citeproctext]{ref-Tyler1985}
Tyler CW \& Apkarian PA (1985).
\href{https://doi.org/10.1016/0042-6989(85)90183-x}{Effects of contrast,
orientation and binocularity in the pattern evoked potential}.
\emph{Vision Res} \textbf{25,} 755--766.

\bibitem[\citeproctext]{ref-Tyler1979}
Tyler CW, Apkarian P, Levi DM \& Nakayama K (1979).
\href{https://www.ncbi.nlm.nih.gov/pubmed/447469}{Rapid assessment of
visual function: An electronic sweep technique for the pattern visual
evoked potential}. \emph{Investigative Ophthalmology \& Visual Science}
\textbf{18,} 703--713.

\bibitem[\citeproctext]{ref-Vilidaite2018}
Vilidaite G, Norcia AM, West RJH, Elliott CJH, Pei F, Wade AR \& Baker
DH (2018). \href{https://doi.org/10.1098/rspb.2018.2255}{Autism sensory
dysfunction in an evolutionarily conserved system}. \emph{Proceedings
Biological Sciences} \textbf{285,} 20182255.

\bibitem[\citeproctext]{ref-Zheng2019}
Zheng X, Xu G, Zhi Y, Wang Y, Han C, Wang B, Zhang S, Zhang K \& Liang R
(2019). \href{https://doi.org/10.1016/j.visres.2019.07.003}{Objective
and quantitative assessment of interocular suppression in strabismic
amblyopia based on steady-state motion visual evoked potentials}.
\emph{Vision Res} \textbf{164,} 44--52.

\end{CSLReferences}

\end{document}

### 2024-04-07 review.qmd
Added amblyopia fits

@@
 Parkinson's
+
+Alex - it would be good to add some CRFs from @Afsari2014 to the figure.
+Is it possible to extract either the fitted parameter values or the raw data for some control flies and some mutants?
+Whatever you think is most compelling.
+
 ```{python}
@@
 
-# Amblyopia parameters would follow a similar structure if they were provided.
-# Since they're not detailed in your R code, I'll move on to the autism plot.
+amblyparams = np.array([
+    [0.39, 37.2, 1.4, 0.22],  # amblyopic eye
+    [2.33, 111.69, 1.4, 0.22]   # fellow eye
+])
+
+# Calculate responses
+respambly = amblyparams[0,0] * ((finecontrasts)**amblyparams[0,2] /
+                (amblyparams[0,1] + (finecontrasts)**amblyparams[0,2])) + amblyparams[0,3]
+
+respfellow = amblyparams[1,0] * ((finecontrasts)**amblyparams[1,2] /
+                (amblyparams[1,1] + (finecontrasts)**amblyparams[1,2])) + amblyparams[1,3]
+
+plt.figure(figsize=(12, 10))
+plt.subplot(2, 2, 2)
+plt.plot(finecontrastsdB, respambly, lw=4, color='red', label='Amblyopic eye')
+plt.plot(finecontrastsdB, respfellow, lw=4, color='black', linestyle='--', label='Fellow eye')
+plt.xticks(ticklocsx, ticklabelsx)
+plt.yticks(np.arange(0, 2, 0.5))
+plt.xlabel('Target contrast (%)')
+plt.ylabel('SSVEP response')
+plt.title('(b) Amblyopia')
+plt.legend()
 
@@
 # To save the figure
-plt.savefig('clinicalplot.pdf')
+plt.savefig('Figures/clinicalplot.pdf')
 ```
@@
 
-ticklocsy <- seq(0,6,2)
+# fits to data from Baker et al.. (2015)
+amblyparams <- matrix(0,nrow=2,ncol=4)
+amblyparams[1,] <- c(0.39,37.2,1.4,0.22)  # amblyopic eye
+amblyparams[2,] <- c(2.33,111.69,1.4,0.22)  # good eye
+
+respambly <- amblyparams[1,1] * (finecontrasts^amblyparams[1,3]/(amblyparams[1,2] + finecontrasts^amblyparams[2,3])) + amblyparams[1,4]
+
+respfellow <- amblyparams[2,1] * (finecontrasts^amblyparams[2,3]/(amblyparams[2,2] + finecontrasts^amblyparams[2,3])) + amblyparams[2,4]
+
+ticklocsy <- seq(0,2,0.5)
 ticklabelsy <- ticklocsy
@@
 par(pty="s")  # make axis square
-plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,6))   
+plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,2))   
 axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
@@
 title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+title(main='(b) Amblyopia',cex.main=2)
 
-title(main='(b) Amblyopia',cex.main=2)
+lines(finecontrastsdB,respambly,lwd=4,col='red')
+lines(finecontrastsdB,respfellow,lwd=4,lty=2,col='black')
+
+legend(-18,2,c('Amblyopic eye','Fellow eye'),lwd=4,lty=c(1,2),col=c('red','black'),box.lwd=2)
 

### 2024-04-07 review.qmd
Fix bugs

--- 2024-04-07 review.qmd
@@
 
-# Amblyopia parameters would follow a similar structure if they were provided.
-# Since they're not detailed in your R code, I'll move on to the autism plot.
+amblyparams = np.array([
+    [0.39, 37.2, 1.4, 0.22],  # amblyopic eye
+    [2.33, 111.69, 1.4, 0.22]   # fellow eye
+])
+
+# Calculate responses
+respambly = amblyparams[0,0] * ((finecontrasts)**amblyparams[0,2] /
+                (amblyparams[0,1] + (finecontrasts)**amblyparams[0,2])) + amblyparams[0,3]
+
+respfellow = amblyparams[1,0] * ((finecontrasts)**amblyparams[1,2] /
+                (amblyparams[1,1] + (finecontrasts)**amblyparams[1,2])) + amblyparams[1,3]
+
+plt.subplot(2, 2, 2)
+plt.plot(finecontrastsdB, respambly, lw=4, color='red', label='Amblyopic eye')
+plt.plot(finecontrastsdB, respfellow, lw=4, color='black', linestyle='--', label='Fellow eye')
+plt.xticks(ticklocsx, ticklabelsx)
+plt.yticks(np.arange(0, 2.5, 0.5))
+plt.xlabel('Target contrast (%)')
+plt.ylabel('SSVEP response')
+plt.title('(b) Amblyopia')
+plt.legend()
 
@@
 # To save the figure
-plt.savefig('clinicalplot.pdf')
+plt.savefig('Figures/clinicalplot.pdf')
+
 ```
 
-<!---
-```{r}
-
-pdf('Figures/clinicalplot.pdf',width=12,height=10)
-par(mfrow=c(2,2))
-
-finecontrastsdB <- seq(-18,36,1)
-finecontrasts <- 10^(finecontrastsdB/20)
-ticklocsx <- seq(-18,36,6)
-ticklabelsx <- c('1/8','1/4','1/2',1,2,4,8,16,32,64)
-
-# fits to data from Tsai et al. (2011)
-epilepsyparams <- matrix(0,nrow=2,ncol=4)
-epilepsyparams[1,] <- c(3.4,0.15,1.4,1)  # controls
-epilepsyparams[2,] <- c(5.3,0.23,1.4,1)  # patients
-
-respEPIcontrols <- epilepsyparams[1,1] * ((finecontrasts/100)^epilepsyparams[1,3]/(epilepsyparams[1,2]^epilepsyparams[1,3] + (finecontrasts/100)^epilepsyparams[1,3])) + epilepsyparams[1,4]
-
-respEPIpatients <- epilepsyparams[2,1] * ((finecontrasts/100)^epilepsyparams[2,3]/(epilepsyparams[2,2]^epilepsyparams[2,3] + (finecontrasts/100)^epilepsyparams[2,3])) + epilepsyparams[2,4]
-
-ticklocsy <- seq(0,6,2)
-ticklabelsy <- ticklocsy
-
-par(pty="s")  # make axis square
-plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,6))   
-axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
-axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
-mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
-mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
-title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
-title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
-title(main='(a) Epilepsy',cex.main=2)
-
-lines(finecontrastsdB,respEPIpatients,lwd=4,col='blue')
-lines(finecontrastsdB,respEPIcontrols,lwd=4,lty=2,col='black')
-
-legend(-18,6,c('Epilepsy patients','Control participants'),lwd=4,lty=c(1,2),col=c('blue','black'),box.lwd=2)
-
-ticklocsy <- seq(0,6,2)
-ticklabelsy <- ticklocsy
-
-par(pty="s")  # make axis square
-plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,6))   
-axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
-axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
-mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
-mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
-title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
-title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
-
-title(main='(b) Amblyopia',cex.main=2)
-
-# fits to data from Vilidaite et al (2018)
-autismparams <- matrix(0,nrow=2,ncol=4)
-autismparams[1,] <- c(0.6253,0.2701,125550,5.6199)  # ASD group
-autismparams[2,] <- c(1.0131,0.3369,3589900,6.8622) # control group
-
-respASD <- (autismparams[1,1]*(finecontrasts^autismparams[1,4])/(autismparams[1,3] + finecontrasts^autismparams[1,4])) + autismparams[1,2]
-respCON <- (autismparams[2,1]*(finecontrasts^autismparams[2,4])/(autismparams[2,3] + finecontrasts^autismparams[2,4])) + autismparams[2,2]
-
-ticklocsy <- seq(0,1.5,0.5)
-ticklabelsy <- ticklocsy
-
-par(pty="s")  # make axis square
-plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,1.5))   
-axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
-axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
-mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
-mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
-title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
-title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
-title(main='(c) Autism',cex.main=2)
-
-lines(finecontrastsdB,respASD,lwd=4,col='darkgreen')
-lines(finecontrastsdB,respCON,lwd=4,lty=2,col='black')
-
-legend(-18,1.5,c('ASD children','Control children'),lwd=4,lty=c(1,2),col=c('darkgreen','black'),box.lwd=2)
-
-ticklocsy <- seq(0,6,2)
-ticklabelsy <- ticklocsy
-
-par(pty="s")  # make axis square
-plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,6))   
-axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
-axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
-mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
-mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
-title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
-title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
-title(main="(d) Parkinson's",cex.main=2)
-
-dev.off()
-
-``` commented out R because of the weird reticulate bug...--->
-
-![caption.](Figures/clinicalplot.pdf){#fig-plotclinical}
+![Example contrast response functions for different clinical conditions.
+Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
+(2011).
+Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al.
+(2015).
+Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al.
+(2018).
+Panel (d)...](Figures/clinicalplot.pdf){#fig-plotclinical}
 

### 2024-04-07 review.qmd
Stopped outputting as pdf for consistency with other figures

@@
 ```{python}
+#| label: fig-plotclinical
+#| fig-cap: "Example contrast response functions for different clinical conditions.
+Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
+(2011).
+Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al.
+(2015).
+Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al.
+(2018).
+Panel (d)..."
+#| fig-cap-location: bottom
+
 import numpy as np
@@
 # To save the figure
-plt.savefig('Figures/clinicalplot.pdf')
+# plt.savefig('Figures/clinicalplot.pdf')
 

### 2024-04-07 review.qmd
Removed duplicate figure caption
Added plt.show

@@
 It is also the case for individuals with idiopathic generalised epilepsy [@Tsai2011], a subtype of epilepsy that has a less obvious link to vision.
-The differences apply across the whole contrast-response function, and so resemble a response gain effect, which might be due to reduced inhibition from neighbouring neurons.
+The differences apply across the whole contrast-response function, and so resemble a response gain effect (see [@fig-plotclinical]a), which might be due to reduced inhibition from neighbouring neurons.
 Differences in SSVEP amplitudes have also been reported in individuals with migraine [@Shibata2008], a condition also associated with cortical hyperexciteability.
@@
 (2018).
-Panel (d)..."
+Panel (d)"
 #| fig-cap-location: bottom
@@
 # plt.savefig('Figures/clinicalplot.pdf')
+plt.show()
 
 ```
-
-![Example contrast response functions for different clinical conditions.
-Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
-(2011).
-Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al.
-(2015).
-Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al.
-(2018).
-Panel (d)...](Figures/clinicalplot.pdf){#fig-plotclinical}
 

### 2024-04-08 review.qmd
Datathief to the rescue! Including parameters from Afsari et al 2014

@@
 A significant body of research into contrast processing is concerned with how these normalization mechanisms depend on colour [@Chen2000], orientation [@Foley1994], eye of origin [@Legge1979; @Baker2007], spatial and temporal frequency [@Meese2009], location [@Petrov2005; @Tadin2003; @Polat1993], age [@Betts2005], and the presence of neurological disorders [@Porciatti2000; @Tsai2011].
-The SSVEP has proven to be invaliable in this research because it provides an objective readout of contrast representation at different stages of the visual system, and allows us to 'tag' the probe and background at separate frequencies.
+The SSVEP has proven to be invaluable in this research because it provides an objective readout of contrast representation at different stages of the visual system, and allows us to 'tag' the probe and background at separate frequencies.
 
@@
 For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency or a low multiple thereof.
-This corresponds closely to similar functions reported by invasive studies measuring single unit activity or local field potentials in the cortex [@Shapley1980; @Morrone1982].
+This corresponds closely to similar functions reported by studies measuring single unit activity or local field potentials in the cortex [@Shapley1980; @Morrone1982].
 However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
@@
 
+#Make a figure with three subplots
 fig, axs = plt.subplots(1,3, figsize=(8, 3))
@@
 
-The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the 2F component contains a slight roll-off at high input contrasts.
+The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the second harmonic (2F) component contains a slight roll-off at high input contrasts.
 This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
@@
 
+# Make a figure with three subplots
 fig, axs = plt.subplots(1,3, figsize=(8, 3))
 
-inputModulation = np.sin(2 * np.pi * inputFrequency_Hz[0] * support_S) * inputContrasts[0]
+# Compute waveforms for the 'probe' and 'mask' inputs at the frequencies defined in inputFrequency_Hz
+probeModulation = np.sin(2 * np.pi * inputFrequency_Hz[0] * support_S) * inputContrasts[0]
 maskModulation =  np.sin(2 * np.pi * inputFrequency_Hz[1] * support_S) * inputContrasts[1]
 
-axs[0].plot(support_S, inputModulation,'k')
+axs[0].plot(support_S, probeModulation,'k')
 axs[0].plot(support_S, maskModulation,'r')
@@
 
-combinedSignal=simpleTransducerFunction(inputModulation, maskModulation, c50,RMax)
+# pass both the probe and the mask through the transducer
+combinedSignal=simpleTransducerFunction(probeModulation, maskModulation, c50,RMax)
+
+# Compute the FT of the combined signals
 FT_combinedSignal=np.abs(np.fft.fft(combinedSignal)/ len(combinedSignal))
@@
 MORE HERE - Although this idea was suggested as early as 1988 (probably earlier) not many people seem to have done it much.
-@Candy2001; @Busse2009; @Tsai2012; @Baker2017
+@Candy2001; @Busse2009; @Tsai2012; @Baker2017 *** SEE ALSO RECENT WORK FROM WINAWER LAB - MUCH OF IT IS 'FAST' VEPs ***
 
@@
 
-Because SSVEP responses at detection threshold are very small, estimating threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
+Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
 This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
@@
 
-Measuring the SSVEP at multiple levels was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
+A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrasr levels.
+This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
 To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990].
@@
 In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
-Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around 1 year [@Norcia1985b; @Hamer1989] compared to around 6-7 years with behavioural measures [@Ellemberg1999; @Atkinson1983].
+Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983].
 At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
@@
 This was subsequently replicated in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
-Interestingly this study replicated its key findings in a *drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.
+Interestingly this study replicated its key findings in a *Drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.
 
@@
 
-# Assuming Parkinson's parameters are similar to those given for epilepsy/autism,
-# but with specifics for Parkinson's. We'll use the epilepsy parameters as a placeholder.
+# PD parameters for Drosophila from Afsari et al HMG 2014
+PDparams2F1 = np.array([
+    [22, 0.7, 2, 1],  # controls
+    [24, 0.4, 2, 1]])   # hLRRK2
 
-# Note: Replace epilepsyparams with Parkinson's specific parameters as necessary.
-respParkControls = epilepsyparams[0,0] * ((finecontrasts/100)**epilepsyparams[0,2] /
-                      (epilepsyparams[0,1]**epilepsyparams[0,2] + (finecontrasts/100)**epilepsyparams[0,2])) + epilepsyparams[0,3]
+PDparams1F1 = np.array([
+    [104, 0.76, 2, 1],  # controls
+    [170, 0.382, 2, 1]])   # hLRRK2
 
-respParkPatients = epilepsyparams[1,0] * ((finecontrasts/100)**epilepsyparams[1,2] /
-                      (epilepsyparams[1,1]**epilepsyparams[1,2] + (finecontrasts/100)**epilepsyparams[1,2])) + epilepsyparams[1,3]
+# Note: Here we plot the 2F1 curves which, we think, most closely correspond to the more central neuronal (as opposed to the photoreceptor-derived) signals
+respParkControls = PDparams2F1[0,0] * ((finecontrasts/100)**PDparams2F1[0,2] /
+                      (PDparams2F1[0,1]**PDparams2F1[0,2] + (finecontrasts/100)**PDparams2F1[0,2])) + PDparams2F1[0,3]
+
+respParkPatients = PDparams2F1[1,0] * ((finecontrasts/100)**PDparams2F1[1,2] /
+                      (PDparams2F1[1,1]**PDparams2F1[1,2] + (finecontrasts/100)**PDparams2F1[1,2])) + PDparams2F1[1,3]
 
 plt.subplot(2, 2, 4)  # Fourth plot for Parkinson's
-plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='Parkinson\'s patients')
-plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='Control participants')
+plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='hLRRK2 transgenic flies')
+plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='TH+ (control) flies')
 plt.xticks(ticklocsx, ticklabelsx)

### 2024-04-08 review.tex
Datathief to the rescue! Including parameters from Afsari et al 2014

@@
 presence of neurological disorders (Porciatti \emph{et al.}, 2000; Tsai
-\emph{et al.}, 2011). The SSVEP has proven to be invaliable in this
+\emph{et al.}, 2011). The SSVEP has proven to be invaluable in this
 research because it provides an objective readout of contrast
@@
 at the stimulus frequency or a low multiple thereof. This corresponds
-closely to similar functions reported by invasive studies measuring
-single unit activity or local field potentials in the cortex (Shapley \&
-Victor, 1980; Morrone \emph{et al.}, 1982). However the SSVEP has the
-advantage that it is non-invasive, and so can be measured in awake,
-behaving human participants.
+closely to similar functions reported by studies measuring single unit
+activity or local field potentials in the cortex (Shapley \& Victor,
+1980; Morrone \emph{et al.}, 1982). However the SSVEP has the advantage
+that it is non-invasive, and so can be measured in awake, behaving human
+participants.
 
@@
 The hyperbolic ratio function is monotonic but the CRF resulting from
-measuring the amplitude of the 2F component contains a slight roll-off
-at high input contrasts. This results from the distortion of the input
-sine waves at high contrast due to a combination of the full-wave
-rectification and saturating non-linearity. Power at other harmonics
-increases, and the total power generated by the input is monotonic. This
-roll-off is often seen in experimental data and has been referred to as
-`supersaturation' (Tyler \& Apkarian, 1985; Peirce, 2007).
+measuring the amplitude of the second harmonic (2F) component contains a
+slight roll-off at high input contrasts. This results from the
+distortion of the input sine waves at high contrast due to a combination
+of the full-wave rectification and saturating non-linearity. Power at
+other harmonics increases, and the total power generated by the input is
+monotonic. This roll-off is often seen in experimental data and has been
+referred to as `supersaturation' (Tyler \& Apkarian, 1985; Peirce,
+2007).
 
@@
 \emph{et al.} (2009); Tsai \emph{et al.} (2012); Baker \& Wade (2017)
+*** SEE ALSO RECENT WORK FROM WINAWER LAB - MUCH OF IT IS `FAST' VEPs
+***
 
@@
 Because SSVEP responses at detection threshold are very small,
-estimating threshold is achieved by measuring the contrast response
+estimating a threshold is achieved by measuring the contrast response
 function at relatively high levels, and extrapolating back along the
@@
 
-Measuring the SSVEP at multiple levels was made faster by the
-development of the `sweep VEP' paradigm in which the stimulus changed
-its contrast, spatial frequency or some other property, throughout a
-trial (Tyler \emph{et al.}, 1979). To avoid hysteresis effects, the
-sweep is sometimes conducted both up and down in the same experiment
-Norcia \& Tyler (1985\emph{a}). The sweep VEP (really, a sweep `SSVEP')
-technique is now commonly-used to obtain a rapid and objective
-measurement of visual acuity. In particular, because of its relative
-speed and simplicity, this technique has now become a standard for
-conducting tests of visual acuity in very young subjects or where
-behavioural tests are not appropriate (Ridder, 2004; Bach \emph{et al.},
-2008; Hoffmann \emph{et al.}, 2017; Bach \& Farmer, 2020).
+A robust estimate of the threshold therefore requires the measurement of
+the SSVEP amplitude at many different super-threshold contrasr levels.
+This was made faster by the development of the `sweep VEP' paradigm in
+which the stimulus changed its contrast, spatial frequency or some other
+property, throughout a trial (Tyler \emph{et al.}, 1979). To avoid
+hysteresis effects, the sweep is sometimes conducted both up and down in
+the same experiment Norcia \& Tyler (1985\emph{a}). The sweep VEP
+(really, a sweep `SSVEP') technique is now commonly-used to obtain a
+rapid and objective measurement of visual acuity. In particular, because
+of its relative speed and simplicity, this technique has now become a
+standard for conducting tests of visual acuity in very young subjects or
+where behavioural tests are not appropriate (Ridder, 2004; Bach \emph{et
+al.}, 2008; Hoffmann \emph{et al.}, 2017; Bach \& Farmer, 2020).
 
@@
 measured by SSVEP reaches adult levels more slowly, but near-adult
-levels are recorded around 1 year (Norcia \& Tyler, 1985\emph{b}; Hamer
-\emph{et al.}, 1989) compared to around 6-7 years with behavioural
-measures (Atkinson \& Braddick, 1983; Ellemberg \emph{et al.}, 1999). At
-least some of this difference is likely due to the relative objectivity
-and high SNR of the SSVEP technique compared to other methods such as
-preferential looking, which require careful measurement of the infant's
-gaze direction. However it should be noted that other groups have
-reported electrophysiological correlates of visual acuity that more
-closely match the behavioural measures (De Vries-Khoe \& Spekreijse,
-1982).
+levels are recorded around one year (Norcia \& Tyler, 1985\emph{b};
+Hamer \emph{et al.}, 1989) compared to around six to seven years with
+behavioural measures (Atkinson \& Braddick, 1983; Ellemberg \emph{et
+al.}, 1999). At least some of this difference is likely due to the
+relative objectivity and high SNR of the SSVEP technique compared to
+other methods such as preferential looking, which require careful
+measurement of the infant's gaze direction. However it should be noted
+that other groups have reported electrophysiological correlates of
+visual acuity that more closely match the behavioural measures (De
+Vries-Khoe \& Spekreijse, 1982).
 
@@
 vision. The differences apply across the whole contrast-response
-function, and so resemble a response gain effect, which might be due to
-reduced inhibition from neighbouring neurons. Differences in SSVEP
-amplitudes have also been reported in individuals with migraine (Shibata
-\emph{et al.}, 2008), a condition also associated with cortical
+function, and so resemble a response gain effect (see
+Figure~\ref{fig-plotclinical}a), which might be due to reduced
+inhibition from neighbouring neurons. Differences in SSVEP amplitudes
+have also been reported in individuals with migraine (Shibata \emph{et
+al.}, 2008), a condition also associated with cortical
 hyperexciteability.
@@
 harmonic (using on/off flicker). Interestingly this study replicated its
-key findings in a \emph{drosophila} genetic model of autism (\emph{Nhe3}
+key findings in a \emph{Drosophila} genetic model of autism (\emph{Nhe3}
 mutations), illustrating the translational potential of the steady-state
@@
 
-\includegraphics{review_files/figure-pdf/cell-8-output-1.pdf}
+Alex - it would be good to add some CRFs from Afsari \emph{et al.}
+(2014) to the figure. Is it possible to extract either the fitted
+parameter values or the raw data for some control flies and some
+mutants? Whatever you think is most compelling.
 
@@
 
-\includegraphics{Figures/clinicalplot.pdf}
+\includegraphics{review_files/figure-pdf/fig-plotclinical-output-1.pdf}
 
@@
 
-\caption{\label{fig-plotclinical}caption.}
+\caption{\label{fig-plotclinical}Example contrast response functions for
+different clinical conditions. Panel (a) shows modelled contrast
+response functions for epilepsy patients (blue) and control participants
+(black), based on the data of Tsai et al.~(2011). Panel (b) shows
+functions for the amblyopic (red) and fellow (black) eyes of adults with
+amblyopia, based on the data of Baker et al.~(2015). Panel (c) shows
+functions for children with (green) and without (black) a diagnosis of
+autism, based on the data of Vilidaite et al.~(2018). Panel (d)}
 
@@
 \begin{CSLReferences}{1}{1}
+\bibitem[\citeproctext]{ref-Afsari2014}
+Afsari F, Christensen KV, Smith GP, Hentzer M, Nippe OM, Elliott CJH \&
+Wade AR (2014). \href{https://doi.org/10.1093/hmg/ddu159}{Abnormal
+visual gain control in a {Parkinson}'s disease model}. \emph{Human
+Molecular Genetics} \textbf{23,} 4465--4478.
+
 \bibitem[\citeproctext]{ref-Albrecht1982}

### 2024-04-08 review.qmd
Added more figure cation, grids

@@
 (2018).
-Panel (d)"
+Panel (d) shows data from *Drosophila melanogaster* (fruit flies) in Afsari et al 2014.
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity compared to control animals"
 #| fig-cap-location: bottom
@@
 plt.legend()
+plt.grid()
 
@@
 plt.legend()
+plt.grid()
+
 # To save the figure

### 2024-04-13 review.qmd
Minor edits

@@
 Contrast is the currency of the early visual system.
-Measuring the way that the computations underlying contrast processing depend on factors such as spatial and temporal frequency, age, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
+Measuring the way that the computations underlying contrast processing depend on factors such as spatial and temporal frequency, age, clinical conditions, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
 One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged' (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
@@
 It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
-Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@McKeefry1996; @Baseler1997; @Di-Russo2001].
+Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@McKeefry1996; @Baseler1997; @Di-Russo2001].
 
-Contrast is relatively simple to define: typically, contrast is specified as the percentage deviation of a uniform stimulus from the background.
+Contrast is relatively simple to define: typically, it is specified as the percentage deviation of a uniform stimulus from the background.
 So, for example, a disk of 100 units of cone activation ($I_{\mathrm{stim}}$) surrounded by a 'background' of 50 units of activation ($I_{\mathrm{background}}$) has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
@@
 #| label: fig-centresurround
-#| fig-cap: The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region"
+#| fig-cap: The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region
 #| fig-cap-location: bottom
@@
 Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction.
-For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency or a low multiple thereof.
+For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency, or a low multiple thereof.
 This corresponds closely to similar functions reported by studies measuring single unit activity or local field potentials in the cortex [@Shapley1980; @Morrone1982].
@@
 
-Previous figures showed how a single-input SSVEP might generate a single set of peaks in the frequency domain response which, in turn, can be used as a read-out of neuronal activity.
+**Previous figures** showed how a single-input SSVEP might generate a single set of peaks in the frequency domain response which, in turn, can be used as a read-out of neuronal activity.
 In @fig-plotspectra, we show the effect of adding a second contrast component: the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
@@
 #| label: fig-plotsinewaves
-#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain"
+#| fig-cap: a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain.
 #| fig-cap-location: bottom
@@
 #| label: fig-plotspectra
-#| fig-cap: "Temporal frequency representations of combining two inputs at frequencies F1=5Hz and F2=7Hz combination.
+#| fig-cap: Temporal frequency representations of combining two inputs at frequencies F1=5Hz and F2=7Hz.
 A pure sine-wave input would generate responses only at F1 and F2.
-However, the rectification that accompanies early visual processing converts this contrast reversal into a more complex frequency-domain representation with power predominantly at the second harmonics 2F1 and 2F2"
+However, the rectification that accompanies early visual processing converts this contrast reversal into a more complex frequency-domain representation with power predominantly at the second harmonics 2F1 and 2F2.
 #| fig-cap-location: bottom
@@
 These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
-Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position.
+Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
 The SSVEP response to a contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
@@
 #| label: fig-contrastresponsefunction
-#| fig-cap: "Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
-Note the 'supersaturation' at high contrast levels."
+#| fig-cap: Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
+Note the 'supersaturation' at high contrast levels.
 #| fig-cap-location: bottom
@@
 #| label: fig-intermodulation
-#| fig-cap: "Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies"
+#| fig-cap: Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies
 #| fig-cap-location: bottom
@@
 The complexity of even a simple simulation of the frequency-domain signal arising from non-linear interactions is intriguing.
-Presumably, the signal measured from early visual cortex contains is the result of a cascade of nonlinear retinal and cortical operations up to that point.
+Presumably, the signal measured from early visual cortex is the result of a cascade of nonlinear retinal and cortical operations up to that point.
 It therefore contains a 'signature' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
@@
 This possibility was recognised in the early days of the technique [@Regan1988].
-Although characterising the complete set of computations along entire processing pathway is problematic, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
+Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
 ...
@@
 @Candy2001; @Busse2009; @Tsai2012; @Baker2017 *** SEE ALSO RECENT WORK FROM WINAWER LAB - MUCH OF IT IS 'FAST' VEPs ***
+
+Adaptation
 
@@
 #| label: fig-sweepvep
-#| fig-cap: "Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts"
+#| fig-cap: Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
 #| fig-cap-location: bottom
@@
 
-Parkinson's
-
-Alex - it would be good to add some CRFs from @Afsari2014 to the figure.
-Is it possible to extract either the fitted parameter values or the raw data for some control flies and some mutants?
-Whatever you think is most compelling.
+Parkinson's @Afsari2014
 
@@
 #| label: fig-plotclinical
-#| fig-cap: "Example contrast response functions for different clinical conditions.
+#| fig-cap: Example contrast response functions for different clinical conditions.
 Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
@@
 Panel (d) shows data from *Drosophila melanogaster* (fruit flies) in Afsari et al 2014.
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity compared to control animals"
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity compared to control animals.
 #| fig-cap-location: bottom

### 2024-04-13 review.qmd
Fixed fig captions

@@
 #| label: fig-centresurround
-#| fig-cap: The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region
+#| fig-cap: "The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region."
 #| fig-cap-location: bottom
@@
 #| label: fig-plotsinewaves
-#| fig-cap: a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain.
+#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain."
 #| fig-cap-location: bottom
@@
 #| label: fig-plotspectra
-#| fig-cap: Temporal frequency representations of combining two inputs at frequencies F1=5Hz and F2=7Hz.
+#| fig-cap: "Temporal frequency representations of combining two inputs at frequencies F1=5Hz and F2=7Hz.
 A pure sine-wave input would generate responses only at F1 and F2.
-However, the rectification that accompanies early visual processing converts this contrast reversal into a more complex frequency-domain representation with power predominantly at the second harmonics 2F1 and 2F2.
+However, the rectification that accompanies early visual processing converts this contrast reversal into a more complex frequency-domain representation with power predominantly at the second harmonics 2F1 and 2F2."
 #| fig-cap-location: bottom
@@
 #| label: fig-contrastresponsefunction
-#| fig-cap: Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
-Note the 'supersaturation' at high contrast levels.
+#| fig-cap: "Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
+Note the 'supersaturation' at high contrast levels."
 #| fig-cap-location: bottom
@@
 #| label: fig-intermodulation
-#| fig-cap: Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies
+#| fig-cap: "Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies."
 #| fig-cap-location: bottom
@@
 #| label: fig-sweepvep
-#| fig-cap: Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
+#| fig-cap: "Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts."
 #| fig-cap-location: bottom
@@
 #| label: fig-plotclinical
-#| fig-cap: Example contrast response functions for different clinical conditions.
+#| fig-cap: "Example contrast response functions for different clinical conditions.
 Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
@@
 Panel (d) shows data from *Drosophila melanogaster* (fruit flies) in Afsari et al 2014.
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity compared to control animals.
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity compared to control animals."
 #| fig-cap-location: bottom

### 2024-04-13 review.qmd
Added a few bits

@@
 
-To illustrate the effect of contrast gain control [@Heeger1992], we include a second component that contributes to the gain pool of the first.
+To illustrate the effect of contrast gain control [@Heeger1992], we include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
+The mask stimulus will suppress the response at the target frequency, reducing its amplitude.
+The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Busse2009].
 At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT (@fig-intermodulation).
@@
 Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
-...
-MORE HERE - Although this idea was suggested as early as 1988 (probably earlier) not many people seem to have done it much.
-@Candy2001; @Busse2009; @Tsai2012; @Baker2017 *** SEE ALSO RECENT WORK FROM WINAWER LAB - MUCH OF IT IS 'FAST' VEPs ***
+For example, @Tsai2012 demonstrated that a gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
+This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.
+Our own work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
+More detailed modelling of intracortical recordings [@Groen2022] has revealed details of the timecourse of gain control effects and adaptation, specifically that normalization is delayed slightly relative to the initial visual response.
 
@@
 
-Parkinson's @Afsari2014
+Recent work on understanding Parkinson's disease has also used *Drosophila* genetic models.
+@Afsari2014 found that mutant flies produced stronger SSVEP responses to flickering lights than control flies (see [@fig-plotclinical]d).
+The authors theorised that differences in early gain control during development might lead to visual deficits later in life.
+Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms manifest.
+The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments.
+SSVEP responses also provide a potential method to diagnose Parkinson's before any symptoms manifest, and to monitor the effect of treatments.
 
 ```{python}
+
 #| label: fig-plotclinical
@@
 (2018).
-Panel (d) shows data from *Drosophila melanogaster* (fruit flies) in Afsari et al 2014.
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity compared to control animals."
+Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al.
+(2014).
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black)."
 #| fig-cap-location: bottom

### 2024-04-20 review.qmd
Added new figures. Removed some html support files

@@
 
+```{python setup}
+
+# all package loading and subfunctions defined here
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib import colormaps
+import scipy as scp
+
+def simpleTransducerFunction(inputContrast, c50, RMax):
+ # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
+ # RMax defines the reponse gain of the transducer (how big the biggest output is). 
+ # c50 defines the conrast sensitivity. Smaller values are more sensitive.
+ # The exponent can also vary - here we fix it to 2
+  expnt=2
+
+  # Contrast cannot be negative. To model the responses of a set of on and off 
+  # contrast detectors we perform full-wave rectification of the signal
+  inputContrast=np.abs(inputContrast)
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
+  return output
+
+def doubleTransducerFunction(inputContrast, maskContrast, c50, RMax):
+ # This version has two inputs, but the second input is purely suppressive
+  expnt=2
+  inputContrast=np.abs(inputContrast)
+  maskContrast=np.abs(maskContrast)
+
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
+  return output
+
+def create_grating(size, frequency, phase, contrast):
+    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
+    grating = np.sin(2 * np.pi * frequency * np.sqrt(x**2 + y**2) + phase)
+    grating = (grating + 1) / 2  # Normalize to [0, 1]
+    grating = grating * contrast + (1 - contrast) / 2  # Adjust contrast
+    return grating
+
+def create_circular_mask(size):
+    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
+    mask = np.sqrt(x**2 + y**2) <= 0.3
+    return mask
+
+```
+
+```{python createcentresurround}
+
+# Set parameters
+size = 256
+frequency = 10
+phase = 0
+contrast_center = 0.5
+contrast_surround_low = 0.1
+contrast_surround_high = 0.9
+
+# Create gratings and circular masks
+grating_center = create_grating(size, frequency, phase, contrast_center)
+grating_surround_low = create_grating(size, frequency, phase, contrast_surround_low)
+grating_surround_high = create_grating(size, frequency, phase, contrast_surround_high)
+mask = create_circular_mask(size)
+
+# Combine gratings and masks
+image_low_contrast = grating_surround_low.copy()
+image_low_contrast[mask] = grating_center[mask]
+image_high_contrast = grating_surround_high.copy()
+image_high_contrast[mask] = grating_center[mask]
+
+# Create subplots
+fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(7, 3))
+
+# Plot low-contrast background
+ax1.imshow(image_low_contrast, cmap='gray', vmin=0, vmax=1)
+ax1.axis('off')
+ax1.set_title('Low-Contrast Background', )
+
+# Plot high-contrast background
+ax2.imshow(image_high_contrast, cmap='gray', vmin=0, vmax=1)
+ax2.axis('off')
+ax2.set_title('High-Contrast Background')
+
+plt.tight_layout()
+plot.savefig('Figures/centresurround.pdf')
+
+```
+
+```{python linearsystem}
+
+inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
+contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+c50=.10
+RMax=10
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal=[]
+
+initialSignal = np.zeros(len(contrastRange)) 
+transducedSignal = np.zeros(len(contrastRange))
+for contrastIndex,thisContrast in enumerate(contrastRange):
+  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
+
+  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+
+  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
+  initialSignal[contrastIndex]=inputAmplitude[0]
+
+   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
+   # using the transducer functiona above.
+  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
+  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
+  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
+  transducedSignal[contrastIndex]=transducedAmplitude[0]
+
+frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
+
+#Make a figure with three subplots
+fig, axs = plt.subplots(1,3, figsize=(8, 3))
+
+axs[0].plot(support_S, inputModulation)
+axs[0].set_title('Input sine wave', fontsize=10)
+
+# Plot the absolute values of the FFT in the second axis as a bar plot
+axs[1].bar(frequencySupport, np.abs(FT_Signal[frequencySupport]))
+axs[1].set_xlabel('Frequency (Hz)')
+axs[1].set_ylabel('Magnitude (a.u.)')
+axs[1].set_title('Fourier spectrum')
+axs[1].grid()
+
+axs[2].plot(contrastRange,initialSignal)
+axs[2].set_xlabel('Contrast')
+axs[2].set_ylabel('Amplitude (a.u.)')
+axs[2].set_title('Linear computation stage', fontsize=10)
+axs[2].grid()
+
+axs[2].grid()
+plt.tight_layout()
+plot.savefig('Figures/linearsystem.pdf')
+
+```
+
+```{python transducerout}
+
+# Define some constants
+inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
+contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+c50=.10
+RMax=10
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal=[]
+
+initialSignal = np.zeros(len(contrastRange)) 
+transducedSignal = np.zeros(len(contrastRange))
+for contrastIndex,thisContrast in enumerate(contrastRange):
+  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
+
+  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+
+  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
+  initialSignal[contrastIndex]=inputAmplitude[0]
+
+   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
+   # using the transducer functiona above.
+  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
+  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
+  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
+  transducedSignal[contrastIndex]=transducedAmplitude[0]
+
+frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
+
+#Make a figure with three subplots
+fig, axs = plt.subplots(1,3, figsize=(8, 3))
+
+axs[0].plot(support_S, transducedModulation)
+axs[0].set_title('Transduced signal', fontsize=10)
+
+# Plot the absolute values of the FFT in the second axis as a bar plot
+axs[1].bar(frequencySupport, np.abs(FT_transducedModulation[frequencySupport]))
+axs[1].set_xlabel('Frequency (Hz)')
+axs[1].set_ylabel('Magnitude (a.u.)')
+axs[1].set_title('Fourier spectrum')
+axs[1].grid()
+
+axs[2].plot(contrastRange,transducedSignal)
+axs[2].set_xlabel('Contrast')
+axs[2].set_ylabel('Amplitude (a.u.)')
+axs[2].set_title('Transducer output', fontsize=10)
+axs[2].grid()
+
+axs[2].grid()
+plt.tight_layout()
+plot.savefig('Figures/transducer.pdf')
+
+```
+
+```{python twoinputsout}
+
+# Define some constants
+inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
+contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
+maskFrequency_Hz=np.array([7]) # We simulate a system with on input at many different contrasts
+maskContrast=np.array(0.3) # contrast of mask signal
+
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+c50=.10
+RMax=10
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal=[]
+
+initialSignal = np.zeros(len(contrastRange)) 
+transducedSignal = np.zeros(len(contrastRange))
+maskedSignal = np.zeros(len(contrastRange))
+for contrastIndex,thisContrast in enumerate(contrastRange):
+  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
+  maskModulation = np.sin(2 * np.pi * maskFrequency_Hz * support_S) * maskContrast
+
+   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
+   # using the transducer functiona above.
+  transducedModulation=doubleTransducerFunction(inputModulation, 0*maskModulation, c50,RMax)
+  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
+  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
+  transducedSignal[contrastIndex]=transducedAmplitude[0]
+
+  transducedModulation=doubleTransducerFunction(inputModulation, maskModulation, c50,RMax)
+  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
+  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
+  maskedSignal[contrastIndex]=transducedAmplitude[0]
+
+frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
+
+#Make a figure with three subplots
+fig, axs = plt.subplots(1,3, figsize=(8, 3))
+
+axs[0].plot(support_S, transducedModulation)
+axs[0].set_title('Transduced signal', fontsize=10)
+
+# Plot the absolute values of the FFT in the second axis as a bar plot
+axs[1].bar(frequencySupport, np.abs(FT_transducedModulation[frequencySupport]))
+axs[1].set_xlabel('Frequency (Hz)')
+axs[1].set_ylabel('Magnitude (a.u.)')
+axs[1].set_title('Fourier spectrum')
+axs[1].grid()
+
+axs[2].plot(contrastRange,transducedSignal)
+axs[2].plot(contrastRange,maskedSignal)
+axs[2].set_xlabel('Contrast')
+axs[2].set_ylabel('Amplitude (a.u.)')
+axs[2].set_title('Transducer output', fontsize=10)
+axs[2].grid()
+
+axs[2].grid()
+plt.tight_layout()
+plot.savefig('Figures/twoinputs.pdf')
+
+```
+
 # Abstract:
@@
 
-```{python}
-#| label: fig-centresurround
-#| fig-cap: "The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region."
-#| fig-cap-location: bottom
-
-import numpy as np
-import matplotlib.pyplot as plt
-
-def create_grating(size, frequency, phase, contrast):
-    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
-    grating = np.sin(2 * np.pi * frequency * np.sqrt(x**2 + y**2) + phase)
-    grating = (grating + 1) / 2  # Normalize to [0, 1]
-    grating = grating * contrast + (1 - contrast) / 2  # Adjust contrast
-    return grating
-
-def create_circular_mask(size):
-    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
-    mask = np.sqrt(x**2 + y**2) <= 0.3
-    return mask
-
-# Set parameters
-size = 256
-frequency = 10
-phase = 0
-contrast_center = 0.5
-contrast_surround_low = 0.1
-contrast_surround_high = 0.9
-
-# Create gratings and circular masks
-grating_center = create_grating(size, frequency, phase, contrast_center)
-grating_surround_low = create_grating(size, frequency, phase, contrast_surround_low)
-grating_surround_high = create_grating(size, frequency, phase, contrast_surround_high)
-mask = create_circular_mask(size)
-
-# Combine gratings and masks
-image_low_contrast = grating_surround_low.copy()
-image_low_contrast[mask] = grating_center[mask]
-image_high_contrast = grating_surround_high.copy()
-image_high_contrast[mask] = grating_center[mask]
-
-# Create subplots
-fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(7, 3))
-
-# Plot low-contrast background
-ax1.imshow(image_low_contrast, cmap='gray', vmin=0, vmax=1)
-ax1.axis('off')
-ax1.set_title('Low-Contrast Background', )
-
-# Plot high-contrast background
-ax2.imshow(image_high_contrast, cmap='gray', vmin=0, vmax=1)
-ax2.axis('off')
-ax2.set_title('High-Contrast Background')
-
-plt.tight_layout()
-plt.show()
-
-```
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/centresurround.pdf){#fig-centresurround}
 
@@
 This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed responses of the local neuronal population [@Heeger1992; @Carandini2011; @Busse2009; @Baker2017].
+
+![Linear.](Figures/linearsystem.pdf){#fig-linear}
+
+![Linear.](Figures/transducer.pdf){#fig-transducer}
+
+![Linear.](Figures/twoinputs.pdf){#fig-twoinputs}
 
@@
 The authors theorised that differences in early gain control during development might lead to visual deficits later in life.
-Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms manifest.
+Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms (tremor, rigidity, slow movement) manifest.
 The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments.
@@
 
+I wonder if this is a good place to mention attention.
+
 # Conclusions

### 2024-04-20 review.qmd
Fixed plot output

@@
 plt.tight_layout()
-plot.savefig('Figures/centresurround.pdf')
+plt.savefig('Figures/centresurround.pdf')
 
@@
 plt.tight_layout()
-plot.savefig('Figures/linearsystem.pdf')
+plt.savefig('Figures/linearsystem.pdf')
 
@@
 plt.tight_layout()
-plot.savefig('Figures/transducer.pdf')
+plt.savefig('Figures/transducer.pdf')
 
@@
 plt.tight_layout()
-plot.savefig('Figures/twoinputs.pdf')
+plt.savefig('Figures/twoinputs.pdf')
 

### 2024-04-20 review.qmd
Suppress code output

@@
   echo: false
+  output: false
+  include: false
 ---
@@
 axs[0].set_title('Input sine wave', fontsize=10)
+axs[0].set_xlabel('Time (s)')
 
@@
 axs[0].set_title('Transduced signal', fontsize=10)
+axs[0].set_xlabel('Time (s)')
 
@@
 
-![Linear.](Figures/linearsystem.pdf){#fig-linear}
-
-![Linear.](Figures/transducer.pdf){#fig-transducer}
-
-![Linear.](Figures/twoinputs.pdf){#fig-twoinputs}
+![Linear](Figures/linearsystem.pdf){#fig-linear}
+
+![Nonliner](Figures/transducer.pdf){#fig-transducer}
+
+![Two inputs](Figures/twoinputs.pdf){#fig-twoinputs}
 

### 2024-04-20 review.qmd
All figures now created before the main text

@@
 import scipy as scp
+from sklearn.linear_model import LinearRegression
 
@@
 
-# Abstract:
-
-Contrast is the currency of the early visual system.
-Measuring the way that the computations underlying contrast processing depend on factors such as spatial and temporal frequency, age, clinical conditions, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
-One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged' (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
-SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high SNR, and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
-Here we describe how SSVEPs have been used to study visual contrast over the past 70 years [@Dawson1954].
-Because our thinking about SSVEPs is well-described by simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.
-This paper can therefore be used both as a review of the use of SSVEP in measuring human contrast processing, and as an interactive learning aid.
-
-# Introduction
-
-Neurons in the visual areas of the brain are primarily responsive to changes in cone photoreceptor activations across time and space.
-This property, referred to as contrast, sets the fundamental limits of our visual abilities, which remain steady over a remarkably wide range of environmental light levels.
-The human response to contrast can be studied using many different techniques.
-Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965], defined as the inverse of the lowest contrast that can be reliably detected.
-But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG.
-Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.
-
-The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2005; @Di-Russo2007].
-The flickering stimulus entrains neural population responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
-Two common variants involve sinusoidal on-off flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
-On-off flicker can drive independent populations of on- and off-cells in the retina once per cycle and can therefore produce a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on.
-Counterphase flicker contains two transients per cycle and therefore does not produce a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
-Because square-waves are spectrally broad-band, square wave flicker tends to produce more complex spectral harmonics than sine-wave flicker.
-
-The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system [@Regan1988].
-SSVEP signals can also be elicited by periodic changes of stimulus properties other than achromatic and chromatic contrast, such as motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
-
-# Why measure responses to contrast?
-
-Contrast is one of the most fundamental pieces of information that the eye transmits to the brain.
-It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
-Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@McKeefry1996; @Baseler1997; @Di-Russo2001].
-
-Contrast is relatively simple to define: typically, it is specified as the percentage deviation of a uniform stimulus from the background.
-So, for example, a disk of 100 units of cone activation ($I_{\mathrm{stim}}$) surrounded by a 'background' of 50 units of activation ($I_{\mathrm{background}}$) has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
-Where patterns are more complex (for example, the sine-wave gratings or Gabor patches common in vision science), the @Michelson1927 definition of contrast is specified by the maximum and minimum excursions from the mean:
-
-\begin{equation}
-\frac {I_{\mathrm{stimmax} }-I_{\mathrm{stimmin} }}{I_{\mathrm{stimmax} }+I_{\mathrm{stimmin} }}.
-\end{equation}
-
-These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance; @Lennie1993) and also to definitions based on cone excitations [@Derrington1984; @MacLeod1979] which are more common in work on chromatic processing.
-
-Although its mathematical definition is straightforward, the computations that underlie contrast processing in the brain have been the subject of intense research for many decades.
-The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
-This normalization, achieved through a computation called 'contrast gain control' [@Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
-As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround).
-
-![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/centresurround.pdf){#fig-centresurround}
-
-A significant body of research into contrast processing is concerned with how these normalization mechanisms depend on colour [@Chen2000], orientation [@Foley1994], eye of origin [@Legge1979; @Baker2007], spatial and temporal frequency [@Meese2009], location [@Petrov2005; @Tadin2003; @Polat1993], age [@Betts2005], and the presence of neurological disorders [@Porciatti2000; @Tsai2011].
-The SSVEP has proven to be invaluable in this research because it provides an objective readout of contrast representation at different stages of the visual system, and allows us to 'tag' the probe and background at separate frequencies.
-
-Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction.
-For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency, or a low multiple thereof.
-This corresponds closely to similar functions reported by studies measuring single unit activity or local field potentials in the cortex [@Shapley1980; @Morrone1982].
-However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
-
-To understand the utility of the contrast SSVEP, it is helpful to identify the cascade of processing stages in the early visual system that give rise to it.
-In the following section we illustrate how a typical SSVEP signal measured over early visual cortex might contain information about a large number of early visual computations.
-
-# Contrast processing - linear and nonlinear
-
-Neurons have a limited dynamic range, yet they can transmit information about visual stimuli that span many orders of magnitude.
-In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear, sigmoidal contrast vs response (CRF) transducer [@Albrecht1982; @Tolhurst1981] and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Carandini1994; @Carandini1998; @Busse2009].
-A neuronal population will therefore span a sensitivity range greater than any individual member.
-
-Individual neurons at multiple stages of the visual hierarchy also change their sensitivity depending on the average spatiotemporal contrast energy of their environment.
-This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed responses of the local neuronal population [@Heeger1992; @Carandini2011; @Busse2009; @Baker2017].
-
-![Linear](Figures/linearsystem.pdf){#fig-linear}
-
-![Nonliner](Figures/transducer.pdf){#fig-transducer}
-
-![Two inputs](Figures/twoinputs.pdf){#fig-twoinputs}
-
-**Previous figures** showed how a single-input SSVEP might generate a single set of peaks in the frequency domain response which, in turn, can be used as a read-out of neuronal activity.
-In @fig-plotspectra, we show the effect of adding a second contrast component: the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
-The nature of the interaction --- and therefore the pattern of intermodulation terms ---  is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan1988]
-
-First, we implement a simple non-linear transducer with a single input to illustrate the typical sigmoidal shape of a contrast response function (@fig-contrastresponsefunction).
-
-Next, we show that adding a second component which contributes to the gain pool of the first, results in a complex pattern of intermodulation terms (@fig-plotspectra).
-
-```{python}
-#| label: fig-plotsinewaves
-#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain."
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-import numpy as np
-import scipy as scp
-
-inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
-
-duration_S=1 # All variables in this code have an _[units] where possible
-
-digitisationRate_Hz=1000
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))
-
-for i, freq in enumerate(inputFrequencies_Hz):
-    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)
-
-# Create a figure with two subplots
-fig, axs = plt.subplots(len(inputFrequencies_Hz), 1, figsize=(8, 6))
-
-# Plot each input modulation in a separate subplot
-for i, freq in enumerate(inputFrequencies_Hz):
-    axs[i].plot(support_S, inputModulations[:, i])
-    axs[i].set_xlabel('Time (s)')
-    axs[i].set_ylabel('Amplitude')
-    axs[i].set_title('Frequency: {} Hz'.format(freq))
-
-# Adjust the spacing between subplots
-plt.tight_layout()
-plt.show()
-
-```
-
-The two inputs are represented here as amplitude modulations across time.
-
-If the two inputs are simply added together, the representation of the resulting signal in the Fourier domain is just the linear sum of the two independent signals (first two panels of @fig-plotspectra).
-However, it is important to take physiology into account.
-If we ignore cells sensitive to contrast transients, contrast is represented as a combination of rectified signals from 'on' and 'off'-sensitive cells (broadly, cells that code positive or negative changes in spatiotemporal luminance).
-To model the population response of these cells to a contrast-reversing luminance sine wave, we therefore introduce an initial full-wave rectification stage that simulates the outputs of two, equal populations of on- and off-polarity detectors.
-Once this is done, we see that the Fourier spectrum of the resulting signal contains power only at the second harmonics and above (final two panels of @fig-plotspectra).
-It is possible to generate first-hamonic responses in contrast-response SSVEPs, but only when the inputs are modulated in an 'on/off' manner rather than contrast reversed.
-
-```{python}
-#| label: fig-plotspectra
-#| fig-cap: "Temporal frequency representations of combining two inputs at frequencies F1=5Hz and F2=7Hz.
-A pure sine-wave input would generate responses only at F1 and F2.
-However, the rectification that accompanies early visual processing converts this contrast reversal into a more complex frequency-domain representation with power predominantly at the second harmonics 2F1 and 2F2."
-#| fig-cap-location: bottom
-
-#| label: plotsinewaves
-#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain"
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-import numpy as np
-import scipy as scp
-
-inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
-
-duration_S=1 # All variables in this code have an _[units] where possible
-
-digitisationRate_Hz=1000
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))
-for i, freq in enumerate(inputFrequencies_Hz):
-    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)
-
-freqCutoff_Hz=30
-combinedSignal=np.sum(inputModulations,axis=1)
-combinedRectifiedSignal=np.sum(np.abs(inputModulations), axis=1)
-FT_combinedSignal=np.fft.fft(combinedSignal)/len(combinedSignal)
-FT_combinedRectifiedSignal=np.fft.fft(combinedRectifiedSignal)/len(combinedSignal)
-fig,axs=plt.subplots(1,4,figsize=(8,3))
-axs[0].plot(support_S,combinedSignal)
-axs[0].set_xlabel('Time (s)')
-axs[0].set_ylabel('Amplitude (a.u.)')
-axs[0].set_title('Input signals:\nlinear combination')
-
-# Create the frequency support for the FFT plot
-frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
-
-# Plot the absolute values of the FFT in the second axis as a bar plot
-axs[1].bar(frequencySupport, np.abs(FT_combinedSignal[frequencySupport]))
-axs[1].set_xlabel('Frequency (Hz)')
-axs[1].set_ylabel('Magnitude (a.u.)')
-axs[1].set_title('FFT\n linear combination')
-# Plot the absolute values of the FFT in the second axis as a bar plot
-
-axs[2].plot(support_S,combinedRectifiedSignal)
-axs[2].set_xlabel('Time (s)')
-axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('Sum rectified inputs')
-
-axs[3].bar(frequencySupport, np.abs(FT_combinedRectifiedSignal[frequencySupport]))
-axs[3].set_xlabel('Frequency (Hz)')
-axs[3].set_ylabel('Magnitude (a.u.)')
-axs[3].set_title('FFT of\n sum rectified inputs')
-
-plt.tight_layout()
-plt.show()
-
-```
-
-Even without considering a spatial component, the early visual system is far more complex than the model here suggests.
-For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code temporal changes in contrast.
-These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
-Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
-The SSVEP response to a contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
-
-```{python}
-#| label: fig-contrastresponsefunction
-#| fig-cap: "Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
-Note the 'supersaturation' at high contrast levels."
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-from matplotlib import colormaps
-import numpy as np
-import scipy as scp
-
-def simpleTransducerFunction(inputContrast, c50, RMax):
- # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
- # RMax defines the reponse gain of the transducer (how big the biggest output is). 
- # c50 defines the conrast sensitivity. Smaller values are more sensitive.
- # The exponent can also vary - here we fix it to 2
-  expnt=2
-
-  # Contrast cannot be negative. To model the responses of a set of on and off 
-  # contrast detectors we perform full-wave rectification of the signal
-  inputContrast=np.abs(inputContrast)
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
-  return output
-
-# Define some constants
-inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
-contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
-duration_S=1 # All variables in this code have an _[units] where possible
-digitisationRate_Hz=1000
-freqCutoff_Hz=30
-c50=.10
-RMax=10
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
-transducedSignal=[]
-
-#Make a figure with three subplots
-fig, axs = plt.subplots(1,3, figsize=(8, 3))
-
-initialSignal = np.zeros(len(contrastRange)) 
-transducedSignal = np.zeros(len(contrastRange))
-for contrastIndex,thisContrast in enumerate(contrastRange):
-  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
-
-  axs[0].plot(support_S, inputModulation)
-
-  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
-
-  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
-  initialSignal[contrastIndex]=inputAmplitude[0]
-
-   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
-   # using the transducer functiona above.
-  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
-  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
-  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequencies_Hz*2])
-  transducedSignal[contrastIndex]=transducedAmplitude[0]
-
-axs[0].set_title('Sine-wave responses at\n different contrast levels', fontsize=10)
-
-axs[1].plot(contrastRange,initialSignal)
-axs[1].set_xlabel('Contrast')
-axs[1].set_ylabel('Amplitude (a.u.)')
-axs[1].set_title('Linear computation stage', fontsize=10)
-axs[1].grid()
-
-axs[2].plot(contrastRange,transducedSignal)
-axs[2].set_xlabel('Contrast')
-axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('After nonlinear transducer', fontsize=10)
-
-axs[2].grid()
-plt.tight_layout()
-plt.show()
-
-```
-
-One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
-In @fig-contrastresponsefunction this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity:
-
-\begin{equation}
-R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},
-\end{equation}
-
-where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the response is at half-maximum) and *n* controls the steepness of the curve.
-
-The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the second harmonic (2F) component contains a slight roll-off at high input contrasts.
-This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
-Power at other harmonics increases, and the total power generated by the input is monotonic.
-This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler1985; @Peirce2007].
-
-To illustrate the effect of contrast gain control [@Heeger1992], we include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
-The mask stimulus will suppress the response at the target frequency, reducing its amplitude.
-The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Busse2009].
-At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT (@fig-intermodulation).
-
-```{python}
-#| label: fig-intermodulation
-#| fig-cap: "Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies."
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-from matplotlib import colormaps
-import numpy as np
-import scipy as scp
-
-def simpleTransducerFunction(inputContrast, maskContrast, c50, RMax):
- # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
- # RMax defines the reponse gain of the transducer (how big the biggest output is). 
- # c50 defines the conrast sensitivity. Smaller values are more sensitive.
- # The exponent can also vary - here we fix it to 2
-  expnt=2
-  inputContrast=np.abs(inputContrast)
-  maskContrast=np.abs(maskContrast)
-
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
-  return output
-
-# Define some constants
-inputFrequency_Hz=np.array([5,7]) # We simulate a system with two inputs at a single contrast
-inputContrasts=np.array([.2 ,.2])
-duration_S=1 # All variables in this code have an _[units] where possible
-digitisationRate_Hz=1000
-freqCutoff_Hz=30
-c50=.10
-RMax=10
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
-transducedSignal=[]
-
-# Make a figure with three subplots
-fig, axs = plt.subplots(1,3, figsize=(8, 3))
-
-# Compute waveforms for the 'probe' and 'mask' inputs at the frequencies defined in inputFrequency_Hz
-probeModulation = np.sin(2 * np.pi * inputFrequency_Hz[0] * support_S) * inputContrasts[0]
-maskModulation =  np.sin(2 * np.pi * inputFrequency_Hz[1] * support_S) * inputContrasts[1]
-
-axs[0].plot(support_S, probeModulation,'k')
-axs[0].plot(support_S, maskModulation,'r')
-axs[0].set_title('Input modulations', fontsize=10)
-axs[0].grid()
-
-# pass both the probe and the mask through the transducer
-combinedSignal=simpleTransducerFunction(probeModulation, maskModulation, c50,RMax)
-
-# Compute the FT of the combined signals
-FT_combinedSignal=np.abs(np.fft.fft(combinedSignal)/ len(combinedSignal))
-axs[1].plot(support_S, combinedSignal,'k')
-axs[1].set_title('Combined signal after transducer', fontsize=10)
-axs[1].grid()
-
-axs[2].bar(np.arange(1,19), FT_combinedSignal[1:19])
-
-axs[2].grid()
-plt.tight_layout()
-plt.show()
-
-```
-
-The complexity of even a simple simulation of the frequency-domain signal arising from non-linear interactions is intriguing.
-Presumably, the signal measured from early visual cortex is the result of a cascade of nonlinear retinal and cortical operations up to that point.
-It therefore contains a 'signature' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
-In principle, that information could be recovered from the SSVEP signal.
-
-This possibility was recognised in the early days of the technique [@Regan1988].
-Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
-For example, @Tsai2012 demonstrated that a gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
-This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.
-Our own work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
-More detailed modelling of intracortical recordings [@Groen2022] has revealed details of the timecourse of gain control effects and adaptation, specifically that normalization is delayed slightly relative to the initial visual response.
-
-Adaptation
-
-# Measuring the development of contrast processing
-
-An early use of the SSVEP was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
-In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
-However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [@Norcia1990; @Tyler1979; @Braddick1986] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
-
-Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
-This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
-
-```{python}
-
-#| label: fig-sweepvep
-#| fig-cap: "Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts."
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-from matplotlib import colormaps
-import numpy as np
-import scipy as scp
-from sklearn.linear_model import LinearRegression
-
-def simpleTransducerFunction(inputContrast, c50, RMax):
- # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
- # RMax defines the reponse gain of the transducer (how big the biggest output is). 
- # c50 defines the conrast sensitivity. Smaller values are more sensitive.
- # The exponent can also vary - here we fix it to 2
-  expnt=2
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
-  return output
+```{python makesweep}
 
@@
 plt.tight_layout()
-plt.show()
+plt.savefig('Figures/sweepplot.pdf')
 
@@
 
-A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrasr levels.
-This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990].
-The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
-In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].
-
-This approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
-In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
-Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983].
-At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
-However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
-
-The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section [@Candy2001; @Pei2017].
-Although contrast gain control is measurable in infants as young as six weeks old [@Candy2001; @Morrone1986; @Skoczenski1998], its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
-
-# Contrast processing in clinical conditions
-
-The SSVEP technique has also been used to study clinical conditions, such as diseases and developmental disorders.
-This can often be informative regarding the underlying mechanism that characterises the condition.
-Here we focus on four conditions, but there is potential to apply the method more broadly, either as a diagnostic technique, or to monitor disease severity and progression, or the efficacy of treatments.
-
-Epilepsy is a neurological condition in which patients experience seizures - episodes of uncontrolled neural activity that can cause unconsciousness, involuntary movements and convulsions, and atypical sensory experiences.
-@Porciatti2000 showed that individuals with photosensitive epilepsy generate larger steady-state signals in response to flickering visual stimuli, and their contrast response functions saturate less than those of healthy controls.
-This is consistent with the idea that epilepsy involves a cortical hyperexciteability that makes seizures more likely.
-It is also the case for individuals with idiopathic generalised epilepsy [@Tsai2011], a subtype of epilepsy that has a less obvious link to vision.
-The differences apply across the whole contrast-response function, and so resemble a response gain effect (see [@fig-plotclinical]a), which might be due to reduced inhibition from neighbouring neurons.
-Differences in SSVEP amplitudes have also been reported in individuals with migraine [@Shibata2008], a condition also associated with cortical hyperexciteability.
-
-Amblyopia is a disorder of binocular vision, characterised by one eye contributing much less to perception than the other.
-This is often due to strabismus (squint) or anisometropia (difference in optical prescription between the eyes) during development.
-Contemporary accounts suggest that the amblyopic eye is suppressed by signals from the fellow eye.
-SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye across the contrast range [@Baker2015; @Lygo2021].
-There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.
-The steady-state approach may be more sensitive and objective than typical acuity measurements, and also has the potential to measure suppression between the eyes directly [e.g.
-@Zheng2019; @Du2023; @Hu2023].
-
-Autism is a condition often associated with differences in vision [@Simmons2009] and other senses [@MacLennan2022].
-@Pei2014 used a sweep-VEP method with counterphase flickering stimuli, and found weaker responses in autistic children at spatial frequencies around 8c/deg, compared with age-matched controls.
-This was subsequently replicated in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
-Interestingly this study replicated its key findings in a *Drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.
-
-Recent work on understanding Parkinson's disease has also used *Drosophila* genetic models.
-@Afsari2014 found that mutant flies produced stronger SSVEP responses to flickering lights than control flies (see [@fig-plotclinical]d).
-The authors theorised that differences in early gain control during development might lead to visual deficits later in life.
-Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms (tremor, rigidity, slow movement) manifest.
-The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments.
-SSVEP responses also provide a potential method to diagnose Parkinson's before any symptoms manifest, and to monitor the effect of treatments.
-
-```{python}
-
-#| label: fig-plotclinical
-#| fig-cap: "Example contrast response functions for different clinical conditions.
-Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
-(2011).
-Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al.
-(2015).
-Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al.
-(2018).
-Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al.
-(2014).
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black)."
-#| fig-cap-location: bottom
-
-import numpy as np
-import matplotlib.pyplot as plt
+```{python makeclinicalplot}
 
@@
 # To save the figure
-# plt.savefig('Figures/clinicalplot.pdf')
+plt.savefig('Figures/clinicalplot.pdf')
+
+```
+
+# Abstract:
+
+Contrast is the currency of the early visual system.
+Measuring the way that the computations underlying contrast processing depend on factors such as spatial and temporal frequency, age, clinical conditions, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
+One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged' (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
+SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high SNR, and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
+Here we describe how SSVEPs have been used to study visual contrast over the past 70 years [@Dawson1954].
+Because our thinking about SSVEPs is well-described by simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.
+This paper can therefore be used both as a review of the use of SSVEP in measuring human contrast processing, and as an interactive learning aid.
+
+# Introduction
+
+Neurons in the visual areas of the brain are primarily responsive to changes in cone photoreceptor activations across time and space.
+This property, referred to as contrast, sets the fundamental limits of our visual abilities, which remain steady over a remarkably wide range of environmental light levels.
+The human response to contrast can be studied using many different techniques.
+Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965], defined as the inverse of the lowest contrast that can be reliably detected.
+But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG.
+Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.
+
+The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2005; @Di-Russo2007].
+The flickering stimulus entrains neural population responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
+Two common variants involve sinusoidal on-off flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
+On-off flicker can drive independent populations of on- and off-cells in the retina once per cycle and can therefore produce a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on.
+Counterphase flicker contains two transients per cycle and therefore does not produce a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
+Because square-waves are spectrally broad-band, square wave flicker tends to produce more complex spectral harmonics than sine-wave flicker.
+
+The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system [@Regan1988].
+SSVEP signals can also be elicited by periodic changes of stimulus properties other than achromatic and chromatic contrast, such as motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
+
+# Why measure responses to contrast?
+
+Contrast is one of the most fundamental pieces of information that the eye transmits to the brain.
+It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
+Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@McKeefry1996; @Baseler1997; @Di-Russo2001].
+
+Contrast is relatively simple to define: typically, it is specified as the percentage deviation of a uniform stimulus from the background.
+So, for example, a disk of 100 units of cone activation ($I_{\mathrm{stim}}$) surrounded by a 'background' of 50 units of activation ($I_{\mathrm{background}}$) has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
+Where patterns are more complex (for example, the sine-wave gratings or Gabor patches common in vision science), the @Michelson1927 definition of contrast is specified by the maximum and minimum excursions from the mean:
+
+\begin{equation}
+\frac {I_{\mathrm{stimmax} }-I_{\mathrm{stimmin} }}{I_{\mathrm{stimmax} }+I_{\mathrm{stimmin} }}.
+\end{equation}
+
+These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance; @Lennie1993) and also to definitions based on cone excitations [@Derrington1984; @MacLeod1979] which are more common in work on chromatic processing.
+
+Although its mathematical definition is straightforward, the computations that underlie contrast processing in the brain have been the subject of intense research for many decades.
+The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
+This normalization, achieved through a computation called 'contrast gain control' [@Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
+As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround).
+
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/centresurround.pdf){#fig-centresurround}
+
+A significant body of research into contrast processing is concerned with how these normalization mechanisms depend on colour [@Chen2000], orientation [@Foley1994], eye of origin [@Legge1979; @Baker2007], spatial and temporal frequency [@Meese2009], location [@Petrov2005; @Tadin2003; @Polat1993], age [@Betts2005], and the presence of neurological disorders [@Porciatti2000; @Tsai2011].
+The SSVEP has proven to be invaluable in this research because it provides an objective readout of contrast representation at different stages of the visual system, and allows us to 'tag' the probe and background at separate frequencies.
+
+Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction.
+For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency, or a low multiple thereof.
+This corresponds closely to similar functions reported by studies measuring single unit activity or local field potentials in the cortex [@Shapley1980; @Morrone1982].
+However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
+
+To understand the utility of the contrast SSVEP, it is helpful to identify the cascade of processing stages in the early visual system that give rise to it.
+In the following section we illustrate how a typical SSVEP signal measured over early visual cortex might contain information about a large number of early visual computations.
+
+# Contrast processing - linear and nonlinear
+
+Neurons have a limited dynamic range, yet they can transmit information about visual stimuli that span many orders of magnitude.
+In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear, sigmoidal contrast vs response (CRF) transducer [@Albrecht1982; @Tolhurst1981] and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Carandini1994; @Carandini1998; @Busse2009].
+A neuronal population will therefore span a sensitivity range greater than any individual member.
+
+Individual neurons at multiple stages of the visual hierarchy also change their sensitivity depending on the average spatiotemporal contrast energy of their environment.
+This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed responses of the local neuronal population [@Heeger1992; @Carandini2011; @Busse2009; @Baker2017].
+
+![Linear](Figures/linearsystem.pdf){#fig-linear}
+
+![Nonliner](Figures/transducer.pdf){#fig-transducer}
+
+![Two inputs](Figures/twoinputs.pdf){#fig-twoinputs}
+
+**Previous figures** showed how a single-input SSVEP might generate a single set of peaks in the frequency domain response which, in turn, can be used as a read-out of neuronal activity.
+In @fig-plotspectra, we show the effect of adding a second contrast component: the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
+The nature of the interaction --- and therefore the pattern of intermodulation terms ---  is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan1988]
+
+First, we implement a simple non-linear transducer with a single input to illustrate the typical sigmoidal shape of a contrast response function (@fig-contrastresponsefunction).
+
+Next, we show that adding a second component which contributes to the gain pool of the first, results in a complex pattern of intermodulation terms (@fig-plotspectra).
+
+```{python}
+#| label: fig-plotsinewaves
+#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain."
+#| fig-cap-location: bottom
+
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy as scp
+
+inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
+
+duration_S=1 # All variables in this code have an _[units] where possible
+
+digitisationRate_Hz=1000
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))
+
+for i, freq in enumerate(inputFrequencies_Hz):
+    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)
+
+# Create a figure with two subplots
+fig, axs = plt.subplots(len(inputFrequencies_Hz), 1, figsize=(8, 6))
+
+# Plot each input modulation in a separate subplot
+for i, freq in enumerate(inputFrequencies_Hz):
+    axs[i].plot(support_S, inputModulations[:, i])
+    axs[i].set_xlabel('Time (s)')
+    axs[i].set_ylabel('Amplitude')
+    axs[i].set_title('Frequency: {} Hz'.format(freq))
+
+# Adjust the spacing between subplots
+plt.tight_layout()
 plt.show()
@@
 ```
+
+The two inputs are represented here as amplitude modulations across time.
+
+If the two inputs are simply added together, the representation of the resulting signal in the Fourier domain is just the linear sum of the two independent signals (first two panels of @fig-plotspectra).
+However, it is important to take physiology into account.
+If we ignore cells sensitive to contrast transients, contrast is represented as a combination of rectified signals from 'on' and 'off'-sensitive cells (broadly, cells that code positive or negative changes in spatiotemporal luminance).
+To model the population response of these cells to a contrast-reversing luminance sine wave, we therefore introduce an initial full-wave rectification stage that simulates the outputs of two, equal populations of on- and off-polarity detectors.
+Once this is done, we see that the Fourier spectrum of the resulting signal contains power only at the second harmonics and above (final two panels of @fig-plotspectra).
+It is possible to generate first-hamonic responses in contrast-response SSVEPs, but only when the inputs are modulated in an 'on/off' manner rather than contrast reversed.
+
+```{python}
+#| label: fig-plotspectra
+#| fig-cap: "Temporal frequency representations of combining two inputs at frequencies F1=5Hz and F2=7Hz.
+A pure sine-wave input would generate responses only at F1 and F2.
+However, the rectification that accompanies early visual processing converts this contrast reversal into a more complex frequency-domain representation with power predominantly at the second harmonics 2F1 and 2F2."
+#| fig-cap-location: bottom
+
+#| label: plotsinewaves
+#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain"
+#| fig-cap-location: bottom
+
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy as scp
+
+inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
+
+duration_S=1 # All variables in this code have an _[units] where possible
+
+digitisationRate_Hz=1000
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))
+for i, freq in enumerate(inputFrequencies_Hz):
+    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)
+
+freqCutoff_Hz=30
+combinedSignal=np.sum(inputModulations,axis=1)
+combinedRectifiedSignal=np.sum(np.abs(inputModulations), axis=1)
+FT_combinedSignal=np.fft.fft(combinedSignal)/len(combinedSignal)
+FT_combinedRectifiedSignal=np.fft.fft(combinedRectifiedSignal)/len(combinedSignal)
+fig,axs=plt.subplots(1,4,figsize=(8,3))
+axs[0].plot(support_S,combinedSignal)
+axs[0].set_xlabel('Time (s)')
+axs[0].set_ylabel('Amplitude (a.u.)')
+axs[0].set_title('Input signals:\nlinear combination')
+
+# Create the frequency support for the FFT plot
+frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
+
+# Plot the absolute values of the FFT in the second axis as a bar plot
+axs[1].bar(frequencySupport, np.abs(FT_combinedSignal[frequencySupport]))
+axs[1].set_xlabel('Frequency (Hz)')
+axs[1].set_ylabel('Magnitude (a.u.)')
+axs[1].set_title('FFT\n linear combination')
+# Plot the absolute values of the FFT in the second axis as a bar plot
+
+axs[2].plot(support_S,combinedRectifiedSignal)
+axs[2].set_xlabel('Time (s)')
+axs[2].set_ylabel('Amplitude (a.u.)')
+axs[2].set_title('Sum rectified inputs')
+
+axs[3].bar(frequencySupport, np.abs(FT_combinedRectifiedSignal[frequencySupport]))
+axs[3].set_xlabel('Frequency (Hz)')
+axs[3].set_ylabel('Magnitude (a.u.)')
+axs[3].set_title('FFT of\n sum rectified inputs')
+
+plt.tight_layout()
+plt.show()
+
+```
+
+Even without considering a spatial component, the early visual system is far more complex than the model here suggests.
+For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code temporal changes in contrast.
+These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
+Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
+The SSVEP response to a contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
+
+```{python}
+#| label: fig-contrastresponsefunction
+#| fig-cap: "Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
+Note the 'supersaturation' at high contrast levels."
+#| fig-cap-location: bottom
+
+import matplotlib.pyplot as plt
+from matplotlib import colormaps
+import numpy as np
+import scipy as scp
+
+def simpleTransducerFunction(inputContrast, c50, RMax):
+ # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
+ # RMax defines the reponse gain of the transducer (how big the biggest output is). 
+ # c50 defines the conrast sensitivity. Smaller values are more sensitive.
+ # The exponent can also vary - here we fix it to 2
+  expnt=2
+
+  # Contrast cannot be negative. To model the responses of a set of on and off 
+  # contrast detectors we perform full-wave rectification of the signal
+  inputContrast=np.abs(inputContrast)
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
+  return output
+
+# Define some constants
+inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
+contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+c50=.10
+RMax=10
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal=[]
+
+#Make a figure with three subplots
+fig, axs = plt.subplots(1,3, figsize=(8, 3))
+
+initialSignal = np.zeros(len(contrastRange)) 
+transducedSignal = np.zeros(len(contrastRange))
+for contrastIndex,thisContrast in enumerate(contrastRange):
+  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
+
+  axs[0].plot(support_S, inputModulation)
+
+  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+
+  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
+  initialSignal[contrastIndex]=inputAmplitude[0]
+
+   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
+   # using the transducer functiona above.
+  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
+  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
+  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequencies_Hz*2])
+  transducedSignal[contrastIndex]=transducedAmplitude[0]
+
+axs[0].set_title('Sine-wave responses at\n different contrast levels', fontsize=10)
+
+axs[1].plot(contrastRange,initialSignal)
+axs[1].set_xlabel('Contrast')
+axs[1].set_ylabel('Amplitude (a.u.)')
+axs[1].set_title('Linear computation stage', fontsize=10)
+axs[1].grid()
+
+axs[2].plot(contrastRange,transducedSignal)
+axs[2].set_xlabel('Contrast')
+axs[2].set_ylabel('Amplitude (a.u.)')
+axs[2].set_title('After nonlinear transducer', fontsize=10)
+
+axs[2].grid()
+plt.tight_layout()
+plt.show()
+
+```
+
+One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
+In @fig-contrastresponsefunction this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity:
+
+\begin{equation}
+R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},
+\end{equation}
+
+where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the response is at half-maximum) and *n* controls the steepness of the curve.
+
+The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the second harmonic (2F) component contains a slight roll-off at high input contrasts.
+This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
+Power at other harmonics increases, and the total power generated by the input is monotonic.
+This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler1985; @Peirce2007].
+
+To illustrate the effect of contrast gain control [@Heeger1992], we include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
+The mask stimulus will suppress the response at the target frequency, reducing its amplitude.
+The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Busse2009].
+At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT (@fig-intermodulation).
+
+```{python}
+#| label: fig-intermodulation
+#| fig-cap: "Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies."
+#| fig-cap-location: bottom
+
+import matplotlib.pyplot as plt
+from matplotlib import colormaps
+import numpy as np
+import scipy as scp
+
+def simpleTransducerFunction(inputContrast, maskContrast, c50, RMax):
+ # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
+ # RMax defines the reponse gain of the transducer (how big the biggest output is). 
+ # c50 defines the conrast sensitivity. Smaller values are more sensitive.
+ # The exponent can also vary - here we fix it to 2
+  expnt=2
+  inputContrast=np.abs(inputContrast)
+  maskContrast=np.abs(maskContrast)
+
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
+  return output
+
+# Define some constants
+inputFrequency_Hz=np.array([5,7]) # We simulate a system with two inputs at a single contrast
+inputContrasts=np.array([.2 ,.2])
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+c50=.10
+RMax=10
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal=[]
+
+# Make a figure with three subplots
+fig, axs = plt.subplots(1,3, figsize=(8, 3))
+
+# Compute waveforms for the 'probe' and 'mask' inputs at the frequencies defined in inputFrequency_Hz
+probeModulation = np.sin(2 * np.pi * inputFrequency_Hz[0] * support_S) * inputContrasts[0]
+maskModulation =  np.sin(2 * np.pi * inputFrequency_Hz[1] * support_S) * inputContrasts[1]
+
+axs[0].plot(support_S, probeModulation,'k')
+axs[0].plot(support_S, maskModulation,'r')
+axs[0].set_title('Input modulations', fontsize=10)
+axs[0].grid()
+
+# pass both the probe and the mask through the transducer
+combinedSignal=simpleTransducerFunction(probeModulation, maskModulation, c50,RMax)
+
+# Compute the FT of the combined signals
+FT_combinedSignal=np.abs(np.fft.fft(combinedSignal)/ len(combinedSignal))
+axs[1].plot(support_S, combinedSignal,'k')
+axs[1].set_title('Combined signal after transducer', fontsize=10)
+axs[1].grid()
+
+axs[2].bar(np.arange(1,19), FT_combinedSignal[1:19])
+
+axs[2].grid()
+plt.tight_layout()
+plt.show()
+
+```
+
+The complexity of even a simple simulation of the frequency-domain signal arising from non-linear interactions is intriguing.
+Presumably, the signal measured from early visual cortex is the result of a cascade of nonlinear retinal and cortical operations up to that point.
+It therefore contains a 'signature' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
+In principle, that information could be recovered from the SSVEP signal.
+
+This possibility was recognised in the early days of the technique [@Regan1988].
+Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
+For example, @Tsai2012 demonstrated that a gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
+This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.
+Our own work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
+More detailed modelling of intracortical recordings [@Groen2022] has revealed details of the timecourse of gain control effects and adaptation, specifically that normalization is delayed slightly relative to the initial visual response.
+
+Adaptation
+
+# Measuring the development of contrast processing
+
+An early use of the SSVEP was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
+In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
+However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [@Norcia1990; @Tyler1979; @Braddick1986] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
+
+Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
+This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
+
+![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.](Figures/sweepplot.pdf){#fig-sweepvep}
+
+A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrasr levels.
+This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
+To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990].
+The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
+In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].
+
+This approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
+In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983].
+At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
+However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
+
+The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section [@Candy2001; @Pei2017].
+Although contrast gain control is measurable in infants as young as six weeks old [@Candy2001; @Morrone1986; @Skoczenski1998], its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
+
+# Contrast processing in clinical conditions
+
+The SSVEP technique has also been used to study clinical conditions, such as diseases and developmental disorders.
+This can often be informative regarding the underlying mechanism that characterises the condition.
+Here we focus on four conditions, but there is potential to apply the method more broadly, either as a diagnostic technique, or to monitor disease severity and progression, or the efficacy of treatments.
+
+Epilepsy is a neurological condition in which patients experience seizures - episodes of uncontrolled neural activity that can cause unconsciousness, involuntary movements and convulsions, and atypical sensory experiences.
+@Porciatti2000 showed that individuals with photosensitive epilepsy generate larger steady-state signals in response to flickering visual stimuli, and their contrast response functions saturate less than those of healthy controls.
+This is consistent with the idea that epilepsy involves a cortical hyperexciteability that makes seizures more likely.
+It is also the case for individuals with idiopathic generalised epilepsy [@Tsai2011], a subtype of epilepsy that has a less obvious link to vision.
+The differences apply across the whole contrast-response function, and so resemble a response gain effect (see [@fig-plotclinical]a), which might be due to reduced inhibition from neighbouring neurons.
+Differences in SSVEP amplitudes have also been reported in individuals with migraine [@Shibata2008], a condition also associated with cortical hyperexciteability.
+
+Amblyopia is a disorder of binocular vision, characterised by one eye contributing much less to perception than the other.
+This is often due to strabismus (squint) or anisometropia (difference in optical prescription between the eyes) during development.
+Contemporary accounts suggest that the amblyopic eye is suppressed by signals from the fellow eye.
+SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye across the contrast range [@Baker2015; @Lygo2021].
+There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.
+The steady-state approach may be more sensitive and objective than typical acuity measurements, and also has the potential to measure suppression between the eyes directly [e.g.
+@Zheng2019; @Du2023; @Hu2023].
+
+Autism is a condition often associated with differences in vision [@Simmons2009] and other senses [@MacLennan2022].
+@Pei2014 used a sweep-VEP method with counterphase flickering stimuli, and found weaker responses in autistic children at spatial frequencies around 8c/deg, compared with age-matched controls.
+This was subsequently replicated in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
+Interestingly this study replicated its key findings in a *Drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.
+
+Recent work on understanding Parkinson's disease has also used *Drosophila* genetic models.
+@Afsari2014 found that mutant flies produced stronger SSVEP responses to flickering lights than control flies (see [@fig-plotclinical]d).
+The authors theorised that differences in early gain control during development might lead to visual deficits later in life.
+Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms (tremor, rigidity, slow movement) manifest.
+The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments.
+SSVEP responses also provide a potential method to diagnose Parkinson's before any symptoms manifest, and to monitor the effect of treatments.
+
+![Example contrast response functions for different clinical conditions.
+Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
+(2011).
+Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al.
+(2015).
+Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al.
+(2018).
+Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al.
+(2014).
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/clinicalplot.pdf){#fig-plotclinical}
 

### 2024-04-20 review.qmd
Sorted out gain control demonstration and text

@@
 
-axs[2].plot(contrastRange,initialSignal)
-axs[2].set_xlabel('Contrast')
+axs[2].plot(contrastRange*100,initialSignal)
+axs[2].set_xlabel('Contrast (%)')
 axs[2].set_ylabel('Amplitude (a.u.)')
@@
 
-axs[2].plot(contrastRange,transducedSignal)
-axs[2].set_xlabel('Contrast')
+axs[2].plot(contrastRange*100,transducedSignal)
+axs[2].set_xlabel('Contrast (%)')
 axs[2].set_ylabel('Amplitude (a.u.)')
@@
 maskFrequency_Hz=np.array([7]) # We simulate a system with on input at many different contrasts
-maskContrast=np.array(0.3) # contrast of mask signal
+maskContrast=np.max(contrastRange) # contrast of mask signal
 
@@
 
-axs[2].plot(contrastRange,transducedSignal)
-axs[2].plot(contrastRange,maskedSignal)
-axs[2].set_xlabel('Contrast')
+axs[2].plot(contrastRange*100,transducedSignal)
+axs[2].plot(contrastRange*100,maskedSignal)
+axs[2].set_xlabel('Contrast (%)')
 axs[2].set_ylabel('Amplitude (a.u.)')
@@
 
-\begin{equation}
-\frac {I_{\mathrm{stimmax} }-I_{\mathrm{stimmin} }}{I_{\mathrm{stimmax} }+I_{\mathrm{stimmin} }}.
-\end{equation}
+$$\frac {I_{\mathrm{stimmax} }-I_{\mathrm{stimmin} }}{I_{\mathrm{stimmax} }+I_{\mathrm{stimmin} }}.$${#eq-michelson}
 
@@
 
-![Linear](Figures/linearsystem.pdf){#fig-linear}
-
-![Nonliner](Figures/transducer.pdf){#fig-transducer}
-
-![Two inputs](Figures/twoinputs.pdf){#fig-twoinputs}
-
-**Previous figures** showed how a single-input SSVEP might generate a single set of peaks in the frequency domain response which, in turn, can be used as a read-out of neuronal activity.
-In @fig-plotspectra, we show the effect of adding a second contrast component: the two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
-The nature of the interaction --- and therefore the pattern of intermodulation terms ---  is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan1988]
-
-First, we implement a simple non-linear transducer with a single input to illustrate the typical sigmoidal shape of a contrast response function (@fig-contrastresponsefunction).
-
-Next, we show that adding a second component which contributes to the gain pool of the first, results in a complex pattern of intermodulation terms (@fig-plotspectra).
-
-```{python}
-#| label: fig-plotsinewaves
-#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain."
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-import numpy as np
-import scipy as scp
-
-inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
-
-duration_S=1 # All variables in this code have an _[units] where possible
-
-digitisationRate_Hz=1000
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))
-
-for i, freq in enumerate(inputFrequencies_Hz):
-    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)
-
-# Create a figure with two subplots
-fig, axs = plt.subplots(len(inputFrequencies_Hz), 1, figsize=(8, 6))
-
-# Plot each input modulation in a separate subplot
-for i, freq in enumerate(inputFrequencies_Hz):
-    axs[i].plot(support_S, inputModulations[:, i])
-    axs[i].set_xlabel('Time (s)')
-    axs[i].set_ylabel('Amplitude')
-    axs[i].set_title('Frequency: {} Hz'.format(freq))
-
-# Adjust the spacing between subplots
-plt.tight_layout()
-plt.show()
-
-```
-
-The two inputs are represented here as amplitude modulations across time.
-
-If the two inputs are simply added together, the representation of the resulting signal in the Fourier domain is just the linear sum of the two independent signals (first two panels of @fig-plotspectra).
+To better understand these processes, we will show how sinusoidal input signals might be processed by the visual system to produce SSVEPs.
+@fig-linear illustrates how sine waves of different contrasts are processed in a linear system.
+The first panel shows the input sine wave, which would be used to modulate stimulus amplitude over time.
+Notice that there are five peaks in the waveform during the one second sample, so the stimulation frequency is 5Hz (F1).
+The second panel shows Fourier transform of the waveform, which contains a substantial peak at this frequency.
+If we parametrically vary the stimulus contrast (i.e.
+the amplitude of the waveform), the amplitude of the F1 component increases linearly with contrast (right panel).
+
+![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
+These calculations assume an entirely linear system.](Figures/linearsystem.pdf){#fig-linear}
+
+Next we can consider the impact of a nonlinearity in processing on the responses.
+One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
+In @fig-transducer this is modeled by a hyperbolic ratio function resulting in a saturating non-linearity:
+
+$$R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},$${#eq-nakarushton}
+
+where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast (or the time-varying waveform), $C_{50}$ is the 'semi-saturation constant' (the point at which the response is at half-maximum) and *n* controls the steepness of the curve.
+
+This nonlinearity has clear effects on signal transduction.
+The sinusoidal waveform is distorted by the nonlinearity, and appears frequency doubled (left panel of @fig-transducer; note that other nonlinearities such as rectification and squaring can have similar effects).
+The distortion is reflected in the Fourier spectrum (middle panel of @fig-transducer), which now includes responses at integer multiples (known as harmonics) of the stimulation frequency (10, 20 and 30Hz), but no response at the fundamental (5Hz).
+Finally, the contrast response function is now nonlinear (right panel of @fig-transducer).
+Although the hyperbolic ratio function (@eq-nakarushton) is monotonic, the CRF resulting from measuring the amplitude of the second harmonic (2F) component contains a slight roll-off at high input contrasts.
+This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
+Power at other harmonics increases, and the total power generated by the input is monotonic.
+This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler1985; @Peirce2007].
+
+![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
+The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
+The contrast response function is now nonlinear, and saturates at high contrasts (right panel).](Figures/transducer.pdf){#fig-transducer}
+
+To illustrate the effect of contrast gain control [@Heeger1992], we next include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
+The mask stimulus will suppress the response at the target frequency, reducing its amplitude.
+The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Busse2009].
+At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the transduced waveforms (left panel of @fig-twoinputs) and in the Fourier spectrum (middle panel of @fig-twoinputs).
+The two components interact generating nonlinear intermodulation terms at sums and differences of the input frequencies (F1, F2).
+The nature of the interaction --- and therefore the pattern of intermodulation terms ---  is determined by the computations happening as the contrast signal moves from retina to cortex [@Regan1988].
+
+![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
+Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/twoinputs.pdf){#fig-twoinputs}
+
+If the two inputs were simply added together, the representation of the resulting signal in the Fourier domain would be the linear sum of the two independent signals (i.e.
+peaks at F1 and F2).
 However, it is important to take physiology into account.
-If we ignore cells sensitive to contrast transients, contrast is represented as a combination of rectified signals from 'on' and 'off'-sensitive cells (broadly, cells that code positive or negative changes in spatiotemporal luminance).
-To model the population response of these cells to a contrast-reversing luminance sine wave, we therefore introduce an initial full-wave rectification stage that simulates the outputs of two, equal populations of on- and off-polarity detectors.
-Once this is done, we see that the Fourier spectrum of the resulting signal contains power only at the second harmonics and above (final two panels of @fig-plotspectra).
-It is possible to generate first-hamonic responses in contrast-response SSVEPs, but only when the inputs are modulated in an 'on/off' manner rather than contrast reversed.
+Contrast gain control reduces the amplitude of the responses via a suppressive process, which can be incorporated into our transducer function via an additional denominator term:
+
+$$R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n + C_{\mathrm{mask}}^n},$${#eq-gaincontrol}
+
+where $C_\mathrm{mask}}$ reflects the contrast of the mask component at a distinct frequency from that of the target.
+The effect of this extra term is to reduce the target response (see right panel of @fig-twoinputs).
+For a linear contrast axis, the contrast response function becomes shallower, whereas on a logarithmic contrast axis it maintains its steepness and shifts to the right.
+
+Even without considering a spatial component, the early visual system is far more complex than the model here suggests.
+For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code temporal changes in contrast.
+These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
+Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
+The SSVEP response to a contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
+
+The complexity of even a simple simulation of the frequency-domain signal arising from non-linear interactions is intriguing.
+Presumably, the signal measured from early visual cortex is the result of a cascade of nonlinear retinal and cortical operations up to that point.
+It therefore contains a 'signature' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
+In principle, that information could be recovered from the SSVEP signal - a possibility recognised in the early days of the technique [@Regan1988].
+Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
+For example, @Tsai2012 demonstrated that a gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
+This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.
+Our own work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
+More detailed modelling of intracortical recordings [@Groen2022] has revealed details of the timecourse of gain control effects and adaptation, specifically that normalization is delayed slightly relative to the initial visual response.
+
+Adaptation
+
+# Measuring the development of contrast processing
+
+An early use of the SSVEP was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
+In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
+However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
+In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [@Norcia1990; @Tyler1979; @Braddick1986] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
+
+Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
+This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
+
+![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.](Figures/sweepplot.pdf){#fig-sweepvep}
+
+A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrasr levels.
+This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
+To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990].
+The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
+In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].
+
+This approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
+In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983].
+At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
+However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
+
+The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section [@Candy2001; @Pei2017].
+Although contrast gain control is measurable in infants as young as six weeks old [@Candy2001; @Morrone1986; @Skoczenski1998], its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
+
+# Contrast processing in clinical conditions
+
+The SSVEP technique has also been used to study clinical conditions, such as diseases and developmental disorders.
+This can often be informative regarding the underlying mechanism that characterises the condition.
+Here we focus on four conditions, but there is potential to apply the method more broadly, either as a diagnostic technique, or to monitor disease severity and progression, or the efficacy of treatments.
+
+Epilepsy is a neurological condition in which patients experience seizures - episodes of uncontrolled neural activity that can cause unconsciousness, involuntary movements and convulsions, and atypical sensory experiences.
+@Porciatti2000 showed that individuals with photosensitive epilepsy generate larger steady-state signals in response to flickering visual stimuli, and their contrast response functions saturate less than those of healthy controls.
+This is consistent with the idea that epilepsy involves a cortical hyperexciteability that makes seizures more likely.
+It is also the case for individuals with idiopathic generalised epilepsy [@Tsai2011], a subtype of epilepsy that has a less obvious link to vision.
+The differences apply across the whole contrast-response function, and so resemble a response gain effect (see [@fig-plotclinical]a), which might be due to reduced inhibition from neighbouring neurons.
+Differences in SSVEP amplitudes have also been reported in individuals with migraine [@Shibata2008], a condition also associated with cortical hyperexciteability.
+
+Amblyopia is a disorder of binocular vision, characterised by one eye contributing much less to perception than the other.
+This is often due to strabismus (squint) or anisometropia (difference in optical prescription between the eyes) during development.
+Contemporary accounts suggest that the amblyopic eye is suppressed by signals from the fellow eye.
+SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye across the contrast range [@Baker2015; @Lygo2021].
+There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.
+The steady-state approach may be more sensitive and objective than typical acuity measurements, and also has the potential to measure suppression between the eyes directly [e.g.
+@Zheng2019; @Du2023; @Hu2023].
+
+Autism is a condition often associated with differences in vision [@Simmons2009] and other senses [@MacLennan2022].
+@Pei2014 used a sweep-VEP method with counterphase flickering stimuli, and found weaker responses in autistic children at spatial frequencies around 8c/deg, compared with age-matched controls.
+This was subsequently replicated in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
+Interestingly this study replicated its key findings in a *Drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.
+
+Recent work on understanding Parkinson's disease has also used *Drosophila* genetic models.
+@Afsari2014 found that mutant flies produced stronger SSVEP responses to flickering lights than control flies (see [@fig-plotclinical]d).
+The authors theorised that differences in early gain control during development might lead to visual deficits later in life.
+Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms (tremor, rigidity, slow movement) manifest.
+The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments.
+SSVEP responses also provide a potential method to diagnose Parkinson's before any symptoms manifest, and to monitor the effect of treatments.
+
+![Example contrast response functions for different clinical conditions.
+Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
+(2011).
+Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al.
+(2015).
+Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al.
+(2018).
+Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al.
+(2014).
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/clinicalplot.pdf){#fig-plotclinical}
+
+# Brain-computer interfaces
+
+I wonder if this is a good place to mention attention.
+
+# Conclusions
+
+*2: Sweeps and CRFs. Measurements of contrast sensitivity, extrapolating
+the sweep to zero response to get t'hold. Infants and adults?*
+
+*3: Measurements of modulation. Figures adapted from other papers:
+Attention to space, attention to features, adaptation (?).
+masking/surround suppression\\*
+
+*4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*
+
+*5: Future directions (decoding in frequency domain?, animals? BCI?)*
+
+[]{#anchor-2}History
+
+[]{#anchor-3}1. The basics of SSVEP and contrast sensitivity including a
+history of both fields
+
+SSVEP (Steady-State Visual Evoked Potential): A continuous electrical
+response evoked in the brain by visual stimuli flickering at a constant
+frequency (Regan, 1966).
+
+Contrast Sensitivity: The ability to detect differences in luminance
+between an object and its background (Campbell \& Green, 1965).
+
+Regan, D. (1966). Some characteristics of average steady-state and
+transient responses evoked by modulated light. Electroencephalography
+and clinical neurophysiology, 20(3), 238-248.
+
+Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
+affecting visual resolution. J Physiology, 181, 576-593.
+
+Check Regan paper for earlier (e.g. EEG refs). Norcia review will be
+helpful!
+
+(From Tyler / Levi / Apkarian paper):
+
+6\. Regan, D.: Rapid methods for refracting the eye and assessing the
+visual acuity in amblyopia using steady-state visual evoked potentials.
+In Desmedt, J.E., editor: Visual Evoked Potentials in Man: New
+Developments, Oxford, 1977, Clarendon Press, pp. 418-426.
+
+7\. Fricker, S.J.: Narrow-band filter techniques for the detection and
+measurement of evoked responses, Electroencephalogr. Clin. Neurophysiol.
+14:411, 1962.
+
+8\. Van der Tweel, L. H., Sem-Jacobsen, C.W., Kamp, A., Van Leeuwen,
+W.S., and Verings, F.T.H.: Objective determination of response to
+modulated light, Acta Physiol. Pharmacol. Neerl. 7:528, 1958.
+
+9\. Regan, D.: Latencies of evoked potentials to flicker and to pattern
+speedily estimated by simultaneous stimulation method,
+Electroencephalogr. Clin. Neurophysiol. 40:654, 1976.
+
+10\. Tyler, C.W., Apkarian, P., and Nakayama, K.: Multiple spatial
+frequency tuning of electrical responses from the human\....
+
+\[Around here perhaps a section about what gain control is, mentioning
+other methods as well including psychophysics, MRI, electrophysiology,
+and other EEG markers including ERPs and evoked gamma band oscillations.
+Maybe outline the Heeger gain control model and its cousins.\]
+
+\[Sure. But in fact gain control is only a small part of this story -
+especially in the early days. They were, I think, more interested in
+using SSVEP to measure absolute contrast sensitivity and to get lower
+bounds on things like infant visual development. Gain control might be
+better as a separate section later.\]
+
+[]{#anchor-4}2. Spatial, temporal frequency and contrast sensitivity
+measurements
+
+Basically using SSVEP to measure a cortical output amplitude for any
+given input contrast. You can vary parameters like SF, TF, position,
+color and of course contrast. Early on people realised that you can
+'sweep' the stimulus to get a CRF. You broadly get a line in log
+contrast space if you do that (Tyler) - then you can extrapolate that
+line down to zero response to estimate the threshold. That doesn't
+\>quite\< work but it's pretty close.
+
+You can also use this for 'difficult' populations like babies. One
+interesting story was about how SSVEP become a replacement for
+preferential looking (which was the other way of looking at infant
+visual development). See e.g. Davida Teller. SSVEPs allowed people to
+make objective measurements of contrast sensitivity development and
+deduce that the visual system was more mature (e.g. more functional) in
+infancy than previously expected. Also measures of colour sensitivity.
+Tinyeyes is based off those measurements. Other people: Tyler, Norcia,
+Gunilla H-P, many of the people at SKERI in the 1980s and 90s. Norcia
+86,88,90 - mentioned in Regan's nice autobiography :
+
+In parallel of course, people were using frequency tagging to do single
+unit work - the 1F vs 2F simple/complex cell classification scheme was
+all about this (Lennie and others).
+
+[]{#anchor-5}3. SSVEP in functional localization
+
+Techniques like fMRI have been combined with SSVEP to achieve more
+precise spatial localization (Di Russo et al., 2007). See however the
+Ales cruciform paper.
+
+Ales and Norcia:\... (showed that people's intuition about V1 upper /
+lower v.f reversing polarity wasn't really correct). This feels a little
+outside our scope though\... We have also combined SSVEP with source
+imaging techniques to probe responses in different cortical locations
+(again, Ales papers, some early stuff from Stan K? Appelbaum, Wade,
+Norcia figure/ground\...) and then a host of later work from that lab
+and others.
+
+[]{#anchor-6}4. SSVEP and contrast gain control including adaptation,
+masking, and attention
+
+\[The story goes: Up to the lte 90s people were primarily interested in
+measuring contrast sensitivity - the shape of the response function was
+assumed to be basically log-linear - and they fit it with straight lines
+to extrapolate back to zero response. But then (once Heeger's 1992 paper
+had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley \&
+Victor, 1981)](https://www.zotero.org/google-docs/?zAmtgj)), people
+started thinking about gain control - Candy and Norcia in about 1999,
+Porciatti, probably a load of Tyler papers that I don't even know
+about\...\]. Probably look in p
+
+And then people worked out that if you can use SSVEP to measure contrast
+responses, you can also use it to measure things that modulate contrast
+responses. These include adaptation, masking, suppression, attention
+(feature and space), clinical things.
+
+[]{#anchor-7}Adaptation:
+
+Continuous exposure to high-contrast patterns reduces contrast
+sensitivity, which can be measured using SSVEP (Ross et al., 1989).
+Others? Baker recent gain control paper is perhaps worth mentioning here
+as a 'confound' of sorts. Engel 2018 [*(Vergeer et al.,
+2018)*](https://www.zotero.org/google-docs/?kdURiS). This paper is
+interesting [*(Rideaux et al.,
+2023)*](https://www.zotero.org/google-docs/?J2EO3d) and \>sort\< of
+SSVEP.
+
+[]{#anchor-8}Masking:
+
+High contrast masks can suppress the visibility of low contrast
+patterns, which has implications in SSVEP amplitude (Haynes et al.,
+2003).
+
+Attention: Directing attention can enhance contrast sensitivity, as
+shown in studies using SSVEP (Müller et al., 2006). Also Tsai
+(dynamics), Baker / Wade (several), Winawer? I think JW has a nice
+dynamic model of normalization with some MEG data. Busse et al cat/human
+comparison. Candy and Norcia 2001 JNS [(Candy et al.,
+2001)](https://www.zotero.org/google-docs/?7hhs5Z)
+
+Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
+adaptation and masking on incremental thresholds for contrast. Vision
+research, 29(2), 205-215.
+
+Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
+Neuromagnetic correlates of perceived contrast in primary visual cortex.
+Journal of Neurophysiology, 89(6), 2655-2666.
+
+Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
+Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
+selective attention on the steady-state visual evoked potential in the
+20--28 Hz range. Cognitive Brain Research, 24(1), 1-13.
+
+5\. Clinical implications
+
+[]{#anchor-9}Clinical applications (this is a whole section)
+
+Not sure exactly how SSVEP used in clinic. mfVEP?
+
+, e.g., in monitoring visual impairments, tracking neuronal diseases, or
+neurofeedback (Norcia et al., 2015).
+
+Tsai epilepsy [(Tsai et al.,
+2011)](https://www.zotero.org/google-docs/?AdhH3f). Other photosensitive
+epilepsy: Porciatti [(Porciatti et al.,
+2000)](https://www.zotero.org/google-docs/?UTBQbV)
+
+Citation:
+
+Migraine (Regan et al). Autism?
+
+Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
+Rossion, B. (2015). The steady-state visual evoked potential in vision
+research: A review. Journal of vision, 15(6), 4-4.
+
+Animal work: Flies (Eliott, West, Himmelberg, Ales, Norcia): Mice/rats
+(probably many - can't think off the top of my head - we were working
+with a mouse EEG person at UCSF about 12 years ago\...), Monkeys
+(Kiorpes?),
+
+[]{#anchor-10}6. Future directions
+
+Use as a readout of modulations. TMS? FUS? Marmite B12 / Fluoxetine /
+amblyopia in Rats,
+
+GABA Huang
+
+There is a lot of SSVEP interest these days because of BCIs. I think
+it's pretty weak but there is \>so much\< of it that it might be worth
+mentioning\...
+
+Advanced signal processing techniques and machine learning can be
+integrated to improve SSVEP-based systems (Zhu et al., 2010).
+
+Exploring new clinical and diagnostic applications, understanding
+neurological diseases, and developing novel therapeutic interventions.
+
+Citation:
+
+Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
+stimulation methods used in SSVEP-based BCIs. Computational intelligence
+and neuroscience, 2010.
+
+# References
+
+<!-- old code segments that are no longer used -->
 
@@
 
-Even without considering a spatial component, the early visual system is far more complex than the model here suggests.
-For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code temporal changes in contrast.
-These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
-Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
-The SSVEP response to a contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
+```{python}
+#| label: fig-plotsinewaves
+#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain."
+#| fig-cap-location: bottom
+
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy as scp
+
+inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
+
+duration_S=1 # All variables in this code have an _[units] where possible
+
+digitisationRate_Hz=1000
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))
+
+for i, freq in enumerate(inputFrequencies_Hz):
+    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)
+
+# Create a figure with two subplots
+fig, axs = plt.subplots(len(inputFrequencies_Hz), 1, figsize=(8, 6))
+
+# Plot each input modulation in a separate subplot
+for i, freq in enumerate(inputFrequencies_Hz):
+    axs[i].plot(support_S, inputModulations[:, i])
+    axs[i].set_xlabel('Time (s)')
+    axs[i].set_ylabel('Amplitude')
+    axs[i].set_title('Frequency: {} Hz'.format(freq))
+
+# Adjust the spacing between subplots
+plt.tight_layout()
+plt.show()
+
+```
 
@@
 
-One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
-In @fig-contrastresponsefunction this is modeled as described above by a hyperbolic ratio function resulting in a saturating non-linearity:
-
-\begin{equation}
-R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},
-\end{equation}
-
-where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast, $C_{50}$ is the 'semi-saturation constant' (the point at which the response is at half-maximum) and *n* controls the steepness of the curve.
-
-The hyperbolic ratio function is monotonic but the CRF resulting from measuring the amplitude of the second harmonic (2F) component contains a slight roll-off at high input contrasts.
-This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
-Power at other harmonics increases, and the total power generated by the input is monotonic.
-This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler1985; @Peirce2007].
-
-To illustrate the effect of contrast gain control [@Heeger1992], we include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
-The mask stimulus will suppress the response at the target frequency, reducing its amplitude.
-The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Busse2009].
-At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT (@fig-intermodulation).
-
 ```{python}
@@
 ```
-
-The complexity of even a simple simulation of the frequency-domain signal arising from non-linear interactions is intriguing.
-Presumably, the signal measured from early visual cortex is the result of a cascade of nonlinear retinal and cortical operations up to that point.
-It therefore contains a 'signature' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
-In principle, that information could be recovered from the SSVEP signal.
-
-This possibility was recognised in the early days of the technique [@Regan1988].
-Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
-For example, @Tsai2012 demonstrated that a gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
-This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.
-Our own work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
-More detailed modelling of intracortical recordings [@Groen2022] has revealed details of the timecourse of gain control effects and adaptation, specifically that normalization is delayed slightly relative to the initial visual response.
-
-Adaptation
-
-# Measuring the development of contrast processing
-
-An early use of the SSVEP was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
-In well-motivated adults, psychophysical measurements of contrast sensitivity remain the gold standard.
-However, it is difficult and time consuming to obtain reliable psychophysical data from infants.
-In these cases, SSVEP measurements represent a fast and efficient method for measuring low-level visual responses [@Norcia1990; @Tyler1979; @Braddick1986] and the high SNR of SSVEP means that infants need only look at the screen for short periods of time.
-
-Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
-This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
-
-![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.](Figures/sweepplot.pdf){#fig-sweepvep}
-
-A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrasr levels.
-This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990].
-The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
-In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].
-
-This approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
-In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
-Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983].
-At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
-However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
-
-The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section [@Candy2001; @Pei2017].
-Although contrast gain control is measurable in infants as young as six weeks old [@Candy2001; @Morrone1986; @Skoczenski1998], its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
-
-# Contrast processing in clinical conditions
-
-The SSVEP technique has also been used to study clinical conditions, such as diseases and developmental disorders.
-This can often be informative regarding the underlying mechanism that characterises the condition.
-Here we focus on four conditions, but there is potential to apply the method more broadly, either as a diagnostic technique, or to monitor disease severity and progression, or the efficacy of treatments.
-
-Epilepsy is a neurological condition in which patients experience seizures - episodes of uncontrolled neural activity that can cause unconsciousness, involuntary movements and convulsions, and atypical sensory experiences.
-@Porciatti2000 showed that individuals with photosensitive epilepsy generate larger steady-state signals in response to flickering visual stimuli, and their contrast response functions saturate less than those of healthy controls.
-This is consistent with the idea that epilepsy involves a cortical hyperexciteability that makes seizures more likely.
-It is also the case for individuals with idiopathic generalised epilepsy [@Tsai2011], a subtype of epilepsy that has a less obvious link to vision.
-The differences apply across the whole contrast-response function, and so resemble a response gain effect (see [@fig-plotclinical]a), which might be due to reduced inhibition from neighbouring neurons.
-Differences in SSVEP amplitudes have also been reported in individuals with migraine [@Shibata2008], a condition also associated with cortical hyperexciteability.
-
-Amblyopia is a disorder of binocular vision, characterised by one eye contributing much less to perception than the other.
-This is often due to strabismus (squint) or anisometropia (difference in optical prescription between the eyes) during development.
-Contemporary accounts suggest that the amblyopic eye is suppressed by signals from the fellow eye.
-SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye across the contrast range [@Baker2015; @Lygo2021].
-There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.
-The steady-state approach may be more sensitive and objective than typical acuity measurements, and also has the potential to measure suppression between the eyes directly [e.g.
-@Zheng2019; @Du2023; @Hu2023].
-
-Autism is a condition often associated with differences in vision [@Simmons2009] and other senses [@MacLennan2022].
-@Pei2014 used a sweep-VEP method with counterphase flickering stimuli, and found weaker responses in autistic children at spatial frequencies around 8c/deg, compared with age-matched controls.
-This was subsequently replicated in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
-Interestingly this study replicated its key findings in a *Drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.
-
-Recent work on understanding Parkinson's disease has also used *Drosophila* genetic models.
-@Afsari2014 found that mutant flies produced stronger SSVEP responses to flickering lights than control flies (see [@fig-plotclinical]d).
-The authors theorised that differences in early gain control during development might lead to visual deficits later in life.
-Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms (tremor, rigidity, slow movement) manifest.
-The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments.
-SSVEP responses also provide a potential method to diagnose Parkinson's before any symptoms manifest, and to monitor the effect of treatments.
-
-![Example contrast response functions for different clinical conditions.
-Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
-(2011).
-Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al.
-(2015).
-Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al.
-(2018).
-Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al.
-(2014).
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/clinicalplot.pdf){#fig-plotclinical}
-
-# Brain-computer interfaces
-
-I wonder if this is a good place to mention attention.
-
-# Conclusions
-
-*2: Sweeps and CRFs. Measurements of contrast sensitivity, extrapolating
-the sweep to zero response to get t'hold. Infants and adults?*
-
-*3: Measurements of modulation. Figures adapted from other papers:
-Attention to space, attention to features, adaptation (?).
-masking/surround suppression\\*
-
-*4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*
-
-*5: Future directions (decoding in frequency domain?, animals? BCI?)*
-
-[]{#anchor-2}History
-
-[]{#anchor-3}1. The basics of SSVEP and contrast sensitivity including a
-history of both fields
-
-SSVEP (Steady-State Visual Evoked Potential): A continuous electrical
-response evoked in the brain by visual stimuli flickering at a constant
-frequency (Regan, 1966).
-
-Contrast Sensitivity: The ability to detect differences in luminance
-between an object and its background (Campbell \& Green, 1965).
-
-Regan, D. (1966). Some characteristics of average steady-state and
-transient responses evoked by modulated light. Electroencephalography
-and clinical neurophysiology, 20(3), 238-248.
-
-Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
-affecting visual resolution. J Physiology, 181, 576-593.
-
-Check Regan paper for earlier (e.g. EEG refs). Norcia review will be
-helpful!
-
-(From Tyler / Levi / Apkarian paper):
-
-6\. Regan, D.: Rapid methods for refracting the eye and assessing the
-visual acuity in amblyopia using steady-state visual evoked potentials.
-In Desmedt, J.E., editor: Visual Evoked Potentials in Man: New
-Developments, Oxford, 1977, Clarendon Press, pp. 418-426.
-
-7\. Fricker, S.J.: Narrow-band filter techniques for the detection and
-measurement of evoked responses, Electroencephalogr. Clin. Neurophysiol.
-14:411, 1962.
-
-8\. Van der Tweel, L. H., Sem-Jacobsen, C.W., Kamp, A., Van Leeuwen,
-W.S., and Verings, F.T.H.: Objective determination of response to
-modulated light, Acta Physiol. Pharmacol. Neerl. 7:528, 1958.
-
-9\. Regan, D.: Latencies of evoked potentials to flicker and to pattern
-speedily estimated by simultaneous stimulation method,
-Electroencephalogr. Clin. Neurophysiol. 40:654, 1976.
-
-10\. Tyler, C.W., Apkarian, P., and Nakayama, K.: Multiple spatial
-frequency tuning of electrical responses from the human\....
-
-\[Around here perhaps a section about what gain control is, mentioning
-other methods as well including psychophysics, MRI, electrophysiology,
-and other EEG markers including ERPs and evoked gamma band oscillations.
-Maybe outline the Heeger gain control model and its cousins.\]
-
-\[Sure. But in fact gain control is only a small part of this story -
-especially in the early days. They were, I think, more interested in
-using SSVEP to measure absolute contrast sensitivity and to get lower
-bounds on things like infant visual development. Gain control might be
-better as a separate section later.\]
-
-[]{#anchor-4}2. Spatial, temporal frequency and contrast sensitivity
-measurements
-
-Basically using SSVEP to measure a cortical output amplitude for any
-given input contrast. You can vary parameters like SF, TF, position,
-color and of course contrast. Early on people realised that you can
-'sweep' the stimulus to get a CRF. You broadly get a line in log
-contrast space if you do that (Tyler) - then you can extrapolate that
-line down to zero response to estimate the threshold. That doesn't
-\>quite\< work but it's pretty close.
-
-You can also use this for 'difficult' populations like babies. One
-interesting story was about how SSVEP become a replacement for
-preferential looking (which was the other way of looking at infant
-visual development). See e.g. Davida Teller. SSVEPs allowed people to
-make objective measurements of contrast sensitivity development and
-deduce that the visual system was more mature (e.g. more functional) in
-infancy than previously expected. Also measures of colour sensitivity.
-Tinyeyes is based off those measurements. Other people: Tyler, Norcia,
-Gunilla H-P, many of the people at SKERI in the 1980s and 90s. Norcia
-86,88,90 - mentioned in Regan's nice autobiography :
-
-In parallel of course, people were using frequency tagging to do single
-unit work - the 1F vs 2F simple/complex cell classification scheme was
-all about this (Lennie and others).
-
-[]{#anchor-5}3. SSVEP in functional localization
-
-Techniques like fMRI have been combined with SSVEP to achieve more
-precise spatial localization (Di Russo et al., 2007). See however the
-Ales cruciform paper.
-
-Ales and Norcia:\... (showed that people's intuition about V1 upper /
-lower v.f reversing polarity wasn't really correct). This feels a little
-outside our scope though\... We have also combined SSVEP with source
-imaging techniques to probe responses in different cortical locations
-(again, Ales papers, some early stuff from Stan K? Appelbaum, Wade,
-Norcia figure/ground\...) and then a host of later work from that lab
-and others.
-
-[]{#anchor-6}4. SSVEP and contrast gain control including adaptation,
-masking, and attention
-
-\[The story goes: Up to the lte 90s people were primarily interested in
-measuring contrast sensitivity - the shape of the response function was
-assumed to be basically log-linear - and they fit it with straight lines
-to extrapolate back to zero response. But then (once Heeger's 1992 paper
-had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley \&
-Victor, 1981)](https://www.zotero.org/google-docs/?zAmtgj)), people
-started thinking about gain control - Candy and Norcia in about 1999,
-Porciatti, probably a load of Tyler papers that I don't even know
-about\...\]. Probably look in p
-
-And then people worked out that if you can use SSVEP to measure contrast
-responses, you can also use it to measure things that modulate contrast
-responses. These include adaptation, masking, suppression, attention
-(feature and space), clinical things.
-
-[]{#anchor-7}Adaptation:
-
-Continuous exposure to high-contrast patterns reduces contrast
-sensitivity, which can be measured using SSVEP (Ross et al., 1989).
-Others? Baker recent gain control paper is perhaps worth mentioning here
-as a 'confound' of sorts. Engel 2018 [*(Vergeer et al.,
-2018)*](https://www.zotero.org/google-docs/?kdURiS). This paper is
-interesting [*(Rideaux et al.,
-2023)*](https://www.zotero.org/google-docs/?J2EO3d) and \>sort\< of
-SSVEP.
-
-[]{#anchor-8}Masking:
-
-High contrast masks can suppress the visibility of low contrast
-patterns, which has implications in SSVEP amplitude (Haynes et al.,
-2003).
-
-Attention: Directing attention can enhance contrast sensitivity, as
-shown in studies using SSVEP (Müller et al., 2006). Also Tsai
-(dynamics), Baker / Wade (several), Winawer? I think JW has a nice
-dynamic model of normalization with some MEG data. Busse et al cat/human
-comparison. Candy and Norcia 2001 JNS [(Candy et al.,
-2001)](https://www.zotero.org/google-docs/?7hhs5Z)
-
-Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
-adaptation and masking on incremental thresholds for contrast. Vision
-research, 29(2), 205-215.
-
-Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
-Neuromagnetic correlates of perceived contrast in primary visual cortex.
-Journal of Neurophysiology, 89(6), 2655-2666.
-
-Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
-Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
-selective attention on the steady-state visual evoked potential in the
-20--28 Hz range. Cognitive Brain Research, 24(1), 1-13.
-
-5\. Clinical implications
-
-[]{#anchor-9}Clinical applications (this is a whole section)
-
-Not sure exactly how SSVEP used in clinic. mfVEP?
-
-, e.g., in monitoring visual impairments, tracking neuronal diseases, or
-neurofeedback (Norcia et al., 2015).
-
-Tsai epilepsy [(Tsai et al.,
-2011)](https://www.zotero.org/google-docs/?AdhH3f). Other photosensitive
-epilepsy: Porciatti [(Porciatti et al.,
-2000)](https://www.zotero.org/google-docs/?UTBQbV)
-
-Citation:
-
-Migraine (Regan et al). Autism?
-
-Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
-Rossion, B. (2015). The steady-state visual evoked potential in vision
-research: A review. Journal of vision, 15(6), 4-4.
-
-Animal work: Flies (Eliott, West, Himmelberg, Ales, Norcia): Mice/rats
-(probably many - can't think off the top of my head - we were working
-with a mouse EEG person at UCSF about 12 years ago\...), Monkeys
-(Kiorpes?),
-
-[]{#anchor-10}6. Future directions
-
-Use as a readout of modulations. TMS? FUS? Marmite B12 / Fluoxetine /
-amblyopia in Rats,
-
-GABA Huang
-
-There is a lot of SSVEP interest these days because of BCIs. I think
-it's pretty weak but there is \>so much\< of it that it might be worth
-mentioning\...
-
-Advanced signal processing techniques and machine learning can be
-integrated to improve SSVEP-based systems (Zhu et al., 2010).
-
-Exploring new clinical and diagnostic applications, understanding
-neurological diseases, and developing novel therapeutic interventions.
-
-Citation:
-
-Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
-stimulation methods used in SSVEP-based BCIs. Computational intelligence
-and neuroscience, 2010.
-
-# References

### 2024-04-20 review.qmd
Fixed equation error
A few minor tweaks

@@
 X = np.log10(contrastRange[:4]).reshape(-1, 1)
+X2 = np.log10(np.logspace(0.1,1.1,nLevels)/100).reshape(-1, 1)
 y = transducedSignal[:4]
@@
 axs.semilogx(contrastRange, transducedSignal, 'o')
-axs.semilogx(contrastRange[:4], reg.predict(X), 'r--', label='Linear Fit')
+axs.semilogx(contrastRange[:4], reg.predict(X), 'r-', label='Linear Fit')
+axs.semilogx(np.logspace(0.1,1.1,nLevels)/100, reg.predict(X2), 'r--')
 axs.semilogx([x_intercept], [0], 'ro', label='Threshold')
@@
 plt.xticks(ticklocsx, ticklabelsx)
-plt.yticks(np.arange(0, 1.6, 0.5))
+plt.yticks(np.arange(0, 2, 0.5))
 plt.xlabel('Target contrast (%)')
@@
 plt.xticks(ticklocsx, ticklabelsx)
-plt.yticks(np.arange(0, 7, 2))
+plt.yticks(np.arange(0, 20, 4))
 plt.xlabel('Target contrast (%)')
@@
 
-where $C_\mathrm{mask}}$ reflects the contrast of the mask component at a distinct frequency from that of the target.
+where $C_\mathrm{mask}$ reflects the contrast of the mask component at a distinct frequency from that of the target.
 The effect of this extra term is to reduce the target response (see right panel of @fig-twoinputs).
@@
 
-![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.](Figures/sweepplot.pdf){#fig-sweepvep}
+![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
+The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/sweepplot.pdf){#fig-sweepvep}
 
@@
 Contemporary accounts suggest that the amblyopic eye is suppressed by signals from the fellow eye.
-SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye across the contrast range [@Baker2015; @Lygo2021].
+SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye (see [@fig-plotclinical]b) across the contrast range [@Baker2015; @Lygo2021].
 There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.

### 2024-04-20 review.qmd
Added brief section on adaptation

@@
 For a linear contrast axis, the contrast response function becomes shallower, whereas on a logarithmic contrast axis it maintains its steepness and shifts to the right.
+Suppressive effects of this kind have been obtained using SSVEP with a variety of different types of mask, including orthogonal overlaid masks, surround masks, and dichoptic masks [for a meta-analysis see @Baker2021].
 
@@
 Our own work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
-More detailed modelling of intracortical recordings [@Groen2022] has revealed details of the timecourse of gain control effects and adaptation, specifically that normalization is delayed slightly relative to the initial visual response.
-
-Adaptation
+More detailed modelling of intracortical recordings [@Groen2022] has revealed details of the timecourse of gain control effects, specifically that normalization is delayed slightly relative to the initial visual response.
+
+Similar changes to the contrast response function might also be obtained using adaptation paradigms, in which the visual system is exposed to high contrast stimuli for long durations.
+Psychophysically, adaptation increases detection thresholds, but has little effect on contrast discrimination performance [@Ross1993], much like pattern masks [@Foley1994].
+Although SSVEP adaptation effects show strong tuning for orientation [@Vergeer2018; @Campbell1970] and spatial frequency [@Mecacci1976], there appear to be no studies measuring the full SSVEP contrast response function before and after an extended period of adaptation.
 

### 2024-04-20 review.qmd
Added first BCI paragraph

@@
 
-I wonder if this is a good place to mention attention.
+One widespread application of the SSVEP technique is in the design of brain-computer interfaces, which seek to control some aspect of a computer using neural signals.
+The high SNR and good frequency resolution of SSVEPs make them an ideal candidate for this approach.
+Typical studies involve presenting an array of stimuli at different flicker frequencies, and having the participant select one either by overt attention (i.e.
+shifting fixation to foveate the selected stimulus) or covert attention (i.e.
+deploying attention to one stimulus whilst keeping fixated).
+EEG signals are highly sensitive to both visual field position [@Ales2010] and attentional state [@Verghese2012; @Lauritzen2010], so the response to the selected stimulus will typically increase relative to the others, allowing it to be identified by an on-line algorithm.
+
+Probably a summary of BCIs here.
 

### 2024-04-21 review.qmd
Minor edits

@@
 
-where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast (or the time-varying waveform), $C_{50}$ is the 'semi-saturation constant' (the point at which the response is at half-maximum) and *n* controls the steepness of the curve.
+where ${R_{\mathrm{max}}}$ describes the maximum response level, $C_{\mathrm{in}}$ is the input contrast (or the time-varying waveform), $C_{50}$ is the 'semi-saturation constant' (the point at which the response is at half-maximum) and *n* controls the steepness of the curve (with a typical value around $n=2$.
 
@@
 The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
-The contrast response function is now nonlinear, and saturates at high contrasts (right panel).](Figures/transducer.pdf){#fig-transducer}
+The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/transducer.pdf){#fig-transducer}
 

### 2024-04-24 review.qmd
Added some stuff on attention

@@
 
+# Attention
+
+Attention exerts a profound influence on performance: Allowing people to attend covertly to a spatial location for example, improves their performance on a target detection task significantly (Posner, https://pubmed.ncbi.nlm.nih.gov/11934448/, https://pubmed.ncbi.nlm.nih.gov/10788636/,  (Morrone, Denti, & Spinelli,
+2004; Pestelli, Ling, & Carrasco, 2009) ).
+In principle, this effect may be driven both by modulation of the underlying signal or by additional decision-theoretic factors such as reduction in spatial uncertainty (https://jov.arvojournals.org/article.aspx?articleid=2121966, https://jov.arvojournals.org/article.aspx?articleid=2192826).
+Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates (Luck,
+Chelazzi, Hillyard, & Desimone, 1997; Marcus & van
+Essen, 2002; McAdams & Maunsell, 1999, Mehta, Ulbert, and
+Schroeder (2000)), but with the advent of spatially-resolved human brain imaging methods in the late 90s it became apparent that spatial attention was linked to frank changes in both fMRI bold responses (Tootell, 98: (Brefczynski &
+De Yoe, 1999; Buracas & Boynton, 2007; Gandhi,
+Heeger, & Boynton, 1999; Kastner, Pinsk, De Weerd,
+Desimone, & Ungerleider, 1999; Li, Lu, Tjan, Dosher, &
+Chu, 2008; Murray, 2008; Silver, Ress, & Heeger, 2007;
+Somers, Dale, Seiffert, & Tootell, 1999; Tootell et al.,
+1998),) and EEG signals (https://pubmed.ncbi.nlm.nih.gov/16221931/).
+
+In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control modulation (Reynolds and Heeger 2009, Boynton Vis Res 2009).
+This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating ...
+(effect of stimulus configuration critical).
+Then...
+talk about Lauriten et al, Verghese (2 papers) maybe mention Crognale / Highsmith, Wang+Wade and Jasna's paper from 2018.
+https://www.sciencedirect.com/science/article/abs/pii/S1053811918303860) Probably lots more Carasco ...
+
 # Brain-computer interfaces
 
-One widespread application of the SSVEP technique is in the design of brain-computer interfaces, which seek to control some aspect of a computer using neural signals.
-The high SNR and good frequency resolution of SSVEPs make them an ideal candidate for this approach.
-Typical studies involve presenting an array of stimuli at different flicker frequencies, and having the participant select one either by overt attention (i.e.
+One widespread recent application of the SSVEP technique is in the design of brain-computer interfaces, which seek to control some aspect of a computer using neural signals.
+The high SNR and frequency resolution of SSVEPs make them an ideal candidate for this approach.
+Typical studies may involve presenting an array of stimuli at different flicker frequencies, and having the participant select one either by overt attention (i.e.
 shifting fixation to foveate the selected stimulus) or covert attention (i.e.

### 2024-04-24 review.qmd
Added some stuff on attention

@@
 
-Attention exerts a profound influence on performance: Allowing people to attend covertly to a spatial location for example, improves their performance on a target detection task significantly (Posner, https://pubmed.ncbi.nlm.nih.gov/11934448/, https://pubmed.ncbi.nlm.nih.gov/10788636/,  (Morrone, Denti, & Spinelli,
+Attention exerts a profound influence on visual performance: For example, allowing people to attend covertly to a spatial location for example, improves their performance on a target detection task significantly (Posner, https://pubmed.ncbi.nlm.nih.gov/11934448/, https://pubmed.ncbi.nlm.nih.gov/10788636/,  (Morrone, Denti, & Spinelli,
 2004; Pestelli, Ling, & Carrasco, 2009) ).

### 2024-04-25 review.qmd
Attention

@@
 Somers, Dale, Seiffert, & Tootell, 1999; Tootell et al.,
-1998),) and EEG signals (https://pubmed.ncbi.nlm.nih.gov/16221931/).
-
+1998),) and EEG signals (https://pubmed.ncbi.nlm.nih.gov/16221931/), .
+
+SSVEP measurements of attentional effects on both luminance and chromatic contrast strongly implicated gain control an an underlying mechanism [@DiRusso2001GainControl] but the literature was.
 In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control modulation (Reynolds and Heeger 2009, Boynton Vis Res 2009).
-This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating ...
+This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating the subtle interaction between the size of the attentional 'spotlight' and the stimulus which explains many apparent contradictions in the literature.
 (effect of stimulus configuration critical).

### 2024-04-25 review.qmd
Attention

@@
 Attention exerts a profound influence on visual performance: For example, allowing people to attend covertly to a spatial location for example, improves their performance on a target detection task significantly (Posner, https://pubmed.ncbi.nlm.nih.gov/11934448/, https://pubmed.ncbi.nlm.nih.gov/10788636/,  (Morrone, Denti, & Spinelli,
-2004; Pestelli, Ling, & Carrasco, 2009) ).
-In principle, this effect may be driven both by modulation of the underlying signal or by additional decision-theoretic factors such as reduction in spatial uncertainty (https://jov.arvojournals.org/article.aspx?articleid=2121966, https://jov.arvojournals.org/article.aspx?articleid=2192826).
+2004; Pestelli, Ling, & Carrasco, 2009) , [@bashinski1980enhancement]).
+
+In principle, this enhancement may be driven both by modulation of the underlying signal or noise characteristics, or by additional decision-theoretic factors such as reduction in spatial uncertainty (https://jov.arvojournals.org/article.aspx?articleid=2121966, https://jov.arvojournals.org/article.aspx?articleid=2192826).
+
 Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates (Luck,
@@
 
-SSVEP measurements of attentional effects on both luminance and chromatic contrast strongly implicated gain control an an underlying mechanism [@DiRusso2001GainControl] but the literature was.
-In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control modulation (Reynolds and Heeger 2009, Boynton Vis Res 2009).
-This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating the subtle interaction between the size of the attentional 'spotlight' and the stimulus which explains many apparent contradictions in the literature.
-(effect of stimulus configuration critical).
+SSVEP and psychophysical measurements of attentional effects on both luminance and chromatic contrast strongly implicated gain control an an underlying mechanism [@DiRusso2001GainControl, @lu1998external].
+In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control computation (Reynolds and Heeger 2009, Boynton Vis Res 2009).
+This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating subtle interactions between the size of the attentional 'spotlight' and the stimulus configuration which rationalise many apparent contradictions in the literature.
+Subsequent papers   (effect of stimulus configuration critical).
 Then...

### 2024-04-25 review.qmd
Attention

@@
 
-SSVEP and psychophysical measurements of attentional effects on both luminance and chromatic contrast strongly implicated gain control an an underlying mechanism [@DiRusso2001GainControl, @lu1998external].
+Electrophysiological and psychophysical measurements of attentional effects on both luminance and chromatic contrast strongly implicated gain control an an underlying mechanism [@dirusso99Attention, @DiRusso2001GainControl, @lu1998external] and demonstrated that these effects differed between chromatic and achromatic pathways [@dirusso1999mcpc].
 In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control computation (Reynolds and Heeger 2009, Boynton Vis Res 2009).

### 2024-04-25 review.qmd
Attention

@@
 For a linear contrast axis, the contrast response function becomes shallower, whereas on a logarithmic contrast axis it maintains its steepness and shifts to the right.
-Suppressive effects of this kind have been obtained using SSVEP with a variety of different types of mask, including orthogonal overlaid masks, surround masks, and dichoptic masks [for a meta-analysis see @Baker2021].
+Suppressive effects of this kind have been obtained using SSVEP with a variety of different types of mask, including orthogonal overlaid masks, surround masks, and dichoptic masks (Burr and Morrone 1987; Ross and Speed 1991, Candy, Norcia, Busse, [@salelkar2020interaction] [for a meta-analysis see @cunningham2017NonLinear, @Baker2021].
 

### 2024-04-27 review.qmd
Updated references

--- 2024-04-25 review.qmd
@@
 For a linear contrast axis, the contrast response function becomes shallower, whereas on a logarithmic contrast axis it maintains its steepness and shifts to the right.
-Suppressive effects of this kind have been obtained using SSVEP with a variety of different types of mask, including orthogonal overlaid masks, surround masks, and dichoptic masks [for a meta-analysis see @Baker2021].
+Suppressive effects of this kind have been obtained using SSVEP with a variety of different types of mask, including orthogonal overlaid masks, surround masks, and dichoptic masks [@Burr1987; @Ross1991; @Candy2001; @Busse2009; @Salelkar2020; @Cunningham2017; for a meta-analysis see @Baker2021].
 
@@
 
-Attention exerts a profound influence on visual performance: For example, allowing people to attend covertly to a spatial location for example, improves their performance on a target detection task significantly (Posner, https://pubmed.ncbi.nlm.nih.gov/11934448/, https://pubmed.ncbi.nlm.nih.gov/10788636/,  (Morrone, Denti, & Spinelli,
-2004; Pestelli, Ling, & Carrasco, 2009) , [@bashinski1980enhancement]).
-
-In principle, this enhancement may be driven both by modulation of the underlying signal or noise characteristics, or by additional decision-theoretic factors such as reduction in spatial uncertainty (https://jov.arvojournals.org/article.aspx?articleid=2121966, https://jov.arvojournals.org/article.aspx?articleid=2192826).
+Attention exerts a profound influence on visual performance: For example, allowing people to attend covertly to a spatial location for example, improves their performance on a target detection task significantly [@Posner1980; @Cameron2002; @Carrasco2000; @Bashinski1980; @Morrone2004; @Pestilli2009].
+
+In principle, this enhancement may be driven both by modulation of the underlying signal or noise characteristics, or by additional decision-theoretic factors such as reduction in spatial uncertainty [@Gould2007; @Petrov2006].
 
@@
 
-SSVEP and psychophysical measurements of attentional effects on both luminance and chromatic contrast strongly implicated gain control an an underlying mechanism [@DiRusso2001GainControl, @lu1998external].
+Electrophysiological and psychophysical measurements of attentional effects on both luminance and chromatic contrast strongly implicated gain control an an underlying mechanism [@DiRusso1999; @DiRusso2001; @Lu1998] and demonstrated that these effects differed between chromatic and achromatic pathways [@DiRusso1999b].
 In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control computation (Reynolds and Heeger 2009, Boynton Vis Res 2009).

### 2024-04-27 review.qmd
Fixed references

@@
 
-Attention exerts a profound influence on visual performance: For example, allowing people to attend covertly to a spatial location for example, improves their performance on a target detection task significantly [@Posner1980; @Cameron2002; @Carrasco2000; @Bashinski1980; @Morrone2004; @Pestilli2009].
-
+Attention exerts a profound influence on visual performance: for example, allowing people to attend covertly to a spatial location for example, improves their performance on a target detection task significantly [@Posner1980; @Cameron2002; @Carrasco2000; @Bashinski1980; @Morrone2004; @Pestilli2009].
 In principle, this enhancement may be driven both by modulation of the underlying signal or noise characteristics, or by additional decision-theoretic factors such as reduction in spatial uncertainty [@Gould2007; @Petrov2006].
-
-Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates (Luck,
-Chelazzi, Hillyard, & Desimone, 1997; Marcus & van
-Essen, 2002; McAdams & Maunsell, 1999, Mehta, Ulbert, and
-Schroeder (2000)), but with the advent of spatially-resolved human brain imaging methods in the late 90s it became apparent that spatial attention was linked to frank changes in both fMRI bold responses (Tootell, 98: (Brefczynski &
-De Yoe, 1999; Buracas & Boynton, 2007; Gandhi,
-Heeger, & Boynton, 1999; Kastner, Pinsk, De Weerd,
-Desimone, & Ungerleider, 1999; Li, Lu, Tjan, Dosher, &
-Chu, 2008; Murray, 2008; Silver, Ress, & Heeger, 2007;
-Somers, Dale, Seiffert, & Tootell, 1999; Tootell et al.,
-1998),) and EEG signals (https://pubmed.ncbi.nlm.nih.gov/16221931/), .
+Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates [@Luck1997; @Marcus2002; @McAdams1999; @Mehta2000a; @Mehta2000b], but with the advent of spatially-resolved human brain imaging methods in the late 1990s it became apparent that spatial attention was linked to frank changes in both fMRI bold responses
+[@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006].
 
 Electrophysiological and psychophysical measurements of attentional effects on both luminance and chromatic contrast strongly implicated gain control an an underlying mechanism [@DiRusso1999; @DiRusso2001; @Lu1998] and demonstrated that these effects differed between chromatic and achromatic pathways [@DiRusso1999b].
-In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control computation (Reynolds and Heeger 2009, Boynton Vis Res 2009).
+In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control computation [@Reynolds2009; @Boynton2009].
 This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating subtle interactions between the size of the attentional 'spotlight' and the stimulus configuration which rationalise many apparent contradictions in the literature.

### 2024-04-28 review.qmd
More attention refs

@@
 In principle, this enhancement may be driven both by modulation of the underlying signal or noise characteristics, or by additional decision-theoretic factors such as reduction in spatial uncertainty [@Gould2007; @Petrov2006].
-Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates [@Luck1997; @Marcus2002; @McAdams1999; @Mehta2000a; @Mehta2000b], but with the advent of spatially-resolved human brain imaging methods in the late 1990s it became apparent that spatial attention was linked to frank changes in both fMRI bold responses
+Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates [@Luck1997; @Marcus2002; @McAdams1999; @Mehta2000a; @Mehta2000b], but with the advent of spatially-resolved human brain imaging methods in the late 1990s it became apparent that spatial attention was linked to frank changes in both fMRI BOLD responses
 [@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006].
@@
 Then...
-talk about Lauriten et al, Verghese (2 papers) maybe mention Crognale / Highsmith, Wang+Wade and Jasna's paper from 2018.
-https://www.sciencedirect.com/science/article/abs/pii/S1053811918303860) Probably lots more Carasco ...
+talk about @Lauritzen2010 @Verghese2012 (1 other Verghese paper) maybe mention @Highsmith2010, @Wang2011 and @Martinovic2018 Probably lots more Carasco ...
 

### 2024-05-05 review.qmd
Editing attention

@@
 
-Attention exerts a profound influence on visual performance: for example, allowing people to attend covertly to a spatial location for example, improves their performance on a target detection task significantly [@Posner1980; @Cameron2002; @Carrasco2000; @Bashinski1980; @Morrone2004; @Pestilli2009].
+Attention exerts a profound influence on visual performance: for example, allowing people to attend covertly to a spatial location improves their performance on a target detection task significantly [@Posner1980; @Cameron2002; @Carrasco2000; @Bashinski1980; @Morrone2004; @Pestilli2009].
 In principle, this enhancement may be driven both by modulation of the underlying signal or noise characteristics, or by additional decision-theoretic factors such as reduction in spatial uncertainty [@Gould2007; @Petrov2006].
@@
 
-Electrophysiological and psychophysical measurements of attentional effects on both luminance and chromatic contrast strongly implicated gain control an an underlying mechanism [@DiRusso1999; @DiRusso2001; @Lu1998] and demonstrated that these effects differed between chromatic and achromatic pathways [@DiRusso1999b].
+Electrophysiological and psychophysical measurements of the effect of attention on both luminance and chromatic contrast have strongly implicated gain control as an underlying mechanism [@DiRusso1999; @DiRusso2001; @Lu1998].
+These effects also appear to differ between chromatic and achromatic pathways [@DiRusso1999b] - perhaps as a result of the different levels of nonlinear gain control in the early precotical magno,  parvo and konio-cellular pathways [DerringtonLennie84, KaplanShapley86, LeePokornyJOSA90, SolomonLennie2005].
 In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control computation [@Reynolds2009; @Boynton2009].
 This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating subtle interactions between the size of the attentional 'spotlight' and the stimulus configuration which rationalise many apparent contradictions in the literature.
-Subsequent papers   (effect of stimulus configuration critical).
-Then...
-talk about @Lauritzen2010 @Verghese2012 (1 other Verghese paper) maybe mention @Highsmith2010, @Wang2011 and @Martinovic2018 Probably lots more Carasco ...
+Direct measurements of attentional modulation of achromatic SSVEP signals are broadly consistent with this model [@Lauritzen2010, @Hou2016, @Martinovic2018] and confirm the relatively weaker role of spatial attention on responses driven by chromatic stimuli - particularly those that isolate the opponent S-(L+M) cone pathway [@Wang2011, @Highsmith2010].
+The gain control model of attention can also be extended to SSVEP studies of feature-based attention which show that the modulatory effects can be targeted to the most informative neuronal populations [@Verghese2012].
 

### 2024-05-05 review.tex
Added more figure cation, grids
Editing attention

@@
 that the computations underlying contrast processing depend on factors
-such as spatial and temporal frequency, age, eccentricity, chromaticity
-and the presence of other stimuli has been a focus of vision science for
-at least 100 years. One of the most productive experimental approaches
-in this field has been the use of the `steady-state visually-evoked
-potential' (SSVEP): a technique where contrast modulating inputs are
-`frequency tagged' (presented at well-defined frequencies and phases)
-and the electrical signals that they generate in the brain are analyzed
-in the temporal frequency domain. SSVEPs have several advantages over
-conventional measures of visually-evoked responses: they have relatively
-unambiguous ouput measures, a high SNR, and they allow us to analyze
-interactions between stimulus components using a convenient mathematical
-framework. Here we describe how SSVEPs have been used to study visual
-contrast over the past 70 years (Dawson, 1954). Because our thinking
-about SSVEPs is well-described by simple mathematical models, we embed
-code that illustrates key steps in the modelling and analysis. This
-paper can therefore be used both as a review of the use of SSVEP in
-measuring human contrast processing, and as an interactive learning aid.
+such as spatial and temporal frequency, age, clinical conditions,
+eccentricity, chromaticity and the presence of other stimuli has been a
+focus of vision science for at least 100 years. One of the most
+productive experimental approaches in this field has been the use of the
+`steady-state visually-evoked potential' (SSVEP): a technique where
+contrast modulating inputs are `frequency tagged' (presented at
+well-defined frequencies and phases) and the electrical signals that
+they generate in the brain are analyzed in the temporal frequency
+domain. SSVEPs have several advantages over conventional measures of
+visually-evoked responses: they have relatively unambiguous ouput
+measures, a high SNR, and they allow us to analyze interactions between
+stimulus components using a convenient mathematical framework. Here we
+describe how SSVEPs have been used to study visual contrast over the
+past 70 years (Dawson, 1954). Because our thinking about SSVEPs is
+well-described by simple mathematical models, we embed code that
+illustrates key steps in the modelling and analysis. This paper can
+therefore be used both as a review of the use of SSVEP in measuring
+human contrast processing, and as an interactive learning aid.
 
@@
 (`temporal contrast'). Cone photoreceptors - which drive precortical
-opponent pathways - contribute to both chromatic and achromatic contrast
-and although most of the research we describe here focuses on achromatic
-contrast, SSVEPs have proven to be an excellent measure of early
-chromatic processing as well (McKeefry \emph{et al.}, 1996; Baseler \&
-Sutter, 1997; Di Russo \emph{et al.}, 2001).
-
-Contrast is relatively simple to define: typically, contrast is
-specified as the percentage deviation of a uniform stimulus from the
-background. So, for example, a disk of 100 units of cone activation
+opponent pathways - contribute to both chromatic and achromatic
+contrast, and although most of the research we describe here focuses on
+achromatic contrast, SSVEPs have proven to be an excellent measure of
+early chromatic processing as well (McKeefry \emph{et al.}, 1996;
+Baseler \& Sutter, 1997; Di Russo \emph{et al.}, 2001\emph{a}).
+
+Contrast is relatively simple to define: typically, it is specified as
+the percentage deviation of a uniform stimulus from the background. So,
+for example, a disk of 100 units of cone activation
 (\(I_{\mathrm{stim}}\)) surrounded by a `background' of 50 units of
@@
 
-\begin{equation}
-\frac {I_{\mathrm{stimmax} }-I_{\mathrm{stimmin} }}{I_{\mathrm{stimmax} }+I_{\mathrm{stimmin} }}.
-\end{equation}
+\begin{equation}\phantomsection\label{eq-michelson}{\frac {I_{\mathrm{stimmax} }-I_{\mathrm{stimmin} }}{I_{\mathrm{stimmax} }+I_{\mathrm{stimmin} }}.}\end{equation}
 
@@
 
-\includegraphics{review_files/figure-pdf/fig-centresurround-output-1.pdf}
+\includegraphics{Figures/centresurround.pdf}
 
@@
 depends on its context. A high contrast surround reduces the apparent
-contrast of the central `probe' region''}
+contrast of the central `probe' region.}
 
@@
 is typically defined as the amplitude of the SSVEP frequency component
-at the stimulus frequency or a low multiple thereof. This corresponds
+at the stimulus frequency, or a low multiple thereof. This corresponds
 closely to similar functions reported by studies measuring single unit
@@
 
-Previous figures showed how a single-input SSVEP might generate a single
-set of peaks in the frequency domain response which, in turn, can be
-used as a read-out of neuronal activity. In
-Figure~\ref{fig-plotspectra}, we show the effect of adding a second
-contrast component: the two components interact generating nonlinear
-intermodulation terms at sums and differences of the input frequencies
-(F1, F2). The nature of the interaction --- and therefore the pattern of
-intermodulation terms --- is determined by the computations happening as
-the contrast signal moves from retina to cortex (Regan \& Regan, 1988)
-
-First, we implement a simple non-linear transducer with a single input
-to illustrate the typical sigmoidal shape of a contrast response
-function (Figure~\ref{fig-contrastresponsefunction}).
-
-Next, we show that adding a second component which contributes to the
-gain pool of the first, results in a complex pattern of intermodulation
-terms (Figure~\ref{fig-plotspectra}).
+To better understand these processes, we will show how sinusoidal input
+signals might be processed by the visual system to produce SSVEPs.
+Figure~\ref{fig-linear} illustrates how sine waves of different
+contrasts are processed in a linear system. The first panel shows the
+input sine wave, which would be used to modulate stimulus amplitude over
+time. Notice that there are five peaks in the waveform during the one
+second sample, so the stimulation frequency is 5Hz (F1). The second
+panel shows Fourier transform of the waveform, which contains a
+substantial peak at this frequency. If we parametrically vary the
+stimulus contrast (i.e.~the amplitude of the waveform), the amplitude of
+the F1 component increases linearly with contrast (right panel).
 
@@
 
-\includegraphics{review_files/figure-pdf/fig-plotsinewaves-output-1.pdf}
+\includegraphics{Figures/linearsystem.pdf}
 
@@
 
-\caption{\label{fig-plotsinewaves}a) A typical SSVEP setup: a screen, a
-grating, a flicker rate: b) The responses from a single electrode (time
-domain) c) Frequency domain}
+\caption{\label{fig-linear}Illustration of a sinusoidal input signal
+(left), its Fourier spectrum (middle), and how the amplitude of the
+first harmonic component increases with contrast (right). These
+calculations assume an entirely linear system.}
 
@@
 
-The two inputs are represented here as amplitude modulations across
-time.
-
-If the two inputs are simply added together, the representation of the
-resulting signal in the Fourier domain is just the linear sum of the two
-independent signals (first two panels of Figure~\ref{fig-plotspectra}).
-However, it is important to take physiology into account. If we ignore
-cells sensitive to contrast transients, contrast is represented as a
-combination of rectified signals from `on' and `off'-sensitive cells
-(broadly, cells that code positive or negative changes in spatiotemporal
-luminance). To model the population response of these cells to a
-contrast-reversing luminance sine wave, we therefore introduce an
-initial full-wave rectification stage that simulates the outputs of two,
-equal populations of on- and off-polarity detectors. Once this is done,
-we see that the Fourier spectrum of the resulting signal contains power
-only at the second harmonics and above (final two panels of
-Figure~\ref{fig-plotspectra}). It is possible to generate first-hamonic
-responses in contrast-response SSVEPs, but only when the inputs are
-modulated in an `on/off' manner rather than contrast reversed.
+Next we can consider the impact of a nonlinearity in processing on the
+responses. One of the simplest nonlinearities is the function that
+describes a cell's response to different levels of contrast. In
+Figure~\ref{fig-transducer} this is modeled by a hyperbolic ratio
+function resulting in a saturating non-linearity:
+
+\begin{equation}\phantomsection\label{eq-nakarushton}{R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},}\end{equation}
+
+where \({R_{\mathrm{max}}}\) describes the maximum response level,
+\(C_{\mathrm{in}}\) is the input contrast (or the time-varying
+waveform), \(C_{50}\) is the `semi-saturation constant' (the point at
+which the response is at half-maximum) and \emph{n} controls the
+steepness of the curve (with a typical value around \(n=2\).
+
+This nonlinearity has clear effects on signal transduction. The
+sinusoidal waveform is distorted by the nonlinearity, and appears
+frequency doubled (left panel of Figure~\ref{fig-transducer}; note that
+other nonlinearities such as rectification and squaring can have similar
+effects). The distortion is reflected in the Fourier spectrum (middle
+panel of Figure~\ref{fig-transducer}), which now includes responses at
+integer multiples (known as harmonics) of the stimulation frequency (10,
+20 and 30Hz), but no response at the fundamental (5Hz). Finally, the
+contrast response function is now nonlinear (right panel of
+Figure~\ref{fig-transducer}). Although the hyperbolic ratio function
+(Equation~\ref{eq-nakarushton}) is monotonic, the CRF resulting from
+measuring the amplitude of the second harmonic (2F) component contains a
+slight roll-off at high input contrasts. This results from the
+distortion of the input sine waves at high contrast due to a combination
+of the full-wave rectification and saturating non-linearity. Power at
+other harmonics increases, and the total power generated by the input is
+monotonic. This roll-off is often seen in experimental data and has been
+referred to as `supersaturation' (Tyler \& Apkarian, 1985; Peirce,
+2007).
 
@@
 
-\includegraphics{review_files/figure-pdf/fig-plotspectra-output-1.pdf}
+\includegraphics{Figures/transducer.pdf}
 
@@
 
-\caption{\label{fig-plotspectra}Temporal frequency representations of
-combining two inputs at frequencies F1=5Hz and F2=7Hz combination. A
-pure sine-wave input would generate responses only at F1 and F2.
-However, the rectification that accompanies early visual processing
-converts this contrast reversal into a more complex frequency-domain
-representation with power predominantly at the second harmonics 2F1 and
-2F2}
+\caption{\label{fig-transducer}Illustration of how a sinusoidal input
+signal is affected by a nonlinear transducer. The input wavefore (left
+panel) is frequency doubled by the squaring operator, and now features
+energy at the second harmonic (10Hz) and its multiple (middle panel).
+The contrast response function at the second harmonic is now nonlinear,
+and saturates at high contrasts (right panel).}
 
 \end{figure}%
+
+To illustrate the effect of contrast gain control (Heeger, 1992), we
+next include a second component (a `mask') that contributes to the gain
+pool of the first (the `target'). The mask stimulus will suppress the
+response at the target frequency, reducing its amplitude. The
+suppression is reciprocal - activity at the mask frequency is also
+reduced by the presence of the target, in a contrast-dependent manner
+(Busse \emph{et al.}, 2009). At a single pair of (matched) contrast
+levels, we see a complex pattern of intermodulation terms in the
+transduced waveforms (left panel of Figure~\ref{fig-twoinputs}) and in
+the Fourier spectrum (middle panel of Figure~\ref{fig-twoinputs}). The
+two components interact generating nonlinear intermodulation terms at
+sums and differences of the input frequencies (F1, F2). The nature of
+the interaction --- and therefore the pattern of intermodulation terms
+--- is determined by the computations happening as the contrast signal
+moves from retina to cortex (Regan \& Regan, 1988).
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{Figures/twoinputs.pdf}
+
+}
+
+\caption{\label{fig-twoinputs}Demonstration of how two inputs of
+different frequencies are processed by a nonlinear transducer.
+Additional harmonics are apparent in the Fourier spectrum (middle
+panel), and the contrast response function is suppressed (right panel).}
+
+\end{figure}%
+
+If the two inputs were simply added together, the representation of the
+resulting signal in the Fourier domain would be the linear sum of the
+two independent signals (i.e.~peaks at F1 and F2). However, it is
+important to take physiology into account. Contrast gain control reduces
+the amplitude of the responses via a suppressive process, which can be
+incorporated into our transducer function via an additional denominator
+term:
+
+\begin{equation}\phantomsection\label{eq-gaincontrol}{R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n + C_{\mathrm{mask}}^n},}\end{equation}
+
+where \(C_\mathrm{mask}\) reflects the contrast of the mask component at
+a distinct frequency from that of the target. The effect of this extra
+term is to reduce the target response (see right panel of
+Figure~\ref{fig-twoinputs}). For a linear contrast axis, the contrast
+response function becomes shallower, whereas on a logarithmic contrast
+axis it maintains its steepness and shifts to the right. Suppressive
+effects of this kind have been obtained using SSVEP with a variety of
+different types of mask, including orthogonal overlaid masks, surround
+masks, and dichoptic masks (Burr \& Morrone, 1987; Ross \& Speed, 1991;
+Candy \emph{et al.}, 2001; Busse \emph{et al.}, 2009; Cunningham
+\emph{et al.}, 2017; Salelkar \& Ray, 2020; for a meta-analysis see
+Baker \emph{et al.}, 2021).
 
@@
 contrast modulation while `complex cells' respond to the presence of
-patterned spatial contrast irrespective of its spatial position. The
-SSVEP response to a contrast-reversing sine-wave grating therefore
-contains information about nonlinear computations performed across a
-range of retinal and cortical cell types.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{review_files/figure-pdf/fig-contrastresponsefunction-output-1.pdf}
-
-}
-
-\caption{\label{fig-contrastresponsefunction}Non-linear contrast
-response function resulting from the population average of two full-wave
-rectified signals passing through a non-linear transducer. Note the
-`supersaturation' at high contrast levels.}
-
-\end{figure}%
-
-One of the simplest nonlinearities is the function that describes a
-cell's response to different levels of contrast. In
-Figure~\ref{fig-contrastresponsefunction} this is modeled as described
-above by a hyperbolic ratio function resulting in a saturating
-non-linearity:
-
-\begin{equation}
-R_{\mathrm{max}} = \frac{C_{\mathrm{in}}^n}{C_{50}^n + C_{\mathrm{in}}^n},
-\end{equation}
-
-where \({R_{\mathrm{max}}}\) describes the maximum response level,
-\(C_{\mathrm{in}}\) is the input contrast, \(C_{50}\) is the
-`semi-saturation constant' (the point at which the response is at
-half-maximum) and \emph{n} controls the steepness of the curve.
-
-The hyperbolic ratio function is monotonic but the CRF resulting from
-measuring the amplitude of the second harmonic (2F) component contains a
-slight roll-off at high input contrasts. This results from the
-distortion of the input sine waves at high contrast due to a combination
-of the full-wave rectification and saturating non-linearity. Power at
-other harmonics increases, and the total power generated by the input is
-monotonic. This roll-off is often seen in experimental data and has been
-referred to as `supersaturation' (Tyler \& Apkarian, 1985; Peirce,
-2007).
-
-To illustrate the effect of contrast gain control (Heeger, 1992), we
-include a second component that contributes to the gain pool of the
-first. At a single pair of (matched) contrast levels, we see a complex
-pattern of intermodulation terms in the FT
-(Figure~\ref{fig-intermodulation}).
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{review_files/figure-pdf/fig-intermodulation-output-1.pdf}
-
-}
-
-\caption{\label{fig-intermodulation}Non-linear combination of two
-independent contrast modulations at F1 and F2 results in
-`intermodulation' terms that appear at low-order sums and differences of
-those frequencies}
-
-\end{figure}%
+patterned spatial contrast irrespective of its spatial position (Hubel
+\& Wiesel, 1962). The SSVEP response to a contrast-reversing sine-wave
+grating therefore contains information about nonlinear computations
+performed across a range of retinal and cortical cell types.
 
@@
 signal arising from non-linear interactions is intriguing. Presumably,
-the signal measured from early visual cortex contains is the result of a
-cascade of nonlinear retinal and cortical operations up to that point.
-It therefore contains a `signature' of the computational nature, order
-and parameters of those operations - including the shape of the
-transducer functions and the computations involved in signal
-combination. In principle, that information could be recovered from the
-SSVEP signal.
-
-This possibility was recognised in the early days of the technique
-(Regan \& Regan, 1988). Although characterising the complete set of
-computations along entire processing pathway is problematic, careful
-parametric variation of the input stimuli does allow us to fit models of
-early visual gain control using SSVEP data. \ldots{} MORE HERE -
-Although this idea was suggested as early as 1988 (probably earlier) not
-many people seem to have done it much. Candy \emph{et al.} (2001); Busse
-\emph{et al.} (2009); Tsai \emph{et al.} (2012); Baker \& Wade (2017)
-*** SEE ALSO RECENT WORK FROM WINAWER LAB - MUCH OF IT IS `FAST' VEPs
-***
+the signal measured from early visual cortex is the result of a cascade
+of nonlinear retinal and cortical operations up to that point. It
+therefore contains a `signature' of the computational nature, order and
+parameters of those operations - including the shape of the transducer
+functions and the computations involved in signal combination. In
+principle, that information could be recovered from the SSVEP signal - a
+possibility recognised in the early days of the technique (Regan \&
+Regan, 1988). Although characterising the complete set of computations
+along entire processing pathway is challenging, careful parametric
+variation of the input stimuli does allow us to fit models of early
+visual gain control using SSVEP data. For example, Tsai \emph{et al.}
+(2012) demonstrated that a gain control model gave a good account of the
+pattern of intermodulation responses produced by two overlaid patterns
+flickering at different frequencies. This was achieved by passing full
+stimulus waveforms through the transducer nonlinearity, and calculating
+the Fourier spectrum of the model output. Our own work on signal
+combination across eyes and space similarly demonstrated close
+correspondence between the predictions of a computational model and
+empirical data in humans (Baker \& Wade, 2017). More detailed modelling
+of intracortical recordings (Groen \emph{et al.}, 2022) has revealed
+details of the timecourse of gain control effects, specifically that
+normalization is delayed slightly relative to the initial visual
+response.
+
+Similar changes to the contrast response function might also be obtained
+using adaptation paradigms, in which the visual system is exposed to
+high contrast stimuli for long durations. Psychophysically, adaptation
+increases detection thresholds, but has little effect on contrast
+discrimination performance (Ross \emph{et al.}, 1993), much like pattern
+masks (Foley, 1994). Although SSVEP adaptation effects show strong
+tuning for orientation (Campbell \& Maffei, 1970; Vergeer \emph{et al.},
+2018) and spatial frequency (Mecacci \& Spinelli, 1976), there appear to
+be no studies measuring the full SSVEP contrast response function before
+and after an extended period of adaptation.
 
@@
 
-\includegraphics{review_files/figure-pdf/fig-sweepvep-output-1.pdf}
+\includegraphics{Figures/sweepplot.pdf}
 
@@
 detection threshold can be estimated from sweep VEP data measured at
-higher contrasts}
+higher contrasts. The solid line is the regression fit to the lowest
+four data points, and the dashed line extrapolates the fit back to
+determine the contrast value when y=0, which gives a threshold estimate
+(red point).}
 
@@
 the difference in neural response to signals in each eye, and typically
-show reduced responses to stimuli in the amblyopic eye across the
-contrast range (Baker \emph{et al.}, 2015; Lygo \emph{et al.}, 2021).
-There are currently many novel binocular treatments for amblyopia under
-development, often involving virtual reality or stereo display systems
-designed to encourage the two eyes to work together. The steady-state
-approach may be more sensitive and objective than typical acuity
-measurements, and also has the potential to measure suppression between
-the eyes directly (e.g. Zheng \emph{et al.}, 2019; Hu \emph{et al.},
-2023; Du \emph{et al.}, 2023).
+show reduced responses to stimuli in the amblyopic eye (see
+Figure~\ref{fig-plotclinical}b) across the contrast range (Baker
+\emph{et al.}, 2015; Lygo \emph{et al.}, 2021). There are currently many
+novel binocular treatments for amblyopia under development, often
+involving virtual reality or stereo display systems designed to
+encourage the two eyes to work together. The steady-state approach may
+be more sensitive and objective than typical acuity measurements, and
+also has the potential to measure suppression between the eyes directly
+(e.g. Zheng \emph{et al.}, 2019; Hu \emph{et al.}, 2023; Du \emph{et
+al.}, 2023).
 
@@
 
-Parkinson's
-
-Alex - it would be good to add some CRFs from Afsari \emph{et al.}
-(2014) to the figure. Is it possible to extract either the fitted
-parameter values or the raw data for some control flies and some
-mutants? Whatever you think is most compelling.
+Recent work on understanding Parkinson's disease has also used
+\emph{Drosophila} genetic models. Afsari \emph{et al.} (2014) found that
+mutant flies produced stronger SSVEP responses to flickering lights than
+control flies (see Figure~\ref{fig-plotclinical}d). The authors
+theorised that differences in early gain control during development
+might lead to visual deficits later in life. Although visual responses
+are a convenient assay of neural function, it is likely that same
+general process applies throughout the whole brain, including in the
+motor system where the core Parkinson's symptoms (tremor, rigidity, slow
+movement) manifest. The SSVEP differences were reduced by a kinase
+inhibitor that targets the dopamine system, demonstrating how model
+organisms can be used to test new pharmacological treatments. SSVEP
+responses also provide a potential method to diagnose Parkinson's before
+any symptoms manifest, and to monitor the effect of treatments.
 
@@
 
-\includegraphics{review_files/figure-pdf/fig-plotclinical-output-1.pdf}
+\includegraphics{Figures/clinicalplot.pdf}
 
@@
 functions for children with (green) and without (black) a diagnosis of
-autism, based on the data of Vilidaite et al.~(2018). Panel (d)}
+autism, based on the data of Vilidaite et al.~(2018). Panel (d) shows
+data from \emph{Drosophila melanogaster} (fruit flies) from Afsari et
+al.~(2014). One day-old flies expressing a human gene linked to
+Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and
+sensitivity (red) compared to control animals (black).}
 
@@
 
+\section{Attention}\label{attention}
+
+Attention exerts a profound influence on visual performance: for
+example, allowing people to attend covertly to a spatial location for
+example, improves their performance on a target detection task
+significantly (Bashinski \& Bacharach, 1980; Posner, 1980; Carrasco
+\emph{et al.}, 2000; Cameron \emph{et al.}, 2002; Morrone \emph{et al.},
+2004; Pestilli \emph{et al.}, 2009). In principle, this enhancement may
+be driven both by modulation of the underlying signal or noise
+characteristics, or by additional decision-theoretic factors such as
+reduction in spatial uncertainty (Petrov \emph{et al.}, 2006; Gould
+\emph{et al.}, 2007). Early experiments with non-human primates showed
+little evidence for attentionally-driven changes in neuronal spike rates
+(Luck \emph{et al.}, 1997; McAdams \& Maunsell, 1999; Mehta \emph{et
+al.}, 2000\emph{a}, 2000\emph{b}; Marcus \& Van Essen, 2002), but with
+the advent of spatially-resolved human brain imaging methods in the late
+1990s it became apparent that spatial attention was linked to frank
+changes in both fMRI BOLD responses (Tootell \emph{et al.}, 1998; Somers
+\emph{et al.}, 1999; Gandhi \emph{et al.}, 1999; Brefczynski \& DeYoe,
+1999; Kastner \emph{et al.}, 1999; Buracas \& Boynton, 2007; Silver
+\emph{et al.}, 2007; Li \emph{et al.}, 2008; Murray, 2008) and EEG
+signals (Ding \emph{et al.}, 2006).
+
+Electrophysiological and psychophysical measurements of attentional
+effects on both luminance and chromatic contrast strongly implicated
+gain control an an underlying mechanism (Lu \& Dosher, 1998; Di Russo \&
+Spinelli, 1999\emph{a}; Di Russo \emph{et al.}, 2001\emph{b}) and
+demonstrated that these effects differed between chromatic and
+achromatic pathways (Di Russo \& Spinelli, 1999\emph{b}). In the late
+2000s a comprehensive theoretical model for attentional modulation was
+developed that framed it as a gain control computation (Reynolds \&
+Heeger, 2009; Boynton, 2009). This framework has proven to be
+influential - explaining a wide range of phenomena from the earlier
+literature and demonstrating subtle interactions between the size of the
+attentional `spotlight' and the stimulus configuration which rationalise
+many apparent contradictions in the literature. Subsequent papers
+(effect of stimulus configuration critical). Then\ldots{} talk about
+Lauritzen \emph{et al.} (2010) Verghese \emph{et al.} (2012) (1 other
+Verghese paper) maybe mention Highsmith \& Crognale (2010), Wang \& Wade
+(2011) and Martinovic \& Andersen (2018) Probably lots more Carasco
+\ldots{}
+
 \section{Brain-computer interfaces}\label{brain-computer-interfaces}
+
+One widespread recent application of the SSVEP technique is in the
+design of brain-computer interfaces, which seek to control some aspect
+of a computer using neural signals. The high SNR and frequency
+resolution of SSVEPs make them an ideal candidate for this approach.
+Typical studies may involve presenting an array of stimuli at different
+flicker frequencies, and having the participant select one either by
+overt attention (i.e.~shifting fixation to foveate the selected
+stimulus) or covert attention (i.e.~deploying attention to one stimulus
+whilst keeping fixated). EEG signals are highly sensitive to both visual
+field position (Ales \emph{et al.}, 2010) and attentional state
+(Lauritzen \emph{et al.}, 2010; Verghese \emph{et al.}, 2012), so the
+response to the selected stimulus will typically increase relative to
+the others, allowing it to be identified by an on-line algorithm.
+
+Probably a summary of BCIs here.
 
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
+\section{References}\label{references}
 
@@
 
+\bibitem[\citeproctext]{ref-Ales2010}
+Ales JM, Yates JL \& Norcia AM (2010).
+\href{https://doi.org/10.1016/j.neuroimage.2010.05.016}{V1 is not
+uniquely identified by polarity reversals of responses to upper and
+lower visual field stimuli}. \emph{Neuroimage} \textbf{52,} 1401--1409.
+
 \bibitem[\citeproctext]{ref-Alpern1971}
@@
 
+\bibitem[\citeproctext]{ref-Baker2021}
+Baker DH, Vilidaite G \& Wade AR (2021).
+\href{https://doi.org/10.1371/journal.pcbi.1009507}{Steady-state
+measures of visual suppression}. \emph{PLoS computational biology}
+\textbf{17,} e1009507.
+
 \bibitem[\citeproctext]{ref-Baker2017}
@@
 
+\bibitem[\citeproctext]{ref-Bashinski1980}
+Bashinski HS \& Bacharach VR (1980). Enhancement of perceptual
+sensitivity as the result of selectively attending to spatial locations.
+\emph{Perception \& psychophysics} \textbf{28,} 241--248.
+
 \bibitem[\citeproctext]{ref-Betts2005}
@@
 
+\bibitem[\citeproctext]{ref-Boynton2009}
+Boynton GM (2009). \href{https://doi.org/10.1016/j.visres.2008.11.001}{A
+framework for describing the effects of attention on visual responses}.
+\emph{Vision Res} \textbf{49,} 1129--1143.
+
 \bibitem[\citeproctext]{ref-Braddick1986}
@@
 
+\bibitem[\citeproctext]{ref-Brefczynski1999}
+Brefczynski JA \& DeYoe EA (1999). \href{https://doi.org/10.1038/7280}{A
+physiological correlate of the 'spotlight' of visual attention}.
+\emph{Nat Neurosci} \textbf{2,} 370--374.
+
+\bibitem[\citeproctext]{ref-Buracas2007}
+Buracas GT \& Boynton GM (2007).
+\href{https://doi.org/10.1523/JNEUROSCI.3162-06.2007}{The effect of
+spatial attention on contrast response functions in human visual
+cortex}. \emph{J Neurosci} \textbf{27,} 93--97.
+
+\bibitem[\citeproctext]{ref-Burr1987}
+Burr DC \& Morrone MC (1987).
+\href{https://doi.org/10.1113/jphysiol.1987.sp016643}{Inhibitory
+interactions in the human vision system revealed in pattern-evoked
+potentials}. \emph{J Physiol} \textbf{389,} 1--21.
+
 \bibitem[\citeproctext]{ref-Busse2009}
@@
 
+\bibitem[\citeproctext]{ref-Cameron2002}
+Cameron EL, Tai JC \& Carrasco M (2002).
+\href{https://doi.org/10.1016/s0042-6989(02)00039-1}{Covert attention
+affects the psychometric function of contrast sensitivity}. \emph{Vision
+Res} \textbf{42,} 949--967.
+
 \bibitem[\citeproctext]{ref-Campbell1965}
@@
 \textbf{181,} 576--593.
+
+\bibitem[\citeproctext]{ref-Campbell1970}
+Campbell FW \& Maffei L (1970).
+\href{https://doi.org/10.1113/jphysiol.1970.sp009085}{Electrophysiological
+evidence for the existence of orientation and size detectors in the
+human visual system}. \emph{J Physiol} \textbf{207,} 635--652.
 
@@
 
+\bibitem[\citeproctext]{ref-Carrasco2000}
+Carrasco M, Penpeci-Talgar C \& Eckstein M (2000).
+\href{https://doi.org/10.1016/s0042-6989(00)00024-9}{Spatial covert
+attention increases contrast sensitivity across the CSF: Support for
+signal enhancement}. \emph{Vision Res} \textbf{40,} 1203--1215.
+
 \bibitem[\citeproctext]{ref-Chen2000}
@@
 measurements}. \emph{Vision Res} \textbf{40,} 773--788.
+
+\bibitem[\citeproctext]{ref-Cunningham2017}
+Cunningham DGM, Baker DH \& Peirce JW (2017).
+\href{https://doi.org/10.1167/17.5.10}{{Measuring nonlinear signal
+combination using EEG}}. \emph{Journal of Vision} \textbf{17,} 10--10.
 
@@
 
+\bibitem[\citeproctext]{ref-DiRusso1999}
+Di Russo F \& Spinelli D (1999\emph{a}).
+\href{https://doi.org/10.1016/S0042-6989(99)00031-0}{Electrophysiological
+evidence for an early attentional mechanism in visual processing in
+humans}. \emph{Vision Research} \textbf{39,} 2975--2985.
+
+\bibitem[\citeproctext]{ref-DiRusso1999b}
+Di Russo F \& Spinelli D (1999\emph{b}). Spatial attention has different
+effects on the magno-and parvocellular pathways. \emph{Neuroreport}
+\textbf{10,} 2755--2762.
+
+\bibitem[\citeproctext]{ref-DiRusso2001}
+Di Russo F, Spinelli D \& Morrone MC (2001\emph{b}).
+\href{https://doi.org/10.1016/S0042-6989(01)00134-1}{Automatic gain
+control contrast mechanisms are modulated by attention in humans:
+Evidence from visual evoked potentials}. \emph{Vision Research}
+\textbf{41,} 2435--2447.
+
 \bibitem[\citeproctext]{ref-Di-Russo2001}
-Di Russo F, Spinelli D \& Morrone MC (2001).
+Di Russo F, Spinelli D \& Morrone MC (2001\emph{a}).
 \href{https://doi.org/10.1016/s0042-6989(01)00134-1}{Automatic gain
@@
 2435--2447.
+
+\bibitem[\citeproctext]{ref-Ding2006}
+Ding J, Sperling G \& Srinivasan R (2006). Attentional modulation of
+{SSVEP} power depends on the network tagged by the flicker frequency.
+\emph{Cerebral cortex} \textbf{16,} 1016--1029.
 
@@
 
+\bibitem[\citeproctext]{ref-Gandhi1999}
+Gandhi SP, Heeger DJ \& Boynton GM (1999).
+\href{https://doi.org/10.1073/pnas.96.6.3314}{Spatial attention affects
+brain activity in human primary visual cortex}. \emph{Proc Natl Acad Sci
+U S A} \textbf{96,} 3314--3319.
+
+\bibitem[\citeproctext]{ref-Gould2007}
+Gould IC, Wolfgang BJ \& Smith PL (2007).
+\href{https://doi.org/10.1167/7.13.4}{Spatial uncertainty explains
+exogenous and endogenous attentional cuing effects in visual signal
+detection}. \emph{J Vis} \textbf{7,} 4.1--17.
+
+\bibitem[\citeproctext]{ref-Groen2022}
+Groen IIA, Piantoni G, Montenegro S, Flinker A, Devore S, Devinsky O,
+Doyle W, Dugan P, Friedman D, Ramsey NF, Petridou N \& Winawer J (2022).
+\href{https://doi.org/10.1523/JNEUROSCI.1812-21.2022}{Temporal dynamics
+of neural responses in human visual cortex}. \emph{J Neurosci}
+\textbf{42,} 7562--7580.
+
 \bibitem[\citeproctext]{ref-Hamer1989}
@@
 
+\bibitem[\citeproctext]{ref-Highsmith2010}
+Highsmith J \& Crognale MA (2010).
+\href{https://doi.org/10.1111/j.1475-1313.2010.00747.x}{Attentional
+shifts have little effect on the waveform of the chromatic onset VEP}.
+\emph{Ophthalmic Physiol Opt} \textbf{30,} 525--533.
+
 \bibitem[\citeproctext]{ref-Hoffmann2017}
@@
 
+\bibitem[\citeproctext]{ref-Hubel1962}
+Hubel DH \& Wiesel TN (1962).
+\href{https://doi.org/10.1113/jphysiol.1962.sp006837}{Receptive fields,
+binocular interaction and functional architecture in the cat's visual
+cortex}. \emph{J Physiol} \textbf{160,} 106--154.
+
+\bibitem[\citeproctext]{ref-Kastner1999}
+Kastner S, Pinsk MA, De Weerd P, Desimone R \& Ungerleider LG (1999).
+\href{https://doi.org/10.1016/s0896-6273(00)80734-5}{Increased activity
+in human visual cortex during directed attention in the absence of
+visual stimulation}. \emph{Neuron} \textbf{22,} 751--761.
+
 \bibitem[\citeproctext]{ref-Kuffler1957}
@@
 683--702.
+
+\bibitem[\citeproctext]{ref-Lauritzen2010}
+Lauritzen TZ, Ales JM \& Wade AR (2010).
+\href{https://doi.org/10.1167/10.14.39}{The effects of visuospatial
+attention measured across visual cortex using source-imaged,
+steady-state {EEG}}. \emph{Journal of Vision} \textbf{10,} 39.
 
@@
 
+\bibitem[\citeproctext]{ref-Li2008}
+Li X, Lu Z-L, Tjan BS, Dosher BA \& Chu W (2008).
+\href{https://doi.org/10.1073/pnas.0801390105}{Blood oxygenation
+level-dependent contrast response functions identify mechanisms of
+covert attention in early visual areas}. \emph{Proc Natl Acad Sci U S A}
+\textbf{105,} 6202--6207.
+
+\bibitem[\citeproctext]{ref-Lu1998}
+Lu Z-L \& Dosher BA (1998). External noise distinguishes attention
+mechanisms. \emph{Vision research} \textbf{38,} 1183--1198.
+
+\bibitem[\citeproctext]{ref-Luck1997}
+Luck SJ, Chelazzi L, Hillyard SA \& Desimone R (1997).
+\href{https://doi.org/10.1152/jn.1997.77.1.24}{Neural mechanisms of
+spatial selective attention in areas V1, V2, and V4 of macaque visual
+cortex}. \emph{J Neurophysiol} \textbf{77,} 24--42.
+
 \bibitem[\citeproctext]{ref-Lygo2021}
@@
 
+\bibitem[\citeproctext]{ref-Marcus2002}
+Marcus DS \& Van Essen DC (2002).
+\href{https://doi.org/10.1152/jn.00916.2001}{Scene segmentation and
+attention in primate cortical areas V1 and V2}. \emph{J Neurophysiol}
+\textbf{88,} 2648--2658.
+
+\bibitem[\citeproctext]{ref-Martinovic2018}
+Martinovic J \& Andersen SK (2018).
+\href{https://doi.org/10.1016/j.neuroimage.2018.04.066}{Cortical
+summation and attentional modulation of combined chromatic and luminance
+signals}. \emph{Neuroimage} \textbf{176,} 390--403.
+
+\bibitem[\citeproctext]{ref-McAdams1999}
+McAdams CJ \& Maunsell JH (1999).
+\href{https://doi.org/10.1523/JNEUROSCI.19-01-00431.1999}{Effects of
+attention on orientation-tuning functions of single neurons in macaque
+cortical area V4}. \emph{J Neurosci} \textbf{19,} 431--441.
+
 \bibitem[\citeproctext]{ref-McKeefry1996}
@@
 
+\bibitem[\citeproctext]{ref-Mecacci1976}
+Mecacci L \& Spinelli D (1976).
+\href{https://doi.org/10.1016/0042-6989(76)90025-0}{The effects of
+spatial frequency adaptation on human evoked potentials}. \emph{Vision
+Research} \textbf{16,} 477--479.
+
 \bibitem[\citeproctext]{ref-Meese2009}
@@
 \emph{J Vis} \textbf{9,} 2.1--15.
+
+\bibitem[\citeproctext]{ref-Mehta2000a}
+Mehta AD, Ulbert I \& Schroeder CE (2000\emph{a}).
+\href{https://doi.org/10.1093/cercor/10.4.343}{Intermodal selective
+attention in monkeys. I: Distribution and timing of effects across
+visual areas}. \emph{Cereb Cortex} \textbf{10,} 343--358.
+
+\bibitem[\citeproctext]{ref-Mehta2000b}
+Mehta AD, Ulbert I \& Schroeder CE (2000\emph{b}).
+\href{https://doi.org/10.1093/cercor/10.4.359}{Intermodal selective
+attention in monkeys. II: Physiological mechanisms of modulation}.
+\emph{Cereb Cortex} \textbf{10,} 359--370.
 
@@
 
+\bibitem[\citeproctext]{ref-Morrone2004}
+Morrone MC, Denti V \& Spinelli D (2004).
+\href{https://doi.org/10.1016/j.visres.2003.10.014}{Different
+attentional resources modulate the gain mechanisms for color and
+luminance contrast}. \emph{Vision Res} \textbf{44,} 1389--1401.
+
 \bibitem[\citeproctext]{ref-Movshon1975}
@@
 \textbf{249,} 445--468.
+
+\bibitem[\citeproctext]{ref-Murray2008}
+Murray SO (2008). \href{https://doi.org/10.1167/8.10.2}{The effects of
+spatial attention in early human visual cortex are stimulus
+independent}. \emph{J Vis} \textbf{8,} 2.1--11.
 
@@
 
+\bibitem[\citeproctext]{ref-Pestilli2009}
+Pestilli F, Ling S \& Carrasco M (2009).
+\href{https://doi.org/10.1016/j.visres.2008.09.018}{A population-coding
+model of attention's influence on contrast response: Estimating neural
+effects from psychophysical data}. \emph{Vision Res} \textbf{49,}
+1144--1153.
+
 \bibitem[\citeproctext]{ref-Petrov2005}
@@
 
+\bibitem[\citeproctext]{ref-Petrov2006}
+Petrov Y, Verghese P \& McKee SP (2006).
+\href{https://doi.org/10.1167/6.2.8}{Collinear facilitation is largely
+uncertainty reduction}. \emph{J Vis} \textbf{6,} 170--178.
+
 \bibitem[\citeproctext]{ref-Polat1993}
@@
 
+\bibitem[\citeproctext]{ref-Posner1980}
+Posner MI (1980).
+\href{https://doi.org/10.1080/00335558008248231}{Orienting of
+attention}. \emph{Q J Exp Psychol} \textbf{32,} 3--25.
+
 \bibitem[\citeproctext]{ref-Regan1966}
@@
 
+\bibitem[\citeproctext]{ref-Reynolds2009}
+Reynolds JH \& Heeger DJ (2009).
+\href{https://doi.org/10.1016/j.neuron.2009.01.002}{The normalization
+model of attention}. \emph{Neuron} \textbf{61,} 168--185.
+
 \bibitem[\citeproctext]{ref-Ridder2004}
@@
 
+\bibitem[\citeproctext]{ref-Ross1991}
+Ross J \& Speed HD (1991).
+\href{https://doi.org/10.1098/rspb.1991.0125}{Contrast adaptation and
+contrast masking in human vision}. \emph{Proc Biol Sci} \textbf{246,}
+61--69.
+
+\bibitem[\citeproctext]{ref-Ross1993}
+Ross J, Speed HD \& Morgan MJ (1993).
+\href{https://doi.org/10.1016/0042-6989(93)90003-f}{The effects of
+adaptation and masking on incremental thresholds for contrast}.
+\emph{Vision Res} \textbf{33,} 2051--2056.
+
+\bibitem[\citeproctext]{ref-Salelkar2020}
+Salelkar S \& Ray S (2020).
+\href{https://doi.org/10.1038/s41598-020-62180-y}{Interaction between
+steady-state visually evoked potentials at nearby flicker frequencies}.
+\emph{Sci Rep} \textbf{10,} 5344.
+
 \bibitem[\citeproctext]{ref-Shapley1980}
@@
 
+\bibitem[\citeproctext]{ref-Silver2007}
+Silver MA, Ress D \& Heeger DJ (2007).
+\href{https://doi.org/10.1152/jn.00677.2006}{Neural correlates of
+sustained spatial attention in human early visual cortex}. \emph{J
+Neurophysiol} \textbf{97,} 229--237.
+
 \bibitem[\citeproctext]{ref-Simmons2009}
@@
 visual sensitivity}. \emph{Nature} \textbf{391,} 697--700.
+
+\bibitem[\citeproctext]{ref-Somers1999}
+Somers DC, Dale AM, Seiffert AE \& Tootell RB (1999).
+\href{https://doi.org/10.1073/pnas.96.4.1663}{Functional MRI reveals
+spatially specific attentional modulation in human primary visual
+cortex}. \emph{Proc Natl Acad Sci U S A} \textbf{96,} 1663--1668.
 
@@
 contrast}. \emph{Exp Brain Res} \textbf{41,} 414--419.
+
+\bibitem[\citeproctext]{ref-Tootell1998}
+Tootell RB, Hadjikhani N, Hall EK, Marrett S, Vanduffel W, Vaughan JT \&
+Dale AM (1998). \href{https://doi.org/10.1016/s0896-6273(00)80659-5}{The
+retinotopy of visual spatial attention}. \emph{Neuron} \textbf{21,}
+1409--1422.
 
@@
 
+\bibitem[\citeproctext]{ref-Vergeer2018}
+Vergeer M, Mesik J, Baek Y, Wilmerding K \& Engel SA (2018).
+\href{https://doi.org/10.1167/18.5.2}{Orientation-selective contrast
+adaptation measured with {SSVEP}}. \emph{Journal of Vision} \textbf{18,}
+2.
+
+\bibitem[\citeproctext]{ref-Verghese2012}
+Verghese P, Kim Y-J \& Wade AR (2012).
+\href{https://doi.org/10.1523/JNEUROSCI.1174-12.2012}{Attention selects
+informative neural populations in human {V1}}. \emph{The Journal of
+Neuroscience: The Official Journal of the Society for Neuroscience}
+\textbf{32,} 16379--16390.
+
 \bibitem[\citeproctext]{ref-Vilidaite2018}
@@
 Biological Sciences} \textbf{285,} 20182255.
+
+\bibitem[\citeproctext]{ref-Wang2011}
+Wang J \& Wade AR (2011).
+\href{https://doi.org/10.1167/11.6.1}{Differential attentional
+modulation of cortical responses to {S}-cone and luminance stimuli}.
+\emph{Journal of Vision} \textbf{11,} 1.
 

### 2024-05-05 review.qmd
References update

@@
 Electrophysiological and psychophysical measurements of the effect of attention on both luminance and chromatic contrast have strongly implicated gain control as an underlying mechanism [@DiRusso1999; @DiRusso2001; @Lu1998].
-These effects also appear to differ between chromatic and achromatic pathways [@DiRusso1999b] - perhaps as a result of the different levels of nonlinear gain control in the early precotical magno,  parvo and konio-cellular pathways [DerringtonLennie84, KaplanShapley86, LeePokornyJOSA90, SolomonLennie2005].
+These effects also appear to differ between chromatic and achromatic pathways [@DiRusso1999b] - perhaps as a result of the different levels of nonlinear gain control in the early pre-cortical magno-, parvo- and konio-cellular pathways [@Derrington1984b; @Kaplan1986; @Lee1990; @Solomon2005].
 In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control computation [@Reynolds2009; @Boynton2009].
@@
 
-Probably a summary of BCIs here.
+Probably a summary of BCIs here @Zhu2010
 

### 2024-05-06 review.qmd
More BCI

@@
 Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates [@Luck1997; @Marcus2002; @McAdams1999; @Mehta2000a; @Mehta2000b], but with the advent of spatially-resolved human brain imaging methods in the late 1990s it became apparent that spatial attention was linked to frank changes in both fMRI BOLD responses
-[@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006].
+[@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006, @Muller1998m @MullerHillyard2000, MorganPNAS1996].
 
 Electrophysiological and psychophysical measurements of the effect of attention on both luminance and chromatic contrast have strongly implicated gain control as an underlying mechanism [@DiRusso1999; @DiRusso2001; @Lu1998].
-These effects also appear to differ between chromatic and achromatic pathways [@DiRusso1999b] - perhaps as a result of the different levels of nonlinear gain control in the early pre-cortical magno-, parvo- and konio-cellular pathways [@Derrington1984b; @Kaplan1986; @Lee1990; @Solomon2005].
+These effects also appear to differ between chromatic and achromatic pathways [@DiRusso1999b] - perhaps as a result of the different levels of nonlinear gain control in the early precotical magno-,  parvo- and konio-cellular pathways [DerringtonLennie84, KaplanShapley86, LeePokornyJOSA90, SolomonLennie2005].
 In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control computation [@Reynolds2009; @Boynton2009].
-This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating subtle interactions between the size of the attentional 'spotlight' and the stimulus configuration which rationalise many apparent contradictions in the literature.
-Direct measurements of attentional modulation of achromatic SSVEP signals are broadly consistent with this model [@Lauritzen2010, @Hou2016, @Martinovic2018] and confirm the relatively weaker role of spatial attention on responses driven by chromatic stimuli - particularly those that isolate the opponent S-(L+M) cone pathway [@Wang2011, @Highsmith2010].
-The gain control model of attention can also be extended to SSVEP studies of feature-based attention which show that the modulatory effects can be targeted to the most informative neuronal populations [@Verghese2012].
+This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating subtle interactions between the size of the attentional 'spotlight' and the stimulus configuration which rationalised many apparent contradictions in the literature.
+Direct measurements of attentional modulation of contrast-driven SSVEP signals are broadly consistent with this model [@Lauritzen2010, @Hou2016, @Martinovic2018] and confirm the relatively weaker role of spatial attention on responses driven by chromatic stimuli - particularly those that isolate the opponent S-(L+M) cone pathway [@Wang2011, @Highsmith2010].
+The gain control model of attention can also be extended to SSVEP studies of feature-based attention which show that the modulatory effects can be targeted selectively to the most informative neuronal populations [@Verghese2012,@Ding2006].
 
@@
 The high SNR and frequency resolution of SSVEPs make them an ideal candidate for this approach.
-Typical studies may involve presenting an array of stimuli at different flicker frequencies, and having the participant select one either by overt attention (i.e.
+Typical studies may involve presenting an array of stimuli at different contrast flicker frequencies, and having the participant select one either by overt attention (i.e.
 shifting fixation to foveate the selected stimulus) or covert attention (i.e.
-deploying attention to one stimulus whilst keeping fixated).
+deploying attention to one stimulus whilst keeping fixated) [Middendorf2000, ].
 EEG signals are highly sensitive to both visual field position [@Ales2010] and attentional state [@Verghese2012; @Lauritzen2010], so the response to the selected stimulus will typically increase relative to the others, allowing it to be identified by an on-line algorithm.
-
-Probably a summary of BCIs here @Zhu2010
+This approach is primarily useful in situations that require the BCI to distinguish from among a small set of possibilties : For example, in early work, visual stimuli representing 'Left' and 'Right' commands in a flight simulator were distinguished robustly [@Middendorf].
+Although SSVEP-based BCI interfaces typically do use contrast flicker, work in this field has largely focused on optimising the stimuli to increase SNR or the range of stimuli that can be frequency-tagged (e.g.
+...
+) and reducing visual fatigue associated with the long-term presentation of arrays of high-contrast flicker (e.g.
+.....).
 

### 2024-05-06 review.qmd

--- 2024-05-05 review.qmd
@@
 Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates [@Luck1997; @Marcus2002; @McAdams1999; @Mehta2000a; @Mehta2000b], but with the advent of spatially-resolved human brain imaging methods in the late 1990s it became apparent that spatial attention was linked to frank changes in both fMRI BOLD responses
-[@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006].
+[@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006, @Muller1998m @MullerHillyard2000, MorganPNAS1996].
 
@@
 The high SNR and frequency resolution of SSVEPs make them an ideal candidate for this approach.
-Typical studies may involve presenting an array of stimuli at different flicker frequencies, and having the participant select one either by overt attention (i.e.
+Typical studies may involve presenting an array of stimuli at different contrast flicker frequencies, and having the participant select one either by overt attention (i.e.
 shifting fixation to foveate the selected stimulus) or covert attention (i.e.
-deploying attention to one stimulus whilst keeping fixated).
+deploying attention to one stimulus whilst keeping fixated) [Middendorf2000, ].
 EEG signals are highly sensitive to both visual field position [@Ales2010] and attentional state [@Verghese2012; @Lauritzen2010], so the response to the selected stimulus will typically increase relative to the others, allowing it to be identified by an on-line algorithm.
-
-Probably a summary of BCIs here @Zhu2010
+This approach is primarily useful in situations that require the BCI to distinguish from among a small set of possibilties : For example, in early work, visual stimuli representing 'Left' and 'Right' commands in a flight simulator were distinguished robustly [@Middendorf].
+Although SSVEP-based BCI interfaces typically do use contrast flicker, work in this field has largely focused on optimising the stimuli to increase SNR or the range of stimuli that can be frequency-tagged (e.g.
+...
+) and reducing visual fatigue associated with the long-term presentation of arrays of high-contrast flicker (e.g.
+.....).
+
+Probably a summary of BCIs here.
 

### 2024-05-06 review.qmd
More

@@
 Presumably, the signal measured from early visual cortex is the result of a cascade of nonlinear retinal and cortical operations up to that point.
-It therefore contains a 'signature' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
+It therefore contains a 'signature' or 'fingerprint' of the computational nature, order and parameters of those operations - including the shape of the transducer functions and the computations involved in signal combination.
 In principle, that information could be recovered from the SSVEP signal - a possibility recognised in the early days of the technique [@Regan1988].
@@
 
-# Attention
+# Attention and arousal
 
@@
 Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates [@Luck1997; @Marcus2002; @McAdams1999; @Mehta2000a; @Mehta2000b], but with the advent of spatially-resolved human brain imaging methods in the late 1990s it became apparent that spatial attention was linked to frank changes in both fMRI BOLD responses
-[@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006, @Muller1998m @MullerHillyard2000, MorganPNAS1996].
+[@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006, @Muller1998m @MullerHillyard2000, @MorganPNAS1996, @Hillyard1998NNS].
 
@@
 
+The SSVEP can also be used to study changes in visual processing \textit{driven} by different behavioural states or overall arousal.
+For example, locomotion has been shown to alter neuronal excitability and spatial normalization in mice [@NeilStryker2010, @Ayaz2013] - running mice have higher visual sensitivity and lower surround suppression compared to stationary mice.
+Although measuring EEG responses from locomoting humans is technically challenging, SSVEP studies (which are able to distinguish broadband noise from input signal effectively) have shown that walking also alters early visual processing although in a manner different to that observed in mice [@Benjamin2018, @CaoPlosBiol2019]
+
+SSVEP measurements are typically used to measure time invariant responses due to sustained attention.
+However, recent work has shown that moderately high modulation frequencies (ca 10Hz) and short analysis windows (1.5s) can also be used to track the dyanmic allocation of attention across a task [@ChotaJCN2024] or attention to moving targets [@deLissaPlos12020].
+The ability to track changes in attention over short time periods is also important if SSVEP is to be used for dynamic readouts - for example in a Brain Computer Interface.
+
 # Brain-computer interfaces
@@
 shifting fixation to foveate the selected stimulus) or covert attention (i.e.
-deploying attention to one stimulus whilst keeping fixated) [Middendorf2000, ].
-EEG signals are highly sensitive to both visual field position [@Ales2010] and attentional state [@Verghese2012; @Lauritzen2010], so the response to the selected stimulus will typically increase relative to the others, allowing it to be identified by an on-line algorithm.
+deploying attention to one stimulus whilst keeping fixated) [@Middendorf].
+BEcause SSVEP signals are highly sensitive to both visual field position [@Ales2010, @dirusso2007_HBM_spatiotemporal] and attentional state [@hillyard1996_PNAS, @Hillard1998_NNS, @Verghese2012; @Lauritzen2010],  the response to the selected stimulus will typically increase relative to the others, allowing it to be identified by an on-line algorithm.
+Because more than one stimulus frequency is typically present, this modulation and the associated changes in gain control, will affect the entire pattern of self- and intermodulation terms allowing the choice to be decoded by a multivariate pattern classifier.
+
 This approach is primarily useful in situations that require the BCI to distinguish from among a small set of possibilties : For example, in early work, visual stimuli representing 'Left' and 'Right' commands in a flight simulator were distinguished robustly [@Middendorf].
-Although SSVEP-based BCI interfaces typically do use contrast flicker, work in this field has largely focused on optimising the stimuli to increase SNR or the range of stimuli that can be frequency-tagged (e.g.
+Although SSVEP-based BCI interfaces typically do use contrast flicker, work in this field has largely focused on optimising the stimuli or decoders to increase decoding performance (e.g.
 ...
 ) and reducing visual fatigue associated with the long-term presentation of arrays of high-contrast flicker (e.g.
-.....).
-
-Probably a summary of BCIs here.
+.....) rather than studying visual contrast processing \textit{per se}.
+We therefore note in passing that using SSVEPs to improve our understanding of early contrast processing may yield benefits to this related field.
 
 # Conclusions
+The SSVEP is a powerful tool for studying contrast processing.
+It provides a high SNR readout of neuronal activity that is unambigously linked to the input.
+It is sensitive to both the amplitude and phase of the input and when combined with source imaging, it can be extracted from different cortical regions allowing researchers to track contrast processing computations across the visual pathway.
+Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants) and at frequencies high enough to be effectively invisible [REF - Some high frequency work...
+- Romanian group].
+The SSVEP is also able 'fingerprint' modulators of the input frequency through the harmonic and intermodulation terms they generate in the output.
+In principle, each nonlinearity in the visual pathway can be identified by
 

### 2024-05-06 review.tex
More

@@
 Attention exerts a profound influence on visual performance: for
-example, allowing people to attend covertly to a spatial location for
-example, improves their performance on a target detection task
-significantly (Bashinski \& Bacharach, 1980; Posner, 1980; Carrasco
-\emph{et al.}, 2000; Cameron \emph{et al.}, 2002; Morrone \emph{et al.},
-2004; Pestilli \emph{et al.}, 2009). In principle, this enhancement may
-be driven both by modulation of the underlying signal or noise
-characteristics, or by additional decision-theoretic factors such as
-reduction in spatial uncertainty (Petrov \emph{et al.}, 2006; Gould
-\emph{et al.}, 2007). Early experiments with non-human primates showed
-little evidence for attentionally-driven changes in neuronal spike rates
-(Luck \emph{et al.}, 1997; McAdams \& Maunsell, 1999; Mehta \emph{et
-al.}, 2000\emph{a}, 2000\emph{b}; Marcus \& Van Essen, 2002), but with
-the advent of spatially-resolved human brain imaging methods in the late
+example, allowing people to attend covertly to a spatial location
+improves their performance on a target detection task significantly
+(Bashinski \& Bacharach, 1980; Posner, 1980; Carrasco \emph{et al.},
+2000; Cameron \emph{et al.}, 2002; Morrone \emph{et al.}, 2004; Pestilli
+\emph{et al.}, 2009). In principle, this enhancement may be driven both
+by modulation of the underlying signal or noise characteristics, or by
+additional decision-theoretic factors such as reduction in spatial
+uncertainty (Petrov \emph{et al.}, 2006; Gould \emph{et al.}, 2007).
+Early experiments with non-human primates showed little evidence for
+attentionally-driven changes in neuronal spike rates (Luck \emph{et
+al.}, 1997; McAdams \& Maunsell, 1999; Mehta \emph{et al.},
+2000\emph{a}, 2000\emph{b}; Marcus \& Van Essen, 2002), but with the
+advent of spatially-resolved human brain imaging methods in the late
 1990s it became apparent that spatial attention was linked to frank
@@
 \emph{et al.}, 2007; Li \emph{et al.}, 2008; Murray, 2008) and EEG
-signals (Ding \emph{et al.}, 2006).
-
-Electrophysiological and psychophysical measurements of attentional
-effects on both luminance and chromatic contrast strongly implicated
-gain control an an underlying mechanism (Lu \& Dosher, 1998; Di Russo \&
-Spinelli, 1999\emph{a}; Di Russo \emph{et al.}, 2001\emph{b}) and
-demonstrated that these effects differed between chromatic and
-achromatic pathways (Di Russo \& Spinelli, 1999\emph{b}). In the late
-2000s a comprehensive theoretical model for attentional modulation was
-developed that framed it as a gain control computation (Reynolds \&
-Heeger, 2009; Boynton, 2009). This framework has proven to be
-influential - explaining a wide range of phenomena from the earlier
-literature and demonstrating subtle interactions between the size of the
-attentional `spotlight' and the stimulus configuration which rationalise
-many apparent contradictions in the literature. Subsequent papers
-(effect of stimulus configuration critical). Then\ldots{} talk about
-Lauritzen \emph{et al.} (2010) Verghese \emph{et al.} (2012) (1 other
-Verghese paper) maybe mention Highsmith \& Crognale (2010), Wang \& Wade
-(2011) and Martinovic \& Andersen (2018) Probably lots more Carasco
-\ldots{}
+signals (\textbf{MullerHillyard2000?}).
+
+Electrophysiological and psychophysical measurements of the effect of
+attention on both luminance and chromatic contrast have strongly
+implicated gain control as an underlying mechanism (Lu \& Dosher, 1998;
+Di Russo \& Spinelli, 1999\emph{a}; Di Russo \emph{et al.},
+2001\emph{b}). These effects also appear to differ between chromatic and
+achromatic pathways (Di Russo \& Spinelli, 1999\emph{b}) - perhaps as a
+result of the different levels of nonlinear gain control in the early
+pre-cortical magno-, parvo- and konio-cellular pathways (Derrington \&
+Lennie, 1984; Kaplan \& Shapley, 1986; Lee \emph{et al.}, 1990; Solomon
+\& Lennie, 2005). In the late 2000s a comprehensive theoretical model
+for attentional modulation was developed that framed it as a gain
+control computation (Reynolds \& Heeger, 2009; Boynton, 2009). This
+framework has proven to be influential - explaining a wide range of
+phenomena from the earlier literature and demonstrating subtle
+interactions between the size of the attentional `spotlight' and the
+stimulus configuration which rationalise many apparent contradictions in
+the literature. Direct measurements of attentional modulation of
+achromatic SSVEP signals are broadly consistent with this model
+Martinovic \& Andersen (2018) and confirm the relatively weaker role of
+spatial attention on responses driven by chromatic stimuli -
+particularly those that isolate the opponent S-(L+M) cone pathway
+Highsmith \& Crognale (2010). The gain control model of attention can
+also be extended to SSVEP studies of feature-based attention which show
+that the modulatory effects can be targeted to the most informative
+neuronal populations (Verghese \emph{et al.}, 2012).
 
@@
 Typical studies may involve presenting an array of stimuli at different
-flicker frequencies, and having the participant select one either by
-overt attention (i.e.~shifting fixation to foveate the selected
-stimulus) or covert attention (i.e.~deploying attention to one stimulus
-whilst keeping fixated). EEG signals are highly sensitive to both visual
-field position (Ales \emph{et al.}, 2010) and attentional state
-(Lauritzen \emph{et al.}, 2010; Verghese \emph{et al.}, 2012), so the
-response to the selected stimulus will typically increase relative to
-the others, allowing it to be identified by an on-line algorithm.
-
-Probably a summary of BCIs here.
+contrast flicker frequencies, and having the participant select one
+either by overt attention (i.e.~shifting fixation to foveate the
+selected stimulus) or covert attention (i.e.~deploying attention to one
+stimulus whilst keeping fixated) (\textbf{Middendorf?}). BEcause EEG
+signals are highly sensitive to both visual field position (Ales
+\emph{et al.}, 2010) and attentional state (Lauritzen \emph{et al.},
+2010; Verghese \emph{et al.}, 2012), the response to the selected
+stimulus will typically increase relative to the others, allowing it to
+be identified by an on-line algorithm. Because more than one stimulus
+frequency is typically present, this modulation and the associated
+changes in gain control, will affect the entire pattern of self- and
+intermodulation terms.
+
+This approach is primarily useful in situations that require the BCI to
+distinguish from among a small set of possibilties : For example, in
+early work, visual stimuli representing `Left' and `Right' commands in a
+flight simulator were distinguished robustly (\textbf{Middendorf?}).
+Although SSVEP-based BCI interfaces typically do use contrast flicker,
+work in this field has largely focused on optimising the stimuli or
+decoders to increase decoding performance (e.g.~\ldots{} ) and reducing
+visual fatigue associated with the long-term presentation of arrays of
+high-contrast flicker (e.g.~\ldots..) rather than studying visual
+contrast processing \textit{per se}.
 
@@
 
+\bibitem[\citeproctext]{ref-Derrington1984b}
+Derrington AM \& Lennie P (1984).
+\href{https://doi.org/10.1113/jphysiol.1984.sp015498}{Spatial and
+temporal contrast sensitivities of neurones in lateral geniculate
+nucleus of macaque}. \emph{J Physiol} \textbf{357,} 219--240.
+
 \bibitem[\citeproctext]{ref-Di-Russo2007}
@@
 
+\bibitem[\citeproctext]{ref-Hou2016}
+Hou C, Kim Y-J, Lai XJ \& Verghese P (2016).
+\href{https://doi.org/10.1167/16.3.16}{Degraded attentional modulation
+of cortical neural populations in strabismic amblyopia}. \emph{Journal
+of Vision} \textbf{16,} 16.
+
 \bibitem[\citeproctext]{ref-Hu2023}
@@
 
+\bibitem[\citeproctext]{ref-Kaplan1986}
+Kaplan E \& Shapley RM (1986).
+\href{https://doi.org/10.1073/pnas.83.8.2755}{The primate retina
+contains two types of ganglion cells, with high and low contrast
+sensitivity}. \emph{Proc Natl Acad Sci U S A} \textbf{83,} 2755--2757.
+
 \bibitem[\citeproctext]{ref-Kastner1999}
@@
 steady-state {EEG}}. \emph{Journal of Vision} \textbf{10,} 39.
+
+\bibitem[\citeproctext]{ref-Lee1990}
+Lee BB, Pokorny J, Smith VC, Martin PR \& Valberg A (1990).
+\href{https://doi.org/10.1364/josaa.7.002223}{Luminance and chromatic
+modulation sensitivity of macaque ganglion cells and human observers}.
+\emph{J Opt Soc Am A} \textbf{7,} 2223--2236.
 
@@
 
+\bibitem[\citeproctext]{ref-Solomon2005}
+Solomon SG \& Lennie P (2005).
+\href{https://doi.org/10.1523/JNEUROSCI.5316-04.2005}{Chromatic gain
+controls in visual cortical neurons}. \emph{J Neurosci} \textbf{25,}
+4779--4792.
+
 \bibitem[\citeproctext]{ref-Somers1999}

### 2024-05-06 review.qmd
Reference update

@@
 The SSVEP can also be used to study changes in visual processing \textit{driven} by different behavioural states or overall arousal.
-For example, locomotion has been shown to alter neuronal excitability and spatial normalization in mice [@NeilStryker2010, @Ayaz2013] - running mice have higher visual sensitivity and lower surround suppression compared to stationary mice.
-Although measuring EEG responses from locomoting humans is technically challenging, SSVEP studies (which are able to distinguish broadband noise from input signal effectively) have shown that walking also alters early visual processing although in a manner different to that observed in mice [@Benjamin2018, @CaoPlosBiol2019]
+For example, locomotion has been shown to alter neuronal excitability and spatial normalization in mice [@Niell2010; @Ayaz2013] - running mice have higher visual sensitivity and lower surround suppression compared to stationary mice.
+Although measuring EEG responses from locomoting humans is technically challenging, SSVEP studies (which are able to distinguish broadband noise from input signal effectively) have shown that walking also alters early visual processing although in a manner different to that observed in mice [@Benjamin2018; @Cao2019].
 
 SSVEP measurements are typically used to measure time invariant responses due to sustained attention.
-However, recent work has shown that moderately high modulation frequencies (ca 10Hz) and short analysis windows (1.5s) can also be used to track the dyanmic allocation of attention across a task [@ChotaJCN2024] or attention to moving targets [@deLissaPlos12020].
+However, recent work has shown that moderately high modulation frequencies (ca 10Hz) and short analysis windows (1.5s) can also be used to track the dynamic allocation of attention across a task [@Chota2024] or attention to moving targets [@Lissa2020].
 The ability to track changes in attention over short time periods is also important if SSVEP is to be used for dynamic readouts - for example in a Brain Computer Interface.
@@
 shifting fixation to foveate the selected stimulus) or covert attention (i.e.
-deploying attention to one stimulus whilst keeping fixated) [@Middendorf].
-BEcause SSVEP signals are highly sensitive to both visual field position [@Ales2010, @dirusso2007_HBM_spatiotemporal] and attentional state [@hillyard1996_PNAS, @Hillard1998_NNS, @Verghese2012; @Lauritzen2010],  the response to the selected stimulus will typically increase relative to the others, allowing it to be identified by an on-line algorithm.
+deploying attention to one stimulus whilst keeping fixated) [@Middendorf2000].
+Because SSVEP signals are highly sensitive to both visual field position [@Ales2010; @Di-Russo2007] and attentional state [@Morgan1996; @Muller1998; @Verghese2012; @Lauritzen2010],  the response to the selected stimulus will typically increase relative to the others, allowing it to be identified by an on-line algorithm.
 Because more than one stimulus frequency is typically present, this modulation and the associated changes in gain control, will affect the entire pattern of self- and intermodulation terms allowing the choice to be decoded by a multivariate pattern classifier.
 
-This approach is primarily useful in situations that require the BCI to distinguish from among a small set of possibilties : For example, in early work, visual stimuli representing 'Left' and 'Right' commands in a flight simulator were distinguished robustly [@Middendorf].
+This approach is primarily useful in situations that require the BCI to distinguish from among a small set of possibilties : For example, in early work, visual stimuli representing 'Left' and 'Right' commands in a flight simulator were distinguished robustly [@Middendorf2000].
 Although SSVEP-based BCI interfaces typically do use contrast flicker, work in this field has largely focused on optimising the stimuli or decoders to increase decoding performance (e.g.
@@
 
-*2: Sweeps and CRFs. Measurements of contrast sensitivity, extrapolating
-the sweep to zero response to get t'hold. Infants and adults?*
-
-*3: Measurements of modulation. Figures adapted from other papers:
-Attention to space, attention to features, adaptation (?).
-masking/surround suppression\\*
-
-*4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*
-
-*5: Future directions (decoding in frequency domain?, animals? BCI?)*
-
-[]{#anchor-2}History
-
-[]{#anchor-3}1. The basics of SSVEP and contrast sensitivity including a
-history of both fields
-
-SSVEP (Steady-State Visual Evoked Potential): A continuous electrical
-response evoked in the brain by visual stimuli flickering at a constant
-frequency (Regan, 1966).
-
-Contrast Sensitivity: The ability to detect differences in luminance
-between an object and its background (Campbell \& Green, 1965).
-
-Regan, D. (1966). Some characteristics of average steady-state and
-transient responses evoked by modulated light. Electroencephalography
-and clinical neurophysiology, 20(3), 238-248.
-
-Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
-affecting visual resolution. J Physiology, 181, 576-593.
-
-Check Regan paper for earlier (e.g. EEG refs). Norcia review will be
-helpful!
-
-(From Tyler / Levi / Apkarian paper):
-
-6\. Regan, D.: Rapid methods for refracting the eye and assessing the
-visual acuity in amblyopia using steady-state visual evoked potentials.
-In Desmedt, J.E., editor: Visual Evoked Potentials in Man: New
-Developments, Oxford, 1977, Clarendon Press, pp. 418-426.
-
-7\. Fricker, S.J.: Narrow-band filter techniques for the detection and
-measurement of evoked responses, Electroencephalogr. Clin. Neurophysiol.
-14:411, 1962.
-
-8\. Van der Tweel, L. H., Sem-Jacobsen, C.W., Kamp, A., Van Leeuwen,
-W.S., and Verings, F.T.H.: Objective determination of response to
-modulated light, Acta Physiol. Pharmacol. Neerl. 7:528, 1958.
-
-9\. Regan, D.: Latencies of evoked potentials to flicker and to pattern
-speedily estimated by simultaneous stimulation method,
-Electroencephalogr. Clin. Neurophysiol. 40:654, 1976.
-
-10\. Tyler, C.W., Apkarian, P., and Nakayama, K.: Multiple spatial
-frequency tuning of electrical responses from the human\....
-
-\[Around here perhaps a section about what gain control is, mentioning
-other methods as well including psychophysics, MRI, electrophysiology,
-and other EEG markers including ERPs and evoked gamma band oscillations.
-Maybe outline the Heeger gain control model and its cousins.\]
-
-\[Sure. But in fact gain control is only a small part of this story -
-especially in the early days. They were, I think, more interested in
-using SSVEP to measure absolute contrast sensitivity and to get lower
-bounds on things like infant visual development. Gain control might be
-better as a separate section later.\]
-
-[]{#anchor-4}2. Spatial, temporal frequency and contrast sensitivity
-measurements
-
-Basically using SSVEP to measure a cortical output amplitude for any
-given input contrast. You can vary parameters like SF, TF, position,
-color and of course contrast. Early on people realised that you can
-'sweep' the stimulus to get a CRF. You broadly get a line in log
-contrast space if you do that (Tyler) - then you can extrapolate that
-line down to zero response to estimate the threshold. That doesn't
-\>quite\< work but it's pretty close.
-
-You can also use this for 'difficult' populations like babies. One
-interesting story was about how SSVEP become a replacement for
-preferential looking (which was the other way of looking at infant
-visual development). See e.g. Davida Teller. SSVEPs allowed people to
-make objective measurements of contrast sensitivity development and
-deduce that the visual system was more mature (e.g. more functional) in
-infancy than previously expected. Also measures of colour sensitivity.
-Tinyeyes is based off those measurements. Other people: Tyler, Norcia,
-Gunilla H-P, many of the people at SKERI in the 1980s and 90s. Norcia
-86,88,90 - mentioned in Regan's nice autobiography :
-
-In parallel of course, people were using frequency tagging to do single
-unit work - the 1F vs 2F simple/complex cell classification scheme was
-all about this (Lennie and others).
-
-[]{#anchor-5}3. SSVEP in functional localization
-
-Techniques like fMRI have been combined with SSVEP to achieve more
-precise spatial localization (Di Russo et al., 2007). See however the
-Ales cruciform paper.
-
-Ales and Norcia:\... (showed that people's intuition about V1 upper /
-lower v.f reversing polarity wasn't really correct). This feels a little
-outside our scope though\... We have also combined SSVEP with source
-imaging techniques to probe responses in different cortical locations
-(again, Ales papers, some early stuff from Stan K? Appelbaum, Wade,
-Norcia figure/ground\...) and then a host of later work from that lab
-and others.
-
-[]{#anchor-6}4. SSVEP and contrast gain control including adaptation,
-masking, and attention
-
-\[The story goes: Up to the lte 90s people were primarily interested in
-measuring contrast sensitivity - the shape of the response function was
-assumed to be basically log-linear - and they fit it with straight lines
-to extrapolate back to zero response. But then (once Heeger's 1992 paper
-had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley \&
-Victor, 1981)](https://www.zotero.org/google-docs/?zAmtgj)), people
-started thinking about gain control - Candy and Norcia in about 1999,
-Porciatti, probably a load of Tyler papers that I don't even know
-about\...\]. Probably look in p
-
-And then people worked out that if you can use SSVEP to measure contrast
-responses, you can also use it to measure things that modulate contrast
-responses. These include adaptation, masking, suppression, attention
-(feature and space), clinical things.
-
-[]{#anchor-7}Adaptation:
-
-Continuous exposure to high-contrast patterns reduces contrast
-sensitivity, which can be measured using SSVEP (Ross et al., 1989).
-Others? Baker recent gain control paper is perhaps worth mentioning here
-as a 'confound' of sorts. Engel 2018 [*(Vergeer et al.,
-2018)*](https://www.zotero.org/google-docs/?kdURiS). This paper is
-interesting [*(Rideaux et al.,
-2023)*](https://www.zotero.org/google-docs/?J2EO3d) and \>sort\< of
-SSVEP.
-
-[]{#anchor-8}Masking:
-
-High contrast masks can suppress the visibility of low contrast
-patterns, which has implications in SSVEP amplitude (Haynes et al.,
-2003).
-
-Attention: Directing attention can enhance contrast sensitivity, as
-shown in studies using SSVEP (Müller et al., 2006). Also Tsai
-(dynamics), Baker / Wade (several), Winawer? I think JW has a nice
-dynamic model of normalization with some MEG data. Busse et al cat/human
-comparison. Candy and Norcia 2001 JNS [(Candy et al.,
-2001)](https://www.zotero.org/google-docs/?7hhs5Z)
-
-Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
-adaptation and masking on incremental thresholds for contrast. Vision
-research, 29(2), 205-215.
-
-Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
-Neuromagnetic correlates of perceived contrast in primary visual cortex.
-Journal of Neurophysiology, 89(6), 2655-2666.
-
-Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
-Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
-selective attention on the steady-state visual evoked potential in the
-20--28 Hz range. Cognitive Brain Research, 24(1), 1-13.
-
-5\. Clinical implications
-
-[]{#anchor-9}Clinical applications (this is a whole section)
-
-Not sure exactly how SSVEP used in clinic. mfVEP?
-
-, e.g., in monitoring visual impairments, tracking neuronal diseases, or
-neurofeedback (Norcia et al., 2015).
-
-Tsai epilepsy [(Tsai et al.,
-2011)](https://www.zotero.org/google-docs/?AdhH3f). Other photosensitive
-epilepsy: Porciatti [(Porciatti et al.,
-2000)](https://www.zotero.org/google-docs/?UTBQbV)
-
-Citation:
-
-Migraine (Regan et al). Autism?
-
-Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
-Rossion, B. (2015). The steady-state visual evoked potential in vision
-research: A review. Journal of vision, 15(6), 4-4.
-
-Animal work: Flies (Eliott, West, Himmelberg, Ales, Norcia): Mice/rats
-(probably many - can't think off the top of my head - we were working
-with a mouse EEG person at UCSF about 12 years ago\...), Monkeys
-(Kiorpes?),
-
-[]{#anchor-10}6. Future directions
-
-Use as a readout of modulations. TMS? FUS? Marmite B12 / Fluoxetine /
-amblyopia in Rats,
-
-GABA Huang
-
-There is a lot of SSVEP interest these days because of BCIs. I think
-it's pretty weak but there is \>so much\< of it that it might be worth
-mentioning\...
-
-Advanced signal processing techniques and machine learning can be
-integrated to improve SSVEP-based systems (Zhu et al., 2010).
-
-Exploring new clinical and diagnostic applications, understanding
-neurological diseases, and developing novel therapeutic interventions.
-
-Citation:
-
-Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
-stimulation methods used in SSVEP-based BCIs. Computational intelligence
-and neuroscience, 2010.
-
 # References
-
-<!-- old code segments that are no longer used -->
-
-```{python}
-#| label: fig-plotspectra
-#| fig-cap: "Temporal frequency representations of combining two inputs at frequencies F1=5Hz and F2=7Hz.
-A pure sine-wave input would generate responses only at F1 and F2.
-However, the rectification that accompanies early visual processing converts this contrast reversal into a more complex frequency-domain representation with power predominantly at the second harmonics 2F1 and 2F2."
-#| fig-cap-location: bottom
-
-#| label: plotsinewaves
-#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain"
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-import numpy as np
-import scipy as scp
-
-inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
-
-duration_S=1 # All variables in this code have an _[units] where possible
-
-digitisationRate_Hz=1000
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))
-for i, freq in enumerate(inputFrequencies_Hz):
-    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)
-
-freqCutoff_Hz=30
-combinedSignal=np.sum(inputModulations,axis=1)
-combinedRectifiedSignal=np.sum(np.abs(inputModulations), axis=1)
-FT_combinedSignal=np.fft.fft(combinedSignal)/len(combinedSignal)
-FT_combinedRectifiedSignal=np.fft.fft(combinedRectifiedSignal)/len(combinedSignal)
-fig,axs=plt.subplots(1,4,figsize=(8,3))
-axs[0].plot(support_S,combinedSignal)
-axs[0].set_xlabel('Time (s)')
-axs[0].set_ylabel('Amplitude (a.u.)')
-axs[0].set_title('Input signals:\nlinear combination')
-
-# Create the frequency support for the FFT plot
-frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
-
-# Plot the absolute values of the FFT in the second axis as a bar plot
-axs[1].bar(frequencySupport, np.abs(FT_combinedSignal[frequencySupport]))
-axs[1].set_xlabel('Frequency (Hz)')
-axs[1].set_ylabel('Magnitude (a.u.)')
-axs[1].set_title('FFT\n linear combination')
-# Plot the absolute values of the FFT in the second axis as a bar plot
-
-axs[2].plot(support_S,combinedRectifiedSignal)
-axs[2].set_xlabel('Time (s)')
-axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('Sum rectified inputs')
-
-axs[3].bar(frequencySupport, np.abs(FT_combinedRectifiedSignal[frequencySupport]))
-axs[3].set_xlabel('Frequency (Hz)')
-axs[3].set_ylabel('Magnitude (a.u.)')
-axs[3].set_title('FFT of\n sum rectified inputs')
-
-plt.tight_layout()
-plt.show()
-
-```
-
-```{python}
-#| label: fig-plotsinewaves
-#| fig-cap: "a) A typical SSVEP setup: a screen, a grating, a flicker rate: b) The responses from a single electrode (time domain) c) Frequency domain."
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-import numpy as np
-import scipy as scp
-
-inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)
-
-duration_S=1 # All variables in this code have an _[units] where possible
-
-digitisationRate_Hz=1000
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-inputModulations = np.zeros((len(support_S), len(inputFrequencies_Hz)))
-
-for i, freq in enumerate(inputFrequencies_Hz):
-    inputModulations[:, i] = np.sin(2 * np.pi * freq * support_S)
-
-# Create a figure with two subplots
-fig, axs = plt.subplots(len(inputFrequencies_Hz), 1, figsize=(8, 6))
-
-# Plot each input modulation in a separate subplot
-for i, freq in enumerate(inputFrequencies_Hz):
-    axs[i].plot(support_S, inputModulations[:, i])
-    axs[i].set_xlabel('Time (s)')
-    axs[i].set_ylabel('Amplitude')
-    axs[i].set_title('Frequency: {} Hz'.format(freq))
-
-# Adjust the spacing between subplots
-plt.tight_layout()
-plt.show()
-
-```
-
-```{python}
-#| label: fig-contrastresponsefunction
-#| fig-cap: "Non-linear contrast response function resulting from the population average of two full-wave rectified signals passing through a non-linear transducer.
-Note the 'supersaturation' at high contrast levels."
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-from matplotlib import colormaps
-import numpy as np
-import scipy as scp
-
-def simpleTransducerFunction(inputContrast, c50, RMax):
- # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
- # RMax defines the reponse gain of the transducer (how big the biggest output is). 
- # c50 defines the conrast sensitivity. Smaller values are more sensitive.
- # The exponent can also vary - here we fix it to 2
-  expnt=2
-
-  # Contrast cannot be negative. To model the responses of a set of on and off 
-  # contrast detectors we perform full-wave rectification of the signal
-  inputContrast=np.abs(inputContrast)
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
-  return output
-
-# Define some constants
-inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
-contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
-duration_S=1 # All variables in this code have an _[units] where possible
-digitisationRate_Hz=1000
-freqCutoff_Hz=30
-c50=.10
-RMax=10
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
-transducedSignal=[]
-
-#Make a figure with three subplots
-fig, axs = plt.subplots(1,3, figsize=(8, 3))
-
-initialSignal = np.zeros(len(contrastRange)) 
-transducedSignal = np.zeros(len(contrastRange))
-for contrastIndex,thisContrast in enumerate(contrastRange):
-  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
-
-  axs[0].plot(support_S, inputModulation)
-
-  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
-
-  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
-  initialSignal[contrastIndex]=inputAmplitude[0]
-
-   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
-   # using the transducer functiona above.
-  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
-  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
-  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequencies_Hz*2])
-  transducedSignal[contrastIndex]=transducedAmplitude[0]
-
-axs[0].set_title('Sine-wave responses at\n different contrast levels', fontsize=10)
-
-axs[1].plot(contrastRange,initialSignal)
-axs[1].set_xlabel('Contrast')
-axs[1].set_ylabel('Amplitude (a.u.)')
-axs[1].set_title('Linear computation stage', fontsize=10)
-axs[1].grid()
-
-axs[2].plot(contrastRange,transducedSignal)
-axs[2].set_xlabel('Contrast')
-axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('After nonlinear transducer', fontsize=10)
-
-axs[2].grid()
-plt.tight_layout()
-plt.show()
-
-```
-
-```{python}
-#| label: fig-intermodulation
-#| fig-cap: "Non-linear combination of two independent contrast modulations at F1 and F2 results in 'intermodulation' terms that appear at low-order sums and differences of those frequencies."
-#| fig-cap-location: bottom
-
-import matplotlib.pyplot as plt
-from matplotlib import colormaps
-import numpy as np
-import scipy as scp
-
-def simpleTransducerFunction(inputContrast, maskContrast, c50, RMax):
- # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
- # RMax defines the reponse gain of the transducer (how big the biggest output is). 
- # c50 defines the conrast sensitivity. Smaller values are more sensitive.
- # The exponent can also vary - here we fix it to 2
-  expnt=2
-  inputContrast=np.abs(inputContrast)
-  maskContrast=np.abs(maskContrast)
-
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
-  return output
-
-# Define some constants
-inputFrequency_Hz=np.array([5,7]) # We simulate a system with two inputs at a single contrast
-inputContrasts=np.array([.2 ,.2])
-duration_S=1 # All variables in this code have an _[units] where possible
-digitisationRate_Hz=1000
-freqCutoff_Hz=30
-c50=.10
-RMax=10
-
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
-
-initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
-transducedSignal=[]
-
-# Make a figure with three subplots
-fig, axs = plt.subplots(1,3, figsize=(8, 3))
-
-# Compute waveforms for the 'probe' and 'mask' inputs at the frequencies defined in inputFrequency_Hz
-probeModulation = np.sin(2 * np.pi * inputFrequency_Hz[0] * support_S) * inputContrasts[0]
-maskModulation =  np.sin(2 * np.pi * inputFrequency_Hz[1] * support_S) * inputContrasts[1]
-
-axs[0].plot(support_S, probeModulation,'k')
-axs[0].plot(support_S, maskModulation,'r')
-axs[0].set_title('Input modulations', fontsize=10)
-axs[0].grid()
-
-# pass both the probe and the mask through the transducer
-combinedSignal=simpleTransducerFunction(probeModulation, maskModulation, c50,RMax)
-
-# Compute the FT of the combined signals
-FT_combinedSignal=np.abs(np.fft.fft(combinedSignal)/ len(combinedSignal))
-axs[1].plot(support_S, combinedSignal,'k')
-axs[1].set_title('Combined signal after transducer', fontsize=10)
-axs[1].grid()
-
-axs[2].bar(np.arange(1,19), FT_combinedSignal[1:19])
-
-axs[2].grid()
-plt.tight_layout()
-plt.show()
-
-```

### 2024-05-06 review.qmd
Minor edits

@@
 Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates [@Luck1997; @Marcus2002; @McAdams1999; @Mehta2000a; @Mehta2000b], but with the advent of spatially-resolved human brain imaging methods in the late 1990s it became apparent that spatial attention was linked to frank changes in both fMRI BOLD responses
-[@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006, @Muller1998m @MullerHillyard2000, @MorganPNAS1996, @Hillyard1998NNS].
+[@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006; @Muller2000; @Morgan1996; @Muller1998].
 
@@
 This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating subtle interactions between the size of the attentional 'spotlight' and the stimulus configuration which rationalise many apparent contradictions in the literature.
-Direct measurements of attentional modulation of achromatic SSVEP signals are broadly consistent with this model [@Lauritzen2010, @Hou2016, @Martinovic2018] and confirm the relatively weaker role of spatial attention on responses driven by chromatic stimuli - particularly those that isolate the opponent S-(L+M) cone pathway [@Wang2011, @Highsmith2010].
+Direct measurements of attentional modulation of achromatic SSVEP signals are broadly consistent with this model [@Lauritzen2010; @Hou2016; @Martinovic2018] and confirm the relatively weaker role of spatial attention on responses driven by chromatic stimuli - particularly those that isolate the opponent S-(L+M) cone pathway [@Wang2011; @Highsmith2010].
 The gain control model of attention can also be extended to SSVEP studies of feature-based attention which show that the modulatory effects can be targeted to the most informative neuronal populations [@Verghese2012].
@@
 For example, locomotion has been shown to alter neuronal excitability and spatial normalization in mice [@Niell2010; @Ayaz2013] - running mice have higher visual sensitivity and lower surround suppression compared to stationary mice.
-Although measuring EEG responses from locomoting humans is technically challenging, SSVEP studies (which are able to distinguish broadband noise from input signal effectively) have shown that walking also alters early visual processing although in a manner different to that observed in mice [@Benjamin2018; @Cao2019].
+Although measuring EEG responses from locomoting humans is technically challenging, SSVEP studies (which are able to distinguish broadband noise from input signal effectively) have shown that walking also alters early visual processing, although in a manner different to that observed in mice [@Benjamin2018; @Cao2019].
 
-SSVEP measurements are typically used to measure time invariant responses due to sustained attention.
+SSVEP measurements are typically used to measure time-invariant responses due to sustained attention.
 However, recent work has shown that moderately high modulation frequencies (ca 10Hz) and short analysis windows (1.5s) can also be used to track the dynamic allocation of attention across a task [@Chota2024] or attention to moving targets [@Lissa2020].
@@
 One widespread recent application of the SSVEP technique is in the design of brain-computer interfaces, which seek to control some aspect of a computer using neural signals.
-The high SNR and frequency resolution of SSVEPs make them an ideal candidate for this approach.
+The high SNR and precise frequency resolution of SSVEPs make them an ideal candidate for this approach.
 Typical studies may involve presenting an array of stimuli at different contrast flicker frequencies, and having the participant select one either by overt attention (i.e.
@@
 Because SSVEP signals are highly sensitive to both visual field position [@Ales2010; @Di-Russo2007] and attentional state [@Morgan1996; @Muller1998; @Verghese2012; @Lauritzen2010], the response to the selected stimulus will typically increase relative to the others, allowing it to be identified by an on-line algorithm.
-Because more than one stimulus frequency is typically present, this modulation and the associated changes in gain control, will affect the entire pattern of self- and intermodulation terms allowing the choice to be decoded by a multivariate pattern classifier.
+Because more than one stimulus frequency is generally present, this modulation and the associated changes in gain control will affect the entire pattern of self- and intermodulation terms, allowing the choice to be decoded by a multivariate pattern classifier.
 
-This approach is primarily useful in situations that require the BCI to distinguish from among a small set of possibilties : For example, in early work, visual stimuli representing 'Left' and 'Right' commands in a flight simulator were distinguished robustly [@Middendorf2000].
+This approach is primarily useful in situations that require the BCI to distinguish from among a small set of possibilties: for example, in early work, visual stimuli representing 'Left' and 'Right' commands in a flight simulator were distinguished robustly [@Middendorf2000].
 Although SSVEP-based BCI interfaces typically do use contrast flicker, work in this field has largely focused on optimising the stimuli or decoders to increase decoding performance (e.g.
@@
 # Conclusions
+
 The SSVEP is a powerful tool for studying contrast processing.
@@
 It is sensitive to both the amplitude and phase of the input and when combined with source imaging, it can be extracted from different cortical regions allowing researchers to track contrast processing computations across the visual pathway.
-Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants) and at frequencies high enough to be effectively invisible [REF - Some high frequency work...
-- Romanian group].
+Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants) and at frequencies high enough to be effectively invisible [@Seijdel2023].
 The SSVEP is also able 'fingerprint' modulators of the input frequency through the harmonic and intermodulation terms they generate in the output.

### 2024-05-06 review.qmd
ALmost done!

--- 2024-05-06 review.qmd
@@
 
-# Conclusions
+# Future directions
 The SSVEP is a powerful tool for studying contrast processing.
@@
 It is sensitive to both the amplitude and phase of the input and when combined with source imaging, it can be extracted from different cortical regions allowing researchers to track contrast processing computations across the visual pathway.
-Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants) and at frequencies high enough to be effectively invisible [REF - Some high frequency work...
+Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants or patients with neurological disorders) and at frequencies high enough to be effectively invisible [REF - Some high frequency work...
 - Romanian group].
-The SSVEP is also able 'fingerprint' modulators of the input frequency through the harmonic and intermodulation terms they generate in the output.
-In principle, each nonlinearity in the visual pathway can be identified by
+The SSVEP is also able to 'fingerprint' modulators of the inputs through the harmonic and intermodulation terms they generate in the output.
+In principle, each nonlinearity in the visual pathway can be identified by its contribution to the frequency spectrum at different recording locations.
+This, it turn, allows researchers to study how contrast processing depends on spatial and temporal context and as well as changes in task, cognitive and behavioural state and arousal.
+
+Although visual neuroscience is a relatively old field, there are still outstanding questions relating to contrast processing that could be addressed by SSVEP methods.
+For example, it is still not completely clear how contrast signals are computed in the human retina: although we have more than a century of electrophysiological data from animals, and the broad structure of cone inputs to retinal ganglion cells is understood [@LiFieldNeuron2014] we are still discovering new aspects of retinal processing that could influence the 'coining' of the visual system's currency [@gollischMeister2010, @WangPuthussery2023Nature, @uprety2022melanopsin].
+The ability to frequency tag both inputs (for example, cone-directed luminance contrast) and modulators (for example, stimuli that selectively drive the intrinsically-photosensitive retinal ganglion cells) is a powerful tool to explore this first stage of image generation.
+
+Later in the visual pathway, we would like to know more about the role of corticothalamic feedback in the LGN.
+Although it is commonly thought of as a simple relay between the eye and the cortex,  the majority of inputs to the LGN come from cortex not the eye, and there is good evidence that contrast processing in the LGN can be altered by top-down signals including attention [@BriggsUsrey2007, @ShermanGuillery1996, @OconnorNatNeuro2002, @GouwsJNS2014].
+Bottom-up inputs to the LGN are segregated by eye and so responses there are often considered to be purely monocular but it is possible that feedback signals allow some binocular computations such as interocular normalisation [@ManyBakerRefs and others] to begin even at this relatively early stage.
+Frequency tagging inputs to different eyes or in different precortical pathways allows us to address neurons in different parts of the LGN.
+These techniques combined with advances in recording technology (for example, sensitive source-imaged EEG and MEG recordings that can resolve subcortical structures [@Attal2012, @AttalPlos12013, @Tesche1996], noninvasive deep-brain stimulation techniques [@MohammadjavadietalSciRep2022] or implanted electrode arrays [@KrolakSalmon2003]) may allow us to study these computations in more detail.
+
+Finally...
+
+(Adaptation, Amblyopia, Neurological disease, Chromatic channels and combinations, joint tuning for color/spatiotemporal frequncy/eye), contributions to gain pools (e.g.
+Solomon and Lennie 2006), inheritance of freqeuncy tags in higher areas, combination of low-level contrast features to make complex stimuli...)
 

### 2024-05-06 review.qmd

--- 2024-05-06 review.qmd
@@
 
-# Conclusions
+# Future directions
+The SSVEP is a powerful tool for studying contrast processing.
+It provides a high SNR readout of neuronal activity that is unambigously linked to the input.
+It is sensitive to both the amplitude and phase of the input and when combined with source imaging, it can be extracted from different cortical regions allowing researchers to track contrast processing computations across the visual pathway.
+Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants or patients with neurological disorders) and at frequencies high enough to be effectively invisible [REF - Some high frequency work...
+- Romanian group].
+The SSVEP is also able to 'fingerprint' modulators of the inputs through the harmonic and intermodulation terms they generate in the output.
+In principle, each nonlinearity in the visual pathway can be identified by its contribution to the frequency spectrum at different recording locations.
+This, it turn, allows researchers to study how contrast processing depends on spatial and temporal context and as well as changes in task, cognitive and behavioural state and arousal.
+
+Although visual neuroscience is a relatively old field, there are still outstanding questions relating to contrast processing that could be addressed by SSVEP methods.
+For example, it is still not completely clear how contrast signals are computed in the human retina: although we have more than a century of electrophysiological data from animals, and the broad structure of cone inputs to retinal ganglion cells is understood [@LiFieldNeuron2014] we are still discovering new aspects of retinal processing that could influence the 'coining' of the visual system's currency [@gollischMeister2010, @WangPuthussery2023Nature, @uprety2022melanopsin].
+The ability to frequency tag both inputs (for example, cone-directed luminance contrast) and modulators (for example, stimuli that selectively drive the intrinsically-photosensitive retinal ganglion cells) is a powerful tool to explore this first stage of image generation.
+
+Later in the visual pathway, we would like to know more about the role of corticothalamic feedback in the LGN.
+Although it is commonly thought of as a simple relay between the eye and the cortex,  the majority of inputs to the LGN come from cortex not the eye, and there is good evidence that contrast processing in the LGN can be altered by top-down signals including attention [@BriggsUsrey2007, @ShermanGuillery1996, @OconnorNatNeuro2002, @GouwsJNS2014].
+Bottom-up inputs to the LGN are segregated by eye and so responses there are often considered to be purely monocular but it is possible that feedback signals allow some binocular computations such as interocular normalisation [@ManyBakerRefs and others] to begin even at this relatively early stage.
+Frequency tagging inputs to different eyes or in different precortical pathways allows us to address neurons in different parts of the LGN.
+These techniques combined with advances in recording technology (for example, sensitive source-imaged EEG and MEG recordings that can resolve subcortical structures [@Attal2012, @AttalPlos12013, @Tesche1996], noninvasive deep-brain stimulation techniques [@MohammadjavadietalSciRep2022] or implanted electrode arrays [@KrolakSalmon2003]) may allow us to study these computations in more detail.
+
+Finally...
+
+(Adaptation, Amblyopia, Neurological disease, Chromatic channels and combinations, joint tuning for color/spatiotemporal frequncy/eye), contributions to gain pools (e.g.
+Solomon and Lennie 2006), inheritance of freqeuncy tags in higher areas, combination of low-level contrast features to make complex stimuli...)
+
+*2: Sweeps and CRFs. Measurements of contrast sensitivity, extrapolating
+the sweep to zero response to get t'hold. Infants and adults?*
+
+*3: Measurements of modulation. Figures adapted from other papers:
+Attention to space, attention to features, adaptation (?).
+masking/surround suppression\\*
+
+*4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*
+
+*5: Future directions (decoding in frequency domain?, animals? BCI?)*
+
+[]{#anchor-2}History
+
+[]{#anchor-3}1. The basics of SSVEP and contrast sensitivity including a
+history of both fields
+
+SSVEP (Steady-State Visual Evoked Potential): A continuous electrical
+response evoked in the brain by visual stimuli flickering at a constant
+frequency (Regan, 1966).
+
+Contrast Sensitivity: The ability to detect differences in luminance
+between an object and its background (Campbell \& Green, 1965).
+
+Regan, D. (1966). Some characteristics of average steady-state and
+transient responses evoked by modulated light. Electroencephalography
+and clinical neurophysiology, 20(3), 238-248.
+
+Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
+affecting visual resolution. J Physiology, 181, 576-593.
+
+Check Regan paper for earlier (e.g. EEG refs). Norcia review will be
+helpful!
+
+(From Tyler / Levi / Apkarian paper):
+
+6\. Regan, D.: Rapid methods for refracting the eye and assessing the
+visual acuity in amblyopia using steady-state visual evoked potentials.
+In Desmedt, J.E., editor: Visual Evoked Potentials in Man: New
+Developments, Oxford, 1977, Clarendon Press, pp. 418-426.
+
+7\. Fricker, S.J.: Narrow-band filter techniques for the detection and
+measurement of evoked responses, Electroencephalogr. Clin. Neurophysiol.
+14:411, 1962.
+
+8\. Van der Tweel, L. H., Sem-Jacobsen, C.W., Kamp, A., Van Leeuwen,
+W.S., and Verings, F.T.H.: Objective determination of response to
+modulated light, Acta Physiol. Pharmacol. Neerl. 7:528, 1958.
+
+9\. Regan, D.: Latencies of evoked potentials to flicker and to pattern
+speedily estimated by simultaneous stimulation method,
+Electroencephalogr. Clin. Neurophysiol. 40:654, 1976.
+
+10\. Tyler, C.W., Apkarian, P., and Nakayama, K.: Multiple spatial
+frequency tuning of electrical responses from the human\....
+
+\[Around here perhaps a section about what gain control is, mentioning
+other methods as well including psychophysics, MRI, electrophysiology,
+and other EEG markers including ERPs and evoked gamma band oscillations.
+Maybe outline the Heeger gain control model and its cousins.\]
+
+\[Sure. But in fact gain control is only a small part of this story -
+especially in the early days. They were, I think, more interested in
+using SSVEP to measure absolute contrast sensitivity and to get lower
+bounds on things like infant visual development. Gain control might be
+better as a separate section later.\]
+
+[]{#anchor-4}2. Spatial, temporal frequency and contrast sensitivity
+measurements
+
+Basically using SSVEP to measure a cortical output amplitude for any
+given input contrast. You can vary parameters like SF, TF, position,
+color and of course contrast. Early on people realised that you can
+'sweep' the stimulus to get a CRF. You broadly get a line in log
+contrast space if you do that (Tyler) - then you can extrapolate that
+line down to zero response to estimate the threshold. That doesn't
+\>quite\< work but it's pretty close.
+
+You can also use this for 'difficult' populations like babies. One
+interesting story was about how SSVEP become a replacement for
+preferential looking (which was the other way of looking at infant
+visual development). See e.g. Davida Teller. SSVEPs allowed people to
+make objective measurements of contrast sensitivity development and
+deduce that the visual system was more mature (e.g. more functional) in
+infancy than previously expected. Also measures of colour sensitivity.
+Tinyeyes is based off those measurements. Other people: Tyler, Norcia,
+Gunilla H-P, many of the people at SKERI in the 1980s and 90s. Norcia
+86,88,90 - mentioned in Regan's nice autobiography :
+
+In parallel of course, people were using frequency tagging to do single
+unit work - the 1F vs 2F simple/complex cell classification scheme was
+all about this (Lennie and others).
+
+[]{#anchor-5}3. SSVEP in functional localization
+
+Techniques like fMRI have been combined with SSVEP to achieve more
+precise spatial localization (Di Russo et al., 2007). See however the
+Ales cruciform paper.
+
+Ales and Norcia:\... (showed that people's intuition about V1 upper /
+lower v.f reversing polarity wasn't really correct). This feels a little
+outside our scope though\... We have also combined SSVEP with source
+imaging techniques to probe responses in different cortical locations
+(again, Ales papers, some early stuff from Stan K? Appelbaum, Wade,
+Norcia figure/ground\...) and then a host of later work from that lab
+and others.
+
+[]{#anchor-6}4. SSVEP and contrast gain control including adaptation,
+masking, and attention
+
+\[The story goes: Up to the lte 90s people were primarily interested in
+measuring contrast sensitivity - the shape of the response function was
+assumed to be basically log-linear - and they fit it with straight lines
+to extrapolate back to zero response. But then (once Heeger's 1992 paper
+had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley \&
+Victor, 1981)](https://www.zotero.org/google-docs/?zAmtgj)), people
+started thinking about gain control - Candy and Norcia in about 1999,
+Porciatti, probably a load of Tyler papers that I don't even know
+about\...\]. Probably look in p
+
+And then people worked out that if you can use SSVEP to measure contrast
+responses, you can also use it to measure things that modulate contrast
+responses. These include adaptation, masking, suppression, attention
+(feature and space), clinical things.
+
+[]{#anchor-7}Adaptation:
+
+Continuous exposure to high-contrast patterns reduces contrast
+sensitivity, which can be measured using SSVEP (Ross et al., 1989).
+Others? Baker recent gain control paper is perhaps worth mentioning here
+as a 'confound' of sorts. Engel 2018 [*(Vergeer et al.,
+2018)*](https://www.zotero.org/google-docs/?kdURiS). This paper is
+interesting [*(Rideaux et al.,
+2023)*](https://www.zotero.org/google-docs/?J2EO3d) and \>sort\< of
+SSVEP.
+
+[]{#anchor-8}Masking:
+
+High contrast masks can suppress the visibility of low contrast
+patterns, which has implications in SSVEP amplitude (Haynes et al.,
+2003).
+
+Attention: Directing attention can enhance contrast sensitivity, as
+shown in studies using SSVEP (Müller et al., 2006). Also Tsai
+(dynamics), Baker / Wade (several), Winawer? I think JW has a nice
+dynamic model of normalization with some MEG data. Busse et al cat/human
+comparison. Candy and Norcia 2001 JNS [(Candy et al.,
+2001)](https://www.zotero.org/google-docs/?7hhs5Z)
+
+Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
+adaptation and masking on incremental thresholds for contrast. Vision
+research, 29(2), 205-215.
+
+Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
+Neuromagnetic correlates of perceived contrast in primary visual cortex.
+Journal of Neurophysiology, 89(6), 2655-2666.
+
+Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
+Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
+selective attention on the steady-state visual evoked potential in the
+20--28 Hz range. Cognitive Brain Research, 24(1), 1-13.
+
+5\. Clinical implications
+
+[]{#anchor-9}Clinical applications (this is a whole section)
+
+Not sure exactly how SSVEP used in clinic. mfVEP?
+
+, e.g., in monitoring visual impairments, tracking neuronal diseases, or
+neurofeedback (Norcia et al., 2015).
+
+Tsai epilepsy [(Tsai et al.,
+2011)](https://www.zotero.org/google-docs/?AdhH3f). Other photosensitive
+epilepsy: Porciatti [(Porciatti et al.,
+2000)](https://www.zotero.org/google-docs/?UTBQbV)
+
+Citation:
+
+Migraine (Regan et al). Autism?
+
+Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
+Rossion, B. (2015). The steady-state visual evoked potential in vision
+research: A review. Journal of vision, 15(6), 4-4.
+
+Animal work: Flies (Eliott, West, Himmelberg, Ales, Norcia): Mice/rats
+(probably many - can't think off the top of my head - we were working
+with a mouse EEG person at UCSF about 12 years ago\...), Monkeys
+(Kiorpes?),
+
+[]{#anchor-10}6. Future directions
+
+Use as a readout of modulations. TMS? FUS? Marmite B12 / Fluoxetine /
+amblyopia in Rats,
+
+GABA Huang
+
+There is a lot of SSVEP interest these days because of BCIs. I think
+it's pretty weak but there is \>so much\< of it that it might be worth
+mentioning\...
+
+Advanced signal processing techniques and machine learning can be
+integrated to improve SSVEP-based systems (Zhu et al., 2010).
+
+Exploring new clinical and diagnostic applications, understanding
+neurological diseases, and developing novel therapeutic interventions.
+
+Citation:
+
+Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
+stimulation methods used in SSVEP-based BCIs. Computational intelligence
+and neuroscience, 2010.
 

### 2024-05-08 review.qmd
All

@@
 ...
-) and reducing visual fatigue associated with the long-term presentation of arrays of high-contrast flicker (e.g.
-.....) rather than studying visual contrast processing \textit{per se}.
+) and reducing visual fatigue associated with the long-term presentation of arrays of high-contrast flicker (@Diez2024AssessmentMedBiolComp) rather than studying visual contrast processing \textit{per se}.
 We therefore note in passing that using SSVEPs to improve our understanding of early contrast processing may yield benefits to this related field.
@@
 It is sensitive to both the amplitude and phase of the input and when combined with source imaging, it can be extracted from different cortical regions allowing researchers to track contrast processing computations across the visual pathway.
-Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants or patients with neurological disorders) and at frequencies high enough to be effectively invisible [REF - Some high frequency work...
-- Romanian group].
+Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants or patients with neurological disorders) and at frequencies high enough to be effectively invisible [@Hermann2001ExpBrRes, @MinarikNeuroimage2023].
 The SSVEP is also able to 'fingerprint' modulators of the inputs through the harmonic and intermodulation terms they generate in the output.
-In principle, each nonlinearity in the visual pathway can be identified by its contribution to the frequency spectrum at different recording locations.
+In principle, each nonlinearity in the visual pathway can be identified by its contribution to the frequency spectrum at different recording locations[@Regan1988].
 This, it turn, allows researchers to study how contrast processing depends on spatial and temporal context and as well as changes in task, cognitive and behavioural state and arousal.
@@
 Although visual neuroscience is a relatively old field, there are still outstanding questions relating to contrast processing that could be addressed by SSVEP methods.
-For example, it is still not completely clear how contrast signals are computed in the human retina: although we have more than a century of electrophysiological data from animals, and the broad structure of cone inputs to retinal ganglion cells is understood [@LiFieldNeuron2014] we are still discovering new aspects of retinal processing that could influence the 'coining' of the visual system's currency [@gollischMeister2010, @WangPuthussery2023Nature, @uprety2022melanopsin].
+For example, it is still not completely clear how contrast signals are computed in the human retina.
+Although we have more than a century of electrophysiological data from animals, and the broad structure of cone inputs to retinal ganglion cells is understood [@LiFieldNeuron2014], we are still discovering new aspects of retinal processing that could influence the 'coining' of the visual system's currency [@gollischMeister2010, @WangPuthussery2023Nature, @uprety2022melanopsin].
 The ability to frequency tag both inputs (for example, cone-directed luminance contrast) and modulators (for example, stimuli that selectively drive the intrinsically-photosensitive retinal ganglion cells) is a powerful tool to explore this first stage of image generation.
@@
 
-Finally...
+Finally, SSVEPs continue to provide a way of studying contrast in the cortex.
+Here, we are often interested in how different visual parameters interact.
+For example, how are signals from different eyes combined to generate both scalar contrast values and also binocular depth cues?
+How does contrast combination depend on the low level properties of the individual inputs such as their retinotopic location, cone contrast, spatiotemporal frequency, eye of origin and orientation?
+Colour vision scientists are particularly interested in how chromatic signals originating in a small number of cone-opponent retinal pathways are transformed into a perceptual colour space where the 'unique hues' appear to be only weakly-related to the early retinal outputs and how these computations are conditioned by the spatial and temporal properties of the scene [@LiConeOppNatComm22, @ConwayUniqueHues,@KanekoCerebCort2020, @GegenfurtnerNatRev2003, @SolomonLennie2007NatRev, @WandellColorPNAS1993, ].
+All of these questions can be addressed by using SSVEP and frequency tagging to examine the computations that combine and transform the inputs at different cortical stages (@WattsBiorxiv2023https://www.biorxiv.org/content/10.1101/2023.12.10.570997v2, @BakerOptimalSignal2017, @ChenGegenfurtnerJOV2021, @KatyalDSciRep2018, @Busse2009).
+The original promise of the SSVEP approach was that the  entire complex-valued frequency spectrum recorded at each lcoation provided detailed information about the processing nonlinearities up to and including that point.
+In principle, this would allow researchers to uniquely fit the parameters of their computations models of visual processing [@Regan1988].
+To date, the complexity of the neuronal computations at even the earliest stage of visual processing have hampered this effort - researchers typically restrict their analyses to the amplitudes of single, low-order frequency components (for example, the inputs frequencies or simple sums and differences of those frequencies).
+As our understanding of the early visual system improves, it is becoming possible to generate more realistic parameterised forward models of signal generation [@CharikerJOV2020, @Winawer, @Baker, @Tsai2012, OTHERS from Winawer].
+Feeding frequency-tagged inputs into these types of model allows us to generate synthetic SSVEP responses that can be compared with those measured in human subject.
+In principle, we are therefore able to use the SSVEP to fit the parameters of early visual processing.
+This approach may also allow us to develop more sensitive tests for the changes in early visual processing that accompany a wide range of neurological diseases and disorders.
 

### 2024-05-08 review.qmd
Subfield

@@
 
-Although visual neuroscience is a relatively old field, there are still outstanding questions relating to contrast processing that could be addressed by SSVEP methods.
+Although visual neuroscience is a relatively old subfield, there are still outstanding questions relating to contrast processing that could be addressed by SSVEP methods.
 For example, it is still not completely clear how contrast signals are computed in the human retina.
@@
 To date, the complexity of the neuronal computations at even the earliest stage of visual processing have hampered this effort - researchers typically restrict their analyses to the amplitudes of single, low-order frequency components (for example, the inputs frequencies or simple sums and differences of those frequencies).
-As our understanding of the early visual system improves, it is becoming possible to generate more realistic parameterised forward models of signal generation [@CharikerJOV2020, @Winawer, @Baker, @Tsai2012, OTHERS from Winawer].
+As our understanding of the early visual system improves, it is becoming possible to generate more realistic parameterised forward models of signal generation [@CharikerJOV2020, @Winawer, @Baker, @Tsai2012, @BrainScore, OTHERS from Winawer].
 Feeding frequency-tagged inputs into these types of model allows us to generate synthetic SSVEP responses that can be compared with those measured in human subject.

### 2024-05-08 review.qmd
First
Fixed refs

@@
 ...
-) and reducing visual fatigue associated with the long-term presentation of arrays of high-contrast flicker (@Diez2024AssessmentMedBiolComp) rather than studying visual contrast processing \textit{per se}.
+) and reducing visual fatigue associated with the long-term presentation of arrays of high-contrast flicker [@Diez2024] rather than studying visual contrast processing \textit{per se}.
 We therefore note in passing that using SSVEPs to improve our understanding of early contrast processing may yield benefits to this related field.
@@
 It is sensitive to both the amplitude and phase of the input and when combined with source imaging, it can be extracted from different cortical regions allowing researchers to track contrast processing computations across the visual pathway.
-Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants or patients with neurological disorders) and at frequencies high enough to be effectively invisible [@Hermann2001ExpBrRes, @MinarikNeuroimage2023].
+Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants or patients with neurological disorders) and at frequencies high enough to be effectively invisible [@Herrmann2001; @Minarik2023; @Seijdel2023].
 The SSVEP is also able to 'fingerprint' modulators of the inputs through the harmonic and intermodulation terms they generate in the output.
-In principle, each nonlinearity in the visual pathway can be identified by its contribution to the frequency spectrum at different recording locations[@Regan1988].
+In principle, each nonlinearity in the visual pathway can be identified by its contribution to the frequency spectrum at different recording locations [@Regan1988].
 This, it turn, allows researchers to study how contrast processing depends on spatial and temporal context and as well as changes in task, cognitive and behavioural state and arousal.
@@
 Although visual neuroscience is a relatively old subfield, there are still outstanding questions relating to contrast processing that could be addressed by SSVEP methods.
-For example, it is still not completely clear how contrast signals are computed in the human retina.
-Although we have more than a century of electrophysiological data from animals, and the broad structure of cone inputs to retinal ganglion cells is understood [@LiFieldNeuron2014], we are still discovering new aspects of retinal processing that could influence the 'coining' of the visual system's currency [@gollischMeister2010, @WangPuthussery2023Nature, @uprety2022melanopsin].
+First, it is still not completely clear how contrast signals are computed in the human retina.
+Although we have more than a century of electrophysiological data from animals, and the broad structure of cone inputs to retinal ganglion cells is understood [@Li2014], we are still discovering new aspects of retinal processing that could influence the 'coining' of the visual system's currency [@Gollisch2010; @Wang2023; @Uprety2022].
 The ability to frequency tag both inputs (for example, cone-directed luminance contrast) and modulators (for example, stimuli that selectively drive the intrinsically-photosensitive retinal ganglion cells) is a powerful tool to explore this first stage of image generation.
@@
 Later in the visual pathway, we would like to know more about the role of corticothalamic feedback in the LGN.
-Although it is commonly thought of as a simple relay between the eye and the cortex,  the majority of inputs to the LGN come from cortex not the eye, and there is good evidence that contrast processing in the LGN can be altered by top-down signals including attention [@BriggsUsrey2007, @ShermanGuillery1996, @OconnorNatNeuro2002, @GouwsJNS2014].
-Bottom-up inputs to the LGN are segregated by eye and so responses there are often considered to be purely monocular but it is possible that feedback signals allow some binocular computations such as interocular normalisation [@ManyBakerRefs and others] to begin even at this relatively early stage.
+Although it is commonly thought of as a simple relay between the eye and the cortex,  the majority of inputs to the LGN come from cortex not the eye, and there is good evidence that contrast processing in the LGN can be altered by top-down signals including attention [@Briggs2007; @Sherman1996; @OConnor2002; @Gouws2014].
+Bottom-up inputs to the LGN are segregated by eye and so responses there are often considered to be purely monocular, but it is possible that feedback signals allow some binocular computations such as interocular normalisation [@Baker2007; @Dougherty2019] to begin even at this relatively early stage.
 Frequency tagging inputs to different eyes or in different precortical pathways allows us to address neurons in different parts of the LGN.
-These techniques combined with advances in recording technology (for example, sensitive source-imaged EEG and MEG recordings that can resolve subcortical structures [@Attal2012, @AttalPlos12013, @Tesche1996], noninvasive deep-brain stimulation techniques [@MohammadjavadietalSciRep2022] or implanted electrode arrays [@KrolakSalmon2003]) may allow us to study these computations in more detail.
+These techniques combined with advances in recording technology (for example, sensitive source-imaged EEG and MEG recordings that can resolve subcortical structures [@Attal2012; @Attal2013; @Tesche1996], noninvasive deep-brain stimulation techniques [@MohammadjavadietalSciRep2022] or implanted electrode arrays [@Krolak-Salmon2003]) may allow us to study these computations in more detail.
 
@@
 How does contrast combination depend on the low level properties of the individual inputs such as their retinotopic location, cone contrast, spatiotemporal frequency, eye of origin and orientation?
-Colour vision scientists are particularly interested in how chromatic signals originating in a small number of cone-opponent retinal pathways are transformed into a perceptual colour space where the 'unique hues' appear to be only weakly-related to the early retinal outputs and how these computations are conditioned by the spatial and temporal properties of the scene [@LiConeOppNatComm22, @ConwayUniqueHues,@KanekoCerebCort2020, @GegenfurtnerNatRev2003, @SolomonLennie2007NatRev, @WandellColorPNAS1993, ].
-All of these questions can be addressed by using SSVEP and frequency tagging to examine the computations that combine and transform the inputs at different cortical stages (@WattsBiorxiv2023https://www.biorxiv.org/content/10.1101/2023.12.10.570997v2, @BakerOptimalSignal2017, @ChenGegenfurtnerJOV2021, @KatyalDSciRep2018, @Busse2009).
+Colour vision scientists are particularly interested in how chromatic signals originating in a small number of cone-opponent retinal pathways are transformed into a perceptual colour space where the 'unique hues' appear to be only weakly-related to the early retinal outputs and how these computations are conditioned by the spatial and temporal properties of the scene [@Li22; @Stoughton2008; @Kaneko2020; @Gegenfurtner2003; @Solomon2007; @Wandell1993].
+All of these questions can be addressed by using SSVEP and frequency tagging to examine the computations that combine and transform the inputs at different cortical stages [@Watts2024; @Baker2017; @Chen2021; @Katyal2018; @Busse2009].
 The original promise of the SSVEP approach was that the  entire complex-valued frequency spectrum recorded at each lcoation provided detailed information about the processing nonlinearities up to and including that point.
@@
 To date, the complexity of the neuronal computations at even the earliest stage of visual processing have hampered this effort - researchers typically restrict their analyses to the amplitudes of single, low-order frequency components (for example, the inputs frequencies or simple sums and differences of those frequencies).
-As our understanding of the early visual system improves, it is becoming possible to generate more realistic parameterised forward models of signal generation [@CharikerJOV2020, @Winawer, @Baker, @Tsai2012, @BrainScore, OTHERS from Winawer].
+As our understanding of the early visual system improves, it is becoming possible to generate more realistic parameterised forward models of signal generation [@Chariker2020; @Groen2022; @Baker2017; @Tsai2012; @Schrimpf2020].
 Feeding frequency-tagged inputs into these types of model allows us to generate synthetic SSVEP responses that can be compared with those measured in human subject.

### 2024-05-08 review.qmd
Two more

@@
 Frequency tagging inputs to different eyes or in different precortical pathways allows us to address neurons in different parts of the LGN.
-These techniques combined with advances in recording technology (for example, sensitive source-imaged EEG and MEG recordings that can resolve subcortical structures [@Attal2012; @Attal2013; @Tesche1996], noninvasive deep-brain stimulation techniques [@MohammadjavadietalSciRep2022] or implanted electrode arrays [@Krolak-Salmon2003]) may allow us to study these computations in more detail.
+These techniques combined with advances in recording technology (for example, sensitive source-imaged EEG and MEG recordings that can resolve subcortical structures [@Attal2012; @Attal2013; @Tesche1996], noninvasive deep-brain stimulation techniques [@Mohammadjavadi2022] or implanted electrode arrays [@Krolak-Salmon2003]) may allow us to study these computations in more detail.
 
@@
 How does contrast combination depend on the low level properties of the individual inputs such as their retinotopic location, cone contrast, spatiotemporal frequency, eye of origin and orientation?
-Colour vision scientists are particularly interested in how chromatic signals originating in a small number of cone-opponent retinal pathways are transformed into a perceptual colour space where the 'unique hues' appear to be only weakly-related to the early retinal outputs and how these computations are conditioned by the spatial and temporal properties of the scene [@Li22; @Stoughton2008; @Kaneko2020; @Gegenfurtner2003; @Solomon2007; @Wandell1993].
+Colour vision scientists are particularly interested in how chromatic signals originating in a small number of cone-opponent retinal pathways are transformed into a perceptual colour space where the 'unique hues' appear to be only weakly-related to the early retinal outputs and how these computations are conditioned by the spatial and temporal properties of the scene [@Li2022; @Stoughton2008; @Kaneko2020; @Gegenfurtner2003; @Solomon2007; @Wandell1993].
 All of these questions can be addressed by using SSVEP and frequency tagging to examine the computations that combine and transform the inputs at different cortical stages [@Watts2024; @Baker2017; @Chen2021; @Katyal2018; @Busse2009].

### 2024-05-09 review.qmd
Code to make R+H plots -

@@
 
-Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
+Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolaKting back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
 This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
@@
 
-(Adaptation, Amblyopia, Neurological disease, Chromatic channels and combinations, joint tuning for color/spatiotemporal frequncy/eye), contributions to gain pools (e.g.
-Solomon and Lennie 2006), inheritance of freqeuncy tags in higher areas, combination of low-level contrast features to make complex stimuli...)
-
-*2: Sweeps and CRFs. Measurements of contrast sensitivity, extrapolating
-the sweep to zero response to get t'hold. Infants and adults?*
-
-*3: Measurements of modulation. Figures adapted from other papers:
-Attention to space, attention to features, adaptation (?).
-masking/surround suppression\\*
-
-*4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)*
-
-*5: Future directions (decoding in frequency domain?, animals? BCI?)*
-
-[]{#anchor-2}History
-
-[]{#anchor-3}1. The basics of SSVEP and contrast sensitivity including a
-history of both fields
-
-SSVEP (Steady-State Visual Evoked Potential): A continuous electrical
-response evoked in the brain by visual stimuli flickering at a constant
-frequency (Regan, 1966).
-
-Contrast Sensitivity: The ability to detect differences in luminance
-between an object and its background (Campbell \& Green, 1965).
-
-Regan, D. (1966). Some characteristics of average steady-state and
-transient responses evoked by modulated light. Electroencephalography
-and clinical neurophysiology, 20(3), 238-248.
-
-Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
-affecting visual resolution. J Physiology, 181, 576-593.
-
-Check Regan paper for earlier (e.g. EEG refs). Norcia review will be
-helpful!
-
-(From Tyler / Levi / Apkarian paper):
-
-6\. Regan, D.: Rapid methods for refracting the eye and assessing the
-visual acuity in amblyopia using steady-state visual evoked potentials.
-In Desmedt, J.E., editor: Visual Evoked Potentials in Man: New
-Developments, Oxford, 1977, Clarendon Press, pp. 418-426.
-
-7\. Fricker, S.J.: Narrow-band filter techniques for the detection and
-measurement of evoked responses, Electroencephalogr. Clin. Neurophysiol.
-14:411, 1962.
-
-8\. Van der Tweel, L. H., Sem-Jacobsen, C.W., Kamp, A., Van Leeuwen,
-W.S., and Verings, F.T.H.: Objective determination of response to
-modulated light, Acta Physiol. Pharmacol. Neerl. 7:528, 1958.
-
-9\. Regan, D.: Latencies of evoked potentials to flicker and to pattern
-speedily estimated by simultaneous stimulation method,
-Electroencephalogr. Clin. Neurophysiol. 40:654, 1976.
-
-10\. Tyler, C.W., Apkarian, P., and Nakayama, K.: Multiple spatial
-frequency tuning of electrical responses from the human\....
-
-\[Around here perhaps a section about what gain control is, mentioning
-other methods as well including psychophysics, MRI, electrophysiology,
-and other EEG markers including ERPs and evoked gamma band oscillations.
-Maybe outline the Heeger gain control model and its cousins.\]
-
-\[Sure. But in fact gain control is only a small part of this story -
-especially in the early days. They were, I think, more interested in
-using SSVEP to measure absolute contrast sensitivity and to get lower
-bounds on things like infant visual development. Gain control might be
-better as a separate section later.\]
-
-[]{#anchor-4}2. Spatial, temporal frequency and contrast sensitivity
-measurements
-
-Basically using SSVEP to measure a cortical output amplitude for any
-given input contrast. You can vary parameters like SF, TF, position,
-color and of course contrast. Early on people realised that you can
-'sweep' the stimulus to get a CRF. You broadly get a line in log
-contrast space if you do that (Tyler) - then you can extrapolate that
-line down to zero response to estimate the threshold. That doesn't
-\>quite\< work but it's pretty close.
-
-You can also use this for 'difficult' populations like babies. One
-interesting story was about how SSVEP become a replacement for
-preferential looking (which was the other way of looking at infant
-visual development). See e.g. Davida Teller. SSVEPs allowed people to
-make objective measurements of contrast sensitivity development and
-deduce that the visual system was more mature (e.g. more functional) in
-infancy than previously expected. Also measures of colour sensitivity.
-Tinyeyes is based off those measurements. Other people: Tyler, Norcia,
-Gunilla H-P, many of the people at SKERI in the 1980s and 90s. Norcia
-86,88,90 - mentioned in Regan's nice autobiography :
-
-In parallel of course, people were using frequency tagging to do single
-unit work - the 1F vs 2F simple/complex cell classification scheme was
-all about this (Lennie and others).
-
-[]{#anchor-5}3. SSVEP in functional localization
-
-Techniques like fMRI have been combined with SSVEP to achieve more
-precise spatial localization (Di Russo et al., 2007). See however the
-Ales cruciform paper.
-
-Ales and Norcia:\... (showed that people's intuition about V1 upper /
-lower v.f reversing polarity wasn't really correct). This feels a little
-outside our scope though\... We have also combined SSVEP with source
-imaging techniques to probe responses in different cortical locations
-(again, Ales papers, some early stuff from Stan K? Appelbaum, Wade,
-Norcia figure/ground\...) and then a host of later work from that lab
-and others.
-
-[]{#anchor-6}4. SSVEP and contrast gain control including adaptation,
-masking, and attention
-
-\[The story goes: Up to the lte 90s people were primarily interested in
-measuring contrast sensitivity - the shape of the response function was
-assumed to be basically log-linear - and they fit it with straight lines
-to extrapolate back to zero response. But then (once Heeger's 1992 paper
-had sunk in - see also stuff like Shapley and Victor 1981 [(Shapley \&
-Victor, 1981)](https://www.zotero.org/google-docs/?zAmtgj)), people
-started thinking about gain control - Candy and Norcia in about 1999,
-Porciatti, probably a load of Tyler papers that I don't even know
-about\...\]. Probably look in p
-
-And then people worked out that if you can use SSVEP to measure contrast
-responses, you can also use it to measure things that modulate contrast
-responses. These include adaptation, masking, suppression, attention
-(feature and space), clinical things.
-
-[]{#anchor-7}Adaptation:
-
-Continuous exposure to high-contrast patterns reduces contrast
-sensitivity, which can be measured using SSVEP (Ross et al., 1989).
-Others? Baker recent gain control paper is perhaps worth mentioning here
-as a 'confound' of sorts. Engel 2018 [*(Vergeer et al.,
-2018)*](https://www.zotero.org/google-docs/?kdURiS). This paper is
-interesting [*(Rideaux et al.,
-2023)*](https://www.zotero.org/google-docs/?J2EO3d) and \>sort\< of
-SSVEP.
-
-[]{#anchor-8}Masking:
-
-High contrast masks can suppress the visibility of low contrast
-patterns, which has implications in SSVEP amplitude (Haynes et al.,
-2003).
-
-Attention: Directing attention can enhance contrast sensitivity, as
-shown in studies using SSVEP (Müller et al., 2006). Also Tsai
-(dynamics), Baker / Wade (several), Winawer? I think JW has a nice
-dynamic model of normalization with some MEG data. Busse et al cat/human
-comparison. Candy and Norcia 2001 JNS [(Candy et al.,
-2001)](https://www.zotero.org/google-docs/?7hhs5Z)
-
-Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
-adaptation and masking on incremental thresholds for contrast. Vision
-research, 29(2), 205-215.
-
-Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
-Neuromagnetic correlates of perceived contrast in primary visual cortex.
-Journal of Neurophysiology, 89(6), 2655-2666.
-
-Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
-Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
-selective attention on the steady-state visual evoked potential in the
-20--28 Hz range. Cognitive Brain Research, 24(1), 1-13.
-
-5\. Clinical implications
-
-[]{#anchor-9}Clinical applications (this is a whole section)
-
-Not sure exactly how SSVEP used in clinic. mfVEP?
-
-, e.g., in monitoring visual impairments, tracking neuronal diseases, or
-neurofeedback (Norcia et al., 2015).
-
-Tsai epilepsy [(Tsai et al.,
-2011)](https://www.zotero.org/google-docs/?AdhH3f). Other photosensitive
-epilepsy: Porciatti [(Porciatti et al.,
-2000)](https://www.zotero.org/google-docs/?UTBQbV)
-
-Citation:
-
-Migraine (Regan et al). Autism?
-
-Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
-Rossion, B. (2015). The steady-state visual evoked potential in vision
-research: A review. Journal of vision, 15(6), 4-4.
-
-Animal work: Flies (Eliott, West, Himmelberg, Ales, Norcia): Mice/rats
-(probably many - can't think off the top of my head - we were working
-with a mouse EEG person at UCSF about 12 years ago\...), Monkeys
-(Kiorpes?),
-
-[]{#anchor-10}6. Future directions
-
-Use as a readout of modulations. TMS? FUS? Marmite B12 / Fluoxetine /
-amblyopia in Rats,
-
-GABA Huang
-
-There is a lot of SSVEP interest these days because of BCIs. I think
-it's pretty weak but there is \>so much\< of it that it might be worth
-mentioning\...
-
-Advanced signal processing techniques and machine learning can be
-integrated to improve SSVEP-based systems (Zhu et al., 2010).
-
-Exploring new clinical and diagnostic applications, understanding
-neurological diseases, and developing novel therapeutic interventions.
-
-Citation:
-
-Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
-stimulation methods used in SSVEP-based BCIs. Computational intelligence
-and neuroscience, 2010.
-
-The SSVEP is a powerful tool for studying contrast processing.
-It provides a high SNR readout of neuronal activity that is unambigously linked to the input.
-It is sensitive to both the amplitude and phase of the input and when combined with source imaging, it can be extracted from different cortical regions allowing researchers to track contrast processing computations across the visual pathway.
-Because of the high SNR, it can be measured in subjects where long recording durations are impractical (for example, infants) and at frequencies high enough to be effectively invisible [@Seijdel2023].
-The SSVEP is also able 'fingerprint' modulators of the input frequency through the harmonic and intermodulation terms they generate in the output.
-In principle, each nonlinearity in the visual pathway can be identified by
-
 # References

### 2024-05-09 review.tex
Code to make R+H plots -

@@
 of nonlinear retinal and cortical operations up to that point. It
-therefore contains a `signature' of the computational nature, order and
-parameters of those operations - including the shape of the transducer
-functions and the computations involved in signal combination. In
-principle, that information could be recovered from the SSVEP signal - a
-possibility recognised in the early days of the technique (Regan \&
-Regan, 1988). Although characterising the complete set of computations
-along entire processing pathway is challenging, careful parametric
-variation of the input stimuli does allow us to fit models of early
-visual gain control using SSVEP data. For example, Tsai \emph{et al.}
-(2012) demonstrated that a gain control model gave a good account of the
-pattern of intermodulation responses produced by two overlaid patterns
-flickering at different frequencies. This was achieved by passing full
-stimulus waveforms through the transducer nonlinearity, and calculating
-the Fourier spectrum of the model output. Our own work on signal
-combination across eyes and space similarly demonstrated close
+therefore contains a `signature' or `fingerprint' of the computational
+nature, order and parameters of those operations - including the shape
+of the transducer functions and the computations involved in signal
+combination. In principle, that information could be recovered from the
+SSVEP signal - a possibility recognised in the early days of the
+technique (Regan \& Regan, 1988). Although characterising the complete
+set of computations along entire processing pathway is challenging,
+careful parametric variation of the input stimuli does allow us to fit
+models of early visual gain control using SSVEP data. For example, Tsai
+\emph{et al.} (2012) demonstrated that a gain control model gave a good
+account of the pattern of intermodulation responses produced by two
+overlaid patterns flickering at different frequencies. This was achieved
+by passing full stimulus waveforms through the transducer nonlinearity,
+and calculating the Fourier spectrum of the model output. Our own work
+on signal combination across eyes and space similarly demonstrated close
 correspondence between the predictions of a computational model and
@@
 estimating a threshold is achieved by measuring the contrast response
-function at relatively high levels, and extrapolating back along the
+function at relatively high levels, and extrapolaKting back along the
 function (either contrast vs response measured at a constant spatial
@@
 
-\section{Attention}\label{attention}
+\section{Attention and arousal}\label{attention-and-arousal}
 
@@
 \emph{et al.}, 2007; Li \emph{et al.}, 2008; Murray, 2008) and EEG
-signals (\textbf{MullerHillyard2000?}).
+signals (Morgan \emph{et al.}, 1996; Müller \emph{et al.}, 1998; Müller
+\& Hillyard, 2000; Ding \emph{et al.}, 2006).
 
@@
 achromatic SSVEP signals are broadly consistent with this model
-Martinovic \& Andersen (2018) and confirm the relatively weaker role of
-spatial attention on responses driven by chromatic stimuli -
-particularly those that isolate the opponent S-(L+M) cone pathway
-Highsmith \& Crognale (2010). The gain control model of attention can
-also be extended to SSVEP studies of feature-based attention which show
-that the modulatory effects can be targeted to the most informative
-neuronal populations (Verghese \emph{et al.}, 2012).
+(Lauritzen \emph{et al.}, 2010; Hou \emph{et al.}, 2016; Martinovic \&
+Andersen, 2018) and confirm the relatively weaker role of spatial
+attention on responses driven by chromatic stimuli - particularly those
+that isolate the opponent S-(L+M) cone pathway (Highsmith \& Crognale,
+2010; Wang \& Wade, 2011). The gain control model of attention can also
+be extended to SSVEP studies of feature-based attention which show that
+the modulatory effects can be targeted to the most informative neuronal
+populations (Verghese \emph{et al.}, 2012).
+
+The SSVEP can also be used to study changes in visual processing
+\textit{driven} by different behavioural states or overall arousal. For
+example, locomotion has been shown to alter neuronal excitability and
+spatial normalization in mice (Niell \& Stryker, 2010; Ayaz \emph{et
+al.}, 2013) - running mice have higher visual sensitivity and lower
+surround suppression compared to stationary mice. Although measuring EEG
+responses from locomoting humans is technically challenging, SSVEP
+studies (which are able to distinguish broadband noise from input signal
+effectively) have shown that walking also alters early visual
+processing, although in a manner different to that observed in mice
+(Benjamin \emph{et al.}, 2018; Cao \& Händel, 2019).
+
+SSVEP measurements are typically used to measure time-invariant
+responses due to sustained attention. However, recent work has shown
+that moderately high modulation frequencies (ca 10Hz) and short analysis
+windows (1.5s) can also be used to track the dynamic allocation of
+attention across a task (Chota \emph{et al.}, 2024) or attention to
+moving targets (Lissa \emph{et al.}, 2020). The ability to track changes
+in attention over short time periods is also important if SSVEP is to be
+used for dynamic readouts - for example in a Brain Computer Interface.
 
@@
 design of brain-computer interfaces, which seek to control some aspect
-of a computer using neural signals. The high SNR and frequency
+of a computer using neural signals. The high SNR and precise frequency
 resolution of SSVEPs make them an ideal candidate for this approach.
@@
 selected stimulus) or covert attention (i.e.~deploying attention to one
-stimulus whilst keeping fixated) (\textbf{Middendorf?}). BEcause EEG
-signals are highly sensitive to both visual field position (Ales
-\emph{et al.}, 2010) and attentional state (Lauritzen \emph{et al.},
-2010; Verghese \emph{et al.}, 2012), the response to the selected
-stimulus will typically increase relative to the others, allowing it to
-be identified by an on-line algorithm. Because more than one stimulus
-frequency is typically present, this modulation and the associated
-changes in gain control, will affect the entire pattern of self- and
-intermodulation terms.
+stimulus whilst keeping fixated) (Middendorf \emph{et al.}, 2000).
+Because SSVEP signals are highly sensitive to both visual field position
+(Di Russo \emph{et al.}, 2007; Ales \emph{et al.}, 2010) and attentional
+state (Morgan \emph{et al.}, 1996; Müller \emph{et al.}, 1998; Lauritzen
+\emph{et al.}, 2010; Verghese \emph{et al.}, 2012), the response to the
+selected stimulus will typically increase relative to the others,
+allowing it to be identified by an on-line algorithm. Because more than
+one stimulus frequency is generally present, this modulation and the
+associated changes in gain control will affect the entire pattern of
+self- and intermodulation terms, allowing the choice to be decoded by a
+multivariate pattern classifier.
 
 This approach is primarily useful in situations that require the BCI to
-distinguish from among a small set of possibilties : For example, in
+distinguish from among a small set of possibilties: for example, in
 early work, visual stimuli representing `Left' and `Right' commands in a
-flight simulator were distinguished robustly (\textbf{Middendorf?}).
-Although SSVEP-based BCI interfaces typically do use contrast flicker,
-work in this field has largely focused on optimising the stimuli or
-decoders to increase decoding performance (e.g.~\ldots{} ) and reducing
-visual fatigue associated with the long-term presentation of arrays of
-high-contrast flicker (e.g.~\ldots..) rather than studying visual
-contrast processing \textit{per se}.
-
-\section{Conclusions}\label{conclusions}
-
-\emph{2: Sweeps and CRFs. Measurements of contrast sensitivity,
-extrapolating the sweep to zero response to get t'hold. Infants and
-adults?}
-
-\emph{3: Measurements of modulation. Figures adapted from other papers:
-Attention to space, attention to features, adaptation (?).
-masking/surround suppression\textbackslash{}}
-
-\emph{4: Clinical (Porciatti / Tsai/ Marmite/ Amblyopia / PD?)}
-
-\emph{5: Future directions (decoding in frequency domain?, animals?
-BCI?)}
-
-\phantomsection\label{anchor-2}{}History
-
-\phantomsection\label{anchor-3}{}1. The basics of SSVEP and contrast
-sensitivity including a history of both fields
-
-SSVEP (Steady-State Visual Evoked Potential): A continuous electrical
-response evoked in the brain by visual stimuli flickering at a constant
-frequency (Regan, 1966).
-
-Contrast Sensitivity: The ability to detect differences in luminance
-between an object and its background (Campbell \& Green, 1965).
-
-Regan, D. (1966). Some characteristics of average steady-state and
-transient responses evoked by modulated light. Electroencephalography
-and clinical neurophysiology, 20(3), 238-248.
-
-Campbell, F. W., \& Green, D. G. (1965). Optical and retinal factors
-affecting visual resolution. J Physiology, 181, 576-593.
-
-Check Regan paper for earlier (e.g.~EEG refs). Norcia review will be
-helpful!
-
-(From Tyler / Levi / Apkarian paper):
-
-6. Regan, D.: Rapid methods for refracting the eye and assessing the
-visual acuity in amblyopia using steady-state visual evoked potentials.
-In Desmedt, J.E., editor: Visual Evoked Potentials in Man: New
-Developments, Oxford, 1977, Clarendon Press, pp.~418-426.
-
-7. Fricker, S.J.: Narrow-band filter techniques for the detection and
-measurement of evoked responses, Electroencephalogr. Clin. Neurophysiol.
-14:411, 1962.
-
-8. Van der Tweel, L. H., Sem-Jacobsen, C.W., Kamp, A., Van Leeuwen,
-W.S., and Verings, F.T.H.: Objective determination of response to
-modulated light, Acta Physiol. Pharmacol. Neerl. 7:528, 1958.
-
-9. Regan, D.: Latencies of evoked potentials to flicker and to pattern
-speedily estimated by simultaneous stimulation method,
-Electroencephalogr. Clin. Neurophysiol. 40:654, 1976.
-
-10. Tyler, C.W., Apkarian, P., and Nakayama, K.: Multiple spatial
-frequency tuning of electrical responses from the human.\ldots{}
-
-{[}Around here perhaps a section about what gain control is, mentioning
-other methods as well including psychophysics, MRI, electrophysiology,
-and other EEG markers including ERPs and evoked gamma band oscillations.
-Maybe outline the Heeger gain control model and its cousins.{]}
-
-{[}Sure. But in fact gain control is only a small part of this story -
-especially in the early days. They were, I think, more interested in
-using SSVEP to measure absolute contrast sensitivity and to get lower
-bounds on things like infant visual development. Gain control might be
-better as a separate section later.{]}
-
-\phantomsection\label{anchor-4}{}2. Spatial, temporal frequency and
-contrast sensitivity measurements
-
-Basically using SSVEP to measure a cortical output amplitude for any
-given input contrast. You can vary parameters like SF, TF, position,
-color and of course contrast. Early on people realised that you can
-`sweep' the stimulus to get a CRF. You broadly get a line in log
-contrast space if you do that (Tyler) - then you can extrapolate that
-line down to zero response to estimate the threshold. That doesn't
-\textgreater quite\textless{} work but it's pretty close.
-
-You can also use this for `difficult' populations like babies. One
-interesting story was about how SSVEP become a replacement for
-preferential looking (which was the other way of looking at infant
-visual development). See e.g.~Davida Teller. SSVEPs allowed people to
-make objective measurements of contrast sensitivity development and
-deduce that the visual system was more mature (e.g.~more functional) in
-infancy than previously expected. Also measures of colour sensitivity.
-Tinyeyes is based off those measurements. Other people: Tyler, Norcia,
-Gunilla H-P, many of the people at SKERI in the 1980s and 90s. Norcia
-86,88,90 - mentioned in Regan's nice autobiography :
-
-In parallel of course, people were using frequency tagging to do single
-unit work - the 1F vs 2F simple/complex cell classification scheme was
-all about this (Lennie and others).
-
-\phantomsection\label{anchor-5}{}3. SSVEP in functional localization
-
-Techniques like fMRI have been combined with SSVEP to achieve more
-precise spatial localization (Di Russo et al., 2007). See however the
-Ales cruciform paper.
-
-Ales and Norcia:... (showed that people's intuition about V1 upper /
-lower v.f reversing polarity wasn't really correct). This feels a little
-outside our scope though... We have also combined SSVEP with source
-imaging techniques to probe responses in different cortical locations
-(again, Ales papers, some early stuff from Stan K? Appelbaum, Wade,
-Norcia figure/ground...) and then a host of later work from that lab and
-others.
-
-\phantomsection\label{anchor-6}{}4. SSVEP and contrast gain control
-including adaptation, masking, and attention
-
-{[}The story goes: Up to the lte 90s people were primarily interested in
-measuring contrast sensitivity - the shape of the response function was
-assumed to be basically log-linear - and they fit it with straight lines
-to extrapolate back to zero response. But then (once Heeger's 1992 paper
-had sunk in - see also stuff like Shapley and Victor 1981
-\href{https://www.zotero.org/google-docs/?zAmtgj}{(Shapley \& Victor,
-1981)}), people started thinking about gain control - Candy and Norcia
-in about 1999, Porciatti, probably a load of Tyler papers that I don't
-even know about...{]}. Probably look in p
-
-And then people worked out that if you can use SSVEP to measure contrast
-responses, you can also use it to measure things that modulate contrast
-responses. These include adaptation, masking, suppression, attention
-(feature and space), clinical things.
-
-\phantomsection\label{anchor-7}{}Adaptation:
-
-Continuous exposure to high-contrast patterns reduces contrast
-sensitivity, which can be measured using SSVEP (Ross et al., 1989).
-Others? Baker recent gain control paper is perhaps worth mentioning here
-as a `confound' of sorts. Engel 2018
-\href{https://www.zotero.org/google-docs/?kdURiS}{\emph{(Vergeer et al.,
-2018)}}. This paper is interesting
-\href{https://www.zotero.org/google-docs/?J2EO3d}{\emph{(Rideaux et al.,
-2023)}} and \textgreater sort\textless{} of SSVEP.
-
-\phantomsection\label{anchor-8}{}Masking:
-
-High contrast masks can suppress the visibility of low contrast
-patterns, which has implications in SSVEP amplitude (Haynes et al.,
-2003).
-
-Attention: Directing attention can enhance contrast sensitivity, as
-shown in studies using SSVEP (Müller et al., 2006). Also Tsai
-(dynamics), Baker / Wade (several), Winawer? I think JW has a nice
-dynamic model of normalization with some MEG data. Busse et al cat/human
-comparison. Candy and Norcia 2001 JNS
-\href{https://www.zotero.org/google-docs/?7hhs5Z}{(Candy et al., 2001)}
-
-Ross, J., Speed, H. D., \& Morgan, M. J. (1989). The effects of
-adaptation and masking on incremental thresholds for contrast. Vision
-research, 29(2), 205-215.
-
-Haynes, J. D., Roth, G., Stadler, M., \& Heinze, H. J. (2003).
-Neuromagnetic correlates of perceived contrast in primary visual cortex.
-Journal of Neurophysiology, 89(6), 2655-2666.
-
-Müller, M. M., Picton, T. W., Valdes-Sosa, P., Riera, J.,
-Teder-Sälejärvi, W. A., \& Hillyard, S. A. (2006). Effects of spatial
-selective attention on the steady-state visual evoked potential in the
-20--28 Hz range. Cognitive Brain Research, 24(1), 1-13.
-
-5. Clinical implications
-
-\phantomsection\label{anchor-9}{}Clinical applications (this is a whole
-section)
-
-Not sure exactly how SSVEP used in clinic. mfVEP?
-
-, e.g., in monitoring visual impairments, tracking neuronal diseases, or
-neurofeedback (Norcia et al., 2015).
-
-Tsai epilepsy \href{https://www.zotero.org/google-docs/?AdhH3f}{(Tsai et
-al., 2011)}. Other photosensitive epilepsy: Porciatti
-\href{https://www.zotero.org/google-docs/?UTBQbV}{(Porciatti et al.,
-2000)}
-
-Citation:
-
-Migraine (Regan et al). Autism?
-
-Norcia, A. M., Appelbaum, L. G., Ales, J. M., Cottereau, B. R., \&
-Rossion, B. (2015). The steady-state visual evoked potential in vision
-research: A review. Journal of vision, 15(6), 4-4.
-
-Animal work: Flies (Eliott, West, Himmelberg, Ales, Norcia): Mice/rats
-(probably many - can't think off the top of my head - we were working
-with a mouse EEG person at UCSF about 12 years ago...), Monkeys
-(Kiorpes?),
-
-\phantomsection\label{anchor-10}{}6. Future directions
-
-Use as a readout of modulations. TMS? FUS? Marmite B12 / Fluoxetine /
-amblyopia in Rats,
-
-GABA Huang
-
-There is a lot of SSVEP interest these days because of BCIs. I think
-it's pretty weak but there is \textgreater so much\textless{} of it that
-it might be worth mentioning...
-
-Advanced signal processing techniques and machine learning can be
-integrated to improve SSVEP-based systems (Zhu et al., 2010).
-
-Exploring new clinical and diagnostic applications, understanding
-neurological diseases, and developing novel therapeutic interventions.
-
-Citation:
-
-Zhu, D., Bieger, J., Molina, G. G., \& Aarts, R. M. (2010). A survey of
-stimulation methods used in SSVEP-based BCIs. Computational intelligence
-and neuroscience, 2010.
-
-\section{References}\label{references}
+flight simulator were distinguished robustly (Middendorf \emph{et al.},
+2000). Although SSVEP-based BCI interfaces typically do use contrast
+flicker, work in this field has largely focused on optimising the
+stimuli or decoders to increase decoding performance (e.g.~\ldots{} )
+and reducing visual fatigue associated with the long-term presentation
+of arrays of high-contrast flicker (Diez \emph{et al.}, 2024) rather
+than studying visual contrast processing \textit{per se}. We therefore
+note in passing that using SSVEPs to improve our understanding of early
+contrast processing may yield benefits to this related field.
+
+\section{Future directions}\label{future-directions}
+
+The SSVEP is a powerful tool for studying contrast processing. It
+provides a high SNR readout of neuronal activity that is unambigously
+linked to the input. It is sensitive to both the amplitude and phase of
+the input and when combined with source imaging, it can be extracted
+from different cortical regions allowing researchers to track contrast
+processing computations across the visual pathway. Because of the high
+SNR, it can be measured in subjects where long recording durations are
+impractical (for example, infants or patients with neurological
+disorders) and at frequencies high enough to be effectively invisible
+(Herrmann, 2001; Seijdel \emph{et al.}, 2023; Minarik \emph{et al.},
+2023). The SSVEP is also able to `fingerprint' modulators of the inputs
+through the harmonic and intermodulation terms they generate in the
+output. In principle, each nonlinearity in the visual pathway can be
+identified by its contribution to the frequency spectrum at different
+recording locations (Regan \& Regan, 1988). This, it turn, allows
+researchers to study how contrast processing depends on spatial and
+temporal context and as well as changes in task, cognitive and
+behavioural state and arousal.
+
+Although visual neuroscience is a relatively old subfield, there are
+still outstanding questions relating to contrast processing that could
+be addressed by SSVEP methods. First, it is still not completely clear
+how contrast signals are computed in the human retina. Although we have
+more than a century of electrophysiological data from animals, and the
+broad structure of cone inputs to retinal ganglion cells is understood
+(Li \emph{et al.}, 2014), we are still discovering new aspects of
+retinal processing that could influence the `coining' of the visual
+system's currency (Gollisch \& Meister, 2010; Uprety \emph{et al.},
+2022; Wang \emph{et al.}, 2023). The ability to frequency tag both
+inputs (for example, cone-directed luminance contrast) and modulators
+(for example, stimuli that selectively drive the
+intrinsically-photosensitive retinal ganglion cells) is a powerful tool
+to explore this first stage of image generation.
+
+Later in the visual pathway, we would like to know more about the role
+of corticothalamic feedback in the LGN. Although it is commonly thought
+of as a simple relay between the eye and the cortex, the majority of
+inputs to the LGN come from cortex not the eye, and there is good
+evidence that contrast processing in the LGN can be altered by top-down
+signals including attention (Sherman \& Guillery, 1996; O'Connor
+\emph{et al.}, 2002; Briggs \& Usrey, 2007; Gouws \emph{et al.}, 2014).
+Bottom-up inputs to the LGN are segregated by eye and so responses there
+are often considered to be purely monocular, but it is possible that
+feedback signals allow some binocular computations such as interocular
+normalisation (Baker \emph{et al.}, 2007; Dougherty \emph{et al.}, 2019)
+to begin even at this relatively early stage. Frequency tagging inputs
+to different eyes or in different precortical pathways allows us to
+address neurons in different parts of the LGN. These techniques combined
+with advances in recording technology (for example, sensitive
+source-imaged EEG and MEG recordings that can resolve subcortical
+structures (Tesche, 1996; Attal \emph{et al.}, 2012; Attal \& Schwartz,
+2013), noninvasive deep-brain stimulation techniques (Mohammadjavadi
+\emph{et al.}, 2022) or implanted electrode arrays (Krolak-Salmon
+\emph{et al.}, 2003)) may allow us to study these computations in more
+detail.
+
+Finally, SSVEPs continue to provide a way of studying contrast in the
+cortex. Here, we are often interested in how different visual parameters
+interact. For example, how are signals from different eyes combined to
+generate both scalar contrast values and also binocular depth cues? How
+does contrast combination depend on the low level properties of the
+individual inputs such as their retinotopic location, cone contrast,
+spatiotemporal frequency, eye of origin and orientation? Colour vision
+scientists are particularly interested in how chromatic signals
+originating in a small number of cone-opponent retinal pathways are
+transformed into a perceptual colour space where the `unique hues'
+appear to be only weakly-related to the early retinal outputs and how
+these computations are conditioned by the spatial and temporal
+properties of the scene (Wandell, 1993; Gegenfurtner, 2003; Solomon \&
+Lennie, 2007; Stoughton \& Conway, 2008; Kaneko \emph{et al.}, 2020; Li
+\emph{et al.}, 2022). All of these questions can be addressed by using
+SSVEP and frequency tagging to examine the computations that combine and
+transform the inputs at different cortical stages (Busse \emph{et al.},
+2009; Baker \& Wade, 2017; Katyal \emph{et al.}, 2018; Chen \&
+Gegenfurtner, 2021; Watts \emph{et al.}, 2024). The original promise of
+the SSVEP approach was that the entire complex-valued frequency spectrum
+recorded at each lcoation provided detailed information about the
+processing nonlinearities up to and including that point. In principle,
+this would allow researchers to uniquely fit the parameters of their
+computations models of visual processing (Regan \& Regan, 1988). To
+date, the complexity of the neuronal computations at even the earliest
+stage of visual processing have hampered this effort - researchers
+typically restrict their analyses to the amplitudes of single, low-order
+frequency components (for example, the inputs frequencies or simple sums
+and differences of those frequencies). As our understanding of the early
+visual system improves, it is becoming possible to generate more
+realistic parameterised forward models of signal generation (Tsai
+\emph{et al.}, 2012; Baker \& Wade, 2017; Chariker \emph{et al.}, 2020;
+Schrimpf \emph{et al.}, 2020; Groen \emph{et al.}, 2022). Feeding
+frequency-tagged inputs into these types of model allows us to generate
+synthetic SSVEP responses that can be compared with those measured in
+human subject. In principle, we are therefore able to use the SSVEP to
+fit the parameters of early visual processing. This approach may also
+allow us to develop more sensitive tests for the changes in early visual
+processing that accompany a wide range of neurological diseases and
+disorders.
+
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 
@@
 
+\bibitem[\citeproctext]{ref-Attal2012}
+Attal Y, Maess B, Friederici A \& David O (2012).
+\href{https://doi.org/10.1515/rns.2011.056}{Head models and dynamic
+causal modeling of subcortical activity using
+magnetoencephalographic/electroencephalographic data}. \emph{Rev
+Neurosci} \textbf{23,} 85--95.
+
+\bibitem[\citeproctext]{ref-Attal2013}
+Attal Y \& Schwartz D (2013).
+\href{https://doi.org/10.1371/journal.pone.0059856}{Assessment of
+subcortical source localization using deep brain activity imaging model
+with minimum norm operators: A MEG study}. \emph{PLoS One} \textbf{8,}
+e59856.
+
+\bibitem[\citeproctext]{ref-Ayaz2013}
+Ayaz A, Saleem AB, Schölvinck ML \& Carandini M (2013).
+\href{https://doi.org/10.1016/j.cub.2013.04.012}{Locomotion controls
+spatial integration in mouse visual cortex}. \emph{Curr Biol}
+\textbf{23,} 890--894.
+
 \bibitem[\citeproctext]{ref-Bach2020}
@@
 
+\bibitem[\citeproctext]{ref-Benjamin2018}
+Benjamin AV, Wailes-Newson K, Ma-Wyatt A, Baker DH \& Wade AR (2018).
+\href{https://doi.org/10.1523/JNEUROSCI.1428-17.2017}{The {Effect} of
+{Locomotion} on {Early} {Visual} {Contrast} {Processing} in {Humans}}.
+\emph{Journal of Neuroscience} \textbf{38,} 3050--3059.
+
 \bibitem[\citeproctext]{ref-Betts2005}
@@
 \emph{Nat Neurosci} \textbf{2,} 370--374.
+
+\bibitem[\citeproctext]{ref-Briggs2007}
+Briggs F \& Usrey WM (2007).
+\href{https://doi.org/10.1523/JNEUROSCI.1035-07.2007}{A fast, reciprocal
+pathway between the lateral geniculate nucleus and visual cortex in the
+macaque monkey}. \emph{J Neurosci} \textbf{27,} 5431--5436.
 
@@
 
+\bibitem[\citeproctext]{ref-Cao2019}
+Cao L \& Händel B (2019).
+\href{https://doi.org/10.1371/journal.pbio.3000511}{Walking enhances
+peripheral visual processing in humans}. \emph{PLoS Biol} \textbf{17,}
+e3000511.
+
 \bibitem[\citeproctext]{ref-Carandini1994}
@@
 
+\bibitem[\citeproctext]{ref-Chariker2020}
+Chariker L, Shapley R \& Young L-S (2020).
+\href{https://doi.org/10.1167/jov.20.4.16}{Contrast response in a
+comprehensive network model of macaque V1}. \emph{J Vis} \textbf{20,}
+16.
+
 \bibitem[\citeproctext]{ref-Chen2000}
@@
 measurements}. \emph{Vision Res} \textbf{40,} 773--788.
+
+\bibitem[\citeproctext]{ref-Chen2021}
+Chen J \& Gegenfurtner KR (2021).
+\href{https://doi.org/10.1167/jov.21.8.12}{Electrophysiological evidence
+for higher-level chromatic mechanisms in humans}. \emph{J Vis}
+\textbf{21,} 12.
+
+\bibitem[\citeproctext]{ref-Chota2024}
+Chota S, Bruat AT, Van der Stigchel S \& Strauch C (2024).
+\href{https://doi.org/10.1162/jocn_a_02107}{Steady-state visual evoked
+potentials reveal dynamic (re)allocation of spatial attention during
+maintenance and utilization of visual working memory}. \emph{J Cogn
+Neurosci} \textbf{36,} 800--814.
 
@@
 
+\bibitem[\citeproctext]{ref-Diez2024}
+Diez P, Orosco L, Garcés Correa A \& Carmona L (2024).
+\href{https://doi.org/10.1007/s11517-023-03000-z}{Assessment of visual
+fatigue in SSVEP-based brain-computer interface: A comprehensive study}.
+\emph{Med Biol Eng Comput} \textbf{62,} 1475--1490.
+
 \bibitem[\citeproctext]{ref-Ding2006}
@@
 
+\bibitem[\citeproctext]{ref-Dougherty2019}
+Dougherty K, Schmid MC \& Maier A (2019).
+\href{https://doi.org/10.1002/cne.24417}{Binocular response modulation
+in the lateral geniculate nucleus}. \emph{J Comp Neurol} \textbf{527,}
+522--534.
+
 \bibitem[\citeproctext]{ref-Du2023}
@@
 
+\bibitem[\citeproctext]{ref-Gegenfurtner2003}
+Gegenfurtner KR (2003). \href{https://doi.org/10.1038/nrn1138}{Cortical
+mechanisms of colour vision}. \emph{Nat Rev Neurosci} \textbf{4,}
+563--572.
+
+\bibitem[\citeproctext]{ref-Gollisch2010}
+Gollisch T \& Meister M (2010).
+\href{https://doi.org/10.1016/j.neuron.2009.12.009}{Eye smarter than
+scientists believed: Neural computations in circuits of the retina}.
+\emph{Neuron} \textbf{65,} 150--164.
+
 \bibitem[\citeproctext]{ref-Gould2007}
@@
 detection}. \emph{J Vis} \textbf{7,} 4.1--17.
+
+\bibitem[\citeproctext]{ref-Gouws2014}
+Gouws AD, Alvarez I, Watson DM, Uesaki M, Rodgers J \& Morland AB
+(2014). \href{https://doi.org/10.1523/JNEUROSCI.0164-14.2014}{On the
+role of suppression in spatial attention: Evidence from negative BOLD in
+human subcortical and cortical structures}. \emph{J Neurosci}
+\textbf{34,} 10347--10360.
 
@@
 
+\bibitem[\citeproctext]{ref-Herrmann2001}
+Herrmann CS (2001). \href{https://doi.org/10.1007/s002210100682}{Human
+EEG responses to 1-100 hz flicker: Resonance phenomena in visual cortex
+and their potential correlation to cognitive phenomena}. \emph{Exp Brain
+Res} \textbf{137,} 346--353.
+
 \bibitem[\citeproctext]{ref-Highsmith2010}
@@
 
+\bibitem[\citeproctext]{ref-Kaneko2020}
+Kaneko S, Kuriki I \& Andersen SK (2020).
+\href{https://doi.org/10.1093/texcom/tgaa059}{Steady-state visual evoked
+potentials elicited from early visual cortex reflect both perceptual
+color space and cone-opponent mechanisms}. \emph{Cereb Cortex Commun}
+\textbf{1,} tgaa059.
+
 \bibitem[\citeproctext]{ref-Kaplan1986}
@@
 visual stimulation}. \emph{Neuron} \textbf{22,} 751--761.
+
+\bibitem[\citeproctext]{ref-Katyal2018}
+Katyal S, Vergeer M, He S, He B \& Engel SA (2018).
+\href{https://doi.org/10.1038/s41598-018-19809-w}{Conflict-sensitive
+neurons gate interocular suppression in human visual cortex}. \emph{Sci
+Rep} \textbf{8,} 1239.
+
+\bibitem[\citeproctext]{ref-Krolak-Salmon2003}
+Krolak-Salmon P, Hénaff M-A, Tallon-Baudry C, Yvert B, Guénot M,
+Vighetto A, Mauguière F \& Bertrand O (2003).
+\href{https://doi.org/10.1002/ana.10403}{Human lateral geniculate
+nucleus and visual cortex respond to screen flicker}. \emph{Ann Neurol}
+\textbf{53,} 73--80.
 
@@
 
+\bibitem[\citeproctext]{ref-Li2022}
+Li P, Garg AK, Zhang LA, Rashid MS \& Callaway EM (2022).
+\href{https://doi.org/10.1038/s41467-022-34020-2}{Cone opponent
+functional domains in primary visual cortex combine signals for color
+appearance mechanisms}. \emph{Nat Commun} \textbf{13,} 6344.
+
+\bibitem[\citeproctext]{ref-Li2014}
+Li PH, Field GD, Greschner M, Ahn D, Gunning DE, Mathieson K, Sher A,
+Litke AM \& Chichilnisky EJ (2014).
+\href{https://doi.org/10.1016/j.neuron.2013.10.043}{Retinal
+representation of the elementary visual signal}. \emph{Neuron}
+\textbf{81,} 130--139.
+
 \bibitem[\citeproctext]{ref-Li2008}
@@
 
+\bibitem[\citeproctext]{ref-Lissa2020}
+Lissa P de, Caldara R, Nicholls V \& Miellet S (2020).
+\href{https://doi.org/10.1371/journal.pone.0236967}{In pursuit of visual
+attention: SSVEP frequency-tagging moving targets}. \emph{PLoS One}
+\textbf{15,} e0236967.
+
 \bibitem[\citeproctext]{ref-Lu1998}
@@
 Press.
+
+\bibitem[\citeproctext]{ref-Middendorf2000}
+Middendorf M, McMillan G, Calhoun G \& Jones KS (2000).
+\href{https://doi.org/10.1109/86.847819}{Brain-computer interfaces based
+on the steady-state visual-evoked response}. \emph{IEEE Trans Rehabil
+Eng} \textbf{8,} 211--214.
+
+\bibitem[\citeproctext]{ref-Minarik2023}
+Minarik T, Berger B \& Jensen O (2023).
+\href{https://doi.org/10.1016/j.neuroimage.2023.120389}{Optimal
+parameters for rapid (invisible) frequency tagging using MEG}.
+\emph{Neuroimage} \textbf{281,} 120389.
+
+\bibitem[\citeproctext]{ref-Mohammadjavadi2022}
+Mohammadjavadi M, Ash RT, Li N, Gaur P, Kubanek J, Saenz Y, Glover GH,
+Popelka GR, Norcia AM \& Pauly KB (2022).
+\href{https://doi.org/10.1038/s41598-022-20554-4}{Transcranial
+ultrasound neuromodulation of the thalamic visual pathway in a large
+animal model and the dose-response relationship with MR-ARFI}. \emph{Sci
+Rep} \textbf{12,} 19588.
+
+\bibitem[\citeproctext]{ref-Morgan1996}
+Morgan ST, Hansen JC \& Hillyard SA (1996).
+\href{https://doi.org/10.1073/pnas.93.10.4770}{Selective attention to
+stimulus location modulates the steady-state visual evoked potential}.
+\emph{Proc Natl Acad Sci U S A} \textbf{93,} 4770--4774.
 
@@
 
+\bibitem[\citeproctext]{ref-Muller2000}
+Müller MM \& Hillyard S (2000).
+\href{https://doi.org/10.1016/s1388-2457(00)00371-0}{Concurrent
+recording of steady-state and transient event-related potentials as
+indices of visual-spatial selective attention}. \emph{Clin Neurophysiol}
+\textbf{111,} 1544--1552.
+
+\bibitem[\citeproctext]{ref-Muller1998}
+Müller MM, Teder-Sälejärvi W \& Hillyard SA (1998).
+\href{https://doi.org/10.1038/2865}{The time course of cortical
+facilitation during cued shifts of spatial attention}. \emph{Nat
+Neurosci} \textbf{1,} 631--634.
+
 \bibitem[\citeproctext]{ref-Murray2008}
@@
 independent}. \emph{J Vis} \textbf{8,} 2.1--11.
+
+\bibitem[\citeproctext]{ref-Niell2010}
+Niell CM \& Stryker MP (2010).
+\href{https://doi.org/10.1016/j.neuron.2010.01.033}{Modulation of visual
+responses by behavioral state in mouse visual cortex}. \emph{Neuron}
+\textbf{65,} 472--479.
 
@@
 
+\bibitem[\citeproctext]{ref-OConnor2002}
+O'Connor DH, Fukui MM, Pinsk MA \& Kastner S (2002).
+\href{https://doi.org/10.1038/nn957}{Attention modulates responses in
+the human lateral geniculate nucleus}. \emph{Nat Neurosci} \textbf{5,}
+1203--1209.
+
 \bibitem[\citeproctext]{ref-Pei2014}
@@
 
+\bibitem[\citeproctext]{ref-Schrimpf2020}
+Schrimpf M, Kubilius J, Lee MJ, Ratan Murty NA, Ajemian R \& DiCarlo JJ
+(2020). \href{https://doi.org/10.1016/j.neuron.2020.07.040}{Integrative
+benchmarking to advance neurally mechanistic models of human
+intelligence}. \emph{Neuron} \textbf{108,} 413--423.
+
+\bibitem[\citeproctext]{ref-Seijdel2023}
+Seijdel N, Marshall TR \& Drijvers L (2023).
+\href{https://doi.org/10.1093/cercor/bhac160}{Rapid invisible frequency
+tagging (RIFT): A promising technique to study neural and cognitive
+processing using naturalistic paradigms}. \emph{Cereb Cortex}
+\textbf{33,} 1626--1629.
+
 \bibitem[\citeproctext]{ref-Shapley1980}
@@
 
+\bibitem[\citeproctext]{ref-Sherman1996}
+Sherman SM \& Guillery RW (1996).
+\href{https://doi.org/10.1152/jn.1996.76.3.1367}{Functional organization
+of thalamocortical relays}. \emph{J Neurophysiol} \textbf{76,}
+1367--1395.
+
 \bibitem[\citeproctext]{ref-Shibata2008}
@@
 
+\bibitem[\citeproctext]{ref-Solomon2007}
+Solomon SG \& Lennie P (2007).
+\href{https://doi.org/10.1038/nrn2094}{The machinery of colour vision}.
+\emph{Nat Rev Neurosci} \textbf{8,} 276--286.
+
 \bibitem[\citeproctext]{ref-Somers1999}
@@
 
+\bibitem[\citeproctext]{ref-Stoughton2008}
+Stoughton CM \& Conway BR (2008).
+\href{https://doi.org/10.1016/j.cub.2008.06.018}{Neural basis for unique
+hues}. \emph{Curr Biol} \textbf{18,} R698--9.
+
 \bibitem[\citeproctext]{ref-Tadin2003}
@@
 \textbf{424,} 312--315.
+
+\bibitem[\citeproctext]{ref-Tesche1996}
+Tesche CD (1996).
+\href{https://www.ncbi.nlm.nih.gov/pubmed/8876995}{Non-invasive imaging
+of neuronal population dynamics in human thalamus}. \emph{Brain Res}
+\textbf{729,} 253--258.
 
@@
 
+\bibitem[\citeproctext]{ref-Uprety2022}
+Uprety S, Adhikari P, Feigl B \& Zele AJ (2022). Melanopsin
+photoreception differentially modulates rod-mediated and cone-mediated
+human temporal vision. \emph{Iscience}.
+
 \bibitem[\citeproctext]{ref-Vergeer2018}
@@
 
+\bibitem[\citeproctext]{ref-Wandell1993}
+Wandell BA (1993). \href{https://doi.org/10.1073/pnas.90.21.9778}{Color
+appearance: The effects of illumination and spatial pattern}. \emph{Proc
+Natl Acad Sci U S A} \textbf{90,} 9778--9784.
+
+\bibitem[\citeproctext]{ref-Wang2023}
+Wang AYM, Kulkarni MM, McLaughlin AJ, Gayet J, Smith BE, Hauptschein M,
+McHugh CF, Yao YY \& Puthussery T (2023).
+\href{https://doi.org/10.1038/s41586-023-06659-4}{An ON-type
+direction-selective ganglion cell in primate retina}. \emph{Nature}
+\textbf{623,} 381--386.
+
 \bibitem[\citeproctext]{ref-Wang2011}
@@
 \emph{Journal of Vision} \textbf{11,} 1.
+
+\bibitem[\citeproctext]{ref-Watts2024}
+Watts +Dylan+J, Rozman +Ana, Somers +Lucy+P, Gunel +Bora, Racey +Chris,
+Barnes +Katie \& Bosten +Jenny+M (2024). Tuning of cortical color
+mechanism revealed using steady-state visually evoked potentials.
+\emph{bioRxiv}.
 

### 2024-05-09 attentionGratings.py
Code to model SSVEp attention

@@ -1,0 +1,245 @@
+#%%
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Circle
+
+def hyper_ratio(params, centerContrast, surroundContrast):
+    # Returns a hyperbolic ratio of the form Rzero + Rmax * Cc^n / (Cc^n + Cs^n + c50^n)
+    # Input params are [Rmax, c50, n, Rzero]
+    # Either the center or the surround contrast or both can vary
+
+    Rmax, c50, n, Rzero = params
+
+    Cc = np.atleast_1d(centerContrast)
+    Cs = np.atleast_1d(surroundContrast)
+
+    # Force inputs to be between 0 and 1
+    
+    Cc[np.where(Cc>1)] = 1
+    Cc[np.where(Cc<0)] = 0
+    Cs[np.where(Cs>1)] = 1
+    Cs[np.where(Cs<0)] = 0
+
+    if np.any(Cc < 0) or np.any(Cc > 1):
+        raise ValueError("Center contrast must be in the range [0,1]")
+    if np.any(Cs < 0) or np.any(Cs > 1):
+        raise ValueError("Surround contrast must be in the range [0,1]")
+    
+    # Ensure Cc and Cs are broadcastable to the same shape : Broadly, are they both vectors of the same shape?
+    # Or is one a scalar and the other a vector?
+    if Cc.shape != Cs.shape:
+        if Cc.size == 1:
+            Cc = np.full(Cs.shape, Cc)
+        elif Cs.size == 1:
+            Cs = np.full(Cc.shape, Cs)
+            
+        else:
+            raise ValueError("Center and surround contrast arrays must be broadcastable to the same shape")
+        
+    # Validate parameters
+    if Rmax <= 0:
+        raise ValueError("Rmax must be positive")
+    if n <= 0:
+        raise ValueError("n must be positive")
+    if Rzero < 0:
+        raise ValueError("Rzero must be non-negative")
+    if c50 <= 0:
+        raise ValueError("c50 must be positive")
+
+    # Calculate the hyperbolic ratio
+   
+    response = Rmax * (Cc**n) / ((Cc)**n+ c50**n+Cs**n)
+    return response
+
+def make_circular_grating(image_size, frequency, radius=None, phase_shift=0, orientation=0,contrast=.5):
+    """
+    Creates a sinusoidal grating of a given size, frequency, phase shift, and orientation.
+    The grating can be masked by a circle of a specified radius.
+
+    Args:
[…]

### 2024-05-10 review.qmd
No changes in text
Fixed a few minor typos

@@
 
-Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolaKting back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
+Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
 This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
@@
 
-A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrasr levels.
+A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrast levels.
 This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979, @Norcia1985a; @Norcia1990].
+To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979; @Norcia1985a; @Norcia1990].
 The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
@@
 The authors theorised that differences in early gain control during development might lead to visual deficits later in life.
-Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms (tremor, rigidity, slow movement) manifest.
+Although visual responses are a convenient assay of neural function, it is likely that the same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms (tremor, rigidity, slow movement) manifest.
 The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments.
@@
 
-Attention exerts a profound influence on visual performance: for example, allowing people to attend covertly to a spatial location improves their performance on a target detection task significantly [@Posner1980; @Cameron2002; @Carrasco2000; @Bashinski1980; @Morrone2004; @Pestilli2009].
+Attention exerts a profound influence on visual performance: for example, instructing people to attend covertly to a spatial location improves their performance on a target detection task significantly [@Posner1980; @Cameron2002; @Carrasco2000; @Bashinski1980; @Morrone2004; @Pestilli2009].
 In principle, this enhancement may be driven both by modulation of the underlying signal or noise characteristics, or by additional decision-theoretic factors such as reduction in spatial uncertainty [@Gould2007; @Petrov2006].
@@
 The high SNR and precise frequency resolution of SSVEPs make them an ideal candidate for this approach.
-Typical studies may involve presenting an array of stimuli at different contrast flicker frequencies, and having the participant select one either by overt attention (i.e.
+Typical studies may involve presenting an array of stimuli at different flicker frequencies, and having the participant select one either by overt attention (i.e.
 shifting fixation to foveate the selected stimulus) or covert attention (i.e.
@@
 This approach is primarily useful in situations that require the BCI to distinguish from among a small set of possibilties: for example, in early work, visual stimuli representing 'Left' and 'Right' commands in a flight simulator were distinguished robustly [@Middendorf2000].
-Although SSVEP-based BCI interfaces typically do use contrast flicker, work in this field has largely focused on optimising the stimuli or decoders to increase decoding performance (e.g.
-...
-) and reducing visual fatigue associated with the long-term presentation of arrays of high-contrast flicker [@Diez2024] rather than studying visual contrast processing \textit{per se}.
+Although SSVEP-based BCI interfaces typically do use contrast flicker, work in this field has largely focused on optimising the stimuli or decoders to increase decoding performance and reducing visual fatigue associated with the long-term presentation of arrays of high-contrast flicker [@Diez2024], rather than studying visual contrast processing \textit{per se}.
 We therefore note in passing that using SSVEPs to improve our understanding of early contrast processing may yield benefits to this related field.
@@
 # Future directions
+
 The SSVEP is a powerful tool for studying contrast processing.
@@
 In principle, each nonlinearity in the visual pathway can be identified by its contribution to the frequency spectrum at different recording locations [@Regan1988].
-This, it turn, allows researchers to study how contrast processing depends on spatial and temporal context and as well as changes in task, cognitive and behavioural state and arousal.
+This, it turn, allows researchers to study how contrast processing depends on spatial and temporal context, as well as changes in task, cognitive and behavioural state, and arousal.
 
@@
 All of these questions can be addressed by using SSVEP and frequency tagging to examine the computations that combine and transform the inputs at different cortical stages [@Watts2024; @Baker2017; @Chen2021; @Katyal2018; @Busse2009].
-The original promise of the SSVEP approach was that the  entire complex-valued frequency spectrum recorded at each lcoation provided detailed information about the processing nonlinearities up to and including that point.
-In principle, this would allow researchers to uniquely fit the parameters of their computations models of visual processing [@Regan1988].
-To date, the complexity of the neuronal computations at even the earliest stage of visual processing have hampered this effort - researchers typically restrict their analyses to the amplitudes of single, low-order frequency components (for example, the inputs frequencies or simple sums and differences of those frequencies).
+The original promise of the SSVEP approach was that the  entire complex-valued frequency spectrum recorded at each location provided detailed information about the processing nonlinearities up to and including that point.
+In principle, this would allow researchers to uniquely fit the parameters of their computational models of visual processing [@Regan1988].
+To date, the complexity of the neuronal computations at even the earliest stage of visual processing have hampered this effort - researchers typically restrict their analyses to the amplitudes of single, low-order frequency components (for example, the input frequencies or simple sums and differences of those frequencies).
 As our understanding of the early visual system improves, it is becoming possible to generate more realistic parameterised forward models of signal generation [@Chariker2020; @Groen2022; @Baker2017; @Tsai2012; @Schrimpf2020].
-Feeding frequency-tagged inputs into these types of model allows us to generate synthetic SSVEP responses that can be compared with those measured in human subject.
+Feeding frequency-tagged inputs into these types of model allows us to generate synthetic SSVEP responses that can be compared with those measured in human subjects.
 In principle, we are therefore able to use the SSVEP to fit the parameters of early visual processing.

### 2024-05-10 review.qmd
Added link to github repo

@@
 This normalization, achieved through a computation called 'contrast gain control' [@Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
-As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround).
+As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround; note that the code used to produce all figures in this review is available in python and R at: https://github.com/wadelab/contrastReviewPaper).
 

### 2024-05-13 review.qmd
Micro edits

@@
 Contrast is the currency of the early visual system.
-Measuring the way that the computations underlying contrast processing depend on factors such as spatial and temporal frequency, age, clinical conditions, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years.
+Measuring the way that the computations underlying contrast processing depend on factors such as spatial and temporal frequency, age, clinical conditions, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for over a century.
 One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged' (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
@@
 Neurons in the visual areas of the brain are primarily responsive to changes in cone photoreceptor activations across time and space.
-This property, referred to as contrast, sets the fundamental limits of our visual abilities, which remain steady over a remarkably wide range of environmental light levels.
+This property, referred to as 'contrast', sets the fundamental limits of our visual abilities, which remain steady over a remarkably wide range of environmental light levels.
 The human response to contrast can be studied using many different techniques.
@@
 It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
-Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@McKeefry1996; @Baseler1997; @Di-Russo2001].
+Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@McKeefry1996; @Di-Russo2001]   (see also [@Baseler1997]).
 
-Contrast is relatively simple to define: typically, it is specified as the percentage deviation of a uniform stimulus from the background.
+Contrast is typically specified as the percentage deviation of a uniform stimulus from the background.
 So, for example, a disk of 100 units of cone activation ($I_{\mathrm{stim}}$) surrounded by a 'background' of 50 units of activation ($I_{\mathrm{background}}$) has a contrast of $\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}$ = 100%.
@@
 Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction.
-For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response' function can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency, or a low multiple thereof.
+For example, by varying the peak stimulus contrast parametrically, a 'contrast vs response'  function (CRF) can be measured - where the 'response' is typically defined as the amplitude of the SSVEP frequency component at the stimulus frequency, or a low multiple thereof.
 This corresponds closely to similar functions reported by studies measuring single unit activity or local field potentials in the cortex [@Shapley1980; @Morrone1982].
@@
 Neurons have a limited dynamic range, yet they can transmit information about visual stimuli that span many orders of magnitude.
-In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear, sigmoidal contrast vs response (CRF) transducer [@Albrecht1982; @Tolhurst1981] and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Carandini1994; @Carandini1998; @Busse2009].
+In the domain of contrast, to some extent this is accomplished at a population level - individual neurons typically implement a non-linear, sigmoidal CRF transducer [@Albrecht1982; @Tolhurst1981] and different neurons exhibit peak sensitivity (defined as the maximum slope of the function) at different contrast levels [@Carandini1994; @Carandini1998; @Busse2009].
 A neuronal population will therefore span a sensitivity range greater than any individual member.
@@
 
-To better understand these processes, we will show how sinusoidal input signals might be processed by the visual system to produce SSVEPs.
+To  understand these processes better, we will show how sinusoidal input signals might be processed by the visual system to produce SSVEPs.
 @fig-linear illustrates how sine waves of different contrasts are processed in a linear system.
@@
 The second panel shows Fourier transform of the waveform, which contains a substantial peak at this frequency.
-If we parametrically vary the stimulus contrast (i.e.
+If we  change the stimulus contrast (i.e.
 the amplitude of the waveform), the amplitude of the F1 component increases linearly with contrast (right panel).
@@
 This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.
-Our own work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
+Our more recent work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
 More detailed modelling of intracortical recordings [@Groen2022] has revealed details of the timecourse of gain control effects, specifically that normalization is delayed slightly relative to the initial visual response.
@@
 This can often be informative regarding the underlying mechanism that characterises the condition.
-Here we focus on four conditions, but there is potential to apply the method more broadly, either as a diagnostic technique, or to monitor disease severity and progression, or the efficacy of treatments.
+Here we focus on four conditions, but there is potential to apply the method more broadly: as a dignostic technique, to monitor disease severity and progression, or to assess the efficacy of treatments.
 
@@
 One widespread recent application of the SSVEP technique is in the design of brain-computer interfaces, which seek to control some aspect of a computer using neural signals.
-The high SNR and precise frequency resolution of SSVEPs make them an ideal candidate for this approach.
+The high SNR and precise frequency resolution of SSVEPs are advantages for this approach.
 Typical studies may involve presenting an array of stimuli at different flicker frequencies, and having the participant select one either by overt attention (i.e.

### 2024-05-13 Rversion/reviewR.qmd
Added R version in subfolder

--- 2024-05-13 review.qmd
@@
 
-```{python setup}
+```{r setup}
 
-# all package loading and subfunctions defined here
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib import colormaps
-import scipy as scp
-from sklearn.linear_model import LinearRegression
+# set up helper functions
 
-def simpleTransducerFunction(inputContrast, c50, RMax):
+simpleTransducerFunction <- function(inputContrast, c50, RMax){
  # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
@@
  # The exponent can also vary - here we fix it to 2
-  expnt=2
+  expnt <- 2
 
@@
   # contrast detectors we perform full-wave rectification of the signal
-  inputContrast=np.abs(inputContrast)
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
-  return output
+  inputContrast <- abs(inputContrast)
+  output <- RMax*(inputContrast^expnt)/(inputContrast^expnt + c50^expnt)
+  return(output)
+}
 
-def doubleTransducerFunction(inputContrast, maskContrast, c50, RMax):
+doubleTransducerFunction <- function(inputContrast, maskContrast, c50, RMax){
  # This version has two inputs, but the second input is purely suppressive
-  expnt=2
-  inputContrast=np.abs(inputContrast)
-  maskContrast=np.abs(maskContrast)
+  expnt <- 2
+  inputContrast <- abs(inputContrast)
+  maskContrast <- abs(maskContrast)
 
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
-  return output
+  output <- RMax*(inputContrast^expnt)/(inputContrast^expnt + maskContrast^expnt + c50^expnt)
+  return(output)  
+}
 
-def create_grating(size, frequency, phase, contrast):
-    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
-    grating = np.sin(2 * np.pi * frequency * np.sqrt(x**2 + y**2) + phase)
-    grating = (grating + 1) / 2  # Normalize to [0, 1]
-    grating = grating * contrast + (1 - contrast) / 2  # Adjust contrast
-    return grating
+create_grating <- function(size, frequency, phase, contrast){
+    x <- matrix(rep(seq(-1,1,length=size),size),nrow=size,ncol=size)
+    y <- t(x)
+    grating <- sin(2 * pi * frequency * sqrt(x^2 + y^2) + phase)
+    grating <- (grating + 1) / 2  # Normalize to [0, 1]
+    grating <- grating * contrast + (1 - contrast) / 2  # Adjust contrast 
+  return(grating)
+}
 
-def create_circular_mask(size):
-    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
-    mask = np.sqrt(x**2 + y**2) <= 0.3
-    return mask
+create_circular_mask <- function(size){
+    x <- matrix(rep(seq(-1,1,length=size),size),nrow=size,ncol=size)
+    y <- t(x)
+    radius <- sqrt(x^2 + y^2)
+    mask <- sqrt(x**2 + y**2) <= 0.3
+  return(mask)
+}
+
+if (!dir.exists('/Figures')){dir.create('/Figures')}
 
@@
 
-```{python createcentresurround}
+```{r createcentresurround}
 
 # Set parameters
-size = 256
-frequency = 10
-phase = 0
-contrast_center = 0.5
-contrast_surround_low = 0.1
-contrast_surround_high = 0.9
+size <- 256
+frequency <- 10
+phase <- 0
+contrast_center <- 0.5
+contrast_surround_low <- 0.1
+contrast_surround_high <- 0.9
 
 # Create gratings and circular masks
-grating_center = create_grating(size, frequency, phase, contrast_center)
-grating_surround_low = create_grating(size, frequency, phase, contrast_surround_low)
-grating_surround_high = create_grating(size, frequency, phase, contrast_surround_high)
-mask = create_circular_mask(size)
+grating_center <- create_grating(size, frequency, phase, contrast_center)
+grating_surround_low <- create_grating(size, frequency, phase, contrast_surround_low)
+grating_surround_high <- create_grating(size, frequency, phase, contrast_surround_high)
+mask <- create_circular_mask(size)
 
 # Combine gratings and masks
-image_low_contrast = grating_surround_low.copy()
-image_low_contrast[mask] = grating_center[mask]
-image_high_contrast = grating_surround_high.copy()
-image_high_contrast[mask] = grating_center[mask]
+image_low_contrast <- grating_surround_low
+image_low_contrast[mask] <- grating_center[mask]
+image_high_contrast <- grating_surround_high
+image_high_contrast[mask] <- grating_center[mask]
 
-# Create subplots
-fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(7, 3))
+pdf('Figures/centresurround.pdf',width=12,height=6.5)
 
-# Plot low-contrast background
-ax1.imshow(image_low_contrast, cmap='gray', vmin=0, vmax=1)
-ax1.axis('off')
-ax1.set_title('Low-Contrast Background', )
+par(mfcol=c(1,2))
 
-# Plot high-contrast background
-ax2.imshow(image_high_contrast, cmap='gray', vmin=0, vmax=1)
-ax2.axis('off')
-ax2.set_title('High-Contrast Background')
+plot(x=NULL,y=NULL,xlim=c(0,1),ylim=c(0,1),axes=FALSE, ann=FALSE)
+image(image_low_contrast,col=gray.colors(256),add=TRUE,useRaster=TRUE)
+title(main='Low-Contrast Background')
 
-plt.tight_layout()
-plt.savefig('Figures/centresurround.pdf')
+plot(x=NULL,y=NULL,xlim=c(0,1),ylim=c(0,1),axes=FALSE, ann=FALSE)
+image(image_high_contrast,col=gray.colors(256),add=TRUE,useRaster=TRUE)
+title(main='High-Contrast Background')
+
+dev.off()
 
@@
 
-```{python linearsystem}
+```{r linearsystem}
 
-inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
-contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
-duration_S=1 # All variables in this code have an _[units] where possible
-digitisationRate_Hz=1000
-freqCutoff_Hz=30
-c50=.10
-RMax=10
+inputFrequency_Hz <- 5 # We simulate a system with on input at many different contrasts
+contrastRange <- (10^seq(0,1.5,length=20))/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S <- 1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz <- 1000
+freqCutoff_Hz <- 30
+c50 <- 0.10
+RMax <- 10
 
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+support_S <- seq(0,duration_S,length=duration_S*digitisationRate_Hz)
 
-initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
-transducedSignal=[]
+initialSignal <- NULL # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal <- NULL
 
-initialSignal = np.zeros(len(contrastRange)) 
-transducedSignal = np.zeros(len(contrastRange))
-for contrastIndex,thisContrast in enumerate(contrastRange):
-  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
+for (contrastIndex in 1:length(contrastRange)){
 
-  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+  inputModulation <- sin(2 * pi * inputFrequency_Hz * support_S) * contrastRange[contrastIndex]
 
-  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
-  initialSignal[contrastIndex]=inputAmplitude[0]
+  FT_Signal <- fft(inputModulation) / length(inputModulation)
+
+  inputAmplitude <- abs(FT_Signal[inputFrequency_Hz+1])
+  initialSignal[contrastIndex] <- inputAmplitude
 
@@
    # using the transducer functiona above.
-  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
-  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
-  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
-  transducedSignal[contrastIndex]=transducedAmplitude[0]
+  transducedModulation <- simpleTransducerFunction(inputModulation, c50, RMax)
+  FT_transducedModulation <- fft(transducedModulation)/ length(transducedModulation)
+  transducedAmplitude <- abs(FT_transducedModulation[1 + inputFrequency_Hz*2])
+  transducedSignal[contrastIndex] <- transducedAmplitude
+}
 
-frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
+frequencySupport <- seq(0,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
 
 #Make a figure with three subplots
-fig, axs = plt.subplots(1,3, figsize=(8, 3))
 
-axs[0].plot(support_S, inputModulation)
-axs[0].set_title('Input sine wave', fontsize=10)
-axs[0].set_xlabel('Time (s)')
+pdf('Figures/linearsystem.pdf',width=12,height=4)
+par(mfrow=c(1,3))
+
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(-0.3,0.3))   
+  axis(1, at=seq(0,1,0.5), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(-3,3,1)/10, tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,1,0.5), side = 1, at=seq(0,1,0.5), line=0.2)     # add the tick labels
+  mtext(text = seq(-3,3,1)/10, side = 2, at=seq(-3,3,1)/10, line=0.2, las=1)  
+title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(main="Input sine wave", cex.main=1.5)
+
+lines(support_S, inputModulation, lwd=3)
 
 # Plot the absolute values of the FFT in the second axis as a bar plot
-axs[1].bar(frequencySupport, np.abs(FT_Signal[frequencySupport]))
-axs[1].set_xlabel('Frequency (Hz)')
-axs[1].set_ylabel('Magnitude (a.u.)')
-axs[1].set_title('Fourier spectrum')
-axs[1].grid()
 
-axs[2].plot(contrastRange*100,initialSignal)
-axs[2].set_xlabel('Contrast (%)')
-axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('Linear computation stage', fontsize=10)
-axs[2].grid()
+  plot(frequencySupport[2:length(frequencySupport)]-0.5,abs(FT_Signal[2:length(frequencySupport)]),type='s',lwd=3,axes=FALSE, ann=FALSE, xlim=c(0,30), ylim=c(0,0.16))
+  axis(1, at=seq(0,30,10), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,0.16,0.04), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,30,10), side = 1, at=seq(0,30,10), line=0.2)     # add the tick labels
+  mtext(text = seq(0,0.16,0.04), side = 2, at=seq(0,0.16,0.04), line=0.2, las=1)  
+title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(ylab="Magnitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
+title(main="Fourier spectrum", cex.main=1.5)
 
-axs[2].grid()
-plt.tight_layout()
-plt.savefig('Figures/linearsystem.pdf')
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,32), ylim=c(0,0.16))   
+  axis(1, at=seq(0,32,4), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,0.16,0.04), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,32,4), side = 1, at=seq(0,32,4), line=0.2)     # add the tick labels
+  mtext(text = seq(0,0.16,0.04), side = 2, at=seq(0,0.16,0.04), line=0.2, las=1)  
+title(xlab="Contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(ylab="Amplitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
+title(main="Linear computation stage", cex.main=1.5)
+
+lines(contrastRange*100,initialSignal,lwd=3)
+
+dev.off()
 
@@
 
-```{python transducerout}
+```{r transducerout}
 
-# Define some constants
-inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
-contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
-duration_S=1 # All variables in this code have an _[units] where possible
-digitisationRate_Hz=1000
-freqCutoff_Hz=30
-c50=.10
-RMax=10
+inputFrequency_Hz <- 5 # We simulate a system with on input at many different contrasts
+contrastRange <- (10^seq(0,1.5,length=20))/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S <- 1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz <- 1000
+freqCutoff_Hz <- 30
+c50 <- 0.10
+RMax <- 10
 
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+support_S <- seq(0,duration_S,length=duration_S*digitisationRate_Hz)
 
-initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
-transducedSignal=[]
+initialSignal <- NULL # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal <- NULL
 
-initialSignal = np.zeros(len(contrastRange)) 
-transducedSignal = np.zeros(len(contrastRange))
-for contrastIndex,thisContrast in enumerate(contrastRange):
-  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
+for (contrastIndex in 1:length(contrastRange)){
 
-  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+  inputModulation <- sin(2 * pi * inputFrequency_Hz * support_S) * contrastRange[contrastIndex]
 
-  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
-  initialSignal[contrastIndex]=inputAmplitude[0]
+  FT_Signal <- fft(inputModulation) / length(inputModulation)
+
+  inputAmplitude <- abs(FT_Signal[inputFrequency_Hz+1])
+  initialSignal[contrastIndex] <- inputAmplitude
 
@@
    # using the transducer functiona above.
-  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
-  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
-  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
-  transducedSignal[contrastIndex]=transducedAmplitude[0]
+  transducedModulation <- simpleTransducerFunction(inputModulation, c50, RMax)
+  FT_transducedModulation <- fft(transducedModulation)/ length(transducedModulation)
+  transducedAmplitude <- abs(FT_transducedModulation[1 + inputFrequency_Hz*2])
+  transducedSignal[contrastIndex] <- transducedAmplitude
+}
 
-frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
+frequencySupport <- seq(0,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
 
 #Make a figure with three subplots
-fig, axs = plt.subplots(1,3, figsize=(8, 3))
 
-axs[0].plot(support_S, transducedModulation)
-axs[0].set_title('Transduced signal', fontsize=10)
-axs[0].set_xlabel('Time (s)')
+pdf('Figures/transducer.pdf',width=12,height=4)
+par(mfrow=c(1,3))
+
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,10))   
+  axis(1, at=seq(0,1,0.5), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,10,2), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,1,0.5), side = 1, at=seq(0,1,0.5), line=0.2)     # add the tick labels
+  mtext(text = seq(0,10,2), side = 2, at=seq(0,10,2), line=0.2, las=1)  
+title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(main="Input sine wave", cex.main=1.5)
+
+lines(support_S, transducedModulation, lwd=3)
 
 # Plot the absolute values of the FFT in the second axis as a bar plot
-axs[1].bar(frequencySupport, np.abs(FT_transducedModulation[frequencySupport]))
-axs[1].set_xlabel('Frequency (Hz)')
-axs[1].set_ylabel('Magnitude (a.u.)')
-axs[1].set_title('Fourier spectrum')
-axs[1].grid()
 
-axs[2].plot(contrastRange*100,transducedSignal)
-axs[2].set_xlabel('Contrast (%)')
-axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('Transducer output', fontsize=10)
-axs[2].grid()
+  plot(frequencySupport[2:length(frequencySupport)]-0.5,abs(FT_transducedModulation[2:length(frequencySupport)]),type='s',lwd=3,axes=FALSE, ann=FALSE, xlim=c(0,30), ylim=c(0,2))
+  axis(1, at=seq(0,30,10), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,2,0.5), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,30,10), side = 1, at=seq(0,30,10), line=0.2)     # add the tick labels
+  mtext(text = seq(0,2,0.5), side = 2, at=seq(0,2,0.5), line=0.2, las=1)  
+title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(ylab="Magnitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
+title(main="Fourier spectrum", cex.main=1.5)
 
-axs[2].grid()
-plt.tight_layout()
-plt.savefig('Figures/transducer.pdf')
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,32), ylim=c(0,2))   
+  axis(1, at=seq(0,32,4), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,2,0.5), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,32,4), side = 1, at=seq(0,32,4), line=0.2)     # add the tick labels
+  mtext(text = seq(0,2,0.5), side = 2, at=seq(0,2,0.5), line=0.2, las=1)  
+title(xlab="Contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(ylab="Amplitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
+title(main="Linear computation stage", cex.main=1.5)
+
+lines(contrastRange*100,transducedSignal,lwd=3)
+
+dev.off()
 
@@
 
-```{python twoinputsout}
+```{r twoinputsout}
 
-# Define some constants
-inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
-contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
-maskFrequency_Hz=np.array([7]) # We simulate a system with on input at many different contrasts
-maskContrast=np.max(contrastRange) # contrast of mask signal
+inputFrequency_Hz <- 5 # We simulate a system with on input at many different contrasts
+contrastRange <- (10^seq(0,1.5,length=20))/100 # Equally spaced log contrast between 1 and 10^1.5
+maskFrequency_Hz <- 7
+maskContrast <- max(contrastRange)
 
-duration_S=1 # All variables in this code have an _[units] where possible
-digitisationRate_Hz=1000
-freqCutoff_Hz=30
-c50=.10
-RMax=10
+duration_S <- 1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz <- 1000
+freqCutoff_Hz <- 30
+c50 <- 0.10
+RMax <- 10
 
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+support_S <- seq(0,duration_S,length=duration_S*digitisationRate_Hz)
 
-initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
-transducedSignal=[]
+initialSignal <- NULL # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal <- NULL
+maskedSignal <- NULL
 
-initialSignal = np.zeros(len(contrastRange)) 
-transducedSignal = np.zeros(len(contrastRange))
-maskedSignal = np.zeros(len(contrastRange))
-for contrastIndex,thisContrast in enumerate(contrastRange):
-  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
-  maskModulation = np.sin(2 * np.pi * maskFrequency_Hz * support_S) * maskContrast
+for (contrastIndex in 1:length(contrastRange)){
+
+  inputModulation <- sin(2 * pi * inputFrequency_Hz * support_S) * contrastRange[contrastIndex]
+  maskModulation <- sin(2 * pi * maskFrequency_Hz * support_S) * maskContrast
 
@@
    # using the transducer functiona above.
-  transducedModulation=doubleTransducerFunction(inputModulation, 0*maskModulation, c50,RMax)
-  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
-  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
-  transducedSignal[contrastIndex]=transducedAmplitude[0]
+  transducedModulation <- doubleTransducerFunction(inputModulation, 0*maskModulation, c50,RMax)
+  FT_transducedModulation <- fft(transducedModulation)/ length(transducedModulation)
+  transducedAmplitude <- abs(FT_transducedModulation[1+inputFrequency_Hz*2])
+  transducedSignal[contrastIndex] <- transducedAmplitude
 
-  transducedModulation=doubleTransducerFunction(inputModulation, maskModulation, c50,RMax)
-  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
-  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
-  maskedSignal[contrastIndex]=transducedAmplitude[0]
+  transducedModulation <- doubleTransducerFunction(inputModulation, maskModulation, c50,RMax)
+  FT_transducedModulation <- fft(transducedModulation)/ length(transducedModulation)
+  transducedAmplitude <- abs(FT_transducedModulation[1+inputFrequency_Hz*2])
+  maskedSignal[contrastIndex] <- transducedAmplitude
+}
 
-frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
+frequencySupport <- seq(0,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
 
 #Make a figure with three subplots
-fig, axs = plt.subplots(1,3, figsize=(8, 3))
 
-axs[0].plot(support_S, transducedModulation)
-axs[0].set_title('Transduced signal', fontsize=10)
-axs[0].set_xlabel('Time (s)')
+pdf('Figures/twoinputs.pdf',width=12,height=4)
+par(mfrow=c(1,3))
+
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,10))   
+  axis(1, at=seq(0,1,0.5), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,10,2), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,1,0.5), side = 1, at=seq(0,1,0.5), line=0.2)     # add the tick labels
+  mtext(text = seq(0,10,2), side = 2, at=seq(0,10,2), line=0.2, las=1)  
+title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(main="Input sine wave", cex.main=1.5)
+
+lines(support_S, transducedModulation, lwd=3)
 
 # Plot the absolute values of the FFT in the second axis as a bar plot
-axs[1].bar(frequencySupport, np.abs(FT_transducedModulation[frequencySupport]))
-axs[1].set_xlabel('Frequency (Hz)')
-axs[1].set_ylabel('Magnitude (a.u.)')
-axs[1].set_title('Fourier spectrum')
-axs[1].grid()
 
-axs[2].plot(contrastRange*100,transducedSignal)
-axs[2].plot(contrastRange*100,maskedSignal)
-axs[2].set_xlabel('Contrast (%)')
-axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('Transducer output', fontsize=10)
-axs[2].grid()
+  plot(frequencySupport[2:length(frequencySupport)]-0.5,abs(FT_transducedModulation[2:length(frequencySupport)]),type='s',lwd=3,axes=FALSE, ann=FALSE, xlim=c(0,30), ylim=c(0,2))
+  axis(1, at=seq(0,30,10), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,2,0.5), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,30,10), side = 1, at=seq(0,30,10), line=0.2)     # add the tick labels
+  mtext(text = seq(0,2,0.5), side = 2, at=seq(0,2,0.5), line=0.2, las=1)  
+title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(ylab="Magnitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
+title(main="Fourier spectrum", cex.main=1.5)
 
-axs[2].grid()
-plt.tight_layout()
-plt.savefig('Figures/twoinputs.pdf')
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,32), ylim=c(0,2))   
+  axis(1, at=seq(0,32,4), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,2,0.5), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,32,4), side = 1, at=seq(0,32,4), line=0.2)     # add the tick labels
+  mtext(text = seq(0,2,0.5), side = 2, at=seq(0,2,0.5), line=0.2, las=1)  
+title(xlab="Contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(ylab="Amplitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
+title(main="Linear computation stage", cex.main=1.5)
+
+lines(contrastRange*100,transducedSignal,lwd=3)
+lines(contrastRange*100,maskedSignal,col='orange',lwd=3)
+
+dev.off()
 
@@
 
-```{python makesweep}
+```{r makesweep}
 
 # Define some constants
-nLevels=8
-internalNoise=1
-inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
-contrastRange=np.logspace(1,1.9,nLevels)/100 # Equally spaced log contrast between 1 and 10^1.5
-duration_S=1 # All variables in this code have an _[units] where possible
-digitisationRate_Hz=1000
-freqCutoff_Hz=30
+nLevels <- 8
+internalNoise <- 1
+inputFrequency_Hz <- 5 # We simulate a system with on input at many different contrasts
+contrastRange <- (10^seq(1,1.9,length=nLevels))/100  # Equally spaced log contrast between 1 and 10^1.9
+duration_S <- 1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz <- 1000
+freqCutoff_Hz <- 30
 
-support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+support_S <- seq(0,duration_S,length=duration_S*digitisationRate_Hz)
 
-fig, axs = plt.subplots(1,1, figsize=(4, 3))
+initialSignal <- NULL
+transducedSignal <- NULL
+for (contrastIndex in 1:length(contrastRange)){
+  inputModulation <- sin(2 * pi * inputFrequency_Hz * support_S) * contrastRange[contrastIndex] + runif(length(support_S))*internalNoise
 
-initialSignal = np.zeros(len(contrastRange))
-transducedSignal = np.zeros(len(contrastRange))
-for contrastIndex,thisContrast in enumerate(contrastRange):
-  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast + np.random.rand(len(support_S))*internalNoise
+  FT_Signal <- fft(inputModulation) / length(inputModulation)
 
-  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+  responseAmplitude <- abs(FT_Signal[inputFrequency_Hz+1])
+  initialSignal[contrastIndex] <- responseAmplitude
+  transducedSignal[contrastIndex] <- simpleTransducerFunction(responseAmplitude, 0.05, 10)
+}
+# Now fit a linear regression to the first 4 contrast levels (before it saturates) and find out where this crosses the Y-axis
+X = log10(contrastRange[1:4])
+X2 = log10((10^seq(0.1,1.1,length=nLevels))/100)
+y <- transducedSignal[1:4]
+reg <- lm(y ~ X)
+x_intercept <- 10 ^ (-reg$coefficients[1] / reg$coefficients[2])
+Yfit <- reg$coefficients[2]*X + reg$coefficients[1]
+Yextrap <- reg$coefficients[2]*X2 + reg$coefficients[1]
 
-  responseAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
-  initialSignal[contrastIndex]=responseAmplitude[0]
-  transducedSignal[contrastIndex]=simpleTransducerFunction(responseAmplitude[0], 0.05, 10)
+pdf('Figures/sweepplot.pdf',width=6,height=6)
 
-# Now fit a linear regression to the first 4 contrast levels (before it saturates) and find out where this crosses the Y-axis
-X = np.log10(contrastRange[:4]).reshape(-1, 1)
-X2 = np.log10(np.logspace(0.1,1.1,nLevels)/100).reshape(-1, 1)
-y = transducedSignal[:4]
-reg = LinearRegression().fit(X, y)
-x_intercept = 10 ** (-reg.intercept_ / reg.coef_[0])
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-2,0), ylim=c(0,10))   
+  axis(1, at=seq(-2,0,1), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,10,2), tck=0.01, lab=F, lwd=2)
+  mtext(text = c(0.01,0.1,1), side = 1, at=seq(-2,0,1), line=0.2)     # add the tick labels
+  mtext(text = seq(0,10,2), side = 2, at=seq(0,10,2), line=0.2, las=1)  
+title(xlab="Contrast", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(ylab="Amplitude (a.u.)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(main=paste("Simulated sweep VEP at",nLevels,"different contrasts"), cex.main=1.5)
 
-axs.set_title(f'Simulated sweep VEP at {nLevels} different contrasts', fontsize=10)
-axs.semilogx(contrastRange, transducedSignal, 'o')
-axs.semilogx(contrastRange[:4], reg.predict(X), 'r-', label='Linear Fit')
-axs.semilogx(np.logspace(0.1,1.1,nLevels)/100, reg.predict(X2), 'r--')
-axs.semilogx([x_intercept], [0], 'ro', label='Threshold')
-axs.set_xlabel('Contrast')
-axs.set_ylabel('Amplitude (a.u.)')
-axs.set_xlim(0.01, 1)
-axs.set_ylim(0, 10)
-axs.legend()
-axs.grid()
-plt.tight_layout()
-plt.savefig('Figures/sweepplot.pdf')
+points(log10(contrastRange),transducedSignal,pch=19,col='cornflowerblue',cex=2)
+points(log10(x_intercept),0,pch=19,col='red',cex=2)
+
+lines(X,Yfit,lwd=3,col='red')
+lines(X2,Yextrap,lwd=3,lty=2,col='red')
+
+legend(-2,10,c('Linear fit','Threshold'),lty=c(1,NA),pch=c(NA,19),lwd=3,col='red',box.lwd=2,cex=1.5)
+
+dev.off()
+
+# axs.semilogx(contrastRange[:4], reg.predict(X), 'r-', label='Linear Fit')
+# axs.semilogx(np.logspace(0.1,1.1,nLevels)/100, reg.predict(X2), 'r--')
 
@@
 
-```{python makeclinicalplot}
+```{r makeclinicalplot}
 
 # Parameters
-finecontrastsdB = np.arange(-18, 37, 1)
-finecontrasts = 10**(finecontrastsdB / 20)
-ticklocsx = np.arange(-18, 37, 6)
-ticklabelsx = ['1/8', '1/4', '1/2', 1, 2, 4, 8, 16, 32, 64]
+finecontrastsdB <- seq(-18,36,1)
+finecontrasts <- 10^(finecontrastsdB / 20)
+ticklocsx <- seq(-18,36,6)
+ticklabelsx <- c('1/8', '1/4', '1/2', 1, 2, 4, 8, 16, 32, 64)
 
-epilepsyparams = np.array([
-    [3.4, 0.15, 1.4, 1],  # controls
-    [5.3, 0.23, 1.4, 1]   # patients
-])
+pdf('Figures/clinicalplot.pdf',width=12,height=12)
+par(mfrow=c(2,2))
+
+epilepsyparams <- matrix(0,nrow=2,ncol=4) 
+epilepsyparams[1,] <- c(3.4, 0.15, 1.4, 1)  # controls
+epilepsyparams[2,] <- c(5.3, 0.23, 1.4, 1)  # patients
 
 # Calculate responses
-respEPIcontrols = epilepsyparams[0,0] * ((finecontrasts/100)**epilepsyparams[0,2] /
-                (epilepsyparams[0,1]**epilepsyparams[0,2] + (finecontrasts/100)**epilepsyparams[0,2])) + epilepsyparams[0,3]
+respEPIcontrols <- epilepsyparams[1,1] * ((finecontrasts/100)^epilepsyparams[1,3] /
+                (epilepsyparams[1,2]^epilepsyparams[1,3] + (finecontrasts/100)^epilepsyparams[1,3])) + epilepsyparams[1,4]
 
-respEPIpatients = epilepsyparams[1,0] * ((finecontrasts/100)**epilepsyparams[1,2] /
-                (epilepsyparams[1,1]**epilepsyparams[1,2] + (finecontrasts/100)**epilepsyparams[1,2])) + epilepsyparams[1,3]
+respEPIpatients <- epilepsyparams[2,1] * ((finecontrasts/100)^epilepsyparams[2,3] /
+                (epilepsyparams[2,2]^epilepsyparams[2,3] + (finecontrasts/100)^epilepsyparams[2,3])) + epilepsyparams[2,4]
 
-plt.figure(figsize=(12, 10))
-plt.subplot(2, 2, 1)
-plt.plot(finecontrastsdB, respEPIpatients, lw=4, color='blue', label='Epilepsy patients')
-plt.plot(finecontrastsdB, respEPIcontrols, lw=4, color='black', linestyle='--', label='Control participants')
-plt.xticks(ticklocsx, ticklabelsx)
-plt.yticks(np.arange(0, 7, 2))
-plt.xlabel('Target contrast (%)')
-plt.ylabel('SSVEP response')
-plt.title('(a) Epilepsy')
-plt.legend()
-plt.grid()
+ticklocsy <- seq(0,6,2)
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,6))   
+  axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
+  mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.2)     # add the tick labels
+  mtext(text = ticklocsy, side = 2, at=ticklocsy, line=0.2, las=1)  
+  title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+  title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+  title(main="(a) Epilepsy", cex.main=1.5)
 
-amblyparams = np.array([
-    [0.39, 37.2, 1.4, 0.22],  # amblyopic eye
-    [2.33, 111.69, 1.4, 0.22]   # fellow eye
-])
+lines(finecontrastsdB, respEPIpatients, lwd=4, col='blue')
+lines(finecontrastsdB, respEPIcontrols, lwd=4, col='black', lty=2)
+
+legend(-18,6,c('Epilepsy patients','Control participants'),lwd=4,lty=1:2,col=c('blue','black'),box.lwd=2,cex=1.5)
+
+amblyparams <- matrix(0,nrow=2,ncol=4) 
+amblyparams[1,] <- c(0.39, 37.2, 1.4, 0.22)  # amblyopic eye
+amblyparams[2,] <- c(2.33, 111.69, 1.4, 0.22)  # fellow eye
 
 # Calculate responses
-respambly = amblyparams[0,0] * ((finecontrasts)**amblyparams[0,2] /
-                (amblyparams[0,1] + (finecontrasts)**amblyparams[0,2])) + amblyparams[0,3]
+respambly <- amblyparams[1,1] * ((finecontrasts)^amblyparams[1,3] /
+                (amblyparams[1,2] + (finecontrasts)^amblyparams[1,3])) + amblyparams[1,4]
 
-respfellow = amblyparams[1,0] * ((finecontrasts)**amblyparams[1,2] /
-                (amblyparams[1,1] + (finecontrasts)**amblyparams[1,2])) + amblyparams[1,3]
+respfellow <- amblyparams[2,1] * ((finecontrasts)^amblyparams[2,3] /
+                (amblyparams[2,2] + (finecontrasts)^amblyparams[2,3])) + amblyparams[2,4]
 
-plt.subplot(2, 2, 2)
-plt.plot(finecontrastsdB, respambly, lw=4, color='red', label='Amblyopic eye')
-plt.plot(finecontrastsdB, respfellow, lw=4, color='black', linestyle='--', label='Fellow eye')
-plt.xticks(ticklocsx, ticklabelsx)
-plt.yticks(np.arange(0, 2.5, 0.5))
-plt.xlabel('Target contrast (%)')
-plt.ylabel('SSVEP response')
-plt.title('(b) Amblyopia')
-plt.legend()
-plt.grid()
+ticklocsy <- seq(0,2,0.5)
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,2))   
+  axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
+  mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.2)     # add the tick labels
+  mtext(text = ticklocsy, side = 2, at=ticklocsy, line=0.2, las=1)  
+  title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)     
+  title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+  title(main="(b) Amblyopia", cex.main=1.5)
+
+lines(finecontrastsdB, respambly, lwd=4, col='red')
+lines(finecontrastsdB, respfellow, lwd=4, col='black', lty=2)
+
+legend(-18,2,c('Amblyopic eye','Fellow eye'),lwd=4,lty=1:2,col=c('red','black'),box.lwd=2,cex=1.5)
 
 # Autism and control groups for the autism plot
-autismparams = np.array([
-    [0.6253, 0.2701, 125550, 5.6199],   # ASD group
-    [1.0131, 0.3369, 3589900, 6.8622]   # control group
-])
 
-respASD = (autismparams[0,0] * (finecontrasts**autismparams[0,3]) /
-           (autismparams[0,2] + finecontrasts**autismparams[0,3])) + autismparams[0,1]
-respCON = (autismparams[1,0] * (finecontrasts**autismparams[1,3]) /
-           (autismparams[1,2] + finecontrasts**autismparams[1,3])) + autismparams[1,1]
+autismparams <- matrix(0,nrow=2,ncol=4) 
+autismparams[1,] <- c(0.6253, 0.2701, 125550, 5.6199)  # ASD group
+autismparams[2,] <- c(1.0131, 0.3369, 3589900, 6.8622)  # control group
 
-plt.subplot(2, 2, 3)  # Assuming this is the third plot for Autism
-plt.plot(finecontrastsdB, respASD, lw=4, color='darkgreen', label='ASD children')
-plt.plot(finecontrastsdB, respCON, lw=4, color='black', linestyle='--', label='Control children')
-plt.xticks(ticklocsx, ticklabelsx)
-plt.yticks(np.arange(0, 2, 0.5))
-plt.xlabel('Target contrast (%)')
-plt.ylabel('SSVEP response')
-plt.title('(c) Autism')
-plt.legend()
-plt.grid()
+# Calculate responses
+respASD <- (autismparams[1,1] * (finecontrasts^autismparams[1,4]) /
+           (autismparams[1,3] + finecontrasts^autismparams[1,4])) + autismparams[1,2]
+respCON <- (autismparams[2,1] * (finecontrasts^autismparams[2,4]) /
+           (autismparams[2,3] + finecontrasts^autismparams[2,4])) + autismparams[2,2]
+
+ticklocsy <- seq(0,1.5,0.5)
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,1.5))   
+  axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
+  mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.2)     # add the tick labels
+  mtext(text = ticklocsy, side = 2, at=ticklocsy, line=0.2, las=1)  
+  title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+  title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+  title(main="(c) Autism", cex.main=1.5)
+
+lines(finecontrastsdB, respASD, lwd=4, col='darkgreen')
+lines(finecontrastsdB, respCON, lwd=4, col='black', lty=2)
+
+legend(-18,1.5,c('ASD children','Control children'),lwd=4,lty=1:2,col=c('darkgreen','black'),box.lwd=2,cex=1.5)
 
 # PD parameters for Drosophila from Afsari et al HMG 2014
-PDparams2F1 = np.array([
-    [22, 0.7, 2, 1],  # controls
-    [24, 0.4, 2, 1]])   # hLRRK2
 
-PDparams1F1 = np.array([
-    [104, 0.76, 2, 1],  # controls
-    [170, 0.382, 2, 1]])   # hLRRK2
+PDparams2F1 <- matrix(0,nrow=2,ncol=4) 
+PDparams2F1[1,] <- c(22, 0.7, 2, 1)  # controls
+PDparams2F1[2,] <- c(24, 0.4, 2, 1)  # hLRRK2
+
+PDparams1F1 <- matrix(0,nrow=2,ncol=4) 
+PDparams1F1[1,] <- c(104, 0.76, 2, 1)  # controls
+PDparams1F1[2,] <- c(170, 0.382, 2, 1)  # hLRRK2
 
 # Note: Here we plot the 2F1 curves which, we think, most closely correspond to the more central neuronal (as opposed to the photoreceptor-derived) signals
-respParkControls = PDparams2F1[0,0] * ((finecontrasts/100)**PDparams2F1[0,2] /
-                      (PDparams2F1[0,1]**PDparams2F1[0,2] + (finecontrasts/100)**PDparams2F1[0,2])) + PDparams2F1[0,3]
+respParkControls <- PDparams2F1[1,1] * ((finecontrasts/100)^PDparams2F1[1,3] /
+                      (PDparams2F1[1,2]^PDparams2F1[1,3] + (finecontrasts/100)^PDparams2F1[1,3])) + PDparams2F1[1,4]
 
-respParkPatients = PDparams2F1[1,0] * ((finecontrasts/100)**PDparams2F1[1,2] /
-                      (PDparams2F1[1,1]**PDparams2F1[1,2] + (finecontrasts/100)**PDparams2F1[1,2])) + PDparams2F1[1,3]
+respParkPatients <- PDparams2F1[2,1] * ((finecontrasts/100)^PDparams2F1[2,3] /
+                      (PDparams2F1[2,2]^PDparams2F1[2,3] + (finecontrasts/100)^PDparams2F1[2,3])) + PDparams2F1[2,4]
 
-plt.subplot(2, 2, 4)  # Fourth plot for Parkinson's
-plt.plot(finecontrastsdB, respParkPatients, lw=4, color='red', label='hLRRK2 transgenic flies')
-plt.plot(finecontrastsdB, respParkControls, lw=4, color='black', linestyle='--', label='TH+ (control) flies')
-plt.xticks(ticklocsx, ticklabelsx)
-plt.yticks(np.arange(0, 20, 4))
-plt.xlabel('Target contrast (%)')
-plt.ylabel('SSVEP response')
-plt.title("(d) Parkinson's")
-plt.legend()
-plt.grid()
+ticklocsy <- seq(0,20,4)
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,36), ylim=c(0,20))   
+  axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
+  mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.2)     # add the tick labels
+  mtext(text = ticklocsy, side = 2, at=ticklocsy, line=0.2, las=1)  
+  title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)       
+  title(ylab="SSVEP response", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
+  title(main="(d) Parkinson's", cex.main=1.5)
 
-# To save the figure
-plt.savefig('Figures/clinicalplot.pdf')
+lines(finecontrastsdB, respParkPatients, lwd=4, col='purple')
+lines(finecontrastsdB, respParkControls, lwd=4, col='black', lty=2)
+
+legend(-18,20,c('hLRRK2 transgenic flies','TH+ (control) files'),lwd=4,lty=1:2,col=c('purple','black'),box.lwd=2,cex=1.5)
+
+dev.off()
 

### 2024-05-13 Rversion/reviewR.tex
Added R version in subfolder

--- 2024-05-09 review.tex
@@
 eccentricity, chromaticity and the presence of other stimuli has been a
-focus of vision science for at least 100 years. One of the most
-productive experimental approaches in this field has been the use of the
+focus of vision science for over a century. One of the most productive
+experimental approaches in this field has been the use of the
 `steady-state visually-evoked potential' (SSVEP): a technique where
@@
 changes in cone photoreceptor activations across time and space. This
-property, referred to as contrast, sets the fundamental limits of our
+property, referred to as `contrast', sets the fundamental limits of our
 visual abilities, which remain steady over a remarkably wide range of
@@
 achromatic contrast, SSVEPs have proven to be an excellent measure of
-early chromatic processing as well (McKeefry \emph{et al.}, 1996;
-Baseler \& Sutter, 1997; Di Russo \emph{et al.}, 2001\emph{a}).
-
-Contrast is relatively simple to define: typically, it is specified as
-the percentage deviation of a uniform stimulus from the background. So,
-for example, a disk of 100 units of cone activation
-(\(I_{\mathrm{stim}}\)) surrounded by a `background' of 50 units of
-activation (\(I_{\mathrm{background}}\)) has a contrast of
+early chromatic processing as well (McKeefry \emph{et al.}, 1996; Di
+Russo \emph{et al.}, 2001\emph{a}) (see also (Baseler \& Sutter, 1997)).
+
+Contrast is typically specified as the percentage deviation of a uniform
+stimulus from the background. So, for example, a disk of 100 units of
+cone activation (\(I_{\mathrm{stim}}\)) surrounded by a `background' of
+50 units of activation (\(I_{\mathrm{background}}\)) has a contrast of
 \(\frac{I_{\mathrm{stim}} - I_{\mathrm{background}}}{I_{\mathrm{background}}}\)
@@
 same grating when superimposed on a high contrast background (see
-Figure~\ref{fig-centresurround}).
+Figure~\ref{fig-centresurround}; note that the code used to produce all
+figures in this review is available in python and R at:
+https://github.com/wadelab/contrastReviewPaper).
 
@@
 example, by varying the peak stimulus contrast parametrically, a
-`contrast vs response' function can be measured - where the `response'
-is typically defined as the amplitude of the SSVEP frequency component
-at the stimulus frequency, or a low multiple thereof. This corresponds
-closely to similar functions reported by studies measuring single unit
-activity or local field potentials in the cortex (Shapley \& Victor,
-1980; Morrone \emph{et al.}, 1982). However the SSVEP has the advantage
-that it is non-invasive, and so can be measured in awake, behaving human
-participants.
+`contrast vs response' function (CRF) can be measured - where the
+`response' is typically defined as the amplitude of the SSVEP frequency
+component at the stimulus frequency, or a low multiple thereof. This
+corresponds closely to similar functions reported by studies measuring
+single unit activity or local field potentials in the cortex (Shapley \&
+Victor, 1980; Morrone \emph{et al.}, 1982). However the SSVEP has the
+advantage that it is non-invasive, and so can be measured in awake,
+behaving human participants.
 
@@
 of contrast, to some extent this is accomplished at a population level -
-individual neurons typically implement a non-linear, sigmoidal contrast
-vs response (CRF) transducer (Tolhurst \emph{et al.}, 1981; Albrecht \&
-Hamilton, 1982) and different neurons exhibit peak sensitivity (defined
-as the maximum slope of the function) at different contrast levels
-(Carandini \& Heeger, 1994; Carandini \emph{et al.}, 1998; Busse
-\emph{et al.}, 2009). A neuronal population will therefore span a
-sensitivity range greater than any individual member.
+individual neurons typically implement a non-linear, sigmoidal CRF
+transducer (Tolhurst \emph{et al.}, 1981; Albrecht \& Hamilton, 1982)
+and different neurons exhibit peak sensitivity (defined as the maximum
+slope of the function) at different contrast levels (Carandini \&
+Heeger, 1994; Carandini \emph{et al.}, 1998; Busse \emph{et al.}, 2009).
+A neuronal population will therefore span a sensitivity range greater
+than any individual member.
 
@@
 
-To better understand these processes, we will show how sinusoidal input
+To understand these processes better, we will show how sinusoidal input
 signals might be processed by the visual system to produce SSVEPs.
@@
 panel shows Fourier transform of the waveform, which contains a
-substantial peak at this frequency. If we parametrically vary the
-stimulus contrast (i.e.~the amplitude of the waveform), the amplitude of
-the F1 component increases linearly with contrast (right panel).
+substantial peak at this frequency. If we change the stimulus contrast
+(i.e.~the amplitude of the waveform), the amplitude of the F1 component
+increases linearly with contrast (right panel).
 
@@
 by passing full stimulus waveforms through the transducer nonlinearity,
-and calculating the Fourier spectrum of the model output. Our own work
-on signal combination across eyes and space similarly demonstrated close
-correspondence between the predictions of a computational model and
-empirical data in humans (Baker \& Wade, 2017). More detailed modelling
-of intracortical recordings (Groen \emph{et al.}, 2022) has revealed
-details of the timecourse of gain control effects, specifically that
-normalization is delayed slightly relative to the initial visual
-response.
+and calculating the Fourier spectrum of the model output. Our more
+recent work on signal combination across eyes and space similarly
+demonstrated close correspondence between the predictions of a
+computational model and empirical data in humans (Baker \& Wade, 2017).
+More detailed modelling of intracortical recordings (Groen \emph{et
+al.}, 2022) has revealed details of the timecourse of gain control
+effects, specifically that normalization is delayed slightly relative to
+the initial visual response.
 
@@
 estimating a threshold is achieved by measuring the contrast response
-function at relatively high levels, and extrapolaKting back along the
+function at relatively high levels, and extrapolating back along the
 function (either contrast vs response measured at a constant spatial
@@
 A robust estimate of the threshold therefore requires the measurement of
-the SSVEP amplitude at many different super-threshold contrasr levels.
+the SSVEP amplitude at many different super-threshold contrast levels.
 This was made faster by the development of the `sweep VEP' paradigm in
@@
 hysteresis effects, the sweep is sometimes conducted both up and down in
-the same experiment Norcia \& Tyler (1985\emph{a}). The sweep VEP
-(really, a sweep `SSVEP') technique is now commonly-used to obtain a
-rapid and objective measurement of visual acuity. In particular, because
-of its relative speed and simplicity, this technique has now become a
-standard for conducting tests of visual acuity in very young subjects or
-where behavioural tests are not appropriate (Ridder, 2004; Bach \emph{et
-al.}, 2008; Hoffmann \emph{et al.}, 2017; Bach \& Farmer, 2020).
+the same experiment (Tyler \emph{et al.}, 1979; Norcia \& Tyler,
+1985\emph{a}; Norcia \emph{et al.}, 1990). The sweep VEP (really, a
+sweep `SSVEP') technique is now commonly-used to obtain a rapid and
+objective measurement of visual acuity. In particular, because of its
+relative speed and simplicity, this technique has now become a standard
+for conducting tests of visual acuity in very young subjects or where
+behavioural tests are not appropriate (Ridder, 2004; Bach \emph{et al.},
+2008; Hoffmann \emph{et al.}, 2017; Bach \& Farmer, 2020).
 
@@
 condition. Here we focus on four conditions, but there is potential to
-apply the method more broadly, either as a diagnostic technique, or to
-monitor disease severity and progression, or the efficacy of treatments.
+apply the method more broadly: as a dignostic technique, to monitor
+disease severity and progression, or to assess the efficacy of
+treatments.
 
@@
 might lead to visual deficits later in life. Although visual responses
-are a convenient assay of neural function, it is likely that same
+are a convenient assay of neural function, it is likely that the same
 general process applies throughout the whole brain, including in the
@@
 Attention exerts a profound influence on visual performance: for
-example, allowing people to attend covertly to a spatial location
+example, instructing people to attend covertly to a spatial location
 improves their performance on a target detection task significantly
@@
 of a computer using neural signals. The high SNR and precise frequency
-resolution of SSVEPs make them an ideal candidate for this approach.
-Typical studies may involve presenting an array of stimuli at different
-contrast flicker frequencies, and having the participant select one
-either by overt attention (i.e.~shifting fixation to foveate the
-selected stimulus) or covert attention (i.e.~deploying attention to one
-stimulus whilst keeping fixated) (Middendorf \emph{et al.}, 2000).
-Because SSVEP signals are highly sensitive to both visual field position
-(Di Russo \emph{et al.}, 2007; Ales \emph{et al.}, 2010) and attentional
-state (Morgan \emph{et al.}, 1996; Müller \emph{et al.}, 1998; Lauritzen
-\emph{et al.}, 2010; Verghese \emph{et al.}, 2012), the response to the
-selected stimulus will typically increase relative to the others,
-allowing it to be identified by an on-line algorithm. Because more than
-one stimulus frequency is generally present, this modulation and the
-associated changes in gain control will affect the entire pattern of
-self- and intermodulation terms, allowing the choice to be decoded by a
+resolution of SSVEPs are advantages for this approach. Typical studies
+may involve presenting an array of stimuli at different flicker
+frequencies, and having the participant select one either by overt
+attention (i.e.~shifting fixation to foveate the selected stimulus) or
+covert attention (i.e.~deploying attention to one stimulus whilst
+keeping fixated) (Middendorf \emph{et al.}, 2000). Because SSVEP signals
+are highly sensitive to both visual field position (Di Russo \emph{et
+al.}, 2007; Ales \emph{et al.}, 2010) and attentional state (Morgan
+\emph{et al.}, 1996; Müller \emph{et al.}, 1998; Lauritzen \emph{et
+al.}, 2010; Verghese \emph{et al.}, 2012), the response to the selected
+stimulus will typically increase relative to the others, allowing it to
+be identified by an on-line algorithm. Because more than one stimulus
+frequency is generally present, this modulation and the associated
+changes in gain control will affect the entire pattern of self- and
+intermodulation terms, allowing the choice to be decoded by a
 multivariate pattern classifier.
@@
 flicker, work in this field has largely focused on optimising the
-stimuli or decoders to increase decoding performance (e.g.~\ldots{} )
-and reducing visual fatigue associated with the long-term presentation
-of arrays of high-contrast flicker (Diez \emph{et al.}, 2024) rather
-than studying visual contrast processing \textit{per se}. We therefore
-note in passing that using SSVEPs to improve our understanding of early
-contrast processing may yield benefits to this related field.
+stimuli or decoders to increase decoding performance and reducing visual
+fatigue associated with the long-term presentation of arrays of
+high-contrast flicker (Diez \emph{et al.}, 2024), rather than studying
+visual contrast processing \textit{per se}. We therefore note in passing
+that using SSVEPs to improve our understanding of early contrast
+processing may yield benefits to this related field.
 
@@
 researchers to study how contrast processing depends on spatial and
-temporal context and as well as changes in task, cognitive and
-behavioural state and arousal.
+temporal context, as well as changes in task, cognitive and behavioural
+state, and arousal.
 
@@
 the SSVEP approach was that the entire complex-valued frequency spectrum
-recorded at each lcoation provided detailed information about the
+recorded at each location provided detailed information about the
 processing nonlinearities up to and including that point. In principle,
 this would allow researchers to uniquely fit the parameters of their
-computations models of visual processing (Regan \& Regan, 1988). To
+computational models of visual processing (Regan \& Regan, 1988). To
 date, the complexity of the neuronal computations at even the earliest
@@
 typically restrict their analyses to the amplitudes of single, low-order
-frequency components (for example, the inputs frequencies or simple sums
+frequency components (for example, the input frequencies or simple sums
 and differences of those frequencies). As our understanding of the early
@@
 synthetic SSVEP responses that can be compared with those measured in
-human subject. In principle, we are therefore able to use the SSVEP to
+human subjects. In principle, we are therefore able to use the SSVEP to
 fit the parameters of early visual processing. This approach may also

### 2024-05-14 attentionGratings.py
Changing attention grating code

@@ -1,245 +1,104 @@
-#%%
+
 import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib.patches import Circle
 
 def hyper_ratio(params, centerContrast, surroundContrast):
-    # Returns a hyperbolic ratio of the form Rzero + Rmax * Cc^n / (Cc^n + Cs^n + c50^n)
-    # Input params are [Rmax, c50, n, Rzero]
-    # Either the center or the surround contrast or both can vary
-
     Rmax, c50, n, Rzero = params
+    Cc = np.clip(centerContrast, 0, 1)
+    Cs = np.clip(surroundContrast, 0, 1)
 
-    Cc = np.atleast_1d(centerContrast)
-    Cs = np.atleast_1d(surroundContrast)
-
-    # Force inputs to be between 0 and 1
-    
-    Cc[np.where(Cc>1)] = 1
-    Cc[np.where(Cc<0)] = 0
-    Cs[np.where(Cs>1)] = 1
-    Cs[np.where(Cs<0)] = 0
-
-    if np.any(Cc < 0) or np.any(Cc > 1):
-        raise ValueError("Center contrast must be in the range [0,1]")
-    if np.any(Cs < 0) or np.any(Cs > 1):
-        raise ValueError("Surround contrast must be in the range [0,1]")
-    
-    # Ensure Cc and Cs are broadcastable to the same shape : Broadly, are they both vectors of the same shape?
-    # Or is one a scalar and the other a vector?
-    if Cc.shape != Cs.shape:
-        if Cc.size == 1:
-            Cc = np.full(Cs.shape, Cc)
-        elif Cs.size == 1:
-            Cs = np.full(Cc.shape, Cs)
-            
-        else:
-            raise ValueError("Center and surround contrast arrays must be broadcastable to the same shape")
-        
     # Validate parameters
-    if Rmax <= 0:
-        raise ValueError("Rmax must be positive")
-    if n <= 0:
-        raise ValueError("n must be positive")
-    if Rzero < 0:
-        raise ValueError("Rzero must be non-negative")
-    if c50 <= 0:
-        raise ValueError("c50 must be positive")
+    if any(p <= 0 for p in [Rmax, c50, n]) or Rzero < 0:
+        raise ValueError("Invalid parameters: Rmax, c50, and n must be positive, Rzero must be non-negative")
 
-    # Calculate the hyperbolic ratio
-   
[…]

### 2024-05-22 review.qmd
Alternative rendering for VNS

@@
     keep-tex: true
+ # docx:
+ #   keep-tex: true
 bibliography: WadeBaker_Review.bib
@@
 ---
+
 Department of Psychology and York Biomedical Research Institute, University of York, UK
 
-## Review for Vis Neurosci
+### Funding statement:
+This work was supported by BBSRC grant BB/V007580/1 to DHB and ARW
 
@@
 ```
+{{< pagebreak >}}
+# Measuring contrast processing in the visual system using SSVEP
 
@@
 This paper can therefore be used both as a review of the use of SSVEP in measuring human contrast processing, and as an interactive learning aid.
+
+Keywords:
+
+* EEG
+* VEP
+* SSVEP 
+* Vision
+* Contrast
+
+{{< pagebreak >}}
 
@@
 
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](){#fig-centresurround}
+:::
+::: {.content-visible unless-format="docx"}
 ![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/centresurround.pdf){#fig-centresurround}
+:::
 
@@
 
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
+![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
+These calculations assume an entirely linear system.](){#fig-linear}
+:::
+::: {.content-visible unless-format="docx"}
 ![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
 These calculations assume an entirely linear system.](Figures/linearsystem.pdf){#fig-linear}
+:::
 
@@
 
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
+![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
+The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
+The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](){#fig-transducer}
+:::
+::: {.content-visible unless-format="docx"}
 ![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
@@
 The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/transducer.pdf){#fig-transducer}
+:::
 
@@
 
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
+![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
+Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](){#fig-twoinputs}
+:::
+::: {.content-visible unless-format="docx"}
 ![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
 Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/twoinputs.pdf){#fig-twoinputs}
+:::
 
@@
 
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
+![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
+The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](){#fig-sweepvep}
+:::
+::: {.content-visible unless-format="docx"}
 ![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
 The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/sweepplot.pdf){#fig-sweepvep}
+:::
 
@@
 
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
+![Example contrast response functions for different clinical conditions.
+Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
+(2011).
+Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al.
+(2015).
+Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al.
+(2018).
+Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al.
+(2014).
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](){#fig-plotclinical}
+:::
+
+::: {.content-visible unless-format="docx"}
 ![Example contrast response functions for different clinical conditions.
@@
 One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/clinicalplot.pdf){#fig-plotclinical}
+:::
 

### 2024-05-22 review.tex
Alternative rendering for VNS

@@
 
-\subsection{Review for Vis Neurosci}\label{review-for-vis-neurosci}
+\subsubsection{Funding statement:}\label{funding-statement}
+
+This work was supported by BBSRC grant BB/V007580/1 to DHB and ARW
+
+\newpage{}
+
+\section{Measuring contrast processing in the visual system using
+SSVEP}\label{measuring-contrast-processing-in-the-visual-system-using-ssvep}
 
@@
 
+Keywords:
+
+\begin{itemize}
+\tightlist
+\item
+  EEG
+\item
+  VEP
+\item
+  SSVEP
+\item
+  Vision
+\item
+  Contrast
+\end{itemize}
+
+\newpage{}
+
 \section{Introduction}\label{introduction}

### 2024-05-22 Figures/convertImages.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+
+# Set the desired DPI and width in pixels
+DPI=300
+WIDTH_CM=14
+
+# Convert width from cm to pixels (since 1 inch = 2.54 cm)
+WIDTH_PIXELS=$(echo "$WIDTH_CM * $DPI / 2.54" | bc)
+
+# Loop through all PDF files in the directory
+for file in *.pdf; do
+  # Extract the base name of the file (without extension)
+  base_name="${file%.pdf}"
+
+  # Convert PDF to TIFF
+  convert -density $DPI "$file" -resize ${WIDTH_PIXELS}x -units PixelsPerInch "${base_name}.tiff"
+done
+

### 2024-05-22 review.qmd
Stupid options for VNS

@@
 format: 
-  pdf:
+  #pdf:
+  #  keep-tex: true
+  docx:
     keep-tex: true
- # docx:
- #   keep-tex: true
 bibliography: WadeBaker_Review.bib
@@
 plt.tight_layout()
-plt.savefig('Figures/centresurround.pdf')
+plt.savefig('Figures/1_centresurround.pdf')
 
@@
 plt.tight_layout()
-plt.savefig('Figures/linearsystem.pdf')
+plt.savefig('Figures/2_linearsystem.pdf')
 
@@
 plt.tight_layout()
-plt.savefig('Figures/transducer.pdf')
+plt.savefig('Figures/3_transducer.pdf')
 
@@
 plt.tight_layout()
-plt.savefig('Figures/twoinputs.pdf')
+plt.savefig('Figures/4_twoinputs.pdf')
 
@@
 plt.tight_layout()
-plt.savefig('Figures/sweepplot.pdf')
+plt.savefig('Figures/5_sweepplot.pdf')
 
@@
 # To save the figure
-plt.savefig('Figures/clinicalplot.pdf')
+plt.savefig('Figures/6_clinicalplot.pdf')
 
@@
 ::: {.content-visible unless-format="docx"}
-![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/centresurround.pdf){#fig-centresurround}
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/1_centresurround.pdf){#fig-centresurround}
 :::
@@
 ![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
-These calculations assume an entirely linear system.](Figures/linearsystem.pdf){#fig-linear}
+These calculations assume an entirely linear system.](Figures/2_linearsystem.pdf){#fig-linear}
 :::
@@
 The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
-The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/transducer.pdf){#fig-transducer}
+The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/3_transducer.pdf){#fig-transducer}
 :::
@@
 ![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
-Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/twoinputs.pdf){#fig-twoinputs}
+Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/4_twoinputs.pdf){#fig-twoinputs}
 :::
@@
 ![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
-The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/sweepplot.pdf){#fig-sweepvep}
+The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/5_sweepplot.pdf){#fig-sweepvep}
 :::
@@
 (2014).
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/clinicalplot.pdf){#fig-plotclinical}
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/6_clinicalplot.pdf){#fig-plotclinical}
 :::

### 2024-05-22 Figures/convertImages.sh

@@ -13,6 +13,7 @@ for file in *.pdf; do
   base_name="${file%.pdf}"
 
   # Convert PDF to TIFF
-  convert -density $DPI "$file" -resize ${WIDTH_PIXELS}x -units PixelsPerInch "${base_name}.tiff"
+  convert -density $DPI "$file" -resize ${WIDTH_PIXELS}x -units PixelsPerInch -compress LZW "${base_name}.tiff"
+
 done
 

### 2024-06-13 review.qmd
figs

--- 2024-05-13 review.qmd
@@
 format:
+  docx: 
+    omitFigures: true
   pdf: 
     keep-tex: true
+
 bibliography: WadeBaker_Review.bib
@@
 
-![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/centresurround.pdf){#fig-centresurround}
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/1_centresurround.pdf){#fig-centresurround}
 
@@
 ![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
-These calculations assume an entirely linear system.](Figures/linearsystem.pdf){#fig-linear}
+These calculations assume an entirely linear system.](Figures/2_linearsystem.pdf){#fig-linear}
 
@@
 The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
-The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/transducer.pdf){#fig-transducer}
+The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/3_transducer.pdf){#fig-transducer}
 
@@
 ![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
-Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/twoinputs.pdf){#fig-twoinputs}
+Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/4_twoinputs.pdf){#fig-twoinputs}
 
@@
 ![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
-The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/sweepplot.pdf){#fig-sweepvep}
+The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/5_sweepplot.pdf){#fig-sweepvep}
 
@@
 (2014).
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/clinicalplot.pdf){#fig-plotclinical}
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/6_clinicalplot.pdf){#fig-plotclinical}
 

### 2024-06-13 review.tex
Stupid options for VNS
figs

@@
 
-\subsubsection{Funding statement:}\label{funding-statement}
-
-This work was supported by BBSRC grant BB/V007580/1 to DHB and ARW
-
-\newpage{}
-
-\section{Measuring contrast processing in the visual system using
-SSVEP}\label{measuring-contrast-processing-in-the-visual-system-using-ssvep}
+\subsection{Review for Vis Neurosci}\label{review-for-vis-neurosci}
+
+\includegraphics{review_files/figure-pdf/cell-3-output-1.pdf}
+
+\includegraphics{review_files/figure-pdf/cell-4-output-1.pdf}
+
+\includegraphics{review_files/figure-pdf/cell-5-output-1.pdf}
+
+\includegraphics{review_files/figure-pdf/cell-6-output-1.pdf}
+
+\includegraphics{review_files/figure-pdf/cell-7-output-1.pdf}
+
+\includegraphics{review_files/figure-pdf/cell-8-output-1.pdf}
 
@@
 human contrast processing, and as an interactive learning aid.
-
-Keywords:
-
-\begin{itemize}
-\tightlist
-\item
-  EEG
-\item
-  VEP
-\item
-  SSVEP
-\item
-  Vision
-\item
-  Contrast
-\end{itemize}
-
-\newpage{}
 
@@
 
-\includegraphics{Figures/centresurround.pdf}
+\includegraphics{Figures/1_centresurround.pdf}
 
@@
 
-\includegraphics{Figures/linearsystem.pdf}
+\includegraphics{Figures/2_linearsystem.pdf}
 
@@
 
-\includegraphics{Figures/transducer.pdf}
+\includegraphics{Figures/3_transducer.pdf}
 
@@
 
-\includegraphics{Figures/twoinputs.pdf}
+\includegraphics{Figures/4_twoinputs.pdf}
 
@@
 
-\includegraphics{Figures/sweepplot.pdf}
+\includegraphics{Figures/5_sweepplot.pdf}
 
@@
 
-\includegraphics{Figures/clinicalplot.pdf}
+\includegraphics{Figures/6_clinicalplot.pdf}
 

### 2024-06-13 review.qmd
Make figs in place

--- 2024-05-22 review.qmd
@@
 format:
-  #pdf:
-  #  keep-tex: true
   docx: 
+    omitFigures: true
+  pdf:
+
     keep-tex: true
+
 bibliography: WadeBaker_Review.bib

### 2024-11-27 responseToReviews.md
Editing review

# Comments to the Author
## General Comments

Wade and Baker have written a very approachable and well-motivated review of the use of VEP measurements of visual contrast. The authors rightly focus on the importance of measuring contrast responses under a wide range of conditions, highlighting the ways in which human evoked responses can be made in informative in diverse contexts. The review should thus be of broad interest.

There are points where the review is overly broad and others that could use more detail. These are described below.

A central part of the review is the role of computational models and normalization models, in particular. The use of the normalization framework for describing the VEP contrast response function (CRF) has been fruitful in terms of mechanistic interpretation and linking to other measurements, such as single-unit physiology and fMRI.  This modeling approach assumes that the CRF reflects the activity of a more or less unitary population, and that the population response resembles that of the underlying individual neurons that make up the population.

There are two edge-cases in the literature that suggest a more nuanced approach will be needed. It would be useful to be clearer on the assumptions of the modeling approach and what the significance is for interpretation of cases where the assumptions don’t hold.

Two-limbed contrast response functions. Campbell and Maffei (J. Physiol,1970) in the earliest measurements of the VEP contrast response function reported that it sometimes had two “limbs” – one at low contrast and a second one with a steeper slope at higher contrasts.  Two-limbed contrast response functions have also been found by Nakayama and Mackeben (Vision Res,1982) in macaque with electrodes on the dura, by Apkarian and Tyler (Vision Res, 1985),  Norcia et al., (Vision Res,1989) and Souza et al., (IOVS, 2007), to name a few.  As noted, the sigmoidal model assumes that the population response is unitary, but the two-limbed functions suggest that it is a mixture of populations. The two limbs may not be apparent when the function is sampled with only a few points over a wide range of contrast.

Harmonics with different CRFs.  Bobak et al., (IOVS, 1986), found that the first harmonic (1f) and second harmonic (2f) of the pattern onset/offset SSVEP had different thresholds but similar slopes once above threshold. Norcia et al., (TVST, 2024) found different thresholds and slope for 1f and 2f responses driven by contrast modulated dynamic noise. This again suggests the presence of multiple underlying populations.  The behavior also raises the question of whether to pool harmonics when estimating the response function. This was alluded to in the simulation, but could be expanded on.

The lack of dynamics in the normalization model, as presented should be discussed, the brief mention of Tsai et al. aside. The authors may wish to cite another framework that explicitly models dynamics (e.g. amplitude and phase) in the context of the contrast response function (Zemon and Gordon, Vision Res, 2006) .

### Response
> We thank the reviewer for these positive comments. In general this was intended to be an overview in the sense of a 'primer' for SSVEP rather than a detailed historical account of the field - hence the inclusion of code examples. However, the reviewer raises some important general issues and we now address them: Specifically, we include references to the idea that SSVEP can engage more than one type of neuron, that the population responses is not trivially derived from what is known about the properties of individual neurons and that fitting with a sigmoidal response function may mask some complexities in the CRF that are more readily seen with denser sampling. We think the Bobak reference is IOVS 1984 and we have also included the Kaestner TVST 2024 reference (thank you for drawing our attention to both of these).

### Specific comments

Adaptation effects.  Adapted contrast response functions have been measured by Nelson et al., (EEG Journal, 1984) by comparing CRFs measured by sweeping contrast from low to high vs high to low. They found that responses at low contrast were smaller in the “down” sweeps and attributed it to differences in prior adaptation.  Other studies also measured adaptation effects on the CRF Heinrich and Bach, (IOVS, 2001), Ross and Speed (1991, Proc Roy Soc). And the hard-to-find Bach, Greenlee and Buhler (Clin Vis Sci, 1996) paper that modeled the effect within a contrast gain control framework.
Gain control development.  Morrone and Burr (Nature, 1986) used frequency tagging to measure orientation dependence of masking in infants and adults, as did Candy et al. The difference in the two studies is that in adults Morrone and Burr found response gain for cross-oriented maskers but Candy et al found contrast gain.   Skoczenski and Norcia used a high contrast 30 Hz dynamic 2D noise masker and variable contrast reversing gratings. They found masking consistent with a contrast gain effect.

>

Page 5 bottom. The reciprocal effect of test and masker was first shown by Regan and Regan  (1988; cited)  for flicker modulation depth and then for gratings by Brown et al., (IOVS,  1999 for dichoptic masking) and  then  by candy et al., (J. Neurosci, 2001) along with Busse et al. (Neuron, 2009) as cited. It’s good to have cited Regan and Regan paper that introduced the notion of fingerprints in the spectrum being tied to the nature of the non-linearity. Maybe expand their “fingerprint” approacg a little more.

Figure 4. Oversaturation occurs because the denominator overtakes the numerator at high contrast as it has two terms in it, not the single term in the numerator. Would be worth showing total power not saturating. Is this a general effect?

Equation 3 is a special case where the neurons in the numerator are so selective that they don’t “see” the masker except by being influenced indirectly by the gain pool.  Putting the test and mask in the numerator and summing them before exponentiating yields robust IM. Weights can be used to reflect the neurons preference for the test vs the mask.  Does equation 3 produce IM?

Page 7 Mckeefry et al., (Vis Neuro, 1996) is a good reference for linking harmonics of on/off to transient sustained responses.

Page 8 figure 5. The extrapolation to zero amplitude method should be attributed to Cambell and Maffei (J Physiol) 1970). Campbell and Kulikowski (J Physiol, 1972) made a detailed comparison to psychophysics.

Figure 5 is an oversimplification. As it doesn’t take into account how the additive noise mixes with the internal VEP response.  This was modeled in Norcia et al., (Vision Res, 1989). The additive noise adds a bias to the measurement, especially if the slope of the response function is shallow. The authors should try to connect the linear regression approach for threshold measurement to the sigmoidal non-linearity, as they are different models of the CRF. In practice the difference at the low contrast end is probably too small to matter and the saturation region is excluded.

Page 9. The sweep method was first developed by David Regan (Vision Res, 1973) for use in objective refraction.

Page 9 “This approach” seems to refer to the sweep VEP, not the method of measuring at multiple contrast levels over a series of trials which was used in the Atkinson/Braddick papers. The sweep method refers to the approach of incrementing or decrementing the stimulus value within a single trial and then averaging the sweeps to get the response function.  Regan argued that this was a better way of estimating response functions because amplitude variations over a session are distributed over the whole response function shifting it up or down instead of distorting its shape (Regan (Vison Res, 1975, color coding).

Page 9 …chromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts (Dobson et al., 1978; Norcia & Tyler, 1985b) with both chromatic and achromatic contrast detection reaching near-adult levels by around six months. Needs references for chromatic sensitivity and disparity sensitivity.

The DeVries Khoe and Spekreijse comparison between VEP and psychophysics was done above age 4 years and for acuity not for contrast sensitivity. The large discrepancies between VEP and FPL are in infancy for contrast sensitivity, as well as for acuity.  In the realm of contrast sensitivity, the differences are very large, up to a factor of 50.  Compare Norcia et al., 1990 Vision Res measurement to any behavioral data. Suggest making this comparison more quantitative.

The authors may wish to cite Norcia et al., (NeuroReport, 2000) for an early measurement of suppression in amblyopia using measurements of the contrast response function in each eye in the presence of a high contrast dichoptic masker in the other.  This study showed that amblyopic observers retain suppressive binocular interactions despite severely reduced stereopsis. The dominant eye exerted a greater suppressive effect on the non-dominant eye, suggesting that it maintains a relatively greater number, or more effective, inhibitory contacts with cortical cells.  Also may wish to cite recent work by Hou and colleagues using a similar approach, including application of normalization models.

Schade (JOSA, 1956) was the first to measure the contrast sensitivity function of the human visual system as noted by Campbell and Robson (J. Physiol, 1968). Campbell and Green (1965, cited) credit Schade for how to make sinewave gratings on an oscilloscope but not for having made the first measurement.

Villadaite, Norcia et al., (cited) did not find spatial frequency specific effects at 8 cpd as in Pei et al,. (cited). So it’s not really a replication of that study, but more of an extension to a measurement of CRF instead of SF tuning.

Page 12, better to state whether attention effects were consistent with response gain or contrast gain to make the result more concrete for the reader.

Discussion of time varying response paragraph on page 12 is not that relevant to contrast response function measurement, nor is the BCI section.  These applications of SSVEP and could be deleted or the relationship to contrast response function made clearer.

Page, 14. It’s not clear from the text how one would study feedback with the SSVEP, please elaborate.

In the last section, it would be worth mentioning that in order to “identify” a non-linear system, one needs multiple inputs. If the system is 2nd order, a minimum of two inputs is needed. If the system non-linearity is higher order than that, then more simultaneous inputs are needed (see Boyd, Tang and Chua 1983, IEEE Trans Circuits and Systems  and  Chua and Liao, 1991, Int. J. Circuit Theory and Applications).

.

## Reviewer: 2

# Comments to the Author
This manuscript is intended to be a review of the field of steady-state visual evoked potentials (ssVEPs) over the past 70 years.  The authors state in the introductory paragraph “Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.” The material included and the references cited, however, are quite limited and omit seminal work in this area.

1. All abbreviations should be defined upon first use.
2. The following statement is unclear and inaccurate: “On-off flicker can drive independent populations of on- and off-cells in the retina once per cycle and can therefore produce a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on. Counterphase flicker contains two transients per cycle and therefore does not produce a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.” The term “on-off flicker” is typically used for simple luminance flicker without spatial pattern contrast. In the context of pattern stimulation, “on-off” can refer to “appearance-disappearance” of a spatial pattern with its contrast modulated in time. When the pattern appears and disappears into a uniform field of equal space-average luminance, the light and dark elements are modulated in counterphase. Any linear contribution to the response elicited by each set of elements (e.g., bars or checks) is expected to be canceled when summed at the electrode site on the scalp. Thus, the response at 1F to this stimulus is in fact a nonlinear response. This type of stimulation does not separate contributions from “independent populations of on- and off-cells.” There are other ssVEP techniques developed that do tap ON and OFF pathways selectively, which are not cited here. Contrast-reversal stimuli which involve symmetrical modulation of light and dark elements cancel odd harmonic frequency components in the response, however, both kinds of stimulus techniques mentioned here elicit only nonlinear responses.
3. The statement “SSVEPs have proven to be an excellent measure of early chromatic processing as well” again fails to cite seminal work in this area.
4. The statement “This normalization, achieved through a computation called ‘contrast gain control’” again neglects to cite the original work that defined “contrast gain control” and the authors ignored a nonlinear biophysical model proposed to explain this phenomenon in ssVEP contrast response functions (which was also applied in developmental work on infants and children).
5. The description of Figure 3 and the nonlinear operations performed are unclear. They state in the text that Equation 2 (“hyperbolic ratio function”) is used as a nonlinear transducer, but the figure legend states that a squaring operator was used for frequency doubling. A squaring operation would only produce a sinusoid of double the frequency of the input. The waveform shown in the left panel appears to be a full-wave rectified signal, which would contain theoretically an infinite number of higher harmonics. The text also states that “This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.” When was full-wave rectification performed?
6. The authors use the term “frequency tagging” which has been used in the literature to refer to stimulating different regions of the stimulus field or fellow eyes with different sinusoidal frequencies and measuring the intermodulation (sum and difference) frequency responses that result from nonlinear interactions. They do not, however, cite the foundational work in this field.
7. There is a large body of work that used ssVEPs to investigate contrast processing developmentally (infants through adulthood) and in various disease states, which is mostly missed in this review article.
8. Figures should be labeled.
Thus, the stated intention of this manuscript (“Here we describe how SSVEPs have been used to study visual contrast over the past 70 years”) fails to be met. Only select articles from a few sources are covered, and therefore, the richness and value of this technique is lost.

### 2024-11-27 review.qmd
Editing review

@@
 This normalization, achieved through a computation called 'contrast gain control' [@Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
-As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround; note that the code used to produce all figures in this review is available in python and R at: https://github.com/wadelab/contrastReviewPaper).
+As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround; note that the code used to produce all figures in this review is available in python and R at: https://github.com/wadelab/contrastReviewPaperVNS).
 
@@
 This "normalisation" process is dynamic and nonlinear and is well-modeled by a hyperbolic ratio function in which the response of each neuron is modulated by a local 'gain pool' composed of the summed responses of the local neuronal population [@Heeger1992; @Carandini2011; @Busse2009; @Baker2017].
+
+An additional complexity is introduced by the fact that the EEG is an average population response and that the visual system contains many different types of neurons.
+As a stimulus changes in, for example, contrast it may selectively activate qualitatively different neuronal populations.
+This type
 

### 2024-11-27 responseToReviews.md
Editing both paper and R2R - now at end of R1/ adaptation

@@
 > We thank the reviewer for these positive comments.
-In general this was intended to be an overview in the sense of a 'primer' for SSVEP rather than a detailed historical account of the field - hence the inclusion of code examples.
+In general this was intended to be an overview in the sense of a 'primer' for SSVEP rather than a detailed historical account of the field - hence the inclusion of code examples and relatively domain-agnostic language.
 However, the reviewer raises some important general issues and we now address them: Specifically, we include references to the idea that SSVEP can engage more than one type of neuron, that the population responses is not trivially derived from what is known about the properties of individual neurons and that fitting with a sigmoidal response function may mask some complexities in the CRF that are more readily seen with denser sampling.
@@
 They found that responses at low contrast were smaller in the “down” sweeps and attributed it to differences in prior adaptation.
+
+> we could not find this reference.
+We did find an excellent (and sometimes humorous) review of SSVEPs by Nelson (Nelson, J.
+I., Seiple, W.
+H., Kupersmith, M.
+J., & Carr, R.
+E.
+(1984).
+Lock-in Techniques for the Swept Stimulus Evoked Potential.
+Journal of Clinical Neurophysiology, 1(4), 409–436.
+doi:10.1097/00004691-198401040-00002) but could not find a mention of adaptation effects.
+
 Other studies also measured adaptation effects on the CRF Heinrich and Bach, (IOVS, 2001), Ross and Speed (1991, Proc Roy Soc).
 And the hard-to-find Bach, Greenlee and Buhler (Clin Vis Sci, 1996) paper that modeled the effect within a contrast gain control framework.
+
+> The Ross and Speed paper measures masking using SSVEP but other adaptation measures are, as far as we can tell, psychophysical.
+Heinrich and Bach are concerned with transient changes in SSVEP amplitude within a recording which may have an adaptation origin but which manifest as response increases over the first few seconds and seems a little mysterious.
+Bach, Greenlee and Buhler is, however, pretty unequivocal and we now include it.
+We like the suggestion that the  'increase' in response is attributableto a shift from a square wave response to a more sinusoidal waveform as the apparent contrast of the stimulus is decreased by adaptation-driven gain control.
+The response power which was distributed across odd harmonics in the original response can now be concentrated in the 1st harmonic after adaptation.
+
 Gain control development.

### 2024-11-27 review.qmd
Editing both paper and R2R - now at end of R1/ adaptation

@@
 format:
-  docx: 
-    omitFigures: true
-  pdf:
-
-    keep-tex: true
+  #docx: 
+    #omitFigures: true
+  pdf: default
+    #keep-tex: true
 
@@
 The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system [@Regan1988].
+Because they can arise from different neuronal computations, different populations and even different stages of the visual system, different harmonics can exhibit different input-output functions.
+For example, 1F and 2F responses can have different thresholds [@BobakIOVS1984] and scalp topographies [@reganEvokedPotentialsSpecific1973] while Kaester et al [@KaestnerTVST2024] show both different CRF slopes and thresholds for 1F and 2F components generated by dynamic noise.
+
 SSVEP signals can also be elicited by periodic changes of stimulus properties other than achromatic and chromatic contrast, such as motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
@@
 However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
+
+Although the SSVEP does seem to reflect the activity of relatively well-tuned neuronal populations , changing some aspect of the stimulus may change the nature of those populations.
+For example, Campbell and Maffei [@Campbell1970] noted that densely-sampled measurements of contrast responses can reveal the presence of two, qualitatively different types of neurons that exhibit different log-linear contrast functions (see also [@Tyler1985, @Nakayama1982, @Norcia1989 and @Souza2007]).
+Modeling such two-limbed contrast response functions using a single sigmoidal function is therefore an approximation.
 
@@
 Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
-The SSVEP response to a contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
+Both the amplitude and the phase of the SSVEP response to a single contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
 
@@
 In principle, that information could be recovered from the SSVEP signal - a possibility recognised in the early days of the technique [@Regan1988].
-Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
-For example, @Tsai2012 demonstrated that a gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
+Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual visual processing (e.g.
+[@Zemon2006Dynamic]) and, by incorporating mathematical models of gain control mechanisms with either instantaneous, or temporally-integrated gain control, it is possible to model gain control using SSVEP data.
+For example, [@Tsai2012] demonstrated that a relatively simple gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
 This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.

### 2024-11-27 review.qmd
Finished adaptation

@@
 Psychophysically, adaptation increases detection thresholds, but has little effect on contrast discrimination performance [@Ross1993], much like pattern masks [@Foley1994].
-Although SSVEP adaptation effects show strong tuning for orientation [@Vergeer2018; @Campbell1970] and spatial frequency [@Mecacci1976], there appear to be no studies measuring the full SSVEP contrast response function before and after an extended period of adaptation.
+Although SSVEP adaptation effects show strong tuning for orientation [@Vergeer2018; @Campbell1970] and spatial frequency [@Mecacci1976], there appear to be very measurements of the full SSVEP contrast response function before and after an extended period of adaptation.
+One exception is a study by Bach et al [@BachGreenleeBuhler1988] that examined the effects of adaptation on SSVEP responses who found a paradoxical increase in the 1F SSVEP amplitude following prolonged exposure to high-contrast stimuli across much of the probe contrast range.
+They explain this as being due to the presence of a nonlinear transducer function that saturates at high probe contrast driving the response to higher harmonic frequencies.
+Adaptation reduces the apparent contrast of the input, leading to less distortion, a more sinusoidal output and a concentration of the frequency-tagged response at lower harmonics.
+This explanation is conceptually related to the 'supersaturation' that we observe in Figure xxx.
 

### 2024-11-28 review.qmd
Tidying

@@
 Adaptation reduces the apparent contrast of the input, leading to less distortion, a more sinusoidal output and a concentration of the frequency-tagged response at lower harmonics.
-This explanation is conceptually related to the 'supersaturation' that we observe in Figure xxx.
+This explanation is conceptually related to the 'supersaturation' effect that we observe in Figure 4.
 
@@
 This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979; @Norcia1985a; @Norcia1990].
+To avoid hysteresis or short-term adaptation effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979; @Norcia1985a; @Norcia1990].
 The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
@@
 The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section [@Candy2001; @Pei2017].
-Although contrast gain control is measurable in infants as young as six weeks old [@Candy2001; @Morrone1986; @Skoczenski1998], its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
+Although response- [@Morrone1986] or contrast gain control [@Candy2001; ; @Skoczenski1998] is measurable in infants as young as six weeks old, its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
 

### 2024-11-28 responseToReviews.md
Tidying
Now on 'thresholds'

@@
 > The Ross and Speed paper measures masking using SSVEP but other adaptation measures are, as far as we can tell, psychophysical.
-Heinrich and Bach are concerned with transient changes in SSVEP amplitude within a recording which may have an adaptation origin but which manifest as response increases over the first few seconds and seems a little mysterious.
+Heinrich and Bach are concerned with transient changes in SSVEP amplitude within a recording which may have an adaptation origin but which manifest as response increases over the first few seconds and seems a little mysterious (we often observe these as well).
 Bach, Greenlee and Buhler is, however, pretty unequivocal and we now include it.
-We like the suggestion that the  'increase' in response is attributableto a shift from a square wave response to a more sinusoidal waveform as the apparent contrast of the stimulus is decreased by adaptation-driven gain control.
+We like the suggestion that the  'increase' in response is attributable to a shift from a square wave response to a more sinusoidal waveform as the apparent contrast of the stimulus is decreased by adaptation-driven gain control.
 The response power which was distributed across odd harmonics in the original response can now be concentrated in the 1st harmonic after adaptation.
@@
 
->
+> We have clarified that these papers measure different types of gain control
 
@@
 (Neuron, 2009) as cited.
+
+> Thanks - we have included these references
+
  It’s good to have cited Regan and Regan paper that introduced the notion of fingerprints in the spectrum being tied to the nature of the non-linearity.
-Maybe expand their “fingerprint” approacg a little more.
+Maybe expand their “fingerprint” approach a little more.
+
+>
 
@@
 
+> We think this topic is probably worth a more detailed treatment than we can provide here.
+Broadly, much of the oversaturation effect can be modeled by an increase in saturation and therefore harmonic distortion - power is transferred from the first to higher harmonics as contrast increases.
+It seems likely that total power does not saturate (perceptually, there appears to be a monotonic increase in contrast for example and discrimination thresholds increase up to 100%) but the issue is complicated by the fact that the distribution of neurons sensitive to different contrast levels might not be uniform and that neurons might alter their response properties to signal with optimum sensitivity around a particular range.
+Our goal here is simply to show people new to the field how the CRF is approximately log-linear with contrast.
+
 Equation 3 is a special case where the neurons in the numerator are so selective that they don’t “see” the masker except by being influenced indirectly by the gain pool.
@@
 
+> In the accompanying code we specifically allow the user to alter the way that normalization is computed and many of the (commented) options produce robust IM.
+We encourage the user to experiment with different configurations to explore their effects.
+We have updated the text to make this more explicit.
+
 Page 7 Mckeefry et al., (Vis Neuro, 1996) is a good reference for linking harmonics of on/off to transient sustained responses.
+
+> Included
 
@@
 Campbell and Kulikowski (J Physiol, 1972) made a detailed comparison to psychophysics.
+
+> We have now included these references
 
@@
 
+> We have now note that it is important to consider noise for situations where the SNR is low or the slope is shallow and we expand on the point that you have to stick to the part of the CRF where the curve is more or less linear - as we showed explicitly in the figure.
+
 Page 9. The sweep method was first developed by David Regan (Vision Res, 1973) for use in objective refraction.
+
+> We have included this reference.
 
@@
 Regan argued that this was a better way of estimating response functions because amplitude variations over a session are distributed over the whole response function shifting it up or down instead of distorting its shape (Regan (Vison Res, 1975, color coding).
+
+> This is a good point but the (simpler) one we are trying to make here is that extrapolating from high stimulus levels down to zero provides a robust estimate of visual sensitivity in subjects where such measurements might otherwise be hard to make.
+We have altered the text to clarify this.
 

### 2024-11-28 review.qmd
Now on 'thresholds'

@@
 The mask stimulus will suppress the response at the target frequency, reducing its amplitude.
-The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Busse2009].
+The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Regan1988, @brownDevelopmentRivalryDichoptic1999, @Candy2001, @Busse2009].
 At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the transduced waveforms (left panel of @fig-twoinputs) and in the Fourier spectrum (middle panel of @fig-twoinputs).
@@
 ![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
-Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](){#fig-twoinputs}
+Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).
+The effects of using different nonlinearities can be explored in the accompanying code.](){#fig-twoinputs}
 :::
@@
 ![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
-Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/4_twoinputs.pdf){#fig-twoinputs}
+Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).
+The effects of using different nonlinearities can be explored in the accompanying code.](Figures/4_twoinputs.pdf){#fig-twoinputs}
 :::
@@
 For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code temporal changes in contrast.
-These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
+These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion [@McKeefry1996].
 Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
@@
 Although SSVEP adaptation effects show strong tuning for orientation [@Vergeer2018; @Campbell1970] and spatial frequency [@Mecacci1976], there appear to be very measurements of the full SSVEP contrast response function before and after an extended period of adaptation.
-One exception is a study by Bach et al [@BachGreenleeBuhler1988] that examined the effects of adaptation on SSVEP responses who found a paradoxical increase in the 1F SSVEP amplitude following prolonged exposure to high-contrast stimuli across much of the probe contrast range.
+One exception is a study by Bach et al [@bachGreenleeBuhler1988] that examined the effects of adaptation on SSVEP responses who found a paradoxical increase in the 1F SSVEP amplitude following prolonged exposure to high-contrast stimuli across much of the probe contrast range.
 They explain this as being due to the presence of a nonlinear transducer function that saturates at high probe contrast driving the response to higher harmonic frequencies.
@@
 
-Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
-This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
+Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels (in the linear part of the log-contrast response function prior to saturation), and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis [@Campbell1970] (see @fig-sweepvep).
+This contrast level has been shown to correspond approximately with psychophysically measured detection thresholds [@campbellVisualEvokedPotential1972, @Norcia1986].
+One factor affecting the accuracy of this technique is the presence of noise in the EEG signal which will increase the apparent amplitude of the SSVEP responses across the measurement range.
+The effect of this noise is particulary apparent for recordings with low SNR or where the slope of the CRF is shallow (and therefore where small changes in the vertical offset of the fit result in large changes in the X-axis intercept).
+Accurate estimates of thresholds therefore require the experimenter to model this noise and compensate for its effects - as detailed in [@Norcia1989].
 
@@
 ![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
-The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](){#fig-sweepvep}
+The solid line is the linear regression fit to the lowest four data points prior to the satuating part of the curve, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](){#fig-sweepvep}
 :::
@@
 ![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
-The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/5_sweepplot.pdf){#fig-sweepvep}
+The solid line is the linear regression fit to the  lowest four data points prior to the satuating part of the curve, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/5_sweepplot.pdf){#fig-sweepvep}
 :::
@@
 A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrast levels.
-This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
+This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@reganRapidObjectiveRefraction1973,@Tyler1979].
 To avoid hysteresis or short-term adaptation effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979; @Norcia1985a; @Norcia1990].
@@
 
-This approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
+The approach of estimating visual thresholds by extrapolating SSVEP responses at higher stimulus levels has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
 In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
@@
 
-The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section [@Candy2001; @Pei2017].
+The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section.
 Although response- [@Morrone1986] or contrast gain control [@Candy2001; ; @Skoczenski1998] is measurable in infants as young as six weeks old, its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].

### 2024-11-28 review.tex
Now on 'thresholds'

--- 2024-05-22 review.tex
@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \makeatother
+
 \ifLuaTeX
@@
 reflect nonlinear processing in the visual system (Regan \& Regan,
-1988). SSVEP signals can also be elicited by periodic changes of
-stimulus properties other than achromatic and chromatic contrast, such
-as motion, stereo depth, and facial identity or expression (see Norcia
-\emph{et al.}, 2015, for an overview); however our focus here is on the
-contrast response.
+1988). Because they can arise from different neuronal computations,
+different populations and even different stages of the visual system,
+different harmonics can exhibit different input-output functions. For
+example, 1F and 2F responses can have different thresholds (Bobak
+\emph{et al.}, 1984) and scalp topographies (Regan, 1973) while Kaester
+et al (Kaestner \emph{et al.}, 2024) show both different CRF slopes and
+thresholds for 1F and 2F components generated by dynamic noise.
+
+SSVEP signals can also be elicited by periodic changes of stimulus
+properties other than achromatic and chromatic contrast, such as motion,
+stereo depth, and facial identity or expression (see Norcia \emph{et
+al.}, 2015, for an overview); however our focus here is on the contrast
+response.
 
@@
 figures in this review is available in python and R at:
-https://github.com/wadelab/contrastReviewPaper).
+https://github.com/wadelab/contrastReviewPaperVNS).
 
@@
 
+Although the SSVEP does seem to reflect the activity of relatively
+well-tuned neuronal populations , changing some aspect of the stimulus
+may change the nature of those populations. For example, Campbell and
+Maffei (Campbell \& Maffei, 1970) noted that densely-sampled
+measurements of contrast responses can reveal the presence of two,
+qualitatively different types of neurons that exhibit different
+log-linear contrast functions (see also Souza \emph{et al.} (2007)).
+Modeling such two-limbed contrast response functions using a single
+sigmoidal function is therefore an approximation.
+
 To understand the utility of the contrast SSVEP, it is helpful to
@@
 Baker \& Wade, 2017).
+
+An additional complexity is introduced by the fact that the EEG is an
+average population response and that the visual system contains many
+different types of neurons. As a stimulus changes in, for example,
+contrast it may selectively activate qualitatively different neuronal
+populations. This type
 
@@
 reduced by the presence of the target, in a contrast-dependent manner
-(Busse \emph{et al.}, 2009). At a single pair of (matched) contrast
+Busse \emph{et al.} (2009). At a single pair of (matched) contrast
 levels, we see a complex pattern of intermodulation terms in the
@@
 Additional harmonics are apparent in the Fourier spectrum (middle
-panel), and the contrast response function is suppressed (right panel).}
+panel), and the contrast response function is suppressed (right panel).
+The effects of using different nonlinearities can be explored in the
+accompanying code.}
 
@@
 will introduce second harmonic components even when the stimulus itself
-is modulated in an on-off fashion. Analogously, in the spatial domain,
-so-called `simple cells' are sensitive to the polarity of a spatial
-contrast modulation while `complex cells' respond to the presence of
-patterned spatial contrast irrespective of its spatial position (Hubel
-\& Wiesel, 1962). The SSVEP response to a contrast-reversing sine-wave
-grating therefore contains information about nonlinear computations
-performed across a range of retinal and cortical cell types.
+is modulated in an on-off fashion (McKeefry \emph{et al.}, 1996).
+Analogously, in the spatial domain, so-called `simple cells' are
+sensitive to the polarity of a spatial contrast modulation while
+`complex cells' respond to the presence of patterned spatial contrast
+irrespective of its spatial position (Hubel \& Wiesel, 1962). Both the
+amplitude and the phase of the SSVEP response to a single
+contrast-reversing sine-wave grating therefore contains information
+about nonlinear computations performed across a range of retinal and
+cortical cell types.
 
@@
 careful parametric variation of the input stimuli does allow us to fit
-models of early visual gain control using SSVEP data. For example, Tsai
-\emph{et al.} (2012) demonstrated that a gain control model gave a good
-account of the pattern of intermodulation responses produced by two
-overlaid patterns flickering at different frequencies. This was achieved
-by passing full stimulus waveforms through the transducer nonlinearity,
-and calculating the Fourier spectrum of the model output. Our more
-recent work on signal combination across eyes and space similarly
-demonstrated close correspondence between the predictions of a
-computational model and empirical data in humans (Baker \& Wade, 2017).
-More detailed modelling of intracortical recordings (Groen \emph{et
-al.}, 2022) has revealed details of the timecourse of gain control
-effects, specifically that normalization is delayed slightly relative to
-the initial visual response.
+models of early visual visual processing (e.g. (Zemon \& Gordon, 2006))
+and, by incorporating mathematical models of gain control mechanisms
+with either instantaneous, or temporally-integrated gain control, it is
+possible to model gain control using SSVEP data. For example, (Tsai
+\emph{et al.}, 2012) demonstrated that a relatively simple gain control
+model gave a good account of the pattern of intermodulation responses
+produced by two overlaid patterns flickering at different frequencies.
+This was achieved by passing full stimulus waveforms through the
+transducer nonlinearity, and calculating the Fourier spectrum of the
+model output. Our more recent work on signal combination across eyes and
+space similarly demonstrated close correspondence between the
+predictions of a computational model and empirical data in humans (Baker
+\& Wade, 2017). More detailed modelling of intracortical recordings
+(Groen \emph{et al.}, 2022) has revealed details of the timecourse of
+gain control effects, specifically that normalization is delayed
+slightly relative to the initial visual response.
 
@@
 2018) and spatial frequency (Mecacci \& Spinelli, 1976), there appear to
-be no studies measuring the full SSVEP contrast response function before
-and after an extended period of adaptation.
+be very measurements of the full SSVEP contrast response function before
+and after an extended period of adaptation. One exception is a study by
+Bach et al (Bach \emph{et al.}, 1988) that examined the effects of
+adaptation on SSVEP responses who found a paradoxical increase in the 1F
+SSVEP amplitude following prolonged exposure to high-contrast stimuli
+across much of the probe contrast range. They explain this as being due
+to the presence of a nonlinear transducer function that saturates at
+high probe contrast driving the response to higher harmonic frequencies.
+Adaptation reduces the apparent contrast of the input, leading to less
+distortion, a more sinusoidal output and a concentration of the
+frequency-tagged response at lower harmonics. This explanation is
+conceptually related to the `supersaturation' effect that we observe in
+Figure 4.
 
@@
 estimating a threshold is achieved by measuring the contrast response
-function at relatively high levels, and extrapolating back along the
-function (either contrast vs response measured at a constant spatial
-frequency or spatial frequency vs response at a constant contrast level)
-to estimate its intercept with the x-axis (see
-Figure~\ref{fig-sweepvep}). This contrast level was shown to correspond
-approximately with psychophysically measured detection thresholds
-(Norcia \emph{et al.}, 1986).
+function at relatively high levels (in the linear part of the
+log-contrast response function prior to saturation), and extrapolating
+back along the function (either contrast vs response measured at a
+constant spatial frequency or spatial frequency vs response at a
+constant contrast level) to estimate its intercept with the x-axis
+(Campbell \& Maffei, 1970) (see Figure~\ref{fig-sweepvep}). This
+contrast level has been shown to correspond approximately with
+psychophysically measured detection thresholds Norcia \emph{et al.}
+(1986). One factor affecting the accuracy of this technique is the
+presence of noise in the EEG signal which will increase the apparent
+amplitude of the SSVEP responses across the measurement range. The
+effect of this noise is particulary apparent for recordings with low SNR
+or where the slope of the CRF is shallow (and therefore where small
+changes in the vertical offset of the fit result in large changes in the
+X-axis intercept). Accurate estimates of thresholds therefore require
+the experimenter to model this noise and compensate for its effects - as
+detailed in (Norcia \emph{et al.}, 1989).
 
@@
 detection threshold can be estimated from sweep VEP data measured at
-higher contrasts. The solid line is the regression fit to the lowest
-four data points, and the dashed line extrapolates the fit back to
-determine the contrast value when y=0, which gives a threshold estimate
-(red point).}
+higher contrasts. The solid line is the linear regression fit to the
+lowest four data points prior to the satuating part of the curve, and
+the dashed line extrapolates the fit back to determine the contrast
+value when y=0, which gives a threshold estimate (red point).}
 
@@
 which the stimulus changed its contrast, spatial frequency or some other
-property, throughout a trial (Tyler \emph{et al.}, 1979). To avoid
-hysteresis effects, the sweep is sometimes conducted both up and down in
-the same experiment (Tyler \emph{et al.}, 1979; Norcia \& Tyler,
-1985\emph{a}; Norcia \emph{et al.}, 1990). The sweep VEP (really, a
-sweep `SSVEP') technique is now commonly-used to obtain a rapid and
-objective measurement of visual acuity. In particular, because of its
-relative speed and simplicity, this technique has now become a standard
-for conducting tests of visual acuity in very young subjects or where
-behavioural tests are not appropriate (Ridder, 2004; Bach \emph{et al.},
-2008; Hoffmann \emph{et al.}, 2017; Bach \& Farmer, 2020).
-
-This approach has revealed much about the development of visual
-abilities in infants (Harris \emph{et al.}, 1976; Atkinson \emph{et
-al.}, 1979; Braddick \emph{et al.}, 1986). In general, SSVEP
-measurements of infant vision have revealed that contrast sensitivity
-for both achromatic and chromatic contrast as well as stereoscopic depth
-perception develops earlier than had been supposed previously based on
-behavioural readouts (Dobson \emph{et al.}, 1978; Norcia \& Tyler,
-1985\emph{b}) with both chromatic and achromatic contrast detection
-reaching near-adult levels by around six months. Spatial acuity as
-measured by SSVEP reaches adult levels more slowly, but near-adult
-levels are recorded around one year (Norcia \& Tyler, 1985\emph{b};
-Hamer \emph{et al.}, 1989) compared to around six to seven years with
-behavioural measures (Atkinson \& Braddick, 1983; Ellemberg \emph{et
-al.}, 1999). At least some of this difference is likely due to the
-relative objectivity and high SNR of the SSVEP technique compared to
+property, throughout a trial Tyler \emph{et al.} (1979). To avoid
+hysteresis or short-term adaptation effects, the sweep is sometimes
+conducted both up and down in the same experiment (Tyler \emph{et al.},
+1979; Norcia \& Tyler, 1985\emph{a}; Norcia \emph{et al.}, 1990). The
+sweep VEP (really, a sweep `SSVEP') technique is now commonly-used to
+obtain a rapid and objective measurement of visual acuity. In
+particular, because of its relative speed and simplicity, this technique
+has now become a standard for conducting tests of visual acuity in very
+young subjects or where behavioural tests are not appropriate (Ridder,
+2004; Bach \emph{et al.}, 2008; Hoffmann \emph{et al.}, 2017; Bach \&
+Farmer, 2020).
+
+The approach of estimating visual thresholds by extrapolating SSVEP
+responses at higher stimulus levels has revealed much about the
+development of visual abilities in infants (Harris \emph{et al.}, 1976;
+Atkinson \emph{et al.}, 1979; Braddick \emph{et al.}, 1986). In general,
+SSVEP measurements of infant vision have revealed that contrast
+sensitivity for both achromatic and chromatic contrast as well as
+stereoscopic depth perception develops earlier than had been supposed
+previously based on behavioural readouts (Dobson \emph{et al.}, 1978;
+Norcia \& Tyler, 1985\emph{b}) with both chromatic and achromatic
+contrast detection reaching near-adult levels by around six months.
+Spatial acuity as measured by SSVEP reaches adult levels more slowly,
+but near-adult levels are recorded around one year (Norcia \& Tyler,
+1985\emph{b}; Hamer \emph{et al.}, 1989) compared to around six to seven
+years with behavioural measures (Atkinson \& Braddick, 1983; Ellemberg
+\emph{et al.}, 1999). At least some of this difference is likely due to
+the relative objectivity and high SNR of the SSVEP technique compared to
 other methods such as preferential looking, which require careful
@@
 The SSVEP technique has also been used to study the development of the
-contrast gain control mechanisms described in the previous section
-(Candy \emph{et al.}, 2001; Pei \emph{et al.}, 2017). Although contrast
-gain control is measurable in infants as young as six weeks old (Morrone
-\& Burr, 1986; Skoczenski \& Norcia, 1998; Candy \emph{et al.}, 2001),
-its development appears to be slower, with adult levels being reached at
-approximately 11 years (Pei \emph{et al.}, 2017).
+contrast gain control mechanisms described in the previous section.
+Although response- (Morrone \& Burr, 1986) or contrast gain control
+{[}Candy \emph{et al.} (2001); ; Skoczenski \& Norcia (1998){]} is
+measurable in infants as young as six weeks old, its development appears
+to be slower, with adult levels being reached at approximately 11 years
+(Pei \emph{et al.}, 2017).
 
@@
 
+\bibitem[\citeproctext]{ref-bachGreenleeBuhler1988}
+Bach M, Greenlee M \& Buhler B (1988). Contrast adaptation can increase
+visually evoked potential amplitude. \emph{Clinical Vision Sciences}
+\textbf{3,} 185--194.
+
 \bibitem[\citeproctext]{ref-Bach2008}
@@
 
+\bibitem[\citeproctext]{ref-BobakIOVS1984}
+Bobak P, Bodis-Wollner I, Harnois C \& Thornton J (1984). VEPs in humans
+reveal high and low spatial contrast mechanisms. \emph{Investigative
+Ophthalmology \& Visual Science} \textbf{25,} 980--983.
+
 \bibitem[\citeproctext]{ref-Boynton2009}
@@
 
+\bibitem[\citeproctext]{ref-brownDevelopmentRivalryDichoptic1999}
+Brown RJ, Candy TR \& Norcia AM (1999).
+\href{https://www.ncbi.nlm.nih.gov/pubmed/10586959}{Development of
+rivalry and dichoptic masking in human infants}. \emph{Investigative
+Ophthalmology \& Visual Science} \textbf{40,} 3324--3333.
+
 \bibitem[\citeproctext]{ref-Buracas2007}
@@
 \textbf{181,} 576--593.
+
+\bibitem[\citeproctext]{ref-campbellVisualEvokedPotential1972}
+Campbell FW \& Kulikowski JJ (1972).
+\href{https://doi.org/10.1113/jphysiol.1972.sp009801}{The visual evoked
+potential as a function of contrast of a grating pattern}. \emph{The
+Journal of Physiology} \textbf{222,} 345--356.
 
@@
 cortex}. \emph{J Physiol} \textbf{160,} 106--154.
+
+\bibitem[\citeproctext]{ref-KaestnerTVST2024}
+Kaestner M, Chen YD, Clement C, Hodges A \& Norcia AM (2024).
+\href{https://doi.org/10.1167/tvst.13.2.21}{Two disparity channels in
+human visual cortex with different contrast and blur sensitivity}.
+\emph{Translational Vision Science \& Technology} \textbf{13,} 21--21.
 
@@
 
+\bibitem[\citeproctext]{ref-Nakayama1982}
+Nakayama K \& Mackeben M (1982).
+\href{https://doi.org/10.1016/0042-6989(82)90138-9}{Steady state visual
+evoked potentials in the alert primate}. \emph{Vision Research}
+\textbf{22,} 1261--1271.
+
 \bibitem[\citeproctext]{ref-Niell2010}
@@
 
+\bibitem[\citeproctext]{ref-Norcia1989}
+Norcia AM, Tyler CW, Hamer RD \& Wesemann W (1989).
+\href{https://doi.org/10.1016/0042-6989(89)90048-5}{Measurement of
+spatial contrast sensitivity with the swept contrast {VEP}}.
+\emph{Vision Research} \textbf{29,} 627--637.
+
 \bibitem[\citeproctext]{ref-OConnor2002}
@@
 
+\bibitem[\citeproctext]{ref-reganRapidObjectiveRefraction1973}
+REGAN D (1973). Rapid {Objective Refraction Using Evoked Brain
+Potentials}. \emph{Investigative Ophthalmology \& Visual Science}
+\textbf{12,} 669--679.
+
+\bibitem[\citeproctext]{ref-reganEvokedPotentialsSpecific1973}
+Regan D (1973).
+\href{https://doi.org/10.1016/0042-6989(73)90237-X}{Evoked potentials
+specific to spatial patterns of luminance and colour}. \emph{Vision
+Research} \textbf{13,} 2381--2402.
+
 \bibitem[\citeproctext]{ref-Regan1988}
@@
 
+\bibitem[\citeproctext]{ref-Souza2007}
+Souza GS, Gomes BD, Saito CA, Silva Filho M da \& Silveira LCL (2007).
+Spatial luminance contrast sensitivity measured with transient {VEP}:
+Comparison with psychophysics and evidence of multiple mechanisms.
+\emph{Investigative ophthalmology \& visual science} \textbf{48,}
+3396--3404.
+
 \bibitem[\citeproctext]{ref-Stoughton2008}
@@
 
+\bibitem[\citeproctext]{ref-Zemon2006Dynamic}
+Zemon V \& Gordon J (2006).
+\href{https://doi.org/10.1016/j.visres.2006.07.007}{Luminance-contrast
+mechanisms in humans: Visual evoked potentials and a nonlinear model}.
+\emph{Vision Research} \textbf{46,} 4163--4180.
+
 \bibitem[\citeproctext]{ref-Zheng2019}

### 2024-11-28 review.qmd
On 'binocular visual development'

@@
 The approach of estimating visual thresholds by extrapolating SSVEP responses at higher stimulus levels has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
-In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+In general, SSVEP measurements of infant vision have revealed that contrast sensitivity and acuity for both achromatic and chromatic stimuli as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b].
+For example, using SSVEPs, Norcia and Tyler [@Norcia1985b] found acuity thresholds in 1 month old infants that were roughly half those estimated previously from behavioural experiments (reviewed in [@brownDevelopmentVisualSensitivity1990a]).
+SSVEP measures of thresholds for detecting purely chromatic stimuli find responses to isoluminant stimuli by around 5 to 6 weeks with contrast detection reaching near-adult levels by around six months [@morroneDevelopmentInfantContrast1993].
 Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983].
 At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
-However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
+In slightly older chilren, other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
+
+The ability to use information from both eyes
 

### 2024-11-28 responseToReviews.md
On 'binocular visual development'
Almost finished with R1

@@
 
+> We have made this more concrete.
+We have taken out the reference to steroscopic depth (because it is not directly related to contrast).
+It is hard for us to directly compare the values for contrast sensitivity from FPL vs VEP - the FPL values in particular seem to span quite a range.
+The figure in Brown 1990 suggests a x2 difference in detection threshold (if we are reading the plot correctly) but we are happy to accept higher values if the reviewer can point us to the relevant FPL literature.
+
 The authors may wish to cite Norcia et al., (NeuroReport, 2000) for an early measurement of suppression in amblyopia using measurements of the contrast response function in each eye in the presence of a high contrast dichoptic masker in the other.
@@
 
+> We think these were both in already but we have explicitly grouped them in the section on amblyopia to make their relevance clearer.
+
 Schade (JOSA, 1956) was the first to measure the contrast sensitivity function of the human visual system as noted by Campbell and Robson (J.
@@
 Campbell and Green (1965, cited) credit Schade for how to make sinewave gratings on an oscilloscope but not for having made the first measurement.
+
+> Included
 
@@
 
+> Noted and altered
+
 Page 12, better to state whether attention effects were consistent with response gain or contrast gain to make the result more concrete for the reader.
+
+> Reynolds and Heeger (and Boynton) point out that the range of different types of change reported in the literature (response, contrast or mixtures of both) is explained by subtle differences in stimulus configuration (for example, the size and location of the 'probe' and any masking components relative to the size of the 'attentional spotlight').
+We now note that the underlying model is one of contrast gain control (multiplicative amplification of the input rather than the output).
 
@@
 
+> Originally, we envisaged this document as a primer on SSVEP contrast measurements for those coming to the field for the first time.
+In particular, we were struck by the large number of recent papers that use multi-frequency SSVEP contrast flicker as an input to BCI systems.
+There have been at least 400 such publications ('SSVEP AND BCI') in the last 5 years alone (compared to around 10 for amblyopia ('SSVEP AND amblyopia')and 80 for ('SSVEP AND clinical').
+Some of these publications measure dynamic changes in covert spatial attention, some measure changes in featural attention, some measure fixation).
+We know from personal experience that many of the researchers using contrast flicker as a BCI input would benefit from a slightly more applied guide such as this and the BCI section adds very little to the overall length of the manuscipt at this point.
+
 Page, 14. It’s not clear from the text how one would study feedback with the SSVEP, please elaborate.
+
+> In the paper we write that we would look for modulation of subcortical responses due to top-down signals.
+The most natural things to look for might be interocular normalization (because eye layers are segregated in the LGN) and attentional gain control (because covert spatial attentional modulation must arise in cortex).
+This the type of feedback we are thinking of - not cortico-cortico feedback (which is hard to isolate with SSVEP)
 

### 2024-11-28 review.qmd
Almost finished with R1

@@
 The human response to contrast can be studied using many different techniques.
-Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965], defined as the inverse of the lowest contrast that can be reliably detected.
+Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965, @schadeOpticalPhotoelectricAnalog1956], defined as the inverse of the lowest contrast that can be reliably detected.
 But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG.
@@
 In general, SSVEP measurements of infant vision have revealed that contrast sensitivity and acuity for both achromatic and chromatic stimuli as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b].
-For example, using SSVEPs, Norcia and Tyler [@Norcia1985b] found acuity thresholds in 1 month old infants that were roughly half those estimated previously from behavioural experiments (reviewed in [@brownDevelopmentVisualSensitivity1990a]).
-SSVEP measures of thresholds for detecting purely chromatic stimuli find responses to isoluminant stimuli by around 5 to 6 weeks with contrast detection reaching near-adult levels by around six months [@morroneDevelopmentInfantContrast1993].
+For example, using SSVEPs, Norcia and Tyler [@Norcia1985b] found acuity thresholds in one month old infants that were roughly half those estimated previously from behavioural experiments (reviewed in [@brownDevelopmentVisualSensitivity1990a]).
+The ability to make very sensitive measurements of early visual function in infants also allows researchers to monitor the trajectory of visual development and identify different temporal epochs of developmental change [@Norcia1990].
+
+SSVEP measures of thresholds for detecting purely chromatic stimuli find responses to isoluminant stimuli by around 5 to six weeks with contrast detection reaching near-adult levels by around six months [@morroneDevelopmentInfantContrast1993].
 Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983].
+
 At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
 In slightly older chilren, other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
-
-The ability to use information from both eyes
 
@@
 SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye (see [@fig-plotclinical]b) across the contrast range [@Baker2015; @Lygo2021].
+Measurments of interocular contrast modulation (for example, [@norciaChangesCorticalActivity2000,@Hou-houchuan2021] can also provide information about the way that amblyopia changes the balance of excitation and inhibition between eyes.
+
 There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.
@@
 @Pei2014 used a sweep-VEP method with counterphase flickering stimuli, and found weaker responses in autistic children at spatial frequencies around 8c/deg, compared with age-matched controls.
-This was subsequently replicated in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
+This was subsequently extended to measurements of contrast sensitivity in a further pediatric sample by @Vilidaite2018 (see [@fig-plotclinical]c), who additionally found weaker responses in autistic adults at the second harmonic (using on/off flicker).
 Interestingly this study replicated its key findings in a *Drosophila* genetic model of autism (*Nhe3* mutations), illustrating the translational potential of the steady-state approach, as well as identifying a possible biomarker for autism.
@@
 These effects also appear to differ between chromatic and achromatic pathways [@DiRusso1999b] - perhaps as a result of the different levels of nonlinear gain control in the early pre-cortical magno-, parvo- and konio-cellular pathways [@Derrington1984b; @Kaplan1986; @Lee1990; @Solomon2005].
-In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a gain control computation [@Reynolds2009; @Boynton2009].
-This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating subtle interactions between the size of the attentional 'spotlight' and the stimulus configuration which rationalise many apparent contradictions in the literature.
+In the late 2000s a comprehensive theoretical model for attentional modulation was developed that framed it as a contrast gain control computation [@Reynolds2009; @Boynton2009].
+This framework has proven to be influential - explaining a wide range of phenomena from the earlier literature and demonstrating subtle interactions between the size of the attentional 'spotlight' and the stimulus configuration which rationalise many apparent contradictions in the literature - in particular reports of a response- rather than a contrast-gain control phenomenon.
 Direct measurements of attentional modulation of achromatic SSVEP signals are broadly consistent with this model [@Lauritzen2010; @Hou2016; @Martinovic2018] and confirm the relatively weaker role of spatial attention on responses driven by chromatic stimuli - particularly those that isolate the opponent S-(L+M) cone pathway [@Wang2011; @Highsmith2010].

### 2024-11-29 responseToReviews.md
Merge branch 'main' of main ithub.com:wadelab/contrastReviewPaper

@@
 In particular, we were struck by the large number of recent papers that use multi-frequency SSVEP contrast flicker as an input to BCI systems.
-There have been at least 400 such publications ('SSVEP AND BCI') in the last 5 years alone (compared to around 10 for amblyopia ('SSVEP AND amblyopia')and 80 for ('SSVEP AND clinical').
+There have been at least 400 such publications ('SSVEP AND BCI') in the last 5 years alone (Kcompared to around 10 for amblyopia ('SSVEP AND amblyopia')and 80 for ('SSVEP AND clinical').
 Some of these publications measure dynamic changes in covert spatial attention, some measure changes in featural attention, some measure fixation).

### 2024-11-29 responseToReviews.md
Now onto the other reviewer. Losing the will to live...

@@
 > We thank the reviewer for these positive comments.
-In general this was intended to be an overview in the sense of a 'primer' for SSVEP rather than a detailed historical account of the field - hence the inclusion of code examples and relatively domain-agnostic language.
-However, the reviewer raises some important general issues and we now address them: Specifically, we include references to the idea that SSVEP can engage more than one type of neuron, that the population responses is not trivially derived from what is known about the properties of individual neurons and that fitting with a sigmoidal response function may mask some complexities in the CRF that are more readily seen with denser sampling.
+In general this was intended to be an overview in the sense of a 'primer' for researchers coming to the SSVEP for the first time rather than a detailed historical account of the field - hence the inclusion of code examples and relatively domain-agnostic language.
+However, the reviewer raises some important general issues and we are happy to address them: Specifically, we include references to the idea that SSVEP can engage more than one type of neuron, that the population responses is not trivially derived from what is known about the properties of individual neurons and that fitting with a sigmoidal response function may mask some complexities in the CRF that are more readily seen with denser sampling.
 We think the Bobak reference is IOVS 1984 and we have also included the Kaestner TVST 2024 reference (thank you for drawing our attention to both of these).
@@
 
->
+> We have expanded a little but are also conscious of limitations to this approach (see also the discussion re unique inputs and system order below).
+We think there might be some ways to relax these constraints a little by incorporating information about likely transducer functions at different stages (essentially fitting parameters rather than performing a full system identification) but this still requires many inputs.
 
@@
 
-.
+> This is a good point and we now include the Boyd reference in the section on fingerprinting.
 
@@
 
+> We thank the reviewer for their helpful comments.
+As we noted above, our intention here was to provide an introduction to the SSVEP technique with appropriate simple code samples that would allow relative newcomers to the field to implement the the technique.
+We are, however, keen to avoid misrepresenting the history of the technique and are happy to include appropriate references.
+
 1. All abbreviations should be defined upon first use.
+
+> Done
+
 2.
@@
 Thus, the response at 1F to this stimulus is in fact a nonlinear response.
+
+> Thank you for this explanation.
+We have now reworded the explanation to make this clearer.
+We did take care to define what we mean by 'on-off' flicker in the preceding sentence but we now clarify the alternative 'appearence/disappearence' terminology.
+The point we were making was that odd harmonic responses (1F) are a feature of this this type of modulation because the responses to the appearance and disappearance of a pattern (or mean field) need not be completely symmetrical (See Zemon, et al 1988 for an explicit demonstration).
+
 This type of stimulation does not separate contributions from “independent populations of on- and off-cells.”   There are other ssVEP techniques developed that do tap ON and OFF pathways selectively, which are not cited here.
 Contrast-reversal stimuli which involve symmetrical modulation of light and dark elements cancel odd harmonic frequency components in the response, however, both kinds of stimulus techniques mentioned here elicit only nonlinear responses.
+
+> We have clearly not been sufficiently clear here.
+We do not think we used the term 'separation' and we agree that a) there are infrequently-used paradigms that can target ON and OFF pathways more effectively (for example, 'sawtooth' temporal modulations that provide asymmetrical onset or offset edges) and b) responses from spatial features of differing polarity will tend to be combined at the scalp as long as the spatial frequency is high enough.
+We have expanded this section to attempt to make the general point clear.
+
 3. The statement “SSVEPs have proven to be an excellent measure of early chromatic processing as well” again fails to cite seminal work in this area.
+
+> We have included three earlier papers by Regan on this topic.
+
 4.
 The statement “This normalization, achieved through a computation called ‘contrast gain control’” again neglects to cite the original work that defined “contrast gain control” and the authors ignored a nonlinear biophysical model proposed to explain this phenomenon in ssVEP contrast response functions (which was also applied in developmental work on infants and children).
+
+>
+
 5.
@@
 The text also states that “This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.” When was full-wave rectification performed?
+
+>
+
 6.
@@
 They do not, however, cite the foundational work in this field.
+
 7.
 There is a large body of work that used ssVEPs to investigate contrast processing developmentally (infants through adulthood) and in various disease states, which is mostly missed in this review article.
+
 8. Figures should be labeled.

### 2024-11-29 review.qmd
Now onto the other reviewer. Losing the will to live...

@@
 ---
-title: "Measuring contrast processing in the visual system using SSVEP"
+title: "Measuring contrast processing in the visual system using the steady state visually evoked potential (SSVEP)"
 author:
@@
 One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged' (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain.
-SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high SNR, and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
+SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high signal to noise ratio (SNR), and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
 Here we describe how SSVEPs have been used to study visual contrast over the past 70 years [@Dawson1954].
@@
 Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965, @schadeOpticalPhotoelectricAnalog1956], defined as the inverse of the lowest contrast that can be reliably detected.
-But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG.
+But neural responses can also be measured more directly using techniques such as functional magnetic resonance imaging (fMRI), magnetoencephalography (MEG), and electroencephalography (EEG).
 Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.
@@
 The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2005; @Di-Russo2007].
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual visually-evoked potentials (VEPs) reveals multiple generators throughout visual cortex [@Di-Russo2005; @Di-Russo2007].
 The flickering stimulus entrains neural population responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
-Two common variants involve sinusoidal on-off flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
-On-off flicker can drive independent populations of on- and off-cells in the retina once per cycle and can therefore produce a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on.
-Counterphase flicker contains two transients per cycle and therefore does not produce a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
-Because square-waves are spectrally broad-band, square wave flicker tends to produce more complex spectral harmonics than sine-wave flicker.
+
+Two common stimulus variants involve sinusoidal on-off (or 'appearance/disappearance') flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
+On-off flicker can excite independent populations of on- and off-cells in the visual system once per cycle.
+For spatial patterns, the contributions of individual cortical cells to these excitations are thought to sum at the scalp and generate an average of their onset responses on each cycle.
+For very low spatial frequencies (including zero) resposes at the scalp will be dominated by a single on- or off- cell type on each half cycle.
+
+If the amplitudes and waveforms of responses to the stimulus appearance and disappearnce are not perfectly balanced, a response at the fundamental flicker frequency, known as 1F, and its integer harmonics can be produced.
+For stimuli of any spatial frequency (including zero), an imbalance can be caused by differences in either the onset- and offset- responses of either on- or off-cells.
+For example, if only on-sensitive neurons were present (firing to the onset, but not the offset of local increases in luminance) the response to the on/off (or 'appearance/disappearance') presentation of a grating of any spatial frequency below the resolution limit would consist entirely of odd harmonics.
+
+By comparison, counterphase flickering patterns generate two essentially identical transients per cycle and therefore does not produce a response at 1F, only even harmonics: 2F, 4F, 6F and so on.
+Because square-waves are spectrally broad-band, square wave flicker (either on/off or contrast reversig) tends to produce additional spectral harmonics compared to sine-wave flicker.
 
@@
 Because they can arise from different neuronal computations, different populations and even different stages of the visual system, different harmonics can exhibit different input-output functions.
-For example, 1F and 2F responses can have different thresholds [@BobakIOVS1984] and scalp topographies [@reganEvokedPotentialsSpecific1973] while Kaester et al [@KaestnerTVST2024] show both different CRF slopes and thresholds for 1F and 2F components generated by dynamic noise.
+For example, 1F and 2F responses can have different thresholds [@BobakIOVS1984] and scalp topographies [@reganEvokedPotentialsSpecific1973] while Kaester et al [@KaestnerTVST2024] show both different response slopes and thresholds for 1F and 2F components generated by dynamic noise.
 
@@
 It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
-Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@McKeefry1996; @Di-Russo2001]   (see also [@Baseler1997]).
+Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@reganEvokedPotentialsSpecific1973, @reganColourCodingPattern1975, @reganEffectStimulusColour1966a, @McKeefry1996; @Di-Russo2001] (see also [@Baseler1997]).
 
@@
 For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code temporal changes in contrast.
-These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion [@McKeefry1996].
+These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the lateral geniculate nucleus (LGN) [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion [@McKeefry1996].
 Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
@@
 In principle, that information could be recovered from the SSVEP signal - a possibility recognised in the early days of the technique [@Regan1988].
-Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual visual processing (e.g.
+Although characterising the complete set of computations along entire processing pathway is challenging, and a full system identification may require the use of multiple different input frequencies [@boydMeasuringVolterraKernels1983], careful parametric variation of the input stimuli does allow us to fit models of early visual visual processing (e.g.
 [@Zemon2006Dynamic]) and, by incorporating mathematical models of gain control mechanisms with either instantaneous, or temporally-integrated gain control, it is possible to model gain control using SSVEP data.
@@
 This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.
+More recent work has used this approach to study binocular integration in both normal subjects and amblyopes[@Hou-houchuan2021,@Hou2020].
 Our more recent work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
@@
 In principle, this enhancement may be driven both by modulation of the underlying signal or noise characteristics, or by additional decision-theoretic factors such as reduction in spatial uncertainty [@Gould2007; @Petrov2006].
-Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates [@Luck1997; @Marcus2002; @McAdams1999; @Mehta2000a; @Mehta2000b], but with the advent of spatially-resolved human brain imaging methods in the late 1990s it became apparent that spatial attention was linked to frank changes in both fMRI BOLD responses
+Early experiments with non-human primates showed little evidence for attentionally-driven changes in neuronal spike rates [@Luck1997; @Marcus2002; @McAdams1999; @Mehta2000a; @Mehta2000b], but with the advent of spatially-resolved human brain imaging methods in the late 1990s it became apparent that spatial attention was linked to frank changes in both fMRI
 [@Tootell1998; @Brefczynski1999; @Buracas2007; @Gandhi1999; @Kastner1999; @Li2008; @Murray2008; @Silver2007; @Somers1999] and EEG signals [@Ding2006; @Muller2000; @Morgan1996; @Muller1998].
@@
 
-One widespread recent application of the SSVEP technique is in the design of brain-computer interfaces, which seek to control some aspect of a computer using neural signals.
+One widespread recent application of the SSVEP technique is in the design of brain-computer interfaces (BCIs), which seek to control some aspect of a computer using neural signals.
 The high SNR and precise frequency resolution of SSVEPs are advantages for this approach.

### 2024-11-29 review.tex
Now onto the other reviewer. Losing the will to live...

@@
 % Make \paragraph and \subparagraph free-standing
-\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
 
@@
 \makeatother
-
 \ifLuaTeX
@@
 \hypersetup{
-  pdftitle={Measuring contrast processing in the visual system using SSVEP},
+  pdftitle={Measuring contrast processing in the visual system using the steady state visually evoked potential (SSVEP)},
   pdfauthor={Alex R. Wade; Daniel H. Baker},
@@
 
-\title{Measuring contrast processing in the visual system using SSVEP}
+\title{Measuring contrast processing in the visual system using the
+steady state visually evoked potential (SSVEP)}
 \author{Alex R. Wade \and Daniel H. Baker}
@@
 visually-evoked responses: they have relatively unambiguous ouput
-measures, a high SNR, and they allow us to analyze interactions between
-stimulus components using a convenient mathematical framework. Here we
-describe how SSVEPs have been used to study visual contrast over the
-past 70 years (Dawson, 1954). Because our thinking about SSVEPs is
-well-described by simple mathematical models, we embed code that
-illustrates key steps in the modelling and analysis. This paper can
-therefore be used both as a review of the use of SSVEP in measuring
-human contrast processing, and as an interactive learning aid.
+measures, a high signal to noise ratio (SNR), and they allow us to
+analyze interactions between stimulus components using a convenient
+mathematical framework. Here we describe how SSVEPs have been used to
+study visual contrast over the past 70 years (Dawson, 1954). Because our
+thinking about SSVEPs is well-described by simple mathematical models,
+we embed code that illustrates key steps in the modelling and analysis.
+This paper can therefore be used both as a review of the use of SSVEP in
+measuring human contrast processing, and as an interactive learning aid.
 
@@
 studied using many different techniques. Early work used psychophysical
-methods to measure contrast sensitivity (Campbell \& Green, 1965),
-defined as the inverse of the lowest contrast that can be reliably
-detected. But neural responses can also be measured more directly using
-techniques such as fMRI, MEG, and EEG. Here we will describe how an EEG
-method known as the steady state visually evoked potential (SSVEP)
+methods to measure contrast sensitivity Schade (1956), defined as the
+inverse of the lowest contrast that can be reliably detected. But neural
+responses can also be measured more directly using techniques such as
+functional magnetic resonance imaging (fMRI), magnetoencephalography
+(MEG), and electroencephalography (EEG). Here we will describe how an
+EEG method known as the steady state visually evoked potential (SSVEP)
 technique has contributed to our understanding of human contrast
@@
 The SSVEP is a continuous electrical response evoked in the brain by
-visual stimuli flickering at a constant frequency (Regan, 1966). For
-contrast-defined stimuli, such as sine-wave gratings, it is strongest at
-the occipital pole, adjacent to the early visual areas that generate the
-signal, although careful analysis of individual VEPs reveals multiple
-generators throughout visual cortex (Di Russo \emph{et al.}, 2005,
-2007). The flickering stimulus entrains neural population responses at
-multiples of the stimulus frequency, so continuous EEG data are
-typically analysed by taking the Fourier transform, and estimating the
-amplitude at these frequencies. Two common variants involve sinusoidal
-on-off flicker, where the stimulus alternates between a blank background
-and the peak contrast, and sinusoidal counterphase flicker, where the
-stimulus alternates in phase (i.e the black regions become white and the
-white regions become black). On-off flicker can drive independent
-populations of on- and off-cells in the retina once per cycle and can
-therefore produce a response at the fundamental flicker frequency, known
-as 1F, and its integer harmonics: 2F, 3F, 4F and so on. Counterphase
-flicker contains two transients per cycle and therefore does not produce
-a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
-Because square-waves are spectrally broad-band, square wave flicker
-tends to produce more complex spectral harmonics than sine-wave flicker.
+visual stimuli flickering at a constant frequency (Regan, 1966\emph{a}).
+For contrast-defined stimuli, such as sine-wave gratings, it is
+strongest at the occipital pole, adjacent to the early visual areas that
+generate the signal, although careful analysis of individual
+visually-evoked potentials (VEPs) reveals multiple generators throughout
+visual cortex (Di Russo \emph{et al.}, 2005, 2007). The flickering
+stimulus entrains neural population responses at multiples of the
+stimulus frequency, so continuous EEG data are typically analysed by
+taking the Fourier transform, and estimating the amplitude at these
+frequencies.
+
+Two common stimulus variants involve sinusoidal on-off (or
+`appearance/disappearance') flicker, where the stimulus alternates
+between a blank background and the peak contrast, and sinusoidal
+counterphase flicker, where the stimulus alternates in phase (i.e the
+black regions become white and the white regions become black). On-off
+flicker can excite independent populations of on- and off-cells in the
+visual system once per cycle. For spatial patterns, the contributions of
+individual cortical cells to these excitations are thought to sum at the
+scalp and generate an average of their onset responses on each cycle.
+For very low spatial frequencies (including zero) resposes at the scalp
+will be dominated by a single on- or off- cell type on each half cycle.
+
+If the amplitudes and waveforms of responses to the stimulus appearance
+and disappearnce are not perfectly balanced, a response at the
+fundamental flicker frequency, known as 1F, and its integer harmonics
+can be produced. For stimuli of any spatial frequency (including zero),
+an imbalance can be caused by differences in either the onset- and
+offset- responses of either on- or off-cells. For example, if only
+on-sensitive neurons were present (firing to the onset, but not the
+offset of local increases in luminance) the response to the on/off (or
+`appearance/disappearance') presentation of a grating of any spatial
+frequency below the resolution limit would consist entirely of odd
+harmonics.
+
+By comparison, counterphase flickering patterns generate two essentially
+identical transients per cycle and therefore does not produce a response
+at 1F, only even harmonics: 2F, 4F, 6F and so on. Because square-waves
+are spectrally broad-band, square wave flicker (either on/off or
+contrast reversig) tends to produce additional spectral harmonics
+compared to sine-wave flicker.
 
@@
 \emph{et al.}, 1984) and scalp topographies (Regan, 1973) while Kaester
-et al (Kaestner \emph{et al.}, 2024) show both different CRF slopes and
-thresholds for 1F and 2F components generated by dynamic noise.
+et al (Kaestner \emph{et al.}, 2024) show both different response slopes
+and thresholds for 1F and 2F components generated by dynamic noise.
 
@@
 achromatic contrast, SSVEPs have proven to be an excellent measure of
-early chromatic processing as well (McKeefry \emph{et al.}, 1996; Di
-Russo \emph{et al.}, 2001\emph{a}) (see also (Baseler \& Sutter, 1997)).
+early chromatic processing as well McKeefry \emph{et al.} (1996) (see
+also (Baseler \& Sutter, 1997)).
 
@@
 temporal changes in contrast. These cells (Kuffler \emph{et al.}, 1957;
-Alpern, 1971), and cells with similar properties in the LGN (Levitt
-\emph{et al.}, 2001) and cortex (Hubel \& Wiesel, 1959; Movshon, 1975),
-will introduce second harmonic components even when the stimulus itself
-is modulated in an on-off fashion (McKeefry \emph{et al.}, 1996).
-Analogously, in the spatial domain, so-called `simple cells' are
-sensitive to the polarity of a spatial contrast modulation while
-`complex cells' respond to the presence of patterned spatial contrast
-irrespective of its spatial position (Hubel \& Wiesel, 1962). Both the
-amplitude and the phase of the SSVEP response to a single
-contrast-reversing sine-wave grating therefore contains information
-about nonlinear computations performed across a range of retinal and
-cortical cell types.
+Alpern, 1971), and cells with similar properties in the lateral
+geniculate nucleus (LGN) (Levitt \emph{et al.}, 2001) and cortex (Hubel
+\& Wiesel, 1959; Movshon, 1975), will introduce second harmonic
+components even when the stimulus itself is modulated in an on-off
+fashion (McKeefry \emph{et al.}, 1996). Analogously, in the spatial
+domain, so-called `simple cells' are sensitive to the polarity of a
+spatial contrast modulation while `complex cells' respond to the
+presence of patterned spatial contrast irrespective of its spatial
+position (Hubel \& Wiesel, 1962). Both the amplitude and the phase of
+the SSVEP response to a single contrast-reversing sine-wave grating
+therefore contains information about nonlinear computations performed
+across a range of retinal and cortical cell types.
 
@@
 technique (Regan \& Regan, 1988). Although characterising the complete
-set of computations along entire processing pathway is challenging,
-careful parametric variation of the input stimuli does allow us to fit
-models of early visual visual processing (e.g. (Zemon \& Gordon, 2006))
-and, by incorporating mathematical models of gain control mechanisms
-with either instantaneous, or temporally-integrated gain control, it is
-possible to model gain control using SSVEP data. For example, (Tsai
-\emph{et al.}, 2012) demonstrated that a relatively simple gain control
-model gave a good account of the pattern of intermodulation responses
-produced by two overlaid patterns flickering at different frequencies.
-This was achieved by passing full stimulus waveforms through the
-transducer nonlinearity, and calculating the Fourier spectrum of the
-model output. Our more recent work on signal combination across eyes and
-space similarly demonstrated close correspondence between the
-predictions of a computational model and empirical data in humans (Baker
-\& Wade, 2017). More detailed modelling of intracortical recordings
-(Groen \emph{et al.}, 2022) has revealed details of the timecourse of
-gain control effects, specifically that normalization is delayed
-slightly relative to the initial visual response.
+set of computations along entire processing pathway is challenging, and
+a full system identification may require the use of multiple different
+input frequencies (Boyd \emph{et al.}, 1983), careful parametric
+variation of the input stimuli does allow us to fit models of early
+visual visual processing (e.g. (Zemon \& Gordon, 2006)) and, by
+incorporating mathematical models of gain control mechanisms with either
+instantaneous, or temporally-integrated gain control, it is possible to
+model gain control using SSVEP data. For example, (Tsai \emph{et al.},
+2012) demonstrated that a relatively simple gain control model gave a
+good account of the pattern of intermodulation responses produced by two
+overlaid patterns flickering at different frequencies. This was achieved
+by passing full stimulus waveforms through the transducer nonlinearity,
+and calculating the Fourier spectrum of the model output. More recent
+work has used this approach to study binocular integration in both
+normal subjects and amblyopesHou \emph{et al.} (2020). Our more recent
+work on signal combination across eyes and space similarly demonstrated
+close correspondence between the predictions of a computational model
+and empirical data in humans (Baker \& Wade, 2017). More detailed
+modelling of intracortical recordings (Groen \emph{et al.}, 2022) has
+revealed details of the timecourse of gain control effects, specifically
+that normalization is delayed slightly relative to the initial visual
+response.
 
@@
 SSVEP measurements of infant vision have revealed that contrast
-sensitivity for both achromatic and chromatic contrast as well as
-stereoscopic depth perception develops earlier than had been supposed
+sensitivity and acuity for both achromatic and chromatic stimuli as well
+as stereoscopic depth perception develops earlier than had been supposed
 previously based on behavioural readouts (Dobson \emph{et al.}, 1978;
-Norcia \& Tyler, 1985\emph{b}) with both chromatic and achromatic
-contrast detection reaching near-adult levels by around six months.
-Spatial acuity as measured by SSVEP reaches adult levels more slowly,
-but near-adult levels are recorded around one year (Norcia \& Tyler,
-1985\emph{b}; Hamer \emph{et al.}, 1989) compared to around six to seven
-years with behavioural measures (Atkinson \& Braddick, 1983; Ellemberg
-\emph{et al.}, 1999). At least some of this difference is likely due to
-the relative objectivity and high SNR of the SSVEP technique compared to
-other methods such as preferential looking, which require careful
-measurement of the infant's gaze direction. However it should be noted
-that other groups have reported electrophysiological correlates of
-visual acuity that more closely match the behavioural measures (De
-Vries-Khoe \& Spekreijse, 1982).
+Norcia \& Tyler, 1985\emph{b}). For example, using SSVEPs, Norcia and
+Tyler (Norcia \& Tyler, 1985\emph{b}) found acuity thresholds in one
+month old infants that were roughly half those estimated previously from
+behavioural experiments (reviewed in (Brown, 1990)). The ability to make
+very sensitive measurements of early visual function in infants also
+allows researchers to monitor the trajectory of visual development and
+identify different temporal epochs of developmental change (Norcia
+\emph{et al.}, 1990).
+
+SSVEP measures of thresholds for detecting purely chromatic stimuli find
+responses to isoluminant stimuli by around 5 to six weeks with contrast
+detection reaching near-adult levels by around six months (Morrone
+\emph{et al.}, 1993). Spatial acuity as measured by SSVEP reaches adult
+levels more slowly, but near-adult levels are recorded around one year
+(Norcia \& Tyler, 1985\emph{b}; Hamer \emph{et al.}, 1989) compared to
+around six to seven years with behavioural measures (Atkinson \&
+Braddick, 1983; Ellemberg \emph{et al.}, 1999).
+
+At least some of this difference is likely due to the relative
+objectivity and high SNR of the SSVEP technique compared to other
+methods such as preferential looking, which require careful measurement
+of the infant's gaze direction. In slightly older chilren, other groups
+have reported electrophysiological correlates of visual acuity that more
+closely match the behavioural measures (De Vries-Khoe \& Spekreijse,
+1982).
 
@@
 Figure~\ref{fig-plotclinical}b) across the contrast range (Baker
-\emph{et al.}, 2015; Lygo \emph{et al.}, 2021). There are currently many
-novel binocular treatments for amblyopia under development, often
-involving virtual reality or stereo display systems designed to
-encourage the two eyes to work together. The steady-state approach may
-be more sensitive and objective than typical acuity measurements, and
-also has the potential to measure suppression between the eyes directly
-(e.g. Zheng \emph{et al.}, 2019; Hu \emph{et al.}, 2023; Du \emph{et
-al.}, 2023).
+\emph{et al.}, 2015; Lygo \emph{et al.}, 2021). Measurments of
+interocular contrast modulation (for example, Hou 侯川 \emph{et al.}
+(2021) can also provide information about the way that amblyopia changes
+the balance of excitation and inhibition between eyes.
+
+There are currently many novel binocular treatments for amblyopia under
+development, often involving virtual reality or stereo display systems
+designed to encourage the two eyes to work together. The steady-state
+approach may be more sensitive and objective than typical acuity
+measurements, and also has the potential to measure suppression between
+the eyes directly (e.g. Zheng \emph{et al.}, 2019; Hu \emph{et al.},
+2023; Du \emph{et al.}, 2023).
 
@@
 children at spatial frequencies around 8c/deg, compared with age-matched
-controls. This was subsequently replicated in a further pediatric sample
-by Vilidaite \emph{et al.} (2018) (see Figure~\ref{fig-plotclinical}c),
-who additionally found weaker responses in autistic adults at the second
-harmonic (using on/off flicker). Interestingly this study replicated its
-key findings in a \emph{Drosophila} genetic model of autism (\emph{Nhe3}
-mutations), illustrating the translational potential of the steady-state
-approach, as well as identifying a possible biomarker for autism.
+controls. This was subsequently extended to measurements of contrast
+sensitivity in a further pediatric sample by Vilidaite \emph{et al.}
+(2018) (see Figure~\ref{fig-plotclinical}c), who additionally found
+weaker responses in autistic adults at the second harmonic (using on/off
+flicker). Interestingly this study replicated its key findings in a
+\emph{Drosophila} genetic model of autism (\emph{Nhe3} mutations),
+illustrating the translational potential of the steady-state approach,
+as well as identifying a possible biomarker for autism.
 
@@
 1990s it became apparent that spatial attention was linked to frank
-changes in both fMRI BOLD responses (Tootell \emph{et al.}, 1998; Somers
-\emph{et al.}, 1999; Gandhi \emph{et al.}, 1999; Brefczynski \& DeYoe,
-1999; Kastner \emph{et al.}, 1999; Buracas \& Boynton, 2007; Silver
-\emph{et al.}, 2007; Li \emph{et al.}, 2008; Murray, 2008) and EEG
-signals (Morgan \emph{et al.}, 1996; Müller \emph{et al.}, 1998; Müller
-\& Hillyard, 2000; Ding \emph{et al.}, 2006).
+changes in both fMRI (Tootell \emph{et al.}, 1998; Somers \emph{et al.},
+1999; Gandhi \emph{et al.}, 1999; Brefczynski \& DeYoe, 1999; Kastner
+\emph{et al.}, 1999; Buracas \& Boynton, 2007; Silver \emph{et al.},
+2007; Li \emph{et al.}, 2008; Murray, 2008) and EEG signals (Morgan
+\emph{et al.}, 1996; Müller \emph{et al.}, 1998; Müller \& Hillyard,
+2000; Ding \emph{et al.}, 2006).
 
@@
 \& Lennie, 2005). In the late 2000s a comprehensive theoretical model
-for attentional modulation was developed that framed it as a gain
-control computation (Reynolds \& Heeger, 2009; Boynton, 2009). This
+for attentional modulation was developed that framed it as a contrast
+gain control computation (Reynolds \& Heeger, 2009; Boynton, 2009). This
 framework has proven to be influential - explaining a wide range of
@@
 stimulus configuration which rationalise many apparent contradictions in
-the literature. Direct measurements of attentional modulation of
-achromatic SSVEP signals are broadly consistent with this model
-(Lauritzen \emph{et al.}, 2010; Hou \emph{et al.}, 2016; Martinovic \&
-Andersen, 2018) and confirm the relatively weaker role of spatial
-attention on responses driven by chromatic stimuli - particularly those
-that isolate the opponent S-(L+M) cone pathway (Highsmith \& Crognale,
-2010; Wang \& Wade, 2011). The gain control model of attention can also
-be extended to SSVEP studies of feature-based attention which show that
-the modulatory effects can be targeted to the most informative neuronal
-populations (Verghese \emph{et al.}, 2012).
+the literature - in particular reports of a response- rather than a
+contrast-gain control phenomenon. Direct measurements of attentional
+modulation of achromatic SSVEP signals are broadly consistent with this
+model (Lauritzen \emph{et al.}, 2010; Hou \emph{et al.}, 2016;
+Martinovic \& Andersen, 2018) and confirm the relatively weaker role of
+spatial attention on responses driven by chromatic stimuli -
+particularly those that isolate the opponent S-(L+M) cone pathway
+(Highsmith \& Crognale, 2010; Wang \& Wade, 2011). The gain control
+model of attention can also be extended to SSVEP studies of
+feature-based attention which show that the modulatory effects can be
+targeted to the most informative neuronal populations (Verghese \emph{et
+al.}, 2012).
 
@@
 One widespread recent application of the SSVEP technique is in the
-design of brain-computer interfaces, which seek to control some aspect
-of a computer using neural signals. The high SNR and precise frequency
-resolution of SSVEPs are advantages for this approach. Typical studies
-may involve presenting an array of stimuli at different flicker
+design of brain-computer interfaces (BCIs), which seek to control some
+aspect of a computer using neural signals. The high SNR and precise
+frequency resolution of SSVEPs are advantages for this approach. Typical
+studies may involve presenting an array of stimuli at different flicker
 frequencies, and having the participant select one either by overt
@@
 
+\bibitem[\citeproctext]{ref-boydMeasuringVolterraKernels1983}
+Boyd S, Tang Y \& Chua L (1983).
+\href{https://doi.org/10.1109/TCS.1983.1085391}{Measuring {Volterra}
+kernels}. \emph{IEEE Transactions on Circuits and Systems} \textbf{30,}
+571--577.
+
 \bibitem[\citeproctext]{ref-Boynton2009}
@@
 macaque monkey}. \emph{J Neurosci} \textbf{27,} 5431--5436.
+
+\bibitem[\citeproctext]{ref-brownDevelopmentVisualSensitivity1990a}
+Brown AM (1990).
+\href{https://doi.org/10.1016/0042-6989(90)90173-I}{Development of
+visual sensitivity to light and color vision in human infants: {A}
+critical review}. \emph{Vision Research} \textbf{30,} 1159--1188.
 
@@
 of Vision} \textbf{16,} 16.
+
+\bibitem[\citeproctext]{ref-Hou2020}
+Hou C, Nicholas SC \& Verghese P (2020).
+\href{https://doi.org/10.1523/JNEUROSCI.2043-19.2020}{Contrast
+{Normalization} {Accounts} for {Binocular} {Interactions} in {Human}
+{Striate} and {Extra}-striate {Visual} {Cortex}}. \emph{The Journal of
+Neuroscience: The Official Journal of the Society for Neuroscience}
+\textbf{40,} 2753--2763.
+
+\bibitem[\citeproctext]{ref-Hou-houchuan2021}
+Hou 侯川 C, Tyson TL, Uner IJ, Nicholas SC \& Verghese P (2021).
+\href{https://doi.org/10.1523/JNEUROSCI.0268-21.2021}{Excitatory
+{Contribution} to {Binocular} {Interactions} in {Human} {Visual}
+{Cortex} {Is} {Reduced} in {Strabismic} {Amblyopia}}. \emph{The Journal
+of Neuroscience: The Official Journal of the Society for Neuroscience}
+\textbf{41,} 8632--8643.
 
@@
 
+\bibitem[\citeproctext]{ref-morroneDevelopmentInfantContrast1993}
+Morrone MC, Burr DC \& Fiorentini A (1993).
+\href{https://doi.org/10.1016/0042-6989(93)90133-H}{Development of
+infant contrast sensitivity to chromatic stimuli}. \emph{Vision
+Research} \textbf{33,} 2535--2552.
+
 \bibitem[\citeproctext]{ref-Morrone1982}
@@
 
+\bibitem[\citeproctext]{ref-norciaChangesCorticalActivity2000}
+Norcia AM, Harrad RA \& Brown RJ (2000). Changes in cortical activity
+during suppression in stereoblindness. \emph{NeuroReport} \textbf{11,}
+1007.
+
 \bibitem[\citeproctext]{ref-Norcia1985b}
@@
 
+\bibitem[\citeproctext]{ref-reganEffectStimulusColour1966a}
+Regan D (1966\emph{b}). \href{https://doi.org/10.1038/2101056a0}{An
+{Effect} of {Stimulus Colour} on {Average Steady-state Potentials}
+evoked in {Man}}. \emph{Nature} \textbf{210,} 1056--1057.
+
 \bibitem[\citeproctext]{ref-Regan1966}
-Regan D (1966). \href{https://doi.org/10.1016/0013-4694(66)90088-5}{Some
-characteristics of average steady-state and transient responses evoked
-by modulated light}. \emph{Electroencephalogr Clin Neurophysiol}
-\textbf{20,} 238--248.
+Regan D (1966\emph{a}).
+\href{https://doi.org/10.1016/0013-4694(66)90088-5}{Some characteristics
+of average steady-state and transient responses evoked by modulated
+light}. \emph{Electroencephalogr Clin Neurophysiol} \textbf{20,}
+238--248.
 
@@
 
+\bibitem[\citeproctext]{ref-reganColourCodingPattern1975}
+Regan D (1975).
+\href{https://doi.org/10.1016/0042-6989(75)90205-9}{Colour coding of
+pattern responses in man investigated by evoked potential feedback and
+direct plot techniques}. \emph{Vision Research} \textbf{15,} 175--183.
+
 \bibitem[\citeproctext]{ref-Regan1988}
@@
 \emph{Sci Rep} \textbf{10,} 5344.
+
+\bibitem[\citeproctext]{ref-schadeOpticalPhotoelectricAnalog1956}
+Schade OH (1956). \href{https://doi.org/10.1364/JOSA.46.000721}{Optical
+and {Photoelectric Analog} of the {Eye}}. \emph{JOSA} \textbf{46,}
+721--739.
 

### 2024-11-30 responseToReviews.md
Now talking about the history of contrast gain control

@@
 As we noted above, our intention here was to provide an introduction to the SSVEP technique with appropriate simple code samples that would allow relative newcomers to the field to implement the the technique.
-We are, however, keen to avoid misrepresenting the history of the technique and are happy to include appropriate references.
+We are, however, keen to avoid misrepresenting the history of SSVEP research and are happy to include appropriate references.
 
@@
 > Thank you for this explanation.
-We have now reworded the explanation to make this clearer.
+We have now reworded the paper to make this clearer.
 We did take care to define what we mean by 'on-off' flicker in the preceding sentence but we now clarify the alternative 'appearence/disappearence' terminology.
@@
 > We have clearly not been sufficiently clear here.
-We do not think we used the term 'separation' and we agree that a) there are infrequently-used paradigms that can target ON and OFF pathways more effectively (for example, 'sawtooth' temporal modulations that provide asymmetrical onset or offset edges) and b) responses from spatial features of differing polarity will tend to be combined at the scalp as long as the spatial frequency is high enough.
+We do not think we used the term 'separation' and we agree that a) there are infrequently-used paradigms that can target ON and OFF pathways more effectively (for example, 'sawtooth' temporal modulations that provide asymmetrical onset or offset transients) and b) responses from spatial features of differing polarity will tend to be combined at the scalp as long as the spatial frequency is high enough.
 We have expanded this section to attempt to make the general point clear.
@@
 
->
+> Contrast gain control is related to 'automatic gain control' - a electrical engineering concept that arose in the early 20th century.
+The concept of modulating the input (rather than the output) of the system became established in the 1930s and 40s (for example in WW2 radar systems) and the specific idea of computing that modulation as a function of the local average signal power seems to have arisen around this time and was refined over the 50s and 60s (for example, in 'sidechain compression').
+In the domain of vision, an early example of 'contrast gain control' (scaling the input - in this case luminance) is found in Sperling and Sondhi 1968 but the first formalization of the concept in the domain of spatial vision that included the critical concepts of an early, divisive gain control signal computed from some additional estimate of local contrast seems to appear in Heeger 1992 (to explain animal data) and Foley 1994 where it is applied to human psychophysical data.
 
-5.
+1.
 The description of Figure 3 and the nonlinear operations performed are unclear.

### 2024-12-01 responseToReviews.md
Adding figure for full wave rectification

@@
 > Contrast gain control is related to 'automatic gain control' - a electrical engineering concept that arose in the early 20th century.
-The concept of modulating the input (rather than the output) of the system became established in the 1930s and 40s (for example in WW2 radar systems) and the specific idea of computing that modulation as a function of the local average signal power seems to have arisen around this time and was refined over the 50s and 60s (for example, in 'sidechain compression').
-In the domain of vision, an early example of 'contrast gain control' (scaling the input - in this case luminance) is found in Sperling and Sondhi 1968 but the first formalization of the concept in the domain of spatial vision that included the critical concepts of an early, divisive gain control signal computed from some additional estimate of local contrast seems to appear in Heeger 1992 (to explain animal data) and Foley 1994 where it is applied to human psychophysical data.
+The concept of modulating the input (rather than the output) of the system became established in the 1930s and 40s (for example in WW2 radar systems).
+The specific idea of computing that modulation as a function of the local average signal power seems to have arisen around this time and was refined over the 50s and 60s (for example, in 'sidechain compression').
+In the domain of vision, an early example of 'contrast gain control' (scaling the input - in this case luminance) is found in Sperling and Sondhi 1968.
+
+The first formalization of the concept in the domain of spatial vision that included the critical concepts of an early, divisive gain control signal computed from some additional estimate of local contrast seems to appear in   a seminal paper by Shapley and Victor, 1978.
+Application to contrast adaptation was demonstrated by Ohzawa, Sclar, Freeman, 1982, 1985] with a demonstration of instantaneous gain control in the human SSVEP by Bobak et al in 1988.
+A synthesis that included both adaptation and masking was proposed by Heeger 1992 while Foley 1994 show an application to human psychophysical data.
+We have now cited these more foundational papers.
+We would be happy to include the references to the non-linear biophysical model used in adult and developmental work if the reviewer can point us to the appropriate literature.
 
@@
 
->
+> The transducer function used to produce Figure 3 is available in the paper and reproduced here:
+
+```python
+def simpleTransducerFunction(inputContrast, c50, RMax):
+ # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
+ # RMax defines the reponse gain of the transducer (how big the biggest output is). 
+ # c50 defines the conrast sensitivity. Smaller values are more sensitive.
+ # The exponent can also vary - here we fix it to 2 but it can be varied.
+  expnt=2
+
+  # Contrast cannot be negative. To model the responses of a set of on and off 
+  # contrast detectors we perform full-wave rectification of the signal
+  inputContrast=np.abs(inputContrast)
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
+  return output
+
+```
+
+The input contrast is full-wave rectified (via the 'abs' operator).
+Here we choose an exponent of '2' for the hyperbolic ratio function so it is *also* squared both in the numerator and the denominator.
+But this parameter is often fitted rather than fixed and the reader is encouraged to experiment with the exponent to examine its effect on the resulting amplitude spectrum.
+While there are, technically, infinite harmonics of this input, they decrease with frequency and we plot only three for reasons of space.
 

### 2024-12-01 review.qmd
Adding figure for full wave rectification

@@
  # c50 defines the conrast sensitivity. Smaller values are more sensitive.
- # The exponent can also vary - here we fix it to 2
+ # The exponent can also vary - here we fix it to 2. Varying it produces different amplitude spectra.
+
   expnt=2
 
-  # Contrast cannot be negative. To model the responses of a set of on and off 
-  # contrast detectors we perform full-wave rectification of the signal
-  inputContrast=np.abs(inputContrast)
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
+  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt) # Here we apply the hyperbolic ratio function
+
   return output
@@
 transducedSignal = np.zeros(len(contrastRange))
+
 for contrastIndex,thisContrast in enumerate(contrastRange):
@@
 ```
+```{python fullwaverectification}
 
+# Define some constants
+inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
+contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S=1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz=1000
+freqCutoff_Hz=30
+c50=.10
+RMax=10
+
+support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)
+
+initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal=[]
+
+initialSignal = np.zeros(len(contrastRange)) 
+transducedSignal = np.zeros(len(contrastRange))
+
+for contrastIndex,thisContrast in enumerate(contrastRange):
+  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
+
+  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
+
+  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
+  initialSignal[contrastIndex]=inputAmplitude[0]
+
+  # We are about to model response to contrast.
+Different cell populations detect positive and negative-going contrast and we can model their combined response by performing full-wave rectification as follows:
+  fullWaveRectifiedSignal=np.abs(inputModulation)
+
+  # This doubles the frequencies of the resulting FT and also introduces harmonics at higher frequencies but it does not alter the general shape of the contrast vs response function.
+
+   # Now compute the FT on the signal after it has passed through full wave rectification
+  FT_transducedModulation=np.fft.fft(fullWaveRectifiedSignal)/ len(fullWaveRectifiedSignal)
+  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
+  transducedSignal[contrastIndex]=transducedAmplitude[0]
+
+frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
+
+#Make a figure with three subplots
+fig, axs = plt.subplots(1,3, figsize=(8, 3))
+
+axs[0].plot(support_S, transducedModulation)
+axs[0].set_title('Transduced signal', fontsize=10)
+axs[0].set_xlabel('Time (s)')
+
+# Plot the absolute values of the FFT in the second axis as a bar plot
+axs[1].bar(frequencySupport, np.abs(FT_transducedModulation[frequencySupport]))
+axs[1].set_xlabel('Frequency (Hz)')
+axs[1].set_ylabel('Magnitude (a.u.)')
+axs[1].set_title('Fourier spectrum')
+axs[1].grid()
+
+axs[2].plot(contrastRange*100,transducedSignal)
+axs[2].set_xlabel('Contrast (%)')
+axs[2].set_ylabel('Amplitude (a.u.)')
+axs[2].set_title('Transducer output', fontsize=10)
+axs[2].grid()
+
+axs[2].grid()
+plt.tight_layout()
+plt.savefig('Figures/3_fullwaverectification.pdf')
+
+```
 ```{python transducerout}
@@
 
-   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
-   # using the transducer functiona above.
-  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
+  # We are about to model response to contrast.
+Different cell populations detect positive and negative-going contrast and we can model their combined response by performing full-wave rectification as follows:
+  fullWaveRectifiedSignal=np.abs(inputModulation)
+
+  # This doubles the frequencies of the resulting FT and also introduces harmonics at higher frequencies but it does not alter the general shape of the contrast vs response function.
+
+   # Now compute the FT on the signal after it has passed through a saturating non-linearity. Do this 
+   # using the transducer function above.
+  transducedModulation=simpleTransducerFunction(fullWaveRectifiedSignal, c50,RMax)
   FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
@@
 plt.tight_layout()
-plt.savefig('Figures/3_transducer.pdf')
+plt.savefig('Figures/4_transducer.pdf')
 
@@
 plt.tight_layout()
-plt.savefig('Figures/4_twoinputs.pdf')
+plt.savefig('Figures/5_twoinputs.pdf')
 
@@
 plt.tight_layout()
-plt.savefig('Figures/5_sweepplot.pdf')
+plt.savefig('Figures/6_sweepplot.pdf')
 
@@
 # To save the figure
-plt.savefig('Figures/6_clinicalplot.pdf')
+plt.savefig('Figures/7_clinicalplot.pdf')
 
@@
 
-Next we can consider the impact of a nonlinearity in processing on the responses.
+Next we can consider the impact of two types of nonlinearity in processing on the responses.
+
+First, we consider the fact that neurons in the early visual system code either positive or negative contrast.
+The responses of each different type of neuron is an increase in firing rate to the increasing levels of the prefered contrast polatiry and these neurons typically exhibit zero (or baseline) response to the non-prefered polarity.
+We can model the combined responses of these two cell populations to a time varying contrast signal by full-wave-rectification.
+The effect of this computation is, effectively, to double the input freuquency in the population measurement and to introduce additional higher (even) harmonics due to the discontinuity at the contrast reversals.
+The effect of full-wave rectification on the SSVEP signal is shown in @fig-fullwaverectification
+
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
+![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
+These calculations assume an entirely linear system.](){#fig-fullwaverectification}
+:::
+::: {.content-visible unless-format="docx"}
+![Illustration of a full-wave rectified sinusoid (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).](Figures/3_fullwaverectification.pdf){#fig-fullwaverectification}
+:::
+
 One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
@@
 The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
-The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/3_transducer.pdf){#fig-transducer}
+The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/4_transducer.pdf_transducer.pdf){#fig-transducer}
 :::

### 2024-12-01 review.tex
Adding figure for full wave rectification

@@
 
-Next we can consider the impact of a nonlinearity in processing on the
-responses. One of the simplest nonlinearities is the function that
-describes a cell's response to different levels of contrast. In
+Next we can consider the impact of two types of nonlinearity in
+processing on the responses.
+
+First, we consider the fact that neurons in the early visual system code
+either positive or negative contrast. The responses of each different
+type of neuron is an increase in firing rate to the increasing levels of
+the prefered contrast polatiry and these neurons typically exhibit zero
+(or baseline) response to the non-prefered polarity. We can model the
+combined responses of these two cell populations to a time varying
+contrast signal by full-wave-rectification. The effect of this
+computation is, effectively, to double the input freuquency in the
+population measurement and to introduce additional higher (even)
+harmonics due to the discontinuity at the contrast reversals. The effect
+of full-wave rectification on the SSVEP signal is shown in
+Figure~\ref{fig-fullwaverectification}
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{Figures/3_fullwaverectification.pdf}
+
+}
+
+\caption{\label{fig-fullwaverectification}Illustration of a full-wave
+rectified sinusoid (left), its Fourier spectrum (middle), and how the
+amplitude of the first harmonic component increases with contrast
+(right).}
+
+\end{figure}%
+
+One of the simplest nonlinearities is the function that describes a
+cell's response to different levels of contrast. In
 Figure~\ref{fig-transducer} this is modeled by a hyperbolic ratio
@@
 
-\includegraphics{Figures/3_transducer.pdf}
+\includegraphics{Figures/4_transducer.pdf_transducer.pdf}
 

### 2024-12-01 Figures/convertImages.sh
Changing figure numbers

@@ -1,19 +1,0 @@
-#!/bin/bash
-
-# Set the desired DPI and width in pixels
-DPI=300
-WIDTH_CM=14
-
-# Convert width from cm to pixels (since 1 inch = 2.54 cm)
-WIDTH_PIXELS=$(echo "$WIDTH_CM * $DPI / 2.54" | bc)
-
-# Loop through all PDF files in the directory
-for file in *.pdf; do
-  # Extract the base name of the file (without extension)
-  base_name="${file%.pdf}"
-
-  # Convert PDF to TIFF
-  convert -density $DPI "$file" -resize ${WIDTH_PIXELS}x -units PixelsPerInch -compress LZW "${base_name}.tiff"
-
-done
-

### 2024-12-01 review.qmd
Changing figure numbers
Fixed new figure rendering

@@
 - FIGURE ABOUT HERE
-
 ![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
@@
 ![Illustration of a full-wave rectified sinusoid (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).](Figures/3_fullwaverectification.pdf){#fig-fullwaverectification}
+
 :::
@@
 The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
-The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/4_transducer.pdf_transducer.pdf){#fig-transducer}
+The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/4_transducer.pdf){#fig-transducer}
 :::
@@
 Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).
-The effects of using different nonlinearities can be explored in the accompanying code.](Figures/4_twoinputs.pdf){#fig-twoinputs}
+The effects of using different nonlinearities can be explored in the accompanying code.](Figures/5_twoinputs.pdf){#fig-twoinputs}
 :::
@@
 ![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
-The solid line is the linear regression fit to the  lowest four data points prior to the satuating part of the curve, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/5_sweepplot.pdf){#fig-sweepvep}
+The solid line is the linear regression fit to the  lowest four data points prior to the satuating part of the curve, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/6_sweepplot.pdf){#fig-sweepvep}
 :::
@@
 (2014).
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/6_clinicalplot.pdf){#fig-plotclinical}
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/7_clinicalplot.pdf){#fig-plotclinical}
 :::

### 2024-12-02 review.qmd
Editing - added separate figures for rectification and hyperbolic rf transducer

@@
 
-  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)
-
-  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
-  initialSignal[contrastIndex]=inputAmplitude[0]
-
   # We are about to model response to contrast.
@@
 
-axs[0].plot(support_S, transducedModulation)
-axs[0].set_title('Transduced signal', fontsize=10)
+axs[0].plot(support_S, fullWaveRectifiedSignal)
+axs[0].set_title('Rectified signal', fontsize=10)
 axs[0].set_xlabel('Time (s)')
+axs[0].set_ylabel('Magnitude (a.u.)')
 
@@
 axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('Transducer output', fontsize=10)
+axs[2].set_title('Contrast response function', fontsize=10)
 axs[2].grid()
@@
 axs[0].set_xlabel('Time (s)')
+axs[0].set_ylabel('Magnitude (a.u.)')
 
@@
 axs[2].set_ylabel('Amplitude (a.u.)')
-axs[2].set_title('Transducer output', fontsize=10)
+axs[2].set_title('Contrast response funcion', fontsize=10)
 axs[2].grid()
@@
 First, we consider the fact that neurons in the early visual system code either positive or negative contrast.
-The responses of each different type of neuron is an increase in firing rate to the increasing levels of the prefered contrast polatiry and these neurons typically exhibit zero (or baseline) response to the non-prefered polarity.
+The responses of each different type of neuron is an increase in firing rate to the increasing levels of the prefered contrast polarity and these neurons typically exhibit zero (or baseline) response to the non-prefered polarity.
 We can model the combined responses of these two cell populations to a time varying contrast signal by full-wave-rectification.
@@
 
-One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
-In @fig-transducer this is modeled by a hyperbolic ratio function resulting in a saturating non-linearity:
+The contrast response function (CRF) is still linear for this model - rectification has doubled the frequency (and introduced higher harmonics) but the response amplitude measured by the second harmonic is still directly proportional to the input amplitude.
+
+Neurons in the visual system typically do not exhibit this perfect response linearity.
+Instead, neuronal response are well-modeled by some form of saturating non-linearity: as the contrast increases, the response of the neuron increases but the response increase per unit contrast is reduced at the high end of the contrast range.
+This saturating nonlinearity is often modeled as a hyperbolic ratio function of the form
 
@@
 This nonlinearity has clear effects on signal transduction.
-The sinusoidal waveform is distorted by the nonlinearity, and appears frequency doubled (left panel of @fig-transducer; note that other nonlinearities such as rectification and squaring can have similar effects).
-The distortion is reflected in the Fourier spectrum (middle panel of @fig-transducer), which now includes responses at integer multiples (known as harmonics) of the stimulation frequency (10, 20 and 30Hz), but no response at the fundamental (5Hz).
-Finally, the contrast response function is now nonlinear (right panel of @fig-transducer).
-Although the hyperbolic ratio function (@eq-nakarushton) is monotonic, the CRF resulting from measuring the amplitude of the second harmonic (2F) component contains a slight roll-off at high input contrasts.
+In @fig-transducer the full-wave rectified waveform (which is combining the response of both 'on' and 'off' neurons) passes through the saturating nonlinearity (left panel of @fig-transducer).
+The frequency doubling seen in @fig-fullwaverectification is still present but the the contrast response function is now nonlinear (right panel of @fig-transducer).
+
+Interestingly, although the hyperbolic ratio function (@eq-nakarushton) is monotonic, the CRF resulting from measuring the amplitude of the second harmonic (2F) component contains a slight roll-off at high input contrasts.
 This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.
@@
 ![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
-The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
-The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](){#fig-transducer}
+The input wavefore (left panel) is frequency doubled by the full-wave rectification (see @fig-fullwaverectification), and now has energy at the second harmonic (10Hz) and higher even harmonics (middle panel).
+Passing the signal through a second nonlinearity (the hyperbolic ratio function) causes the contrast response function at the second harmonic (2F) to become nonlinear and it and saturates (or even 'supersaturates') at high contrasts (right panel).](){#fig-transducer}
 :::
@@
 ![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
-The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
-The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/4_transducer.pdf){#fig-transducer}
+The input wavefore (left panel) is frequency doubled by the full-wave rectification (see @fig-fullwaverectification), and now has energy at the second harmonic (10Hz) and higher even harmonics (middle panel).
+Passing the signal through a second nonlinearity (the hyperbolic ratio function) causes the contrast response function at the second harmonic (2F) to become nonlinear and it and saturates (or even 'supersaturates') at high contrasts (right panel).](Figures/4_transducer.pdf){#fig-transducer}
 :::

### 2024-12-02 responseToReviews.md
At the end. I think it needs checking..

@@
 
-> The Ross and Speed paper measures masking using SSVEP but other adaptation measures are, as far as we can tell, psychophysical.
+> The Ross and Speed paper measures masking using SSVEP but the other adaptation measures are, as far as we can tell, psychophysical.
 Heinrich and Bach are concerned with transient changes in SSVEP amplitude within a recording which may have an adaptation origin but which manifest as response increases over the first few seconds and seems a little mysterious (we often observe these as well).
@@
 > We think this topic is probably worth a more detailed treatment than we can provide here.
-Broadly, much of the oversaturation effect can be modeled by an increase in saturation and therefore harmonic distortion - power is transferred from the first to higher harmonics as contrast increases.
+Broadly, much of the oversaturation effect can be modeled by an increase in saturation and therefore harmonic distortion - power is transferred from the first to higher harmonics as contrast increases and the output becomes more 'square-wave-like'.
 It seems likely that total power does not saturate (perceptually, there appears to be a monotonic increase in contrast for example and discrimination thresholds increase up to 100%) but the issue is complicated by the fact that the distribution of neurons sensitive to different contrast levels might not be uniform and that neurons might alter their response properties to signal with optimum sensitivity around a particular range.
@@
 > We have now note that it is important to consider noise for situations where the SNR is low or the slope is shallow and we expand on the point that you have to stick to the part of the CRF where the curve is more or less linear - as we showed explicitly in the figure.
+The code also includes an explicit 'noise baseline' that can be used to demonstrate changes in the intercept as a function of SNR.
 
@@
 > We have clearly not been sufficiently clear here.
-We do not think we used the term 'separation' and we agree that a) there are infrequently-used paradigms that can target ON and OFF pathways more effectively (for example, 'sawtooth' temporal modulations that provide asymmetrical onset or offset transients) and b) responses from spatial features of differing polarity will tend to be combined at the scalp as long as the spatial frequency is high enough.
+We do not think we suggested any form of 'separation' and we agree that a) there are infrequently-used paradigms that can target ON and OFF pathways more effectively (for example, 'sawtooth' temporal modulations that provide asymmetrical onset or offset transients) and b) responses from spatial features of differing polarity will tend to be combined at the scalp as long as the spatial frequency is high enough.
 We have expanded this section to attempt to make the general point clear.
@@
 We have now cited these more foundational papers.
-We would be happy to include the references to the non-linear biophysical model used in adult and developmental work if the reviewer can point us to the appropriate literature.
+We would be happy to include the relevant references to the non-linear biophysical model used in adult and developmental work if the reviewer can point us to the appropriate literature.
 
@@
 
-> The transducer function used to produce Figure 3 is available in the paper and reproduced here:
+> We agree that we went too fast here and the description of the signal processing was confusing.
+We now break out the two stages into (Figure 3) full wave rectification only (to model the contributions of on and off cells) - explicitly pointing out that this produces only even harmonics but does not alter the linearity of the CRF.
+We then show (Figure 4) the effect of passing this full wave rectified signal through a hyperbolic ratio function and illustrate the resulting saturation of the CRF (Figure 4).
+We think that breaking the model out in this way makes the contribution of each part clearer and allows the reader to play with the code for separate stages.
+The input contrast is squared in Figure 4 simply because the exponent we choose in the saturating nonlinearity is '2' but this is a parameter that is often fitted rather than fixed and the reader is encouraged to experiment with the exponent to examine its effect on the resulting amplitude spectrum.
 
-```python
-def simpleTransducerFunction(inputContrast, c50, RMax):
- # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
- # RMax defines the reponse gain of the transducer (how big the biggest output is). 
- # c50 defines the conrast sensitivity. Smaller values are more sensitive.
- # The exponent can also vary - here we fix it to 2 but it can be varied.
-  expnt=2
-
-  # Contrast cannot be negative. To model the responses of a set of on and off 
-  # contrast detectors we perform full-wave rectification of the signal
-  inputContrast=np.abs(inputContrast)
-  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
-  return output
-
-```
-
-The input contrast is full-wave rectified (via the 'abs' operator).
-Here we choose an exponent of '2' for the hyperbolic ratio function so it is *also* squared both in the numerator and the denominator.
-But this parameter is often fitted rather than fixed and the reader is encouraged to experiment with the exponent to examine its effect on the resulting amplitude spectrum.
-While there are, technically, infinite harmonics of this input, they decrease with frequency and we plot only three for reasons of space.
+While there are, technically, infinite even harmonics in both Figure 3 and Figure 4, they decrease with frequency and we plot only three for reasons of space.
 
@@
 
+> As far as we are aware, the term refers to stimulating some *aspect* of the visual scene (a position or a feature) - not necessarily just a spatial region.
+There need not be more than one region (it is interesting, for example, to frequency tag a single feature or region and measure the effect of attending towards or away from it) and frequency tagged components can exist in the same location (for example, see the work on contrast normalization described above where the 'probe' and 'mask' have different tags and we explicitly examine the intermodulation terms).
+In fact much of this paper is concerned with discussing precisely these intermodulation terms and their origins.
+However, intermodulation is also not a 'necessary' part of a frequency tagging experiment: beautiful early work by Morgan et al (PNAS 1996) uses spatial frequency tagging to examine attentional modulation purely in terms of the amplification or suppression of the F1 and F2 components.
+The tag need not even be visible - there is a fashion at the moment for 'rapid invisible frequency tagging' (RIFT) using flicker close to or beyond the CFF limit.
+In this sense, steady state paradigms are almost by definition frequency tagging paradigms because at least one feature or position is 'tagged'.
+In terms of using multiple input frequencies to characterise interactions between visual features, we think that Regan and Regan may have had the most comprehensive and some of the earliest descriptions of the method and we have credited them extensively throughout the paper.
+We are also happy to consider additional works if the reviewer can suggest them.
+
 7.
@@
 
+> We touch on both these areas but including them in detail would require us to extend what is already becoming a very long paper.
+
 8. Figures should be labeled.
+> It is possible that our text compilation stage failed to auto-generate the figure labels - we apologise if this is the case.
+The current version seems to be working correctly (in that each Figure has a label saying 'Figure 1', 'Figure 2' etc which is referenced in the main text.
+)
+
 Thus, the stated intention of this manuscript (“Here we describe how SSVEPs have been used to study visual contrast over the past 70 years”) fails to be met.
 Only select articles from a few sources are covered, and therefore, the richness and value of this technique is lost.
+
+> The original document had well over 100 references and the current bibliography now contains 156 entries.
+Although we have tried to be fair, we have of necessity 'selected' from the literature and have doubtless omitted many important foundational works.
+We are happy to take suggestions if parts of the field have been misrepresented.
+
+The original purpose of the document (now long eclipsed) was to serve as a convenient primer on the history of the technique and an interactive visualisation of key concepts implemented in Python and R using a novel 'executable paper' format.
+We think that getting this balance right means that we cannot include a comprehensive history of the entire field - in effect we must continue to be 'selective' and we hope that the relatively few papers we have cited here will serve as useful signposts to the richer literature.

### 2024-12-02 review.qmd
At the end. I think it needs checking..

@@
   output=RMax*(inputContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
+  # Changing this equation can produce a different pattern of intermodulation terms in Figure 5.
+  #output=RMax*(inputContrast**expnt + maskContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
   return output
@@
 nLevels=8
-internalNoise=1
+spontaneousNoise=.1 # Changing the level of background noise can alter the slope of the regression line and therefore the estimate of the threshold
+
 inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
-contrastRange=np.logspace(1,1.9,nLevels)/100 # Equally spaced log contrast between 1 and 10^1.5
+contrastRange=np.logspace(.5,1.9,nLevels)/100 # Equally spaced log contrast between 1 and 10^1.5
 duration_S=1 # All variables in this code have an _[units] where possible
@@
 for contrastIndex,thisContrast in enumerate(contrastRange):
-  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast + np.random.rand(len(support_S))*internalNoise
+  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast + np.random.rand(len(support_S))*spontaneousNoise
 
@@
 The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
-This normalization, achieved through a computation called 'contrast gain control' [@Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
+This normalization, achieved through a computation called 'contrast gain control' [@ohzawaContrastGainControl1982, @ohzawaContrastGainControl1985, @bobakCorticalContrastGain1988a, @Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
 As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround; note that the code used to produce all figures in this review is available in python and R at: https://github.com/wadelab/contrastReviewPaperVNS).
@@
 
-To illustrate the effect of contrast gain control [@Heeger1992], we next include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
+To illustrate the effect of contrast gain control [@ohzawaContrastGainControl1982, @Heeger1992], we next include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
 The mask stimulus will suppress the response at the target frequency, reducing its amplitude.

### 2024-12-03 responseToReviews.md
Minor typo edits, and hopefully fixed autobuild

@@
 
-> We have now note that it is important to consider noise for situations where the SNR is low or the slope is shallow and we expand on the point that you have to stick to the part of the CRF where the curve is more or less linear - as we showed explicitly in the figure.
+> We have now noted that it is important to consider noise for situations where the SNR is low or the slope is shallow and we expand on the point that you have to stick to the part of the CRF where the curve is more or less linear - as we showed explicitly in the figure.
 The code also includes an explicit 'noise baseline' that can be used to demonstrate changes in the intercept as a function of SNR.
@@
 In particular, we were struck by the large number of recent papers that use multi-frequency SSVEP contrast flicker as an input to BCI systems.
-There have been at least 400 such publications ('SSVEP AND BCI') in the last 5 years alone (Kcompared to around 10 for amblyopia ('SSVEP AND amblyopia')and 80 for ('SSVEP AND clinical').
-Some of these publications measure dynamic changes in covert spatial attention, some measure changes in featural attention, some measure fixation).
+There have been at least 400 such publications ('SSVEP AND BCI') in the last 5 years alone (compared to around 10 for amblyopia ('SSVEP AND amblyopia') and 80 for ('SSVEP AND clinical').
+Some of these publications measure dynamic changes in covert spatial attention, some measure changes in featural attention, some measure fixation.
 We know from personal experience that many of the researchers using contrast flicker as a BCI input would benefit from a slightly more applied guide such as this and the BCI section adds very little to the overall length of the manuscipt at this point.
@@
 The most natural things to look for might be interocular normalization (because eye layers are segregated in the LGN) and attentional gain control (because covert spatial attentional modulation must arise in cortex).
-This the type of feedback we are thinking of - not cortico-cortico feedback (which is hard to isolate with SSVEP)
+This is the type of feedback we are thinking of - not cortico-cortico feedback (which is hard to isolate with SSVEP).
 
@@
 > We thank the reviewer for their helpful comments.
-As we noted above, our intention here was to provide an introduction to the SSVEP technique with appropriate simple code samples that would allow relative newcomers to the field to implement the the technique.
+As we noted above, our intention here was to provide an introduction to the SSVEP technique with appropriate simple code samples that would allow relative newcomers to the field to implement the technique.
 We are, however, keen to avoid misrepresenting the history of SSVEP research and are happy to include appropriate references.
@@
 We did take care to define what we mean by 'on-off' flicker in the preceding sentence but we now clarify the alternative 'appearence/disappearence' terminology.
-The point we were making was that odd harmonic responses (1F) are a feature of this this type of modulation because the responses to the appearance and disappearance of a pattern (or mean field) need not be completely symmetrical (See Zemon, et al 1988 for an explicit demonstration).
+The point we were making was that odd harmonic responses (1F) are a feature of this type of modulation because the responses to the appearance and disappearance of a pattern (or mean field) need not be completely symmetrical (see Zemon, et al 1988 for an explicit demonstration).
 
@@
 
-> We have clearly not been sufficiently clear here.
+> Apologies that we have not been sufficiently clear here.
 We do not think we suggested any form of 'separation' and we agree that a) there are infrequently-used paradigms that can target ON and OFF pathways more effectively (for example, 'sawtooth' temporal modulations that provide asymmetrical onset or offset transients) and b) responses from spatial features of differing polarity will tend to be combined at the scalp as long as the spatial frequency is high enough.
@@
 
-The first formalization of the concept in the domain of spatial vision that included the critical concepts of an early, divisive gain control signal computed from some additional estimate of local contrast seems to appear in a seminal paper by Shapley and Victor, 1978.
-Application to contrast adaptation was demonstrated by Ohzawa, Sclar, Freeman, 1982, 1985] with a demonstration of instantaneous gain control in the human SSVEP by Bobak et al in 1988.
-A synthesis that included both adaptation and masking was proposed by Heeger 1992 while Foley 1994 show an application to human psychophysical data.
+> The first formalization of the concept in the domain of spatial vision that included the critical concepts of an early, divisive gain control signal computed from some additional estimate of local contrast seems to appear in a seminal paper by Shapley and Victor, (1978).
+Application to contrast adaptation was demonstrated by Ohzawa, Sclar, Freeman, (1982, 1985) with a demonstration of instantaneous gain control in the human SSVEP by Bobak et al in 1988.
+A synthesis that included both adaptation and masking was proposed by Heeger 1992, while Foley 1994 showed an application to human psychophysical data.
 We have now cited these more foundational papers.
@@
 
-1.
+5.
 The description of Figure 3 and the nonlinear operations performed are unclear.
@@
 
-> We agree that we went too fast here and the description of the signal processing was confusing.
+> We see that we went too fast here and the description of the signal processing was confusing.
 We now break out the two stages into (Figure 3) full wave rectification only (to model the contributions of on and off cells) - explicitly pointing out that this produces only even harmonics but does not alter the linearity of the CRF.
@@
 
-While there are, technically, infinite even harmonics in both Figure 3 and Figure 4, they decrease with frequency and we plot only three for reasons of space.
+> While there are, technically, infinite even harmonics in both Figure 3 and Figure 4, they decrease with frequency and we plot only three for reasons of space.
 
@@
 In fact much of this paper is concerned with discussing precisely these intermodulation terms and their origins.
-However, intermodulation is also not a 'necessary' part of a frequency tagging experiment: beautiful early work by Morgan et al (PNAS 1996) uses spatial frequency tagging to examine attentional modulation purely in terms of the amplification or suppression of the F1 and F2 components.
+However, intermodulation is also not a 'necessary' part of a frequency tagging experiment: beautiful early work by Morgan et al (PNAS 1996) uses frequency tagging over space to examine attentional modulation purely in terms of the amplification or suppression of the F1 and F2 components.
 The tag need not even be visible - there is a fashion at the moment for 'rapid invisible frequency tagging' (RIFT) using flicker close to or beyond the CFF limit.
@@
 8. Figures should be labeled.
+
 > It is possible that our text compilation stage failed to auto-generate the figure labels - we apologise if this is the case.
-The current version seems to be working correctly (in that each Figure has a label saying 'Figure 1', 'Figure 2' etc which is referenced in the main text.
-)
+The current version seems to be working correctly (in that each Figure has a label saying 'Figure 1', 'Figure 2' etc which is referenced in the main text).
 
@@
 
-The original purpose of the document (now long eclipsed) was to serve as a convenient primer on the history of the technique and an interactive visualisation of key concepts implemented in Python and R using a novel 'executable paper' format.
+> The original purpose of the document (now long eclipsed) was to serve as a convenient primer on the history of the technique and an interactive visualisation of key concepts implemented in Python and R using a novel 'executable paper' format.
 We think that getting this balance right means that we cannot include a comprehensive history of the entire field - in effect we must continue to be 'selective' and we hope that the relatively few papers we have cited here will serve as useful signposts to the richer literature.

### 2024-12-03 review.qmd
Minor typo edits, and hopefully fixed autobuild

@@
 
+if not os.path.exists('Figures/'):
+  os.makedirs('Figures/')
 ```
@@
 ```
+
 ```{python fullwaverectification}
@@
 ```
+
 ```{python transducerout}
@@
 The human response to contrast can be studied using many different techniques.
-Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965, @schadeOpticalPhotoelectricAnalog1956], defined as the inverse of the lowest contrast that can be reliably detected.
+Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965; @schadeOpticalPhotoelectricAnalog1956], defined as the inverse of the lowest contrast that can be reliably detected.
 But neural responses can also be measured more directly using techniques such as functional magnetic resonance imaging (fMRI), magnetoencephalography (MEG), and electroencephalography (EEG).
@@
 For spatial patterns, the contributions of individual cortical cells to these excitations are thought to sum at the scalp and generate an average of their onset responses on each cycle.
-For very low spatial frequencies (including zero) resposes at the scalp will be dominated by a single on- or off- cell type on each half cycle.
+For very low spatial frequencies (including zero), resposes at the scalp will be dominated by a single on- or off- cell type on each half cycle.
 
@@
 
-By comparison, counterphase flickering patterns generate two essentially identical transients per cycle and therefore does not produce a response at 1F, only even harmonics: 2F, 4F, 6F and so on.
-Because square-waves are spectrally broad-band, square wave flicker (either on/off or contrast reversig) tends to produce additional spectral harmonics compared to sine-wave flicker.
+By comparison, counterphase flickering patterns generate two essentially identical transients per cycle and therefore do not produce a response at 1F, only even harmonics: 2F, 4F, 6F and so on.
+Because square-waves are spectrally broad-band, square wave flicker (either on/off or contrast reversing) tends to produce additional spectral harmonics compared to sine-wave flicker.
 
@@
 It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
-Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@reganEvokedPotentialsSpecific1973, @reganColourCodingPattern1975, @reganEffectStimulusColour1966a, @McKeefry1996; @Di-Russo2001] (see also [@Baseler1997]).
+Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@reganEvokedPotentialsSpecific1973; @reganColourCodingPattern1975; @reganEffectStimulusColour1966a; @McKeefry1996; @Di-Russo2001] (see also [@Baseler1997]).
 
@@
 
-These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance; @Lennie1993) and also to definitions based on cone excitations [@Derrington1984; @MacLeod1979] which are more common in work on chromatic processing.
+These contrast definitions are appropriate both to photometric measures of stimulus contrast [for example, luminance; @Lennie1993] and also to definitions based on cone excitations [@Derrington1984; @MacLeod1979] which are more common in work on chromatic processing.
 
@@
 The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
-This normalization, achieved through a computation called 'contrast gain control' [@ohzawaContrastGainControl1982, @ohzawaContrastGainControl1985, @bobakCorticalContrastGain1988a, @Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
+This normalization, achieved through a computation called 'contrast gain control' [@ohzawaContrastGainControl1982; @ohzawaContrastGainControl1985; @bobakCorticalContrastGain1988a; @Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
 As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround; note that the code used to produce all figures in this review is available in python and R at: https://github.com/wadelab/contrastReviewPaperVNS).
@@
 
-Although the SSVEP does seem to reflect the activity of relatively well-tuned neuronal populations , changing some aspect of the stimulus may change the nature of those populations.
-For example, Campbell and Maffei [@Campbell1970] noted that densely-sampled measurements of contrast responses can reveal the presence of two, qualitatively different types of neurons that exhibit different log-linear contrast functions (see also [@Tyler1985, @Nakayama1982, @Norcia1989 and @Souza2007]).
+Although the SSVEP does seem to reflect the activity of relatively well-tuned neuronal populations, changing some aspect of the stimulus may change the nature of those populations.
+For example, @Campbell1970 noted that densely-sampled measurements of contrast responses can reveal the presence of two qualitatively different types of neurons that exhibit different log-linear contrast functions [see also @Tyler1985; @Nakayama1982; @Norcia1989; @Souza2007].
 Modeling such two-limbed contrast response functions using a single sigmoidal function is therefore an approximation.
@@
 An additional complexity is introduced by the fact that the EEG is an average population response and that the visual system contains many different types of neurons.
-As a stimulus changes in, for example, contrast it may selectively activate qualitatively different neuronal populations.
-This type
+As a stimulus changes in (for example) contrast, it may selectively activate qualitatively different neuronal populations.
 
@@
 Notice that there are five peaks in the waveform during the one second sample, so the stimulation frequency is 5Hz (F1).
-The second panel shows Fourier transform of the waveform, which contains a substantial peak at this frequency.
+The second panel shows the Fourier transform of the waveform, which contains a substantial peak at this frequency.
 If we change the stimulus contrast (i.e.
@@
 First, we consider the fact that neurons in the early visual system code either positive or negative contrast.
-The responses of each different type of neuron is an increase in firing rate to the increasing levels of the prefered contrast polarity and these neurons typically exhibit zero (or baseline) response to the non-prefered polarity.
+The response of each different type of neuron is an increase in firing rate to the increasing levels of the preferred contrast polarity, and these neurons typically exhibit zero (or baseline) response to the non-preferred polarity.
 We can model the combined responses of these two cell populations to a time varying contrast signal by full-wave-rectification.
-The effect of this computation is, effectively, to double the input freuquency in the population measurement and to introduce additional higher (even) harmonics due to the discontinuity at the contrast reversals.
+The effect of this computation is, effectively, to double the input frequency in the population measurement and to introduce additional higher (even) harmonics due to the discontinuity at the contrast reversals.
 The effect of full-wave rectification on the SSVEP signal is shown in @fig-fullwaverectification
@@
 Instead, neuronal response are well-modeled by some form of saturating non-linearity: as the contrast increases, the response of the neuron increases but the response increase per unit contrast is reduced at the high end of the contrast range.
-This saturating nonlinearity is often modeled as a hyperbolic ratio function of the form
+This saturating nonlinearity is often modeled as a hyperbolic ratio function of the form:
 
@@
 ![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
-The input wavefore (left panel) is frequency doubled by the full-wave rectification (see @fig-fullwaverectification), and now has energy at the second harmonic (10Hz) and higher even harmonics (middle panel).
+The input waveform (left panel) is frequency doubled by the full-wave rectification (see @fig-fullwaverectification), and now has energy at the second harmonic (10Hz) and higher even harmonics (middle panel).
 Passing the signal through a second nonlinearity (the hyperbolic ratio function) causes the contrast response function at the second harmonic (2F) to become nonlinear and it and saturates (or even 'supersaturates') at high contrasts (right panel).](){#fig-transducer}
@@
 ![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
-The input wavefore (left panel) is frequency doubled by the full-wave rectification (see @fig-fullwaverectification), and now has energy at the second harmonic (10Hz) and higher even harmonics (middle panel).
+The input waveform (left panel) is frequency doubled by the full-wave rectification (see @fig-fullwaverectification), and now has energy at the second harmonic (10Hz) and higher even harmonics (middle panel).
 Passing the signal through a second nonlinearity (the hyperbolic ratio function) causes the contrast response function at the second harmonic (2F) to become nonlinear and it and saturates (or even 'supersaturates') at high contrasts (right panel).](Figures/4_transducer.pdf){#fig-transducer}
@@
 
-To illustrate the effect of contrast gain control [@ohzawaContrastGainControl1982, @Heeger1992], we next include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
+To illustrate the effect of contrast gain control [@ohzawaContrastGainControl1982; @Heeger1992], we next include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
 The mask stimulus will suppress the response at the target frequency, reducing its amplitude.
-The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Regan1988, @brownDevelopmentRivalryDichoptic1999, @Candy2001, @Busse2009].
+The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Regan1988; @brownDevelopmentRivalryDichoptic1999; @Candy2001; @Busse2009].
 At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the transduced waveforms (left panel of @fig-twoinputs) and in the Fourier spectrum (middle panel of @fig-twoinputs).
@@
 
-where $C_\mathrm{mask}$ reflects the contrast of the mask component at a distinct frequency from that of the target.
+where $C_\mathrm{mask}$ reflects the contrast of the mask component at a frequency distinct from that of the target.
 The effect of this extra term is to reduce the target response (see right panel of @fig-twoinputs).
@@
 Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
-Both the amplitude and the phase of the SSVEP response to a single contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
+Both the amplitude and the phase of the SSVEP response to a single contrast-reversing sine-wave grating therefore contain information about nonlinear computations performed across a range of retinal and cortical cell types.
 
@@
 In principle, that information could be recovered from the SSVEP signal - a possibility recognised in the early days of the technique [@Regan1988].
-Although characterising the complete set of computations along entire processing pathway is challenging, and a full system identification may require the use of multiple different input frequencies [@boydMeasuringVolterraKernels1983], careful parametric variation of the input stimuli does allow us to fit models of early visual visual processing (e.g.
-[@Zemon2006Dynamic]) and, by incorporating mathematical models of gain control mechanisms with either instantaneous, or temporally-integrated gain control, it is possible to model gain control using SSVEP data.
-For example, [@Tsai2012] demonstrated that a relatively simple gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
+Although characterising the complete set of computations along the entire processing pathway is challenging, and a full system identification may require the use of multiple different input frequencies [@boydMeasuringVolterraKernels1983], careful parametric variation of the input stimuli does allow us to fit models of early visual visual processing [e.g.
+@Zemon2006Dynamic] and, by incorporating mathematical models of gain control mechanisms with either instantaneous, or temporally-integrated gain control, it is possible to model gain control using SSVEP data.
+For example, @Tsai2012 demonstrated that a relatively simple gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
 This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.
-More recent work has used this approach to study binocular integration in both normal subjects and amblyopes[@Hou-houchuan2021,@Hou2020].
+More recent work has used this approach to study binocular integration in both normal subjects and amblyopes [@Hou-houchuan2021; @Hou2020].
 Our more recent work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
@@
 Psychophysically, adaptation increases detection thresholds, but has little effect on contrast discrimination performance [@Ross1993], much like pattern masks [@Foley1994].
-Although SSVEP adaptation effects show strong tuning for orientation [@Vergeer2018; @Campbell1970] and spatial frequency [@Mecacci1976], there appear to be very measurements of the full SSVEP contrast response function before and after an extended period of adaptation.
-One exception is a study by Bach et al [@bachGreenleeBuhler1988] that examined the effects of adaptation on SSVEP responses who found a paradoxical increase in the 1F SSVEP amplitude following prolonged exposure to high-contrast stimuli across much of the probe contrast range.
-They explain this as being due to the presence of a nonlinear transducer function that saturates at high probe contrast driving the response to higher harmonic frequencies.
+Although SSVEP adaptation effects show strong tuning for orientation [@Vergeer2018; @Campbell1970] and spatial frequency [@Mecacci1976], there appear to be very few measurements of the full SSVEP contrast response function before and after an extended period of adaptation.
+One exception is a study by @bachGreenleeBuhler1988 that examined the effects of adaptation on SSVEP responses, and found a paradoxical increase in the 1F SSVEP amplitude following prolonged exposure to high-contrast stimuli across much of the probe contrast range.
+They explain this as being due to the presence of a nonlinear transducer function that saturates at high probe contrasts, driving the response to higher harmonic frequencies.
 Adaptation reduces the apparent contrast of the input, leading to less distortion, a more sinusoidal output and a concentration of the frequency-tagged response at lower harmonics.
-This explanation is conceptually related to the 'supersaturation' effect that we observe in Figure 4.
+This explanation is conceptually related to the 'supersaturation' effect that we observed in Figure 4.
 
@@
 Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels (in the linear part of the log-contrast response function prior to saturation), and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis [@Campbell1970] (see @fig-sweepvep).
-This contrast level has been shown to correspond approximately with psychophysically measured detection thresholds [@campbellVisualEvokedPotential1972, @Norcia1986].
+This contrast level has been shown to correspond approximately with psychophysically measured detection thresholds [@campbellVisualEvokedPotential1972; @Norcia1986].
 One factor affecting the accuracy of this technique is the presence of noise in the EEG signal which will increase the apparent amplitude of the SSVEP responses across the measurement range.
@@
 A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrast levels.
-This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@reganRapidObjectiveRefraction1973,@Tyler1979].
+This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@reganRapidObjectiveRefraction1973; @Tyler1979].
 To avoid hysteresis or short-term adaptation effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979; @Norcia1985a; @Norcia1990].
@@
 The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section.
-Although response- [@Morrone1986] or contrast gain control [@Candy2001; ; @Skoczenski1998] is measurable in infants as young as six weeks old, its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
+Although response- [@Morrone1986] or contrast gain control [@Candy2001; @Skoczenski1998] is measurable in infants as young as six weeks old, its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
 
@@
 SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye (see [@fig-plotclinical]b) across the contrast range [@Baker2015; @Lygo2021].
-Measurments of interocular contrast modulation (for example, [@norciaChangesCorticalActivity2000,@Hou-houchuan2021] can also provide information about the way that amblyopia changes the balance of excitation and inhibition between eyes.
+Measurments of interocular contrast modulation (for example, [@norciaChangesCorticalActivity2000; @Hou-houchuan2021] can also provide information about the way that amblyopia changes the balance of excitation and inhibition between eyes.
 

### 2024-12-03 review.qmd
Added os package

@@
 # all package loading and subfunctions defined here
+import os
 import numpy as np
@@
 The effect of this noise is particulary apparent for recordings with low SNR or where the slope of the CRF is shallow (and therefore where small changes in the vertical offset of the fit result in large changes in the X-axis intercept).
-Accurate estimates of thresholds therefore require the experimenter to model this noise and compensate for its effects - as detailed in [@Norcia1989].
+Accurate estimates of thresholds therefore require the experimenter to model this noise and compensate for its effects, as detailed in @Norcia1989.
 

### 2024-12-03 Rversion/reviewR.qmd
Propagated changes to R version

@@
     keep-tex: true
-bibliography: WadeBaker_Review.bib
-csl: visual-neuroscience.csl
+bibliography: ../WadeBaker_Review.bib
+csl: ../visual-neuroscience.csl
 execute:
@@
 
-pdf('Figures/centresurround.pdf',width=12,height=6.5)
+pdf('Figures/1_centresurround.pdf',width=12,height=6.5)
 
@@
 
-pdf('Figures/linearsystem.pdf',width=12,height=4)
+pdf('Figures/2_linearsystem.pdf',width=12,height=4)
 par(mfrow=c(1,3))
@@
 lines(contrastRange*100,initialSignal,lwd=3)
+
+dev.off()
+
+```
+
+```{r fullwaverectification}
+
+inputFrequency_Hz <- 5 # We simulate a system with on input at many different contrasts
+contrastRange <- (10^seq(0,1.5,length=20))/100 # Equally spaced log contrast between 1 and 10^1.5
+duration_S <- 1 # All variables in this code have an _[units] where possible
+digitisationRate_Hz <- 1000
+freqCutoff_Hz <- 30
+c50 <- 0.10
+RMax <- 10
+
+support_S <- seq(0,duration_S,length=duration_S*digitisationRate_Hz)
+
+initialSignal <- NULL # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
+transducedSignal <- NULL
+
+for (contrastIndex in 1:length(contrastRange)){
+
+  inputModulation <- sin(2 * pi * inputFrequency_Hz * support_S) * contrastRange[contrastIndex]
+
+    # We are about to model response to contrast.
+Different cell populations detect positive and negative-going contrast and we can model their combined response by performing full-wave rectification as follows:
+  fullWaveRectifiedSignal <- abs(inputModulation)
+
+  # This doubles the frequencies of the resulting FT and also introduces harmonics at higher frequencies but it does not alter the general shape of the contrast vs response function.
+
+  # Now compute the FT on the signal after it has passed through full wave rectification
+  FT_transducedModulation <- fft(fullWaveRectifiedSignal) / length(fullWaveRectifiedSignal)
+
+  transducedAmplitude <- abs(FT_transducedModulation[2*inputFrequency_Hz+1])
+  transducedSignal[contrastIndex] <- transducedAmplitude
+
+}
+
+frequencySupport <- seq(0,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point
+
+#Make a figure with three subplots
+
+pdf('Figures/3_fullwaverectification.pdf',width=12,height=4)
+par(mfrow=c(1,3))
+
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,0.35))   
+  axis(1, at=seq(0,1,0.5), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,3,1)/10, tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,1,0.5), side = 1, at=seq(0,1,0.5), line=0.2)     # add the tick labels
+  mtext(text = seq(0,3,1)/10, side = 2, at=seq(0,3,1)/10, line=0.2, las=1)  
+title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(main="Rectified signal", cex.main=1.5)
+
+lines(support_S, fullWaveRectifiedSignal, lwd=3)
+
+# Plot the absolute values of the FFT in the second axis as a bar plot
+  plot(frequencySupport[2:length(frequencySupport)]-0.5,abs(FT_transducedModulation[2:length(frequencySupport)]),type='s',lwd=3,axes=FALSE, ann=FALSE, xlim=c(0,30), ylim=c(0,0.08))
+  axis(1, at=seq(0,30,10), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,0.08,0.02), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,30,10), side = 1, at=seq(0,30,10), line=0.2)     # add the tick labels
+  mtext(text = seq(0,0.08,0.02), side = 2, at=seq(0,0.08,0.02), line=0.2, las=1)  
+title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(ylab="Magnitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
+title(main="Fourier spectrum", cex.main=1.5)
+
+  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,32), ylim=c(0,0.08))   
+  axis(1, at=seq(0,32,4), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
+  axis(2, at=seq(0,0.08,0.02), tck=0.01, lab=F, lwd=2)
+  mtext(text = seq(0,32,4), side = 1, at=seq(0,32,4), line=0.2)     # add the tick labels
+  mtext(text = seq(0,0.08,0.02), side = 2, at=seq(0,0.08,0.02), line=0.2, las=1)  
+title(xlab="Contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
+title(ylab="Amplitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
+title(main="Contrast response function", cex.main=1.5)
+
+lines(contrastRange*100,transducedSignal,lwd=3)
 
@@
 
-pdf('Figures/transducer.pdf',width=12,height=4)
+pdf('Figures/4_transducer.pdf',width=12,height=4)
 par(mfrow=c(1,3))
@@
 
-pdf('Figures/twoinputs.pdf',width=12,height=4)
+pdf('Figures/5_twoinputs.pdf',width=12,height=4)
 par(mfrow=c(1,3))
@@
 
-pdf('Figures/sweepplot.pdf',width=6,height=6)
+pdf('Figures/6_sweepplot.pdf',width=6,height=6)
 
@@
 
-pdf('Figures/clinicalplot.pdf',width=12,height=12)
+pdf('Figures/7_clinicalplot.pdf',width=12,height=12)
 par(mfrow=c(2,2))
@@
 The human response to contrast can be studied using many different techniques.
-Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965], defined as the inverse of the lowest contrast that can be reliably detected.
-But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG.
+Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965; @schadeOpticalPhotoelectricAnalog1956], defined as the inverse of the lowest contrast that can be reliably detected.
+But neural responses can also be measured more directly using techniques such as functional magnetic resonance imaging (fMRI), magnetoencephalography (MEG), and electroencephalography (EEG).
 Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.
@@
 The SSVEP is a continuous electrical response evoked in the brain by visual stimuli flickering at a constant frequency [@Regan1966].
-For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual VEPs reveals multiple generators throughout visual cortex [@Di-Russo2005; @Di-Russo2007].
+For contrast-defined stimuli, such as sine-wave gratings, it is strongest at the occipital pole, adjacent to the early visual areas that generate the signal, although careful analysis of individual visually-evoked potentials (VEPs) reveals multiple generators throughout visual cortex [@Di-Russo2005; @Di-Russo2007].
 The flickering stimulus entrains neural population responses at multiples of the stimulus frequency, so continuous EEG data are typically analysed by taking the Fourier transform, and estimating the amplitude at these frequencies.
-Two common variants involve sinusoidal on-off flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
-On-off flicker can drive independent populations of on- and off-cells in the retina once per cycle and can therefore produce a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on.
-Counterphase flicker contains two transients per cycle and therefore does not produce a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
-Because square-waves are spectrally broad-band, square wave flicker tends to produce more complex spectral harmonics than sine-wave flicker.
+
+Two common stimulus variants involve sinusoidal on-off (or 'appearance/disappearance') flicker, where the stimulus alternates between a blank background and the peak contrast, and sinusoidal counterphase flicker, where the stimulus alternates in phase (i.e the black regions become white and the white regions become black).
+On-off flicker can excite independent populations of on- and off-cells in the visual system once per cycle.
+For spatial patterns, the contributions of individual cortical cells to these excitations are thought to sum at the scalp and generate an average of their onset responses on each cycle.
+For very low spatial frequencies (including zero), resposes at the scalp will be dominated by a single on- or off- cell type on each half cycle.
+
+If the amplitudes and waveforms of responses to the stimulus appearance and disappearnce are not perfectly balanced, a response at the fundamental flicker frequency, known as 1F, and its integer harmonics can be produced.
+For stimuli of any spatial frequency (including zero), an imbalance can be caused by differences in either the onset- and offset- responses of either on- or off-cells.
+For example, if only on-sensitive neurons were present (firing to the onset, but not the offset of local increases in luminance) the response to the on/off (or 'appearance/disappearance') presentation of a grating of any spatial frequency below the resolution limit would consist entirely of odd harmonics.
+
+By comparison, counterphase flickering patterns generate two essentially identical transients per cycle and therefore do not produce a response at 1F, only even harmonics: 2F, 4F, 6F and so on.
+Because square-waves are spectrally broad-band, square wave flicker (either on/off or contrast reversing) tends to produce additional spectral harmonics compared to sine-wave flicker.
 
 The higher harmonics of the steady-state signal are generally thought to reflect nonlinear processing in the visual system [@Regan1988].
+Because they can arise from different neuronal computations, different populations and even different stages of the visual system, different harmonics can exhibit different input-output functions.
+For example, 1F and 2F responses can have different thresholds [@BobakIOVS1984] and scalp topographies [@reganEvokedPotentialsSpecific1973] while @KaestnerTVST2024 show both different response slopes and thresholds for 1F and 2F components generated by dynamic noise.
+
 SSVEP signals can also be elicited by periodic changes of stimulus properties other than achromatic and chromatic contrast, such as motion, stereo depth, and facial identity or expression [see @Norcia2015, for an overview]; however our focus here is on the contrast response.
@@
 It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
-Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@McKeefry1996; @Di-Russo2001]   (see also [@Baseler1997]).
+Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@reganEvokedPotentialsSpecific1973; @reganColourCodingPattern1975; @reganEffectStimulusColour1966a; @McKeefry1996; @Di-Russo2001] [see also @Baseler1997].
 
@@
 The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment.
-This normalization, achieved through a computation called 'contrast gain control' [@Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
-As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround; note that the code used to produce all figures in this review is available in python and R at: https://github.com/wadelab/contrastReviewPaper).
+This normalization, achieved through a computation called 'contrast gain control' [@ohzawaContrastGainControl1982; @ohzawaContrastGainControl1985; @bobakCorticalContrastGain1988a; @Foley1994; @Heeger1992; @Carandini2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth.
+As an example, a grating placed at the centre of a low-contrast background typically appears more intense than the same grating when superimposed on a high contrast background (see @fig-centresurround; note that the code used to produce all figures in this review is available in python and R at: https://github.com/wadelab/contrastReviewPaperVNS).
 
-![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/centresurround.pdf){#fig-centresurround}
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](){#fig-centresurround}
+:::
+::: {.content-visible unless-format="docx"}
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/1_centresurround.pdf){#fig-centresurround}
+:::
 
@@
 However the SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants.
+
+Although the SSVEP does seem to reflect the activity of relatively well-tuned neuronal populations, changing some aspect of the stimulus may change the nature of those populations.
+For example, @Campbell1970 noted that densely-sampled measurements of contrast responses can reveal the presence of two qualitatively different types of neurons that exhibit different log-linear contrast functions [see also @Tyler1985; @Nakayama1982; @Norcia1989; @Souza2007].
+Modeling such two-limbed contrast response functions using a single sigmoidal function is therefore an approximation.
 
@@
 
+An additional complexity is introduced by the fact that the EEG is an average population response and that the visual system contains many different types of neurons.
+As a stimulus changes in (for example) contrast, it may selectively activate qualitatively different neuronal populations.
+
 To  understand these processes better, we will show how sinusoidal input signals might be processed by the visual system to produce SSVEPs.
@@
 
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
 ![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
-These calculations assume an entirely linear system.](Figures/linearsystem.pdf){#fig-linear}
+These calculations assume an entirely linear system.](){#fig-linear}
+:::
+::: {.content-visible unless-format="docx"}
+![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
+These calculations assume an entirely linear system.](Figures/2_linearsystem.pdf){#fig-linear}
+:::
 
-Next we can consider the impact of a nonlinearity in processing on the responses.
-One of the simplest nonlinearities is the function that describes a cell's response to different levels of contrast.
-In @fig-transducer this is modeled by a hyperbolic ratio function resulting in a saturating non-linearity:
+Next we can consider the impact of two types of nonlinearity in processing on the responses.
+
+First, we consider the fact that neurons in the early visual system code either positive or negative contrast.
+The response of each different type of neuron is an increase in firing rate to the increasing levels of the preferred contrast polarity, and these neurons typically exhibit zero (or baseline) response to the non-preferred polarity.
+We can model the combined responses of these two cell populations to a time varying contrast signal by full-wave-rectification.
+The effect of this computation is, effectively, to double the input frequency in the population measurement and to introduce additional higher (even) harmonics due to the discontinuity at the contrast reversals.
+The effect of full-wave rectification on the SSVEP signal is shown in @fig-fullwaverectification
+
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
+These calculations assume an entirely linear system.](){#fig-fullwaverectification}
+:::
+::: {.content-visible unless-format="docx"}
+![Illustration of a full-wave rectified sinusoid (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).](Figures/3_fullwaverectification.pdf){#fig-fullwaverectification}
+
+:::
+
+The contrast response function (CRF) is still linear for this model - rectification has doubled the frequency (and introduced higher harmonics) but the response amplitude measured by the second harmonic is still directly proportional to the input amplitude.
+
+Neurons in the visual system typically do not exhibit this perfect response linearity.
+Instead, neuronal response are well-modeled by some form of saturating non-linearity: as the contrast increases, the response of the neuron increases but the response increase per unit contrast is reduced at the high end of the contrast range.
+This saturating nonlinearity is often modeled as a hyperbolic ratio function of the form:
 
@@
 
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
 ![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
-The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel).
-The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/transducer.pdf){#fig-transducer}
+The input waveform (left panel) is frequency doubled by the full-wave rectification (see @fig-fullwaverectification), and now has energy at the second harmonic (10Hz) and higher even harmonics (middle panel).
+Passing the signal through a second nonlinearity (the hyperbolic ratio function) causes the contrast response function at the second harmonic (2F) to become nonlinear and it and saturates (or even 'supersaturates') at high contrasts (right panel).](){#fig-transducer}
+:::
+::: {.content-visible unless-format="docx"}
+![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
+The input waveform (left panel) is frequency doubled by the full-wave rectification (see @fig-fullwaverectification), and now has energy at the second harmonic (10Hz) and higher even harmonics (middle panel).
+Passing the signal through a second nonlinearity (the hyperbolic ratio function) causes the contrast response function at the second harmonic (2F) to become nonlinear and it and saturates (or even 'supersaturates') at high contrasts (right panel).](Figures/4_transducer.pdf){#fig-transducer}
+:::
 
-To illustrate the effect of contrast gain control [@Heeger1992], we next include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
+To illustrate the effect of contrast gain control [@ohzawaContrastGainControl1982; @Heeger1992], we next include a second component (a 'mask') that contributes to the gain pool of the first (the 'target').
 The mask stimulus will suppress the response at the target frequency, reducing its amplitude.
-The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Busse2009].
+The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Regan1988; @brownDevelopmentRivalryDichoptic1999; @Candy2001; @Busse2009].
 At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the transduced waveforms (left panel of @fig-twoinputs) and in the Fourier spectrum (middle panel of @fig-twoinputs).
@@
 
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
+
 ![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
-Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/twoinputs.pdf){#fig-twoinputs}
+Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).
+The effects of using different nonlinearities can be explored in the accompanying code.](){#fig-twoinputs}
+:::
+::: {.content-visible unless-format="docx"}
+![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
+Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).
+The effects of using different nonlinearities can be explored in the accompanying code.](Figures/5_twoinputs.pdf){#fig-twoinputs}
+:::
 
@@
 For example as well as cells that code positive or negative contrast in a more or less continuous manner, the retina also contains 'transient' cells that code temporal changes in contrast.
-These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the LGN [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion.
+These cells [@Kuffler1957; @Alpern1971], and cells with similar properties in the lateral geniculate nucleus (LGN) [@Levitt2001] and cortex [@Movshon1975; @Hubel1959], will introduce second harmonic components even when the stimulus itself is modulated in an on-off fashion [@McKeefry1996].
 Analogously, in the spatial domain, so-called 'simple cells' are sensitive to the polarity of a spatial contrast modulation while 'complex cells' respond to the presence of patterned spatial contrast irrespective of its spatial position [@Hubel1962].
-The SSVEP response to a contrast-reversing sine-wave grating therefore contains information about nonlinear computations performed across a range of retinal and cortical cell types.
+Both the amplitude and the phase of the SSVEP response to a single contrast-reversing sine-wave grating therefore contain information about nonlinear computations performed across a range of retinal and cortical cell types.
 
@@
 In principle, that information could be recovered from the SSVEP signal - a possibility recognised in the early days of the technique [@Regan1988].
-Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data.
-For example, @Tsai2012 demonstrated that a gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
+Although characterising the complete set of computations along the entire processing pathway is challenging, and a full system identification may require the use of multiple different input frequencies [@boydMeasuringVolterraKernels1983], careful parametric variation of the input stimuli does allow us to fit models of early visual visual processing [e.g.
+@Zemon2006Dynamic] and, by incorporating mathematical models of gain control mechanisms with either instantaneous, or temporally-integrated gain control, it is possible to model gain control using SSVEP data.
+For example, @Tsai2012 demonstrated that a relatively simple gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies.
 This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output.
+More recent work has used this approach to study binocular integration in both normal subjects and amblyopes [@Hou-houchuan2021; @Hou2020].
 Our more recent work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017].
@@
 Psychophysically, adaptation increases detection thresholds, but has little effect on contrast discrimination performance [@Ross1993], much like pattern masks [@Foley1994].
-Although SSVEP adaptation effects show strong tuning for orientation [@Vergeer2018; @Campbell1970] and spatial frequency [@Mecacci1976], there appear to be no studies measuring the full SSVEP contrast response function before and after an extended period of adaptation.
+Although SSVEP adaptation effects show strong tuning for orientation [@Vergeer2018; @Campbell1970] and spatial frequency [@Mecacci1976], there appear to be very few measurements of the full SSVEP contrast response function before and after an extended period of adaptation.
+One exception is a study by @bachGreenleeBuhler1988 that examined the effects of adaptation on SSVEP responses, and found a paradoxical increase in the 1F SSVEP amplitude following prolonged exposure to high-contrast stimuli across much of the probe contrast range.
+They explain this as being due to the presence of a nonlinear transducer function that saturates at high probe contrasts, driving the response to higher harmonic frequencies.
+Adaptation reduces the apparent contrast of the input, leading to less distortion, a more sinusoidal output and a concentration of the frequency-tagged response at lower harmonics.
+This explanation is conceptually related to the 'supersaturation' effect that we observed in Figure 4.
 
@@
 
-Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels, and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis (see @fig-sweepvep).
-This contrast level was shown to correspond approximately with psychophysically measured detection thresholds [@Norcia1986].
+Because SSVEP responses at detection threshold are very small, estimating a threshold is achieved by measuring the contrast response function at relatively high levels (in the linear part of the log-contrast response function prior to saturation), and extrapolating back along the function (either contrast vs response measured at a constant spatial frequency or spatial frequency vs response at a constant contrast level) to estimate its intercept with the x-axis [@Campbell1970] (see @fig-sweepvep).
+This contrast level has been shown to correspond approximately with psychophysically measured detection thresholds [@campbellVisualEvokedPotential1972; @Norcia1986].
+One factor affecting the accuracy of this technique is the presence of noise in the EEG signal which will increase the apparent amplitude of the SSVEP responses across the measurement range.
+The effect of this noise is particulary apparent for recordings with low SNR or where the slope of the CRF is shallow (and therefore where small changes in the vertical offset of the fit result in large changes in the X-axis intercept).
+Accurate estimates of thresholds therefore require the experimenter to model this noise and compensate for its effects, as detailed in @Norcia1989.
+
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
 
 ![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
-The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/sweepplot.pdf){#fig-sweepvep}
+The solid line is the linear regression fit to the lowest four data points prior to the satuating part of the curve, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](){#fig-sweepvep}
+:::
+::: {.content-visible unless-format="docx"}
+![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
+The solid line is the linear regression fit to the  lowest four data points prior to the satuating part of the curve, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/6_sweepplot.pdf){#fig-sweepvep}
+:::
 
 A robust estimate of the threshold therefore requires the measurement of the SSVEP amplitude at many different super-threshold contrast levels.
-This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@Tyler1979].
-To avoid hysteresis effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979; @Norcia1985a; @Norcia1990].
-The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
+This was made faster by the development of the ‘sweep VEP’ paradigm in which the stimulus changed its contrast, spatial frequency or some other property, throughout a trial [@reganRapidObjectiveRefraction1973; @Tyler1979].
+To avoid hysteresis or short-term adaptation effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979; @Norcia1985a; @Norcia1990].
+The sweep VEP (really, a 'sweep SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
 In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].
 
-This approach has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
-In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+The approach of estimating visual thresholds by extrapolating SSVEP responses at higher stimulus levels has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
+In general, SSVEP measurements of infant vision have revealed that contrast sensitivity and acuity for both achromatic and chromatic stimuli develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b].
+For example, using SSVEPs, @Norcia1985b found acuity thresholds in one month old infants that were roughly half those estimated previously from behavioural experiments [reviewed in @brownDevelopmentVisualSensitivity1990a].
+The ability to make very sensitive measurements of early visual function in infants also allows researchers to monitor the trajectory of visual development and identify different temporal epochs of developmental change [@Norcia1990].
+
+SSVEP measures of thresholds for detecting purely chromatic stimuli find responses to isoluminant stimuli by around 5-6 weeks postnatally, with contrast detection reaching near-adult levels by around six months [@morroneDevelopmentInfantContrast1993].
 Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983].
-At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
-However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
 
-The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section [@Candy2001; @Pei2017].
-Although contrast gain control is measurable in infants as young as six weeks old [@Candy2001; @Morrone1986; @Skoczenski1998], its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
+At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which requires careful measurement of the infant's gaze direction.
+In slightly older chilren, other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
+
+The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section.
+Although response gain [@Morrone1986] or contrast gain control [@Candy2001; @Skoczenski1998] is measurable in infants as young as six weeks old, its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
 
@@
 SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye (see [@fig-plotclinical]b) across the contrast range [@Baker2015; @Lygo2021].
+Measurements of interocular contrast modulation [for example, @norciaChangesCorticalActivity2000; @Hou-houchuan2021] can also provide information about the way that amblyopia changes the balance of excitation and inhibition between eyes.
 There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.
@@
 SSVEP responses also provide a potential method to diagnose Parkinson's before any symptoms manifest, and to monitor the effect of treatments.
+
+::: {.content-visible when-format="docx"}
+- FIGURE ABOUT HERE
 
@@
 (2014).
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/clinicalplot.pdf){#fig-plotclinical}
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](){#fig-plotclinical}
+:::
+
+::: {.content-visible unless-format="docx"}
+![Example contrast response functions for different clinical conditions.
+Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al.
+(2011).
+Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al.
+(2015).
+Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al.
+(2018).
+Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al.
+(2014).
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/7_clinicalplot.pdf){#fig-plotclinical}
+:::
 
@@
 Later in the visual pathway, we would like to know more about the role of corticothalamic feedback in the LGN.
-Although it is commonly thought of as a simple relay between the eye and the cortex,  the majority of inputs to the LGN come from cortex not the eye, and there is good evidence that contrast processing in the LGN can be altered by top-down signals including attention [@Briggs2007; @Sherman1996; @OConnor2002; @Gouws2014].
+Although it is commonly thought of as a simple relay between the eye and the cortex, the majority of inputs to the LGN come from cortex rather than the eye, and there is good evidence that contrast processing in the LGN can be altered by top-down signals including attention [@Briggs2007; @Sherman1996; @OConnor2002; @Gouws2014].
 Bottom-up inputs to the LGN are segregated by eye and so responses there are often considered to be purely monocular, but it is possible that feedback signals allow some binocular computations such as interocular normalisation [@Baker2007; @Dougherty2019] to begin even at this relatively early stage.
@@
 All of these questions can be addressed by using SSVEP and frequency tagging to examine the computations that combine and transform the inputs at different cortical stages [@Watts2024; @Baker2017; @Chen2021; @Katyal2018; @Busse2009].
+
 The original promise of the SSVEP approach was that the  entire complex-valued frequency spectrum recorded at each location provided detailed information about the processing nonlinearities up to and including that point.
@@
 Feeding frequency-tagged inputs into these types of model allows us to generate synthetic SSVEP responses that can be compared with those measured in human subjects.
-In principle, we are therefore able to use the SSVEP to fit the parameters of early visual processing.
+In principle, we are therefore able to use the SSVEP to derive the parameters of early visual processing.
 This approach may also allow us to develop more sensitive tests for the changes in early visual processing that accompany a wide range of neurological diseases and disorders.

### 2024-12-03 Rversion/reviewR.tex
Propagated changes to R version

@@
 visually-evoked responses: they have relatively unambiguous ouput
-measures, a high SNR, and they allow us to analyze interactions between
-stimulus components using a convenient mathematical framework. Here we
-describe how SSVEPs have been used to study visual contrast over the
-past 70 years (Dawson, 1954). Because our thinking about SSVEPs is
-well-described by simple mathematical models, we embed code that
-illustrates key steps in the modelling and analysis. This paper can
-therefore be used both as a review of the use of SSVEP in measuring
-human contrast processing, and as an interactive learning aid.
+measures, a high signal to noise ratio (SNR), and they allow us to
+analyze interactions between stimulus components using a convenient
+mathematical framework. Here we describe how SSVEPs have been used to
+study visual contrast over the past 70 years (Dawson, 1954). Because our
+thinking about SSVEPs is well-described by simple mathematical models,
+we embed code that illustrates key steps in the modelling and analysis.
+This paper can therefore be used both as a review of the use of SSVEP in
+measuring human contrast processing, and as an interactive learning aid.
+
+Keywords:
+
+\begin{itemize}
+\tightlist
+\item
+  EEG
+\item
+  VEP
+\item
+  SSVEP
+\item
+  Vision
+\item
+  Contrast
+\end{itemize}
+
+\newpage{}
 
@@
 studied using many different techniques. Early work used psychophysical
-methods to measure contrast sensitivity (Campbell \& Green, 1965),
-defined as the inverse of the lowest contrast that can be reliably
-detected. But neural responses can also be measured more directly using
-techniques such as fMRI, MEG, and EEG. Here we will describe how an EEG
-method known as the steady state visually evoked potential (SSVEP)
-technique has contributed to our understanding of human contrast
-processing in health, disease and throughout development.
+methods to measure contrast sensitivity (Schade, 1956; Campbell \&
+Green, 1965), defined as the inverse of the lowest contrast that can be
+reliably detected. But neural responses can also be measured more
+directly using techniques such as functional magnetic resonance imaging
+(fMRI), magnetoencephalography (MEG), and electroencephalography (EEG).
+Here we will describe how an EEG method known as the steady state
+visually evoked potential (SSVEP) technique has contributed to our
+understanding of human contrast processing in health, disease and
+throughout development.
 
 The SSVEP is a continuous electrical response evoked in the brain by
-visual stimuli flickering at a constant frequency (Regan, 1966). For
-contrast-defined stimuli, such as sine-wave gratings, it is strongest at
-the occipital pole, adjacent to the early visual areas that generate the
-signal, although careful analysis of individual VEPs reveals multiple
-generators throughout visual cortex (Di Russo \emph{et al.}, 2005,
-2007). The flickering stimulus entrains neural population responses at
-multiples of the stimulus frequency, so continuous EEG data are
-typically analysed by taking the Fourier transform, and estimating the
-amplitude at these frequencies. Two common variants involve sinusoidal
-on-off flicker, where the stimulus alternates between a blank background
-and the peak contrast, and sinusoidal counterphase flicker, where the
-stimulus alternates in phase (i.e the black regions become white and the
-white regions become black). On-off flicker can drive independent
-populations of on- and off-cells in the retina once per cycle and can
-therefore produce a response at the fundamental flicker frequency, known
-as 1F, and its integer harmonics: 2F, 3F, 4F and so on. Counterphase
-flicker contains two transients per cycle and therefore does not produce
-a response at 1F, only at its even harmonics: 2F, 4F, 6F and so on.
-Because square-waves are spectrally broad-band, square wave flicker
-tends to produce more complex spectral harmonics than sine-wave flicker.
+visual stimuli flickering at a constant frequency (Regan, 1966\emph{a}).
+For contrast-defined stimuli, such as sine-wave gratings, it is
+strongest at the occipital pole, adjacent to the early visual areas that
+generate the signal, although careful analysis of individual
+visually-evoked potentials (VEPs) reveals multiple generators throughout
+visual cortex (Di Russo \emph{et al.}, 2005, 2007). The flickering
+stimulus entrains neural population responses at multiples of the
+stimulus frequency, so continuous EEG data are typically analysed by
+taking the Fourier transform, and estimating the amplitude at these
+frequencies.
+
+Two common stimulus variants involve sinusoidal on-off (or
+`appearance/disappearance') flicker, where the stimulus alternates
+between a blank background and the peak contrast, and sinusoidal
+counterphase flicker, where the stimulus alternates in phase (i.e the
+black regions become white and the white regions become black). On-off
+flicker can excite independent populations of on- and off-cells in the
+visual system once per cycle. For spatial patterns, the contributions of
+individual cortical cells to these excitations are thought to sum at the
+scalp and generate an average of their onset responses on each cycle.
+For very low spatial frequencies (including zero), resposes at the scalp
+will be dominated by a single on- or off- cell type on each half cycle.
+
+If the amplitudes and waveforms of responses to the stimulus appearance
+and disappearnce are not perfectly balanced, a response at the
+fundamental flicker frequency, known as 1F, and its integer harmonics
+can be produced. For stimuli of any spatial frequency (including zero),
+an imbalance can be caused by differences in either the onset- and
+offset- responses of either on- or off-cells. For example, if only
+on-sensitive neurons were present (firing to the onset, but not the
+offset of local increases in luminance) the response to the on/off (or
+`appearance/disappearance') presentation of a grating of any spatial
+frequency below the resolution limit would consist entirely of odd
+harmonics.
+
+By comparison, counterphase flickering patterns generate two essentially
+identical transients per cycle and therefore do not produce a response
+at 1F, only even harmonics: 2F, 4F, 6F and so on. Because square-waves
+are spectrally broad-band, square wave flicker (either on/off or
+contrast reversing) tends to produce additional spectral harmonics
+compared to sine-wave flicker.
 
@@
 reflect nonlinear processing in the visual system (Regan \& Regan,
-1988). SSVEP signals can also be elicited by periodic changes of
-stimulus properties other than achromatic and chromatic contrast, such
-as motion, stereo depth, and facial identity or expression (see Norcia
-\emph{et al.}, 2015, for an overview); however our focus here is on the
-contrast response.
+1988). Because they can arise from different neuronal computations,
+different populations and even different stages of the visual system,
+different harmonics can exhibit different input-output functions. For
+example, 1F and 2F responses can have different thresholds (Bobak
+\emph{et al.}, 1984) and scalp topographies (Regan, 1973\emph{a}) while
+Kaestner \emph{et al.} (2024) show both different response slopes and
+thresholds for 1F and 2F components generated by dynamic noise.
+
+SSVEP signals can also be elicited by periodic changes of stimulus
+properties other than achromatic and chromatic contrast, such as motion,
+stereo depth, and facial identity or expression (see Norcia \emph{et
+al.}, 2015, for an overview); however our focus here is on the contrast
+response.
 
@@
 achromatic contrast, SSVEPs have proven to be an excellent measure of
-early chromatic processing as well (McKeefry \emph{et al.}, 1996; Di
-Russo \emph{et al.}, 2001\emph{a}) (see also (Baseler \& Sutter, 1997)).
+early chromatic processing as well (Regan, 1966\emph{b}, 1973\emph{a},
+1975; McKeefry \emph{et al.}, 1996; Di Russo \emph{et al.},
+2001\emph{a}) (see also Baseler \& Sutter, 1997).
 
@@
 environment. This normalization, achieved through a computation called
-`contrast gain control' (Heeger, 1992; Foley, 1994; Carandini \& Heeger,
+`contrast gain control' (Ohzawa \emph{et al.}, 1982, 1985; Bobak
+\emph{et al.}, 1988; Heeger, 1992; Foley, 1994; Carandini \& Heeger,
 2011) maximises the sensitivity of the visual system by making optimal
@@
 
+Although the SSVEP does seem to reflect the activity of relatively
+well-tuned neuronal populations, changing some aspect of the stimulus
+may change the nature of those populations. For example, Campbell \&
+Maffei (1970) noted that densely-sampled measurements of contrast
+responses can reveal the presence of two qualitatively different types
+of neurons that exhibit different log-linear contrast functions
+(Nakayama \& Mackeben, 1982; see also Tyler \& Apkarian, 1985; Norcia
+\emph{et al.}, 1989; Souza \emph{et al.}, 2007). Modeling such
+two-limbed contrast response functions using a single sigmoidal function
+is therefore an approximation.
+
 To understand the utility of the contrast SSVEP, it is helpful to
@@
 
+An additional complexity is introduced by the fact that the EEG is an
+average population response and that the visual system contains many
+different types of neurons. As a stimulus changes in (for example)
+contrast, it may selectively activate qualitatively different neuronal
+populations.
+
 To understand these processes better, we will show how sinusoidal input
@@
 second sample, so the stimulation frequency is 5Hz (F1). The second
-panel shows Fourier transform of the waveform, which contains a
+panel shows the Fourier transform of the waveform, which contains a
 substantial peak at this frequency. If we change the stimulus contrast
@@
 
-Next we can consider the impact of a nonlinearity in processing on the
-responses. One of the simplest nonlinearities is the function that
-describes a cell's response to different levels of contrast. In
-Figure~\ref{fig-transducer} this is modeled by a hyperbolic ratio
-function resulting in a saturating non-linearity:
+Next we can consider the impact of two types of nonlinearity in
+processing on the responses.
+
+First, we consider the fact that neurons in the early visual system code
+either positive or negative contrast. The response of each different
+type of neuron is an increase in firing rate to the increasing levels of
+the preferred contrast polarity, and these neurons typically exhibit
+zero (or baseline) response to the non-preferred polarity. We can model
+the combined responses of these two cell populations to a time varying
+contrast signal by full-wave-rectification. The effect of this
+computation is, effectively, to double the input frequency in the
+population measurement and to introduce additional higher (even)
+harmonics due to the discontinuity at the contrast reversals. The effect
+of full-wave rectification on the SSVEP signal is shown in
+Figure~\ref{fig-fullwaverectification}
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{Figures/3_fullwaverectification.pdf}
+
+}
+
+\caption{\label{fig-fullwaverectification}Illustration of a full-wave
+rectified sinusoid (left), its Fourier spectrum (middle), and how the
+amplitude of the first harmonic component increases with contrast
+(right).}
+
+\end{figure}%
+
+The contrast response function (CRF) is still linear for this model -
+rectification has doubled the frequency (and introduced higher
+harmonics) but the response amplitude measured by the second harmonic is
+still directly proportional to the input amplitude.
+
+Neurons in the visual system typically do not exhibit this perfect
+response linearity. Instead, neuronal response are well-modeled by some
+form of saturating non-linearity: as the contrast increases, the
+response of the neuron increases but the response increase per unit
+contrast is reduced at the high end of the contrast range. This
+saturating nonlinearity is often modeled as a hyperbolic ratio function
+of the form:
 
@@
 
-This nonlinearity has clear effects on signal transduction. The
-sinusoidal waveform is distorted by the nonlinearity, and appears
-frequency doubled (left panel of Figure~\ref{fig-transducer}; note that
-other nonlinearities such as rectification and squaring can have similar
-effects). The distortion is reflected in the Fourier spectrum (middle
-panel of Figure~\ref{fig-transducer}), which now includes responses at
-integer multiples (known as harmonics) of the stimulation frequency (10,
-20 and 30Hz), but no response at the fundamental (5Hz). Finally, the
-contrast response function is now nonlinear (right panel of
-Figure~\ref{fig-transducer}). Although the hyperbolic ratio function
+This nonlinearity has clear effects on signal transduction. In
+Figure~\ref{fig-transducer} the full-wave rectified waveform (which is
+combining the response of both `on' and `off' neurons) passes through
+the saturating nonlinearity (left panel of Figure~\ref{fig-transducer}).
+The frequency doubling seen in Figure~\ref{fig-fullwaverectification} is
+still present but the the contrast response function is now nonlinear
+(right panel of Figure~\ref{fig-transducer}).
+
+Interestingly, although the hyperbolic ratio function
 (Equation~\ref{eq-nakarushton}) is monotonic, the CRF resulting from
@@
 
-\includegraphics{Figures/transducer.pdf}
+\includegraphics{Figures/4_transducer.pdf}
 
@@
 \caption{\label{fig-transducer}Illustration of how a sinusoidal input
-signal is affected by a nonlinear transducer. The input wavefore (left
-panel) is frequency doubled by the squaring operator, and now features
-energy at the second harmonic (10Hz) and its multiple (middle panel).
-The contrast response function at the second harmonic is now nonlinear,
-and saturates at high contrasts (right panel).}
+signal is affected by a nonlinear transducer. The input waveform (left
+panel) is frequency doubled by the full-wave rectification (see
+Figure~\ref{fig-fullwaverectification}), and now has energy at the
+second harmonic (10Hz) and higher even harmonics (middle panel). Passing
+the signal through a second nonlinearity (the hyperbolic ratio function)
+causes the contrast response function at the second harmonic (2F) to
+become nonlinear and it and saturates (or even `supersaturates') at high
+contrasts (right panel).}
 
@@
 
-To illustrate the effect of contrast gain control (Heeger, 1992), we
-next include a second component (a `mask') that contributes to the gain
-pool of the first (the `target'). The mask stimulus will suppress the
-response at the target frequency, reducing its amplitude. The
-suppression is reciprocal - activity at the mask frequency is also
-reduced by the presence of the target, in a contrast-dependent manner
-(Busse \emph{et al.}, 2009). At a single pair of (matched) contrast
-levels, we see a complex pattern of intermodulation terms in the
-transduced waveforms (left panel of Figure~\ref{fig-twoinputs}) and in
-the Fourier spectrum (middle panel of Figure~\ref{fig-twoinputs}). The
-two components interact generating nonlinear intermodulation terms at
-sums and differences of the input frequencies (F1, F2). The nature of
-the interaction --- and therefore the pattern of intermodulation terms
---- is determined by the computations happening as the contrast signal
-moves from retina to cortex (Regan \& Regan, 1988).
+To illustrate the effect of contrast gain control (Ohzawa \emph{et al.},
+1982; Heeger, 1992), we next include a second component (a `mask') that
+contributes to the gain pool of the first (the `target'). The mask
+stimulus will suppress the response at the target frequency, reducing
+its amplitude. The suppression is reciprocal - activity at the mask
+frequency is also reduced by the presence of the target, in a
+contrast-dependent manner (Regan \& Regan, 1988; Brown \emph{et al.},
+1999; Candy \emph{et al.}, 2001; Busse \emph{et al.}, 2009). At a single
+pair of (matched) contrast levels, we see a complex pattern of
+intermodulation terms in the transduced waveforms (left panel of
+Figure~\ref{fig-twoinputs}) and in the Fourier spectrum (middle panel of
+Figure~\ref{fig-twoinputs}). The two components interact generating
+nonlinear intermodulation terms at sums and differences of the input
+frequencies (F1, F2). The nature of the interaction --- and therefore
+the pattern of intermodulation terms --- is determined by the
+computations happening as the contrast signal moves from retina to
+cortex (Regan \& Regan, 1988).
 
@@
 
-\includegraphics{Figures/twoinputs.pdf}
+\includegraphics{Figures/5_twoinputs.pdf}
 
@@
 where \(C_\mathrm{mask}\) reflects the contrast of the mask component at
-a distinct frequency from that of the target. The effect of this extra
+a frequency distinct from that of the target. The effect of this extra
 term is to reduce the target response (see right panel of
@@
 temporal changes in contrast. These cells (Kuffler \emph{et al.}, 1957;
-Alpern, 1971), and cells with similar properties in the LGN (Levitt
-\emph{et al.}, 2001) and cortex (Hubel \& Wiesel, 1959; Movshon, 1975),
-will introduce second harmonic components even when the stimulus itself
-is modulated in an on-off fashion. Analogously, in the spatial domain,
-so-called `simple cells' are sensitive to the polarity of a spatial
-contrast modulation while `complex cells' respond to the presence of
-patterned spatial contrast irrespective of its spatial position (Hubel
-\& Wiesel, 1962). The SSVEP response to a contrast-reversing sine-wave
-grating therefore contains information about nonlinear computations
-performed across a range of retinal and cortical cell types.
+Alpern, 1971), and cells with similar properties in the lateral
+geniculate nucleus (LGN) (Levitt \emph{et al.}, 2001) and cortex (Hubel
+\& Wiesel, 1959; Movshon, 1975), will introduce second harmonic
+components even when the stimulus itself is modulated in an on-off
+fashion (McKeefry \emph{et al.}, 1996). Analogously, in the spatial
+domain, so-called `simple cells' are sensitive to the polarity of a
+spatial contrast modulation while `complex cells' respond to the
+presence of patterned spatial contrast irrespective of its spatial
+position (Hubel \& Wiesel, 1962). Both the amplitude and the phase of
+the SSVEP response to a single contrast-reversing sine-wave grating
+therefore contain information about nonlinear computations performed
+across a range of retinal and cortical cell types.
 
@@
 technique (Regan \& Regan, 1988). Although characterising the complete
-set of computations along entire processing pathway is challenging,
-careful parametric variation of the input stimuli does allow us to fit
-models of early visual gain control using SSVEP data. For example, Tsai
-\emph{et al.} (2012) demonstrated that a gain control model gave a good
-account of the pattern of intermodulation responses produced by two
+set of computations along the entire processing pathway is challenging,
+and a full system identification may require the use of multiple
+different input frequencies (Boyd \emph{et al.}, 1983), careful
+parametric variation of the input stimuli does allow us to fit models of
+early visual visual processing (e.g. Zemon \& Gordon, 2006) and, by
+incorporating mathematical models of gain control mechanisms with either
+instantaneous, or temporally-integrated gain control, it is possible to
+model gain control using SSVEP data. For example, Tsai \emph{et al.}
+(2012) demonstrated that a relatively simple gain control model gave a
+good account of the pattern of intermodulation responses produced by two
 overlaid patterns flickering at different frequencies. This was achieved
 by passing full stimulus waveforms through the transducer nonlinearity,
-and calculating the Fourier spectrum of the model output. Our more
-recent work on signal combination across eyes and space similarly
-demonstrated close correspondence between the predictions of a
-computational model and empirical data in humans (Baker \& Wade, 2017).
-More detailed modelling of intracortical recordings (Groen \emph{et
-al.}, 2022) has revealed details of the timecourse of gain control
-effects, specifically that normalization is delayed slightly relative to
-the initial visual response.
+and calculating the Fourier spectrum of the model output. More recent
+work has used this approach to study binocular integration in both
+normal subjects and amblyopes (Hou \emph{et al.}, 2020; Hou 侯川
+\emph{et al.}, 2021). Our more recent work on signal combination across
+eyes and space similarly demonstrated close correspondence between the
+predictions of a computational model and empirical data in humans (Baker
+\& Wade, 2017). More detailed modelling of intracortical recordings
+(Groen \emph{et al.}, 2022) has revealed details of the timecourse of
+gain control effects, specifically that normalization is delayed
+slightly relative to the initial visual response.
 
@@
 2018) and spatial frequency (Mecacci \& Spinelli, 1976), there appear to
-be no studies measuring the full SSVEP contrast response function before
-and after an extended period of adaptation.
+be very few measurements of the full SSVEP contrast response function
+before and after an extended period of adaptation. One exception is a
+study by Bach \emph{et al.} (1988) that examined the effects of
+adaptation on SSVEP responses, and found a paradoxical increase in the
+1F SSVEP amplitude following prolonged exposure to high-contrast stimuli
+across much of the probe contrast range. They explain this as being due
+to the presence of a nonlinear transducer function that saturates at
+high probe contrasts, driving the response to higher harmonic
+frequencies. Adaptation reduces the apparent contrast of the input,
+leading to less distortion, a more sinusoidal output and a concentration
+of the frequency-tagged response at lower harmonics. This explanation is
+conceptually related to the `supersaturation' effect that we observed in
+Figure 4.
 
@@
 estimating a threshold is achieved by measuring the contrast response
-function at relatively high levels, and extrapolating back along the
-function (either contrast vs response measured at a constant spatial
-frequency or spatial frequency vs response at a constant contrast level)
-to estimate its intercept with the x-axis (see
-Figure~\ref{fig-sweepvep}). This contrast level was shown to correspond
-approximately with psychophysically measured detection thresholds
-(Norcia \emph{et al.}, 1986).
+function at relatively high levels (in the linear part of the
+log-contrast response function prior to saturation), and extrapolating
+back along the function (either contrast vs response measured at a
+constant spatial frequency or spatial frequency vs response at a
+constant contrast level) to estimate its intercept with the x-axis
+(Campbell \& Maffei, 1970) (see Figure~\ref{fig-sweepvep}). This
+contrast level has been shown to correspond approximately with
+psychophysically measured detection thresholds (Campbell \& Kulikowski,
+1972; Norcia \emph{et al.}, 1986). One factor affecting the accuracy of
+this technique is the presence of noise in the EEG signal which will
+increase the apparent amplitude of the SSVEP responses across the
+measurement range. The effect of this noise is particulary apparent for
+recordings with low SNR or where the slope of the CRF is shallow (and
+therefore where small changes in the vertical offset of the fit result
+in large changes in the X-axis intercept). Accurate estimates of
+thresholds therefore require the experimenter to model this noise and
+compensate for its effects, as detailed in Norcia \emph{et al.} (1989).
 
@@
 
-\includegraphics{Figures/sweepplot.pdf}
+\includegraphics{Figures/6_sweepplot.pdf}
 
@@
 which the stimulus changed its contrast, spatial frequency or some other
-property, throughout a trial (Tyler \emph{et al.}, 1979). To avoid
-hysteresis effects, the sweep is sometimes conducted both up and down in
-the same experiment (Tyler \emph{et al.}, 1979; Norcia \& Tyler,
-1985\emph{a}; Norcia \emph{et al.}, 1990). The sweep VEP (really, a
-sweep `SSVEP') technique is now commonly-used to obtain a rapid and
-objective measurement of visual acuity. In particular, because of its
-relative speed and simplicity, this technique has now become a standard
-for conducting tests of visual acuity in very young subjects or where
-behavioural tests are not appropriate (Ridder, 2004; Bach \emph{et al.},
-2008; Hoffmann \emph{et al.}, 2017; Bach \& Farmer, 2020).
-
-This approach has revealed much about the development of visual
-abilities in infants (Harris \emph{et al.}, 1976; Atkinson \emph{et
-al.}, 1979; Braddick \emph{et al.}, 1986). In general, SSVEP
-measurements of infant vision have revealed that contrast sensitivity
-for both achromatic and chromatic contrast as well as stereoscopic depth
-perception develops earlier than had been supposed previously based on
-behavioural readouts (Dobson \emph{et al.}, 1978; Norcia \& Tyler,
-1985\emph{b}) with both chromatic and achromatic contrast detection
-reaching near-adult levels by around six months. Spatial acuity as
-measured by SSVEP reaches adult levels more slowly, but near-adult
-levels are recorded around one year (Norcia \& Tyler, 1985\emph{b};
-Hamer \emph{et al.}, 1989) compared to around six to seven years with
-behavioural measures (Atkinson \& Braddick, 1983; Ellemberg \emph{et
-al.}, 1999). At least some of this difference is likely due to the
-relative objectivity and high SNR of the SSVEP technique compared to
-other methods such as preferential looking, which require careful
-measurement of the infant's gaze direction. However it should be noted
-that other groups have reported electrophysiological correlates of
-visual acuity that more closely match the behavioural measures (De
-Vries-Khoe \& Spekreijse, 1982).
+property, throughout a trial (Regan, 1973\emph{b}; Tyler \emph{et al.},
+1979). To avoid hysteresis or short-term adaptation effects, the sweep
+is sometimes conducted both up and down in the same experiment (Tyler
+\emph{et al.}, 1979; Norcia \& Tyler, 1985\emph{a}; Norcia \emph{et
+al.}, 1990). The sweep VEP (really, a `sweep SSVEP') technique is now
+commonly-used to obtain a rapid and objective measurement of visual
+acuity. In particular, because of its relative speed and simplicity,
+this technique has now become a standard for conducting tests of visual
+acuity in very young subjects or where behavioural tests are not
+appropriate (Ridder, 2004; Bach \emph{et al.}, 2008; Hoffmann \emph{et
+al.}, 2017; Bach \& Farmer, 2020).
+
+The approach of estimating visual thresholds by extrapolating SSVEP
+responses at higher stimulus levels has revealed much about the
+development of visual abilities in infants (Harris \emph{et al.}, 1976;
+Atkinson \emph{et al.}, 1979; Braddick \emph{et al.}, 1986). In general,
+SSVEP measurements of infant vision have revealed that contrast
+sensitivity and acuity for both achromatic and chromatic stimuli
+develops earlier than had been supposed previously based on behavioural
+readouts (Dobson \emph{et al.}, 1978; Norcia \& Tyler, 1985\emph{b}).
+For example, using SSVEPs, Norcia \& Tyler (1985\emph{b}) found acuity
+thresholds in one month old infants that were roughly half those
+estimated previously from behavioural experiments (reviewed in Brown,
+1990). The ability to make very sensitive measurements of early visual
+function in infants also allows researchers to monitor the trajectory of
+visual development and identify different temporal epochs of
+developmental change (Norcia \emph{et al.}, 1990).
+
+SSVEP measures of thresholds for detecting purely chromatic stimuli find
+responses to isoluminant stimuli by around 5-6 weeks postnatally, with
+contrast detection reaching near-adult levels by around six months
+(Morrone \emph{et al.}, 1993). Spatial acuity as measured by SSVEP
+reaches adult levels more slowly, but near-adult levels are recorded
+around one year (Norcia \& Tyler, 1985\emph{b}; Hamer \emph{et al.},
+1989) compared to around six to seven years with behavioural measures
+(Atkinson \& Braddick, 1983; Ellemberg \emph{et al.}, 1999).
+
+At least some of this difference is likely due to the relative
+objectivity and high SNR of the SSVEP technique compared to other
+methods such as preferential looking, which requires careful measurement
+of the infant's gaze direction. In slightly older chilren, other groups
+have reported electrophysiological correlates of visual acuity that more
+closely match the behavioural measures (De Vries-Khoe \& Spekreijse,
+1982).
 
 The SSVEP technique has also been used to study the development of the
-contrast gain control mechanisms described in the previous section
-(Candy \emph{et al.}, 2001; Pei \emph{et al.}, 2017). Although contrast
-gain control is measurable in infants as young as six weeks old (Morrone
-\& Burr, 1986; Skoczenski \& Norcia, 1998; Candy \emph{et al.}, 2001),
-its development appears to be slower, with adult levels being reached at
-approximately 11 years (Pei \emph{et al.}, 2017).
+contrast gain control mechanisms described in the previous section.
+Although response gain (Morrone \& Burr, 1986) or contrast gain control
+(Skoczenski \& Norcia, 1998; Candy \emph{et al.}, 2001) is measurable in
+infants as young as six weeks old, its development appears to be slower,
+with adult levels being reached at approximately 11 years (Pei \emph{et
+al.}, 2017).
 
@@
 Figure~\ref{fig-plotclinical}b) across the contrast range (Baker
-\emph{et al.}, 2015; Lygo \emph{et al.}, 2021). There are currently many
-novel binocular treatments for amblyopia under development, often
-involving virtual reality or stereo display systems designed to
-encourage the two eyes to work together. The steady-state approach may
-be more sensitive and objective than typical acuity measurements, and
-also has the potential to measure suppression between the eyes directly
-(e.g. Zheng \emph{et al.}, 2019; Hu \emph{et al.}, 2023; Du \emph{et
-al.}, 2023).
+\emph{et al.}, 2015; Lygo \emph{et al.}, 2021). Measurements of
+interocular contrast modulation (for example, Norcia \emph{et al.},
+2000; Hou 侯川 \emph{et al.}, 2021) can also provide information about
+the way that amblyopia changes the balance of excitation and inhibition
+between eyes. There are currently many novel binocular treatments for
+amblyopia under development, often involving virtual reality or stereo
+display systems designed to encourage the two eyes to work together. The
+steady-state approach may be more sensitive and objective than typical
+acuity measurements, and also has the potential to measure suppression
+between the eyes directly (e.g. Zheng \emph{et al.}, 2019; Hu \emph{et
+al.}, 2023; Du \emph{et al.}, 2023).
 
@@
 
-\includegraphics{Figures/clinicalplot.pdf}
+\includegraphics{Figures/7_clinicalplot.pdf}
 
@@
 of as a simple relay between the eye and the cortex, the majority of
-inputs to the LGN come from cortex not the eye, and there is good
-evidence that contrast processing in the LGN can be altered by top-down
-signals including attention (Sherman \& Guillery, 1996; O'Connor
-\emph{et al.}, 2002; Briggs \& Usrey, 2007; Gouws \emph{et al.}, 2014).
-Bottom-up inputs to the LGN are segregated by eye and so responses there
-are often considered to be purely monocular, but it is possible that
-feedback signals allow some binocular computations such as interocular
-normalisation (Baker \emph{et al.}, 2007; Dougherty \emph{et al.}, 2019)
-to begin even at this relatively early stage. Frequency tagging inputs
-to different eyes or in different precortical pathways allows us to
-address neurons in different parts of the LGN. These techniques combined
-with advances in recording technology (for example, sensitive
-source-imaged EEG and MEG recordings that can resolve subcortical
-structures (Tesche, 1996; Attal \emph{et al.}, 2012; Attal \& Schwartz,
-2013), noninvasive deep-brain stimulation techniques (Mohammadjavadi
-\emph{et al.}, 2022) or implanted electrode arrays (Krolak-Salmon
-\emph{et al.}, 2003)) may allow us to study these computations in more
-detail.
+inputs to the LGN come from cortex rather than the eye, and there is
+good evidence that contrast processing in the LGN can be altered by
+top-down signals including attention (Sherman \& Guillery, 1996;
+O'Connor \emph{et al.}, 2002; Briggs \& Usrey, 2007; Gouws \emph{et
+al.}, 2014). Bottom-up inputs to the LGN are segregated by eye and so
+responses there are often considered to be purely monocular, but it is
+possible that feedback signals allow some binocular computations such as
+interocular normalisation (Baker \emph{et al.}, 2007; Dougherty \emph{et
+al.}, 2019) to begin even at this relatively early stage. Frequency
+tagging inputs to different eyes or in different precortical pathways
+allows us to address neurons in different parts of the LGN. These
+techniques combined with advances in recording technology (for example,
+sensitive source-imaged EEG and MEG recordings that can resolve
+subcortical structures (Tesche, 1996; Attal \emph{et al.}, 2012; Attal
+\& Schwartz, 2013), noninvasive deep-brain stimulation techniques
+(Mohammadjavadi \emph{et al.}, 2022) or implanted electrode arrays
+(Krolak-Salmon \emph{et al.}, 2003)) may allow us to study these
+computations in more detail.
 
@@
 2009; Baker \& Wade, 2017; Katyal \emph{et al.}, 2018; Chen \&
-Gegenfurtner, 2021; Watts \emph{et al.}, 2024). The original promise of
-the SSVEP approach was that the entire complex-valued frequency spectrum
-recorded at each location provided detailed information about the
-processing nonlinearities up to and including that point. In principle,
-this would allow researchers to uniquely fit the parameters of their
-computational models of visual processing (Regan \& Regan, 1988). To
-date, the complexity of the neuronal computations at even the earliest
-stage of visual processing have hampered this effort - researchers
-typically restrict their analyses to the amplitudes of single, low-order
-frequency components (for example, the input frequencies or simple sums
-and differences of those frequencies). As our understanding of the early
-visual system improves, it is becoming possible to generate more
-realistic parameterised forward models of signal generation (Tsai
-\emph{et al.}, 2012; Baker \& Wade, 2017; Chariker \emph{et al.}, 2020;
-Schrimpf \emph{et al.}, 2020; Groen \emph{et al.}, 2022). Feeding
+Gegenfurtner, 2021; Watts \emph{et al.}, 2024).
+
+The original promise of the SSVEP approach was that the entire
+complex-valued frequency spectrum recorded at each location provided
+detailed information about the processing nonlinearities up to and
+including that point. In principle, this would allow researchers to
+uniquely fit the parameters of their computational models of visual
+processing (Regan \& Regan, 1988). To date, the complexity of the
+neuronal computations at even the earliest stage of visual processing
+have hampered this effort - researchers typically restrict their
+analyses to the amplitudes of single, low-order frequency components
+(for example, the input frequencies or simple sums and differences of
+those frequencies). As our understanding of the early visual system
+improves, it is becoming possible to generate more realistic
+parameterised forward models of signal generation (Tsai \emph{et al.},
+2012; Baker \& Wade, 2017; Chariker \emph{et al.}, 2020; Schrimpf
+\emph{et al.}, 2020; Groen \emph{et al.}, 2022). Feeding
 frequency-tagged inputs into these types of model allows us to generate
@@
 human subjects. In principle, we are therefore able to use the SSVEP to
-fit the parameters of early visual processing. This approach may also
+derive the parameters of early visual processing. This approach may also
 allow us to develop more sensitive tests for the changes in early visual
@@
 
+\bibitem[\citeproctext]{ref-BobakIOVS1984}
+Bobak P, Bodis-Wollner I, Harnois C \& Thornton J (1984). VEPs in humans
+reveal high and low spatial contrast mechanisms. \emph{Investigative
+Ophthalmology \& Visual Science} \textbf{25,} 980--983.
+
+\bibitem[\citeproctext]{ref-bobakCorticalContrastGain1988a}
+Bobak P, Bodis-Wollner I \& Marx MS (1988).
+\href{https://doi.org/10.1113/jphysiol.1988.sp017340}{Cortical contrast
+gain control in human spatial vision.} \emph{The Journal of Physiology}
+\textbf{405,} 421--437.
+
+\bibitem[\citeproctext]{ref-boydMeasuringVolterraKernels1983}
+Boyd S, Tang Y \& Chua L (1983).
+\href{https://doi.org/10.1109/TCS.1983.1085391}{Measuring {Volterra}
+kernels}. \emph{IEEE Transactions on Circuits and Systems} \textbf{30,}
+571--577.
+
 \bibitem[\citeproctext]{ref-Boynton2009}
@@
 
+\bibitem[\citeproctext]{ref-brownDevelopmentVisualSensitivity1990a}
+Brown AM (1990).
+\href{https://doi.org/10.1016/0042-6989(90)90173-I}{Development of
+visual sensitivity to light and color vision in human infants: {A}
+critical review}. \emph{Vision Research} \textbf{30,} 1159--1188.
+
+\bibitem[\citeproctext]{ref-brownDevelopmentRivalryDichoptic1999}
+Brown RJ, Candy TR \& Norcia AM (1999).
+\href{https://www.ncbi.nlm.nih.gov/pubmed/10586959}{Development of
+rivalry and dichoptic masking in human infants}. \emph{Investigative
+Ophthalmology \& Visual Science} \textbf{40,} 3324--3333.
+
 \bibitem[\citeproctext]{ref-Buracas2007}
@@
 
+\bibitem[\citeproctext]{ref-Hou2020}
+Hou C, Nicholas SC \& Verghese P (2020).
+\href{https://doi.org/10.1523/JNEUROSCI.2043-19.2020}{Contrast
+{Normalization} {Accounts} for {Binocular} {Interactions} in {Human}
+{Striate} and {Extra}-striate {Visual} {Cortex}}. \emph{The Journal of
+Neuroscience: The Official Journal of the Society for Neuroscience}
+\textbf{40,} 2753--2763.
+
+\bibitem[\citeproctext]{ref-Hou-houchuan2021}
+Hou 侯川 C, Tyson TL, Uner IJ, Nicholas SC \& Verghese P (2021).
+\href{https://doi.org/10.1523/JNEUROSCI.0268-21.2021}{Excitatory
+{Contribution} to {Binocular} {Interactions} in {Human} {Visual}
+{Cortex} {Is} {Reduced} in {Strabismic} {Amblyopia}}. \emph{The Journal
+of Neuroscience: The Official Journal of the Society for Neuroscience}
+\textbf{41,} 8632--8643.
+
 \bibitem[\citeproctext]{ref-Hu2023}
@@
 \textbf{15,} 4--4.
+
+\bibitem[\citeproctext]{ref-norciaChangesCorticalActivity2000}
+Norcia AM, Harrad RA \& Brown RJ (2000). Changes in cortical activity
+during suppression in stereoblindness. \emph{NeuroReport} \textbf{11,}
+1007.
 
@@
 1203--1209.
+
+\bibitem[\citeproctext]{ref-ohzawaContrastGainControl1982}
+Ohzawa I, Sclar G \& Freeman RD (1982).
+\href{https://doi.org/10.1038/298266a0}{Contrast gain control in the cat
+visual cortex}. \emph{Nature} \textbf{298,} 266--268.
+
+\bibitem[\citeproctext]{ref-ohzawaContrastGainControl1985}
+Ohzawa I, Sclar G \& Freeman RD (1985).
+\href{https://doi.org/10.1152/jn.1985.54.3.651}{Contrast gain control in
+the cat's visual system}. \emph{Journal of Neurophysiology} \textbf{54,}
+651--667.
 
@@
 
+\bibitem[\citeproctext]{ref-reganEffectStimulusColour1966a}
+Regan D (1966\emph{b}). \href{https://doi.org/10.1038/2101056a0}{An
+{Effect} of {Stimulus Colour} on {Average Steady-state Potentials}
+evoked in {Man}}. \emph{Nature} \textbf{210,} 1056--1057.
+
 \bibitem[\citeproctext]{ref-Regan1966}
-Regan D (1966). \href{https://doi.org/10.1016/0013-4694(66)90088-5}{Some
-characteristics of average steady-state and transient responses evoked
-by modulated light}. \emph{Electroencephalogr Clin Neurophysiol}
-\textbf{20,} 238--248.
+Regan D (1966\emph{a}).
+\href{https://doi.org/10.1016/0013-4694(66)90088-5}{Some characteristics
+of average steady-state and transient responses evoked by modulated
+light}. \emph{Electroencephalogr Clin Neurophysiol} \textbf{20,}
+238--248.
+
+\bibitem[\citeproctext]{ref-reganRapidObjectiveRefraction1973}
+Regan D (1973\emph{b}). Rapid {Objective Refraction Using Evoked Brain
+Potentials}. \emph{Investigative Ophthalmology \& Visual Science}
+\textbf{12,} 669--679.
+
+\bibitem[\citeproctext]{ref-reganEvokedPotentialsSpecific1973}
+Regan D (1973\emph{a}).
+\href{https://doi.org/10.1016/0042-6989(73)90237-X}{Evoked potentials
+specific to spatial patterns of luminance and colour}. \emph{Vision
+Research} \textbf{13,} 2381--2402.
+
+\bibitem[\citeproctext]{ref-reganColourCodingPattern1975}
+Regan D (1975).
+\href{https://doi.org/10.1016/0042-6989(75)90205-9}{Colour coding of
+pattern responses in man investigated by evoked potential feedback and
+direct plot techniques}. \emph{Vision Research} \textbf{15,} 175--183.
 

### 2024-12-03 review.qmd
Propagated changes to R version

@@
 Because they can arise from different neuronal computations, different populations and even different stages of the visual system, different harmonics can exhibit different input-output functions.
-For example, 1F and 2F responses can have different thresholds [@BobakIOVS1984] and scalp topographies [@reganEvokedPotentialsSpecific1973] while Kaester et al [@KaestnerTVST2024] show both different response slopes and thresholds for 1F and 2F components generated by dynamic noise.
+For example, 1F and 2F responses can have different thresholds [@BobakIOVS1984] and scalp topographies [@reganEvokedPotentialsSpecific1973] while @KaestnerTVST2024 show both different response slopes and thresholds for 1F and 2F components generated by dynamic noise.
 
@@
 It can be defined as the change in cone photoreceptor activity over space ('spatial contrast') or time ('temporal contrast').
-Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@reganEvokedPotentialsSpecific1973; @reganColourCodingPattern1975; @reganEffectStimulusColour1966a; @McKeefry1996; @Di-Russo2001] (see also [@Baseler1997]).
+Cone photoreceptors - which drive precortical opponent pathways -  contribute to both chromatic and achromatic contrast, and although most of the research we describe here focuses on achromatic contrast, SSVEPs have proven to be an excellent measure of early chromatic processing as well [@reganEvokedPotentialsSpecific1973; @reganColourCodingPattern1975; @reganEffectStimulusColour1966a; @McKeefry1996; @Di-Russo2001] [see also @Baseler1997].
 
@@
 
-These contrast definitions are appropriate both to photometric measures of stimulus contrast [for example, luminance; @Lennie1993] and also to definitions based on cone excitations [@Derrington1984; @MacLeod1979] which are more common in work on chromatic processing.
+These contrast definitions are appropriate both to photometric measures of stimulus contrast (for example, luminance; @Lennie1993) and also to definitions based on cone excitations [@Derrington1984; @MacLeod1979] which are more common in work on chromatic processing.
 
@@
 To avoid hysteresis or short-term adaptation effects, the sweep is sometimes conducted both up and down in the same experiment [@Tyler1979; @Norcia1985a; @Norcia1990].
-The sweep VEP (really, a sweep 'SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
+The sweep VEP (really, a 'sweep SSVEP') technique is now commonly-used to obtain a rapid and objective measurement of visual acuity.
 In particular, because of its relative speed and simplicity, this technique has now become a standard for conducting tests of visual acuity in very young subjects or where behavioural tests are not appropriate [@Ridder2004; @Hoffmann2017; @Bach2008; @Bach2020].
@@
 The approach of estimating visual thresholds by extrapolating SSVEP responses at higher stimulus levels has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979].
-In general, SSVEP measurements of infant vision have revealed that contrast sensitivity and acuity for both achromatic and chromatic stimuli as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b].
-For example, using SSVEPs, Norcia and Tyler [@Norcia1985b] found acuity thresholds in one month old infants that were roughly half those estimated previously from behavioural experiments (reviewed in [@brownDevelopmentVisualSensitivity1990a]).
+In general, SSVEP measurements of infant vision have revealed that contrast sensitivity and acuity for both achromatic and chromatic stimuli develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b].
+For example, using SSVEPs, @Norcia1985b found acuity thresholds in one month old infants that were roughly half those estimated previously from behavioural experiments [reviewed in @brownDevelopmentVisualSensitivity1990a].
 The ability to make very sensitive measurements of early visual function in infants also allows researchers to monitor the trajectory of visual development and identify different temporal epochs of developmental change [@Norcia1990].
 
-SSVEP measures of thresholds for detecting purely chromatic stimuli find responses to isoluminant stimuli by around 5 to six weeks with contrast detection reaching near-adult levels by around six months [@morroneDevelopmentInfantContrast1993].
+SSVEP measures of thresholds for detecting purely chromatic stimuli find responses to isoluminant stimuli by around 5-6 weeks postnatally, with contrast detection reaching near-adult levels by around six months [@morroneDevelopmentInfantContrast1993].
 Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983].
 
-At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.
+At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which requires careful measurement of the infant's gaze direction.
 In slightly older chilren, other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].
@@
 The SSVEP technique has also been used to study the development of the contrast gain control mechanisms described in the previous section.
-Although response- [@Morrone1986] or contrast gain control [@Candy2001; @Skoczenski1998] is measurable in infants as young as six weeks old, its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
+Although response gain [@Morrone1986] or contrast gain control [@Candy2001; @Skoczenski1998] is measurable in infants as young as six weeks old, its development appears to be slower, with adult levels being reached at approximately 11 years [@Pei2017].
 
@@
 SSVEPs provide a convenient and objective method to characterise the difference in neural response to signals in each eye, and typically show reduced responses to stimuli in the amblyopic eye (see [@fig-plotclinical]b) across the contrast range [@Baker2015; @Lygo2021].
-Measurments of interocular contrast modulation (for example, [@norciaChangesCorticalActivity2000; @Hou-houchuan2021] can also provide information about the way that amblyopia changes the balance of excitation and inhibition between eyes.
-
+Measurements of interocular contrast modulation [for example, @norciaChangesCorticalActivity2000; @Hou-houchuan2021] can also provide information about the way that amblyopia changes the balance of excitation and inhibition between eyes.
 There are currently many novel binocular treatments for amblyopia under development, often involving virtual reality or stereo display systems designed to encourage the two eyes to work together.

### 2024-12-03 responseToReviews.md
Response to rev micro updates
Fixing up R2R formatting (use pandoc to make PDF)

@@
 
-Wade and Baker have written a very approachable and well-motivated review of the use of VEP measurements of visual contrast.
+**Wade and Baker have written a very approachable and well-motivated review of the use of VEP measurements of visual contrast.
 The authors rightly focus on the importance of measuring contrast responses under a wide range of conditions, highlighting the ways in which human evoked responses can be made in informative in diverse contexts.
 The review should thus be of broad interest.
-
 There are points where the review is overly broad and others that could use more detail. These are described below.
-
 A central part of the review is the role of computational models and normalization models, in particular.
@@
 This modeling approach assumes that the CRF reflects the activity of a more or less unitary population, and that the population response resembles that of the underlying individual neurons that make up the population.
-
 There are two edge-cases in the literature that suggest a more nuanced approach will be needed.
 It would be useful to be clearer on the assumptions of the modeling approach and what the significance is for interpretation of cases where the assumptions don’t hold.
-
 Two-limbed contrast response functions.
@@
 The two limbs may not be apparent when the function is sampled with only a few points over a wide range of contrast.
-
 Harmonics with different CRFs.
@@
 This was alluded to in the simulation, but could be expanded on.
-
 The lack of dynamics in the normalization model, as presented should be discussed, the brief mention of Tsai et al.
@@
 The authors may wish to cite another framework that explicitly models dynamics (e.g.
-amplitude and phase) in the context of the contrast response function (Zemon and Gordon, Vision Res, 2006) .
+amplitude and phase) in the context of the contrast response function (Zemon and Gordon, Vision Res, 2006)**
 
 ### Response
-> We thank the reviewer for these positive comments.
+We thank the reviewer for these positive comments.
 In general this was intended to be an overview in the sense of a 'primer' for researchers coming to the SSVEP for the first time rather than a detailed historical account of the field - hence the inclusion of code examples and relatively domain-agnostic language.
@@
 We think the Bobak reference is IOVS 1984 and we have also included the Kaestner TVST 2024 reference (thank you for drawing our attention to both of these).
-
 ### Specific comments
 
-Adaptation effects.
+**Adaptation effects.
 Adapted contrast response functions have been measured by Nelson et al., (EEG Journal, 1984) by comparing CRFs measured by sweeping contrast from low to high vs high to low.
-They found that responses at low contrast were smaller in the “down” sweeps and attributed it to differences in prior adaptation.
+They found that responses at low contrast were smaller in the “down” sweeps and attributed it to differences in prior adaptation**
 
-> we could not find this reference.
+We could not find this reference.
 We did find an excellent (and sometimes humorous) review of SSVEPs by Nelson (Nelson, J.
@@
 
-Other studies also measured adaptation effects on the CRF Heinrich and Bach, (IOVS, 2001), Ross and Speed (1991, Proc Roy Soc).
-And the hard-to-find Bach, Greenlee and Buhler (Clin Vis Sci, 1996) paper that modeled the effect within a contrast gain control framework.
+**Other studies also measured adaptation effects on the CRF Heinrich and Bach, (IOVS, 2001), Ross and Speed (1991, Proc Roy Soc).
+And the hard-to-find Bach, Greenlee and Buhler (Clin Vis Sci, 1996) paper that modeled the effect within a contrast gain control framework.**
 
-> The Ross and Speed paper measures masking using SSVEP but the other adaptation measures are, as far as we can tell, psychophysical.
+The Ross and Speed paper measures masking using SSVEP but the other adaptation measures are, as far as we can tell, psychophysical.
 Heinrich and Bach are concerned with transient changes in SSVEP amplitude within a recording which may have an adaptation origin but which manifest as response increases over the first few seconds and seems a little mysterious (we often observe these as well).
@@
 
-Gain control development.
+**Gain control development.
 Morrone and Burr (Nature, 1986) used frequency tagging to measure orientation dependence of masking in infants and adults, as did Candy et al.
@@
 Skoczenski and Norcia used a high contrast 30 Hz dynamic 2D noise masker and variable contrast reversing gratings.
-They found masking consistent with a contrast gain effect.
+They found masking consistent with a contrast gain effect.**
 
-> We have clarified that these papers measure different types of gain control
+We have clarified that these papers measure different types of gain control
 
-Page 5 bottom.
+**Page 5 bottom.
 The reciprocal effect of test and masker was first shown by Regan and Regan  (1988; cited)  for flicker modulation depth and then for gratings by Brown et al., (IOVS,  1999 for dichoptic masking) and  then  by candy et al., (J.
 Neurosci, 2001) along with Busse et al.
-(Neuron, 2009) as cited.
+(Neuron, 2009) as cited**
 
-> Thanks - we have included these references
+Thanks - we have included these references
 
-It’s good to have cited Regan and Regan paper that introduced the notion of fingerprints in the spectrum being tied to the nature of the non-linearity.
-Maybe expand their “fingerprint” approach a little more.
+**It’s good to have cited Regan and Regan paper that introduced the notion of fingerprints in the spectrum being tied to the nature of the non-linearity.
+Maybe expand their “fingerprint” approach a little more.**
 
-> We have expanded a little but are also conscious of limitations to this approach (see also the discussion re unique inputs and system order below).
+We have expanded a little but are also conscious of limitations to this approach (see also the discussion re unique inputs and system order below).
 We think there might be some ways to relax these constraints a little by incorporating information about likely transducer functions at different stages (essentially fitting parameters rather than performing a full system identification) but this still requires many inputs.
 
-Figure 4.
+**Figure 4.
 Oversaturation occurs because the denominator overtakes the numerator at high contrast as it has two terms in it, not the single term in the numerator.
 Would be worth showing total power not saturating.
-Is this a general effect?
+Is this a general effect?**
 
-> We think this topic is probably worth a more detailed treatment than we can provide here.
+We think this topic is probably worth a more detailed treatment than we can provide here.
 Broadly, much of the oversaturation effect can be modeled by an increase in saturation and therefore harmonic distortion - power is transferred from the first to higher harmonics as contrast increases and the output becomes more 'square-wave-like'.
@@
 
-Equation 3 is a special case where the neurons in the numerator are so selective that they don’t “see” the masker except by being influenced indirectly by the gain pool.
+**Equation 3 is a special case where the neurons in the numerator are so selective that they don’t “see” the masker except by being influenced indirectly by the gain pool.
 Putting the test and mask in the numerator and summing them before exponentiating yields robust IM.
 Weights can be used to reflect the neurons preference for the test vs the mask.
-Does equation 3 produce IM?
+Does equation 3 produce IM?**
 
-> In the accompanying code we specifically allow the user to alter the way that normalization is computed and many of the (commented) options produce robust IM.
+In the accompanying code we specifically allow the user to alter the way that normalization is computed and many of the (commented) options produce robust IM.
 We encourage the user to experiment with different configurations to explore their effects.
@@
 
-Page 7 Mckeefry et al., (Vis Neuro, 1996) is a good reference for linking harmonics of on/off to transient sustained responses.
+**Page 7 Mckeefry et al., (Vis Neuro, 1996) is a good reference for linking harmonics of on/off to transient sustained responses.**
 
-> Included
+Included
 
-Page 8 figure 5.
+**Page 8 figure 5.
 The extrapolation to zero amplitude method should be attributed to Cambell and Maffei (J Physiol) 1970).
-Campbell and Kulikowski (J Physiol, 1972) made a detailed comparison to psychophysics.
+Campbell and Kulikowski (J Physiol, 1972) made a detailed comparison to psychophysics.**
 
-> We have now included these references
+We have now included these references
 
-Figure 5 is an oversimplification.
+**Figure 5 is an oversimplification.
 As it doesn’t take into account how the additive noise mixes with the internal VEP response.
@@
 The authors should try to connect the linear regression approach for threshold measurement to the sigmoidal non-linearity, as they are different models of the CRF.
-In practice the difference at the low contrast end is probably too small to matter and the saturation region is excluded.
+In practice the difference at the low contrast end is probably too small to matter and the saturation region is excluded.**
 
-> We have now noted that it is important to consider noise for situations where the SNR is low or the slope is shallow and we expand on the point that you have to stick to the part of the CRF where the curve is more or less linear - as we showed explicitly in the figure.
+We have now noted that it is important to consider noise for situations where the SNR is low or the slope is shallow and we expand on the point that you have to stick to the part of the CRF where the curve is more or less linear - as we showed explicitly in the figure.
 The code also includes an explicit 'noise baseline' that can be used to demonstrate changes in the intercept as a function of SNR.
 
-Page 9. The sweep method was first developed by David Regan (Vision Res, 1973) for use in objective refraction.
+**Page 9. The sweep method was first developed by David Regan (Vision Res, 1973) for use in objective refraction.**
 
-> We have included this reference.
+We have included this reference.
 
-Page 9 “This approach” seems to refer to the sweep VEP, not the method of measuring at multiple contrast levels over a series of trials which was used in the Atkinson/Braddick papers.
+**Page 9 “This approach” seems to refer to the sweep VEP, not the method of measuring at multiple contrast levels over a series of trials which was used in the Atkinson/Braddick papers.
 The sweep method refers to the approach of incrementing or decrementing the stimulus value within a single trial and then averaging the sweeps to get the response function.
-Regan argued that this was a better way of estimating response functions because amplitude variations over a session are distributed over the whole response function shifting it up or down instead of distorting its shape (Regan (Vison Res, 1975, color coding).
+Regan argued that this was a better way of estimating response functions because amplitude variations over a session are distributed over the whole response function shifting it up or down instead of distorting its shape (Regan (Vison Res, 1975, color coding).**
 
-> This is a good point but the (simpler) one we are trying to make here is that extrapolating from high stimulus levels down to zero provides a robust estimate of visual sensitivity in subjects where such measurements might otherwise be hard to make.
+This is a good point but the (simpler) one we are trying to make here is that extrapolating from high stimulus levels down to zero provides a robust estimate of visual sensitivity in subjects where such measurements might otherwise be hard to make.
 We have altered the text to clarify this.
 
-Page 9 …chromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts (Dobson et al., 1978; Norcia & Tyler, 1985b) with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
+**Page 9 …chromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts (Dobson et al., 1978; Norcia & Tyler, 1985b) with both chromatic and achromatic contrast detection reaching near-adult levels by around six months.
 Needs references for chromatic sensitivity and disparity sensitivity.
-
 The DeVries Khoe and Spekreijse comparison between VEP and psychophysics was done above age 4 years and for acuity not for contrast sensitivity.
@@
 Compare Norcia et al., 1990 Vision Res measurement to any behavioral data.
-Suggest making this comparison more quantitative.
+Suggest making this comparison more quantitative.**
 
-> We have made this more concrete.
+We have made this more concrete.
 We have taken out the reference to steroscopic depth (because it is not directly related to contrast).
@@
 
-The authors may wish to cite Norcia et al., (NeuroReport, 2000) for an early measurement of suppression in amblyopia using measurements of the contrast response function in each eye in the presence of a high contrast dichoptic masker in the other.
+**The authors may wish to cite Norcia et al., (NeuroReport, 2000) for an early measurement of suppression in amblyopia using measurements of the contrast response function in each eye in the presence of a high contrast dichoptic masker in the other.
 This study showed that amblyopic observers retain suppressive binocular interactions despite severely reduced stereopsis.
 The dominant eye exerted a greater suppressive effect on the non-dominant eye, suggesting that it maintains a relatively greater number, or more effective, inhibitory contacts with cortical cells.
-Also may wish to cite recent work by Hou and colleagues using a similar approach, including application of normalization models.
+Also may wish to cite recent work by Hou and colleagues using a similar approach, including application of normalization models.**
 
-> We think these were both in already but we have explicitly grouped them in the section on amblyopia to make their relevance clearer.
+We think these were both in already but we have explicitly grouped them in the section on amblyopia to make their relevance clearer.
 
-Schade (JOSA, 1956) was the first to measure the contrast sensitivity function of the human visual system as noted by Campbell and Robson (J.
+**Schade (JOSA, 1956) was the first to measure the contrast sensitivity function of the human visual system as noted by Campbell and Robson (J.
 Physiol, 1968).
-Campbell and Green (1965, cited) credit Schade for how to make sinewave gratings on an oscilloscope but not for having made the first measurement.
+Campbell and Green (1965, cited) credit Schade for how to make sinewave gratings on an oscilloscope but not for having made the first measurement.**
 
-> Included
+Included
 
-Villadaite, Norcia et al., (cited) did not find spatial frequency specific effects at 8 cpd as in Pei et al,.
+**Villadaite, Norcia et al., (cited) did not find spatial frequency specific effects at 8 cpd as in Pei et al,.
 (cited).
-So it’s not really a replication of that study, but more of an extension to a measurement of CRF instead of SF tuning.
+So it’s not really a replication of that study, but more of an extension to a measurement of CRF instead of SF tuning.**
 
-> Noted and altered
+Noted and altered
 
-Page 12, better to state whether attention effects were consistent with response gain or contrast gain to make the result more concrete for the reader.
+**Page 12, better to state whether attention effects were consistent with response gain or contrast gain to make the result more concrete for the reader.**
 
-> Reynolds and Heeger (and Boynton) point out that the range of different types of change reported in the literature (response, contrast or mixtures of both) is explained by subtle differences in stimulus configuration (for example, the size and location of the 'probe' and any masking components relative to the size of the 'attentional spotlight').
+Reynolds and Heeger (and Boynton) point out that the range of different types of change reported in the literature (response, contrast or mixtures of both) is explained by subtle differences in stimulus configuration (for example, the size and location of the 'probe' and any masking components relative to the size of the 'attentional spotlight').
 We now note that the underlying model is one of contrast gain control (multiplicative amplification of the input rather than the output).
 
-Discussion of time varying response paragraph on page 12 is not that relevant to contrast response function measurement, nor is the BCI section.
-These applications of SSVEP and could be deleted or the relationship to contrast response function made clearer.
+**Discussion of time varying response paragraph on page 12 is not that relevant to contrast response function measurement, nor is the BCI section.
+These applications of SSVEP and could be deleted or the relationship to contrast response function made clearer.**
 
-> Originally, we envisaged this document as a primer on SSVEP contrast measurements for those coming to the field for the first time.
+Originally, we envisaged this document as a primer on SSVEP contrast measurements for those coming to the field for the first time.
 In particular, we were struck by the large number of recent papers that use multi-frequency SSVEP contrast flicker as an input to BCI systems.
@@
 
-Page, 14. It’s not clear from the text how one would study feedback with the SSVEP, please elaborate.
+**Page, 14. It’s not clear from the text how one would study feedback with the SSVEP, please elaborate.**
 
-> In the paper we write that we would look for modulation of subcortical responses due to top-down signals.
+In the paper we write that we would look for modulation of subcortical responses due to top-down signals.
 The most natural things to look for might be interocular normalization (because eye layers are segregated in the LGN) and attentional gain control (because covert spatial attentional modulation must arise in cortex).
-This is the type of feedback we are thinking of - not cortico-cortico feedback (which is hard to isolate with SSVEP).
+This is the type of feedback we are thinking of - not cortico-cortico feedback (which is hard to isolate with SSVEP).**
 
-In the last section, it would be worth mentioning that in order to “identify” a non-linear system, one needs multiple inputs.
+**In the last section, it would be worth mentioning that in order to “identify” a non-linear system, one needs multiple inputs.
 If the system is 2nd order, a minimum of two inputs is needed.
@@
 J.
-Circuit Theory and Applications).
+Circuit Theory and Applications).**
 
-> This is a good point and we now include the Boyd reference in the section on fingerprinting.
+This is a good point and we now include the Boyd reference in the section on fingerprinting.
 
@@
 # Comments to the Author
-This manuscript is intended to be a review of the field of steady-state visual evoked potentials (ssVEPs) over the past 70 years.
-The authors state in the introductory paragraph “Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.” The material included and the references cited, however, are quite limited and omit seminal work in this area.
+**This manuscript is intended to be a review of the field of steady-state visual evoked potentials (ssVEPs) over the past 70 years.
+The authors state in the introductory paragraph “Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.” The material included and the references cited, however, are quite limited and omit seminal work in this area.**
 
-> We thank the reviewer for their helpful comments.
+We thank the reviewer for their helpful comments.
 As we noted above, our intention here was to provide an introduction to the SSVEP technique with appropriate simple code samples that would allow relative newcomers to the field to implement the technique.
@@
 
-1. All abbreviations should be defined upon first use.
+**1. All abbreviations should be defined upon first use.**
 
-> Done
+Done
 
-2.
+**2.
 The following statement is unclear and inaccurate: “On-off flicker can drive independent populations of on- and off-cells in the retina once per cycle and can therefore produce a response at the fundamental flicker frequency, known as 1F, and its integer harmonics: 2F, 3F, 4F and so on.
@@
 Any linear contribution to the response elicited by each set of elements (e.g., bars or checks) is expected to be canceled when summed at the electrode site on the scalp.
-Thus, the response at 1F to this stimulus is in fact a nonlinear response.
+Thus, the response at 1F to this stimulus is in fact a nonlinear response.**
 
-> Thank you for this explanation.
+Thank you for this explanation.
 We have now reworded the paper to make this clearer.
@@
 
-> We see that we went too fast here and the description of the signal processing was confusing.
+> We agree that we went too fast here and the description of the signal processing was confusing.
 We now break out the two stages into (Figure 3) full wave rectification only (to model the contributions of on and off cells) - explicitly pointing out that this produces only even harmonics but does not alter the linearity of the CRF.

### 2024-12-03 review.qmd
Fixing up R2R formatting (use pandoc to make PDF)

@@
 format:
-  #docx: 
-    #omitFigures: true
+  docx: 
+    omitFigures: true
   pdf: default
@@
   output=RMax*(inputContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
-  # Changing this equation can produce a different pattern of intermodulation terms in Figure 5.
+
+  # Changing this equation can produce a different pattern of intermodulation terms in Figure 5. For example, use the equation below to include both mask and probe in the numerator.
+
   #output=RMax*(inputContrast**expnt + maskContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
+
   return output

### 2024-12-03 responseToReviews.md
fixed up r2r

@@
 The authors rightly focus on the importance of measuring contrast responses under a wide range of conditions, highlighting the ways in which human evoked responses can be made in informative in diverse contexts.
-The review should thus be of broad interest.
-There are points where the review is overly broad and others that could use more detail. These are described below.
-A central part of the review is the role of computational models and normalization models, in particular.
+The review should thus be of broad interest.**
+
+**There are points where the review is overly broad and others that could use more detail. These are described below.**
+
+**A central part of the review is the role of computational models and normalization models, in particular.
 The use of the normalization framework for describing the VEP contrast response function (CRF) has been fruitful in terms of mechanistic interpretation and linking to other measurements, such as single-unit physiology and fMRI.
-This modeling approach assumes that the CRF reflects the activity of a more or less unitary population, and that the population response resembles that of the underlying individual neurons that make up the population.
-There are two edge-cases in the literature that suggest a more nuanced approach will be needed.
-It would be useful to be clearer on the assumptions of the modeling approach and what the significance is for interpretation of cases where the assumptions don’t hold.
-Two-limbed contrast response functions.
+This modeling approach assumes that the CRF reflects the activity of a more or less unitary population, and that the population response resembles that of the underlying individual neurons that make up the population.**
+
+**There are two edge-cases in the literature that suggest a more nuanced approach will be needed.
+It would be useful to be clearer on the assumptions of the modeling approach and what the significance is for interpretation of cases where the assumptions don’t hold.**
+
+**Two-limbed contrast response functions.
 Campbell and Maffei (J.
@@
 As noted, the sigmoidal model assumes that the population response is unitary, but the two-limbed functions suggest that it is a mixture of populations.
-The two limbs may not be apparent when the function is sampled with only a few points over a wide range of contrast.
-Harmonics with different CRFs.
+The two limbs may not be apparent when the function is sampled with only a few points over a wide range of contrast.**
+
+**Harmonics with different CRFs.
 Bobak et al., (IOVS, 1986), found that the first harmonic (1f) and second harmonic (2f) of the pattern onset/offset SSVEP had different thresholds but similar slopes once above threshold.
@@
 The behavior also raises the question of whether to pool harmonics when estimating the response function.
-This was alluded to in the simulation, but could be expanded on.
-The lack of dynamics in the normalization model, as presented should be discussed, the brief mention of Tsai et al.
+This was alluded to in the simulation, but could be expanded on.**
+
+**The lack of dynamics in the normalization model, as presented should be discussed, the brief mention of Tsai et al.
 aside.
@@
 We think the Bobak reference is IOVS 1984 and we have also included the Kaestner TVST 2024 reference (thank you for drawing our attention to both of these).
+
 ### Specific comments
@@
 
-We could not find this reference.
-We did find an excellent (and sometimes humorous) review of SSVEPs by Nelson (Nelson, J.
+We looked hard but could not find this reference.
+We did find an excellent (and unexpectedly humorous) review of SSVEPs by Nelson (Nelson, J.
 I., Seiple, W.
@@
 In the accompanying code we specifically allow the user to alter the way that normalization is computed and many of the (commented) options produce robust IM.
+It is possible to confirm that Eq3 does produce IM.
 We encourage the user to experiment with different configurations to explore their effects.
@@
 
-This type of stimulation does not separate contributions from “independent populations of on- and off-cells.”   There are other ssVEP techniques developed that do tap ON and OFF pathways selectively, which are not cited here.
-Contrast-reversal stimuli which involve symmetrical modulation of light and dark elements cancel odd harmonic frequency components in the response, however, both kinds of stimulus techniques mentioned here elicit only nonlinear responses.
+**This type of stimulation does not separate contributions from “independent populations of on- and off-cells.”   There are other ssVEP techniques developed that do tap ON and OFF pathways selectively, which are not cited here.
+Contrast-reversal stimuli which involve symmetrical modulation of light and dark elements cancel odd harmonic frequency components in the response, however, both kinds of stimulus techniques mentioned here elicit only nonlinear responses.**
 
-> Apologies that we have not been sufficiently clear here.
+Apologies that we have not been sufficiently clear here.
 We do not think we suggested any form of 'separation' and we agree that a) there are infrequently-used paradigms that can target ON and OFF pathways more effectively (for example, 'sawtooth' temporal modulations that provide asymmetrical onset or offset transients) and b) responses from spatial features of differing polarity will tend to be combined at the scalp as long as the spatial frequency is high enough.
@@
 
-3. The statement “SSVEPs have proven to be an excellent measure of early chromatic processing as well” again fails to cite seminal work in this area.
+**3. The statement “SSVEPs have proven to be an excellent measure of early chromatic processing as well” again fails to cite seminal work in this area.**
 
-> We have included three earlier papers by Regan on this topic.
+We have included three earlier papers by Regan on this topic.
 
-4.
-The statement “This normalization, achieved through a computation called ‘contrast gain control’” again neglects to cite the original work that defined “contrast gain control” and the authors ignored a nonlinear biophysical model proposed to explain this phenomenon in ssVEP contrast response functions (which was also applied in developmental work on infants and children).
+**4.
+The statement “This normalization, achieved through a computation called ‘contrast gain control’” again neglects to cite the original work that defined “contrast gain control” and the authors ignored a nonlinear biophysical model proposed to explain this phenomenon in ssVEP contrast response functions (which was also applied in developmental work on infants and children).**
 
-> Contrast gain control is related to 'automatic gain control' - a electrical engineering concept that arose in the early 20th century.
+Contrast gain control is related to 'automatic gain control' - a electrical engineering concept that arose in the early 20th century.
 The concept of modulating the input (rather than the output) of the system became established in the 1930s and 40s (for example in WW2 radar systems).
@@
 
-> The first formalization of the concept in the domain of spatial vision that included the critical concepts of an early, divisive gain control signal computed from some additional estimate of local contrast seems to appear in a seminal paper by Shapley and Victor, (1978).
+The first formalization of the concept in the domain of spatial vision that included the critical concepts of an early, divisive gain control signal computed from some additional estimate of local contrast seems to appear in a seminal paper by Shapley and Victor, (1978).
 Application to contrast adaptation was demonstrated by Ohzawa, Sclar, Freeman, (1982, 1985) with a demonstration of instantaneous gain control in the human SSVEP by Bobak et al in 1988.
@@
 
-5.
+**5.
 The description of Figure 3 and the nonlinear operations performed are unclear.
@@
 The waveform shown in the left panel appears to be a full-wave rectified signal, which would contain theoretically an infinite number of higher harmonics.
-The text also states that “This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.” When was full-wave rectification performed?
+The text also states that “This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity.” When was full-wave rectification performed?**
 
-> We agree that we went too fast here and the description of the signal processing was confusing.
+We agree that we went too fast here and the description of the signal processing was confusing.
 We now break out the two stages into (Figure 3) full wave rectification only (to model the contributions of on and off cells) - explicitly pointing out that this produces only even harmonics but does not alter the linearity of the CRF.
@@
 
-> While there are, technically, infinite even harmonics in both Figure 3 and Figure 4, they decrease with frequency and we plot only three for reasons of space.
+While there are, technically, infinite even harmonics in both Figure 3 and Figure 4, they decrease with frequency and we plot only three for reasons of space.
 
-6.
+**6.
 The authors use the term “frequency tagging” which has been used in the literature to refer to stimulating different regions of the stimulus field or fellow eyes with different sinusoidal frequencies and measuring the intermodulation (sum and difference) frequency responses that result from nonlinear interactions.
-They do not, however, cite the foundational work in this field.
+They do not, however, cite the foundational work in this field.**
 
-> As far as we are aware, the term refers to stimulating some *aspect* of the visual scene (a position or a feature) - not necessarily just a spatial region.
+As far as we are aware, the term refers to stimulating some *aspect* of the visual scene (a position or a feature) - not necessarily just a spatial region.
 There need not be more than one region (it is interesting, for example, to frequency tag a single feature or region and measure the effect of attending towards or away from it) and frequency tagged components can exist in the same location (for example, see the work on contrast normalization described above where the 'probe' and 'mask' have different tags and we explicitly examine the intermodulation terms).
@@
 
-7.
-There is a large body of work that used ssVEPs to investigate contrast processing developmentally (infants through adulthood) and in various disease states, which is mostly missed in this review article.
+**7.
+There is a large body of work that used ssVEPs to investigate contrast processing developmentally (infants through adulthood) and in various disease states, which is mostly missed in this review article.**
 
-> We touch on both these areas but including them in detail would require us to extend what is already becoming a very long paper.
+We touch on both these areas but including them in detail would require us to extend what is already becoming a very long paper.
 
-8. Figures should be labeled.
+**8. Figures should be labeled.**
 
-> It is possible that our text compilation stage failed to auto-generate the figure labels - we apologise if this is the case.
+It is possible that our text compilation stage failed to auto-generate the figure labels - we apologise if this is the case.
 The current version seems to be working correctly (in that each Figure has a label saying 'Figure 1', 'Figure 2' etc which is referenced in the main text).
 
-Thus, the stated intention of this manuscript (“Here we describe how SSVEPs have been used to study visual contrast over the past 70 years”) fails to be met.
-Only select articles from a few sources are covered, and therefore, the richness and value of this technique is lost.
+**Thus, the stated intention of this manuscript (“Here we describe how SSVEPs have been used to study visual contrast over the past 70 years”) fails to be met.
+Only select articles from a few sources are covered, and therefore, the richness and value of this technique is lost.**
 
-> The original document had well over 100 references and the current bibliography now contains 156 entries.
+The original document had well over 100 references and the current bibliography now contains 156 entries.
 Although we have tried to be fair, we have of necessity 'selected' from the literature and have doubtless omitted many important foundational works.
@@
 
-> The original purpose of the document (now long eclipsed) was to serve as a convenient primer on the history of the technique and an interactive visualisation of key concepts implemented in Python and R using a novel 'executable paper' format.
+The original purpose of the document (now long eclipsed) was to serve as a convenient primer on the history of the technique and an interactive visualisation of key concepts implemented in Python and R using a novel 'executable paper' format.
 We think that getting this balance right means that we cannot include a comprehensive history of the entire field - in effect we must continue to be 'selective' and we hope that the relatively few papers we have cited here will serve as useful signposts to the richer literature.

### 2025-01-30 review.qmd
Forcing .docx to make figures. But figs not rendering properly in .pdf on website! Fix...

@@
   docx: 
-    omitFigures: true
+    omitFigures: false
   pdf: default
@@
 ::: {.content-visible when-format="docx"}
-- FIGURE ABOUT HERE
-
-![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](){#fig-centresurround}
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/1_centresurround.pdf){#fig-centresurround}
 :::
@@
 ::: {.content-visible when-format="docx"}
-- FIGURE ABOUT HERE
-
 ![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
-These calculations assume an entirely linear system.](){#fig-linear}
+These calculations assume an entirely linear system.](Figures/2_linearsystem.pdf){#fig-linear}
 :::
@@
 ::: {.content-visible when-format="docx"}
-- FIGURE ABOUT HERE
-![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).
-These calculations assume an entirely linear system.](){#fig-fullwaverectification}
+![Illustration of a full-wave rectified sinusoid (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right).](Figures/3_fullwaverectification.pdf){#fig-fullwaverectification}
 :::
@@
 ::: {.content-visible when-format="docx"}
-- FIGURE ABOUT HERE
-
 ![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer.
 The input waveform (left panel) is frequency doubled by the full-wave rectification (see @fig-fullwaverectification), and now has energy at the second harmonic (10Hz) and higher even harmonics (middle panel).
-Passing the signal through a second nonlinearity (the hyperbolic ratio function) causes the contrast response function at the second harmonic (2F) to become nonlinear and it and saturates (or even 'supersaturates') at high contrasts (right panel).](){#fig-transducer}
+Passing the signal through a second nonlinearity (the hyperbolic ratio function) causes the contrast response function at the second harmonic (2F) to become nonlinear and it and saturates (or even 'supersaturates') at high contrasts (right panel).](Figures/4_transducer.pdf){#fig-transducer}
 :::
@@
 ::: {.content-visible when-format="docx"}
-- FIGURE ABOUT HERE
-
 ![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer.
 Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).
-The effects of using different nonlinearities can be explored in the accompanying code.](){#fig-twoinputs}
+The effects of using different nonlinearities can be explored in the accompanying code.](Figures/5_twoinputs.pdf){#fig-twoinputs}
 :::
@@
 ::: {.content-visible when-format="docx"}
-- FIGURE ABOUT HERE
-
 ![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts.
-The solid line is the linear regression fit to the lowest four data points prior to the satuating part of the curve, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](){#fig-sweepvep}
+The solid line is the linear regression fit to the  lowest four data points prior to the satuating part of the curve, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/6_sweepplot.pdf){#fig-sweepvep}
 :::
@@
 ::: {.content-visible when-format="docx"}
-- FIGURE ABOUT HERE
-
 ![Example contrast response functions for different clinical conditions.
@@
 (2014).
-One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](){#fig-plotclinical}
+One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/7_clinicalplot.pdf){#fig-plotclinical}
 :::

### 2025-01-30 review.qmd
Getting there... replaced  .pdf with .png but still not quite working

@@
 format:
+  latex: default
   docx: 
@@
 plt.tight_layout()
-plt.savefig('Figures/1_centresurround.pdf')
+plt.savefig('Figures/1_centresurround.png', dpi=300)
 
@@
 plt.tight_layout()
-plt.savefig('Figures/2_linearsystem.pdf')
+plt.savefig('Figures/2_linearsystem.png', dpi=300)
 
@@
 plt.tight_layout()
-plt.savefig('Figures/3_fullwaverectification.pdf')
+plt.savefig('Figures/3_fullwaverectification.png', dpi=300)
 
@@
 plt.tight_layout()
-plt.savefig('Figures/4_transducer.pdf')
+plt.savefig('Figures/4_transducer.png', dpi=300)
 
@@
 plt.tight_layout()
-plt.savefig('Figures/5_twoinputs.pdf')
+plt.savefig('Figures/5_twoinputs.png', dpi=300)
 
@@
 plt.tight_layout()
-plt.savefig('Figures/6_sweepplot.pdf')
+plt.savefig('Figures/6_sweepplot.png', dpi=300)
 
@@
 # To save the figure
-plt.savefig('Figures/7_clinicalplot.pdf')
+plt.savefig('Figures/7_clinicalplot.png', dpi=300)
 
@@
 ::: {.content-visible when-format="docx"}
-![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/1_centresurround.pdf){#fig-centresurround}
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/1_centresurround.png){#fig-centresurround}
 :::
 ::: {.content-visible unless-format="docx"}
-![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/1_centresurround.pdf){#fig-centresurround}
+![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/1_centresurround.png){#fig-centresurround}
 :::

### 2025-01-30 review.tex
Getting there... replaced  .pdf with .png but still not quite working

@@
 studied using many different techniques. Early work used psychophysical
-methods to measure contrast sensitivity Schade (1956), defined as the
-inverse of the lowest contrast that can be reliably detected. But neural
-responses can also be measured more directly using techniques such as
-functional magnetic resonance imaging (fMRI), magnetoencephalography
-(MEG), and electroencephalography (EEG). Here we will describe how an
-EEG method known as the steady state visually evoked potential (SSVEP)
-technique has contributed to our understanding of human contrast
-processing in health, disease and throughout development.
+methods to measure contrast sensitivity (Schade, 1956; Campbell \&
+Green, 1965), defined as the inverse of the lowest contrast that can be
+reliably detected. But neural responses can also be measured more
+directly using techniques such as functional magnetic resonance imaging
+(fMRI), magnetoencephalography (MEG), and electroencephalography (EEG).
+Here we will describe how an EEG method known as the steady state
+visually evoked potential (SSVEP) technique has contributed to our
+understanding of human contrast processing in health, disease and
+throughout development.
 
@@
 scalp and generate an average of their onset responses on each cycle.
-For very low spatial frequencies (including zero) resposes at the scalp
+For very low spatial frequencies (including zero), resposes at the scalp
 will be dominated by a single on- or off- cell type on each half cycle.
@@
 By comparison, counterphase flickering patterns generate two essentially
-identical transients per cycle and therefore does not produce a response
+identical transients per cycle and therefore do not produce a response
 at 1F, only even harmonics: 2F, 4F, 6F and so on. Because square-waves
 are spectrally broad-band, square wave flicker (either on/off or
-contrast reversig) tends to produce additional spectral harmonics
+contrast reversing) tends to produce additional spectral harmonics
 compared to sine-wave flicker.
@@
 example, 1F and 2F responses can have different thresholds (Bobak
-\emph{et al.}, 1984) and scalp topographies (Regan, 1973) while Kaester
-et al (Kaestner \emph{et al.}, 2024) show both different response slopes
-and thresholds for 1F and 2F components generated by dynamic noise.
+\emph{et al.}, 1984) and scalp topographies (Regan, 1973\emph{a}) while
+Kaestner \emph{et al.} (2024) show both different response slopes and
+thresholds for 1F and 2F components generated by dynamic noise.
 
@@
 achromatic contrast, SSVEPs have proven to be an excellent measure of
-early chromatic processing as well McKeefry \emph{et al.} (1996) (see
-also (Baseler \& Sutter, 1997)).
+early chromatic processing as well (Regan, 1966\emph{b}, 1973\emph{a},
+1975; McKeefry \emph{et al.}, 1996; Di Russo \emph{et al.},
+2001\emph{a}) (see also Baseler \& Sutter, 1997).
 
@@
 
-\includegraphics{Figures/1_centresurround.pdf}
+\includegraphics{Figures/1_centresurround.png}
 
@@
 Although the SSVEP does seem to reflect the activity of relatively
-well-tuned neuronal populations , changing some aspect of the stimulus
-may change the nature of those populations. For example, Campbell and
-Maffei (Campbell \& Maffei, 1970) noted that densely-sampled
-measurements of contrast responses can reveal the presence of two,
-qualitatively different types of neurons that exhibit different
-log-linear contrast functions (see also Souza \emph{et al.} (2007)).
-Modeling such two-limbed contrast response functions using a single
-sigmoidal function is therefore an approximation.
+well-tuned neuronal populations, changing some aspect of the stimulus
+may change the nature of those populations. For example, Campbell \&
+Maffei (1970) noted that densely-sampled measurements of contrast
+responses can reveal the presence of two qualitatively different types
+of neurons that exhibit different log-linear contrast functions
+(Nakayama \& Mackeben, 1982; see also Tyler \& Apkarian, 1985; Norcia
+\emph{et al.}, 1989; Souza \emph{et al.}, 2007). Modeling such
+two-limbed contrast response functions using a single sigmoidal function
+is therefore an approximation.
 
@@
 average population response and that the visual system contains many
-different types of neurons. As a stimulus changes in, for example,
-contrast it may selectively activate qualitatively different neuronal
-populations. This type
+different types of neurons. As a stimulus changes in (for example)
+contrast, it may selectively activate qualitatively different neuronal
+populations.
 
@@
 First, we consider the fact that neurons in the early visual system code
-either positive or negative contrast. The responses of each different
+either positive or negative contrast. The response of each different
 type of neuron is an increase in firing rate to the increasing levels of
-the prefered contrast polatiry and these neurons typically exhibit zero
-(or baseline) response to the non-prefered polarity. We can model the
-combined responses of these two cell populations to a time varying
+the preferred contrast polarity, and these neurons typically exhibit
+zero (or baseline) response to the non-preferred polarity. We can model
+the combined responses of these two cell populations to a time varying
 contrast signal by full-wave-rectification. The effect of this
-computation is, effectively, to double the input freuquency in the
+computation is, effectively, to double the input frequency in the
 population measurement and to introduce additional higher (even)
@@
 
-One of the simplest nonlinearities is the function that describes a
-cell's response to different levels of contrast. In
-Figure~\ref{fig-transducer} this is modeled by a hyperbolic ratio
-function resulting in a saturating non-linearity:
+The contrast response function (CRF) is still linear for this model -
+rectification has doubled the frequency (and introduced higher
+harmonics) but the response amplitude measured by the second harmonic is
+still directly proportional to the input amplitude.
+
+Neurons in the visual system typically do not exhibit this perfect
+response linearity. Instead, neuronal response are well-modeled by some
+form of saturating non-linearity: as the contrast increases, the
+response of the neuron increases but the response increase per unit
+contrast is reduced at the high end of the contrast range. This
+saturating nonlinearity is often modeled as a hyperbolic ratio function
+of the form:
 
@@
 
-\includegraphics{Figures/4_transducer.pdf_transducer.pdf}
+\includegraphics{Figures/4_transducer.pdf}
 
@@
 
-To illustrate the effect of contrast gain control (Heeger, 1992), we
-next include a second component (a `mask') that contributes to the gain
-pool of the first (the `target'). The mask stimulus will suppress the
-response at the target frequency, reducing its amplitude. The
-suppression is reciprocal - activity at the mask frequency is also
-reduced by the presence of the target, in a contrast-dependent manner
-Busse \emph{et al.} (2009). At a single pair of (matched) contrast
-levels, we see a complex pattern of intermodulation terms in the
-transduced waveforms (left panel of Figure~\ref{fig-twoinputs}) and in
-the Fourier spectrum (middle panel of Figure~\ref{fig-twoinputs}). The
-two components interact generating nonlinear intermodulation terms at
-sums and differences of the input frequencies (F1, F2). The nature of
-the interaction --- and therefore the pattern of intermodulation terms
---- is determined by the computations happening as the contrast signal
-moves from retina to cortex (Regan \& Regan, 1988).
+To illustrate the effect of contrast gain control (Ohzawa \emph{et al.},
+1982; Heeger, 1992), we next include a second component (a `mask') that
+contributes to the gain pool of the first (the `target'). The mask
+stimulus will suppress the response at the target frequency, reducing
+its amplitude. The suppression is reciprocal - activity at the mask
+frequency is also reduced by the presence of the target, in a
+contrast-dependent manner (Regan \& Regan, 1988; Brown \emph{et al.},
+1999; Candy \emph{et al.}, 2001; Busse \emph{et al.}, 2009). At a single
+pair of (matched) contrast levels, we see a complex pattern of
+intermodulation terms in the transduced waveforms (left panel of
+Figure~\ref{fig-twoinputs}) and in the Fourier spectrum (middle panel of
+Figure~\ref{fig-twoinputs}). The two components interact generating
+nonlinear intermodulation terms at sums and differences of the input
+frequencies (F1, F2). The nature of the interaction --- and therefore
+the pattern of intermodulation terms --- is determined by the
+computations happening as the contrast signal moves from retina to
+cortex (Regan \& Regan, 1988).
 
@@
 
-\includegraphics{Figures/4_twoinputs.pdf}
+\includegraphics{Figures/5_twoinputs.pdf}
 
@@
 the SSVEP response to a single contrast-reversing sine-wave grating
-therefore contains information about nonlinear computations performed
+therefore contain information about nonlinear computations performed
 across a range of retinal and cortical cell types.
@@
 technique (Regan \& Regan, 1988). Although characterising the complete
-set of computations along entire processing pathway is challenging, and
-a full system identification may require the use of multiple different
-input frequencies (Boyd \emph{et al.}, 1983), careful parametric
-variation of the input stimuli does allow us to fit models of early
-visual visual processing (e.g. (Zemon \& Gordon, 2006)) and, by
+set of computations along the entire processing pathway is challenging,
+and a full system identification may require the use of multiple
+different input frequencies (Boyd \emph{et al.}, 1983), careful
+parametric variation of the input stimuli does allow us to fit models of
+early visual visual processing (e.g. Zemon \& Gordon, 2006) and, by
 incorporating mathematical models of gain control mechanisms with either
 instantaneous, or temporally-integrated gain control, it is possible to
-model gain control using SSVEP data. For example, (Tsai \emph{et al.},
-2012) demonstrated that a relatively simple gain control model gave a
+model gain control using SSVEP data. For example, Tsai \emph{et al.}
+(2012) demonstrated that a relatively simple gain control model gave a
 good account of the pattern of intermodulation responses produced by two
@@
 work has used this approach to study binocular integration in both
-normal subjects and amblyopesHou \emph{et al.} (2020). Our more recent
-work on signal combination across eyes and space similarly demonstrated
-close correspondence between the predictions of a computational model
-and empirical data in humans (Baker \& Wade, 2017). More detailed
-modelling of intracortical recordings (Groen \emph{et al.}, 2022) has
-revealed details of the timecourse of gain control effects, specifically
-that normalization is delayed slightly relative to the initial visual
-response.
+normal subjects and amblyopes (Hou \emph{et al.}, 2020; Hou 侯川
+\emph{et al.}, 2021). Our more recent work on signal combination across
+eyes and space similarly demonstrated close correspondence between the
+predictions of a computational model and empirical data in humans (Baker
+\& Wade, 2017). More detailed modelling of intracortical recordings
+(Groen \emph{et al.}, 2022) has revealed details of the timecourse of
+gain control effects, specifically that normalization is delayed
+slightly relative to the initial visual response.
 
@@
 2018) and spatial frequency (Mecacci \& Spinelli, 1976), there appear to
-be very measurements of the full SSVEP contrast response function before
-and after an extended period of adaptation. One exception is a study by
-Bach et al (Bach \emph{et al.}, 1988) that examined the effects of
-adaptation on SSVEP responses who found a paradoxical increase in the 1F
-SSVEP amplitude following prolonged exposure to high-contrast stimuli
+be very few measurements of the full SSVEP contrast response function
+before and after an extended period of adaptation. One exception is a
+study by Bach \emph{et al.} (1988) that examined the effects of
+adaptation on SSVEP responses, and found a paradoxical increase in the
+1F SSVEP amplitude following prolonged exposure to high-contrast stimuli
 across much of the probe contrast range. They explain this as being due
 to the presence of a nonlinear transducer function that saturates at
-high probe contrast driving the response to higher harmonic frequencies.
-Adaptation reduces the apparent contrast of the input, leading to less
-distortion, a more sinusoidal output and a concentration of the
-frequency-tagged response at lower harmonics. This explanation is
-conceptually related to the `supersaturation' effect that we observe in
+high probe contrasts, driving the response to higher harmonic
+frequencies. Adaptation reduces the apparent contrast of the input,
+leading to less distortion, a more sinusoidal output and a concentration
+of the frequency-tagged response at lower harmonics. This explanation is
+conceptually related to the `supersaturation' effect that we observed in
 Figure 4.
@@
 contrast level has been shown to correspond approximately with
-psychophysically measured detection thresholds Norcia \emph{et al.}
-(1986). One factor affecting the accuracy of this technique is the
-presence of noise in the EEG signal which will increase the apparent
-amplitude of the SSVEP responses across the measurement range. The
-effect of this noise is particulary apparent for recordings with low SNR
-or where the slope of the CRF is shallow (and therefore where small
-changes in the vertical offset of the fit result in large changes in the
-X-axis intercept). Accurate estimates of thresholds therefore require
-the experimenter to model this noise and compensate for its effects - as
-detailed in (Norcia \emph{et al.}, 1989).
+psychophysically measured detection thresholds (Campbell \& Kulikowski,
+1972; Norcia \emph{et al.}, 1986). One factor affecting the accuracy of
+this technique is the presence of noise in the EEG signal which will
+increase the apparent amplitude of the SSVEP responses across the
+measurement range. The effect of this noise is particulary apparent for
+recordings with low SNR or where the slope of the CRF is shallow (and
+therefore where small changes in the vertical offset of the fit result
+in large changes in the X-axis intercept). Accurate estimates of
+thresholds therefore require the experimenter to model this noise and
+compensate for its effects, as detailed in Norcia \emph{et al.} (1989).
 
@@
 
-\includegraphics{Figures/5_sweepplot.pdf}
+\includegraphics{Figures/6_sweepplot.pdf}
 
@@
 which the stimulus changed its contrast, spatial frequency or some other
-property, throughout a trial Tyler \emph{et al.} (1979). To avoid
-hysteresis or short-term adaptation effects, the sweep is sometimes
-conducted both up and down in the same experiment (Tyler \emph{et al.},
-1979; Norcia \& Tyler, 1985\emph{a}; Norcia \emph{et al.}, 1990). The
-sweep VEP (really, a sweep `SSVEP') technique is now commonly-used to
-obtain a rapid and objective measurement of visual acuity. In
-particular, because of its relative speed and simplicity, this technique
-has now become a standard for conducting tests of visual acuity in very
-young subjects or where behavioural tests are not appropriate (Ridder,
-2004; Bach \emph{et al.}, 2008; Hoffmann \emph{et al.}, 2017; Bach \&
-Farmer, 2020).
+property, throughout a trial (Regan, 1973\emph{b}; Tyler \emph{et al.},
+1979). To avoid hysteresis or short-term adaptation effects, the sweep
+is sometimes conducted both up and down in the same experiment (Tyler
+\emph{et al.}, 1979; Norcia \& Tyler, 1985\emph{a}; Norcia \emph{et
+al.}, 1990). The sweep VEP (really, a `sweep SSVEP') technique is now
+commonly-used to obtain a rapid and objective measurement of visual
+acuity. In particular, because of its relative speed and simplicity,
+this technique has now become a standard for conducting tests of visual
+acuity in very young subjects or where behavioural tests are not
+appropriate (Ridder, 2004; Bach \emph{et al.}, 2008; Hoffmann \emph{et
+al.}, 2017; Bach \& Farmer, 2020).
 
@@
 SSVEP measurements of infant vision have revealed that contrast
-sensitivity and acuity for both achromatic and chromatic stimuli as well
-as stereoscopic depth perception develops earlier than had been supposed
-previously based on behavioural readouts (Dobson \emph{et al.}, 1978;
-Norcia \& Tyler, 1985\emph{b}). For example, using SSVEPs, Norcia and
-Tyler (Norcia \& Tyler, 1985\emph{b}) found acuity thresholds in one
-month old infants that were roughly half those estimated previously from
-behavioural experiments (reviewed in (Brown, 1990)). The ability to make
-very sensitive measurements of early visual function in infants also
-allows researchers to monitor the trajectory of visual development and
-identify different temporal epochs of developmental change (Norcia
-\emph{et al.}, 1990).
+sensitivity and acuity for both achromatic and chromatic stimuli
+develops earlier than had been supposed previously based on behavioural
+readouts (Dobson \emph{et al.}, 1978; Norcia \& Tyler, 1985\emph{b}).
+For example, using SSVEPs, Norcia \& Tyler (1985\emph{b}) found acuity
+thresholds in one month old infants that were roughly half those
+estimated previously from behavioural experiments (reviewed in Brown,
+1990). The ability to make very sensitive measurements of early visual
+function in infants also allows researchers to monitor the trajectory of
+visual development and identify different temporal epochs of
+developmental change (Norcia \emph{et al.}, 1990).
 
 SSVEP measures of thresholds for detecting purely chromatic stimuli find
-responses to isoluminant stimuli by around 5 to six weeks with contrast
-detection reaching near-adult levels by around six months (Morrone
-\emph{et al.}, 1993). Spatial acuity as measured by SSVEP reaches adult
-levels more slowly, but near-adult levels are recorded around one year
-(Norcia \& Tyler, 1985\emph{b}; Hamer \emph{et al.}, 1989) compared to
-around six to seven years with behavioural measures (Atkinson \&
-Braddick, 1983; Ellemberg \emph{et al.}, 1999).
+responses to isoluminant stimuli by around 5-6 weeks postnatally, with
+contrast detection reaching near-adult levels by around six months
+(Morrone \emph{et al.}, 1993). Spatial acuity as measured by SSVEP
+reaches adult levels more slowly, but near-adult levels are recorded
+around one year (Norcia \& Tyler, 1985\emph{b}; Hamer \emph{et al.},
+1989) compared to around six to seven years with behavioural measures
+(Atkinson \& Braddick, 1983; Ellemberg \emph{et al.}, 1999).
 
@@
 objectivity and high SNR of the SSVEP technique compared to other
-methods such as preferential looking, which require careful measurement
+methods such as preferential looking, which requires careful measurement
 of the infant's gaze direction. In slightly older chilren, other groups
@@
 contrast gain control mechanisms described in the previous section.
-Although response- (Morrone \& Burr, 1986) or contrast gain control
-{[}Candy \emph{et al.} (2001); ; Skoczenski \& Norcia (1998){]} is
-measurable in infants as young as six weeks old, its development appears
-to be slower, with adult levels being reached at approximately 11 years
-(Pei \emph{et al.}, 2017).
+Although response gain (Morrone \& Burr, 1986) or contrast gain control
+(Skoczenski \& Norcia, 1998; Candy \emph{et al.}, 2001) is measurable in
+infants as young as six weeks old, its development appears to be slower,
+with adult levels being reached at approximately 11 years (Pei \emph{et
+al.}, 2017).
 
@@
 Figure~\ref{fig-plotclinical}b) across the contrast range (Baker
-\emph{et al.}, 2015; Lygo \emph{et al.}, 2021). Measurments of
-interocular contrast modulation (for example, Hou 侯川 \emph{et al.}
-(2021) can also provide information about the way that amblyopia changes
-the balance of excitation and inhibition between eyes.
-
-There are currently many novel binocular treatments for amblyopia under
-development, often involving virtual reality or stereo display systems
-designed to encourage the two eyes to work together. The steady-state
-approach may be more sensitive and objective than typical acuity
-measurements, and also has the potential to measure suppression between
-the eyes directly (e.g. Zheng \emph{et al.}, 2019; Hu \emph{et al.},
-2023; Du \emph{et al.}, 2023).
+\emph{et al.}, 2015; Lygo \emph{et al.}, 2021). Measurements of
+interocular contrast modulation (for example, Norcia \emph{et al.},
+2000; Hou 侯川 \emph{et al.}, 2021) can also provide information about
+the way that amblyopia changes the balance of excitation and inhibition
+between eyes. There are currently many novel binocular treatments for
+amblyopia under development, often involving virtual reality or stereo
+display systems designed to encourage the two eyes to work together. The
+steady-state approach may be more sensitive and objective than typical
+acuity measurements, and also has the potential to measure suppression
+between the eyes directly (e.g. Zheng \emph{et al.}, 2019; Hu \emph{et
+al.}, 2023; Du \emph{et al.}, 2023).
 
@@
 
-\includegraphics{Figures/6_clinicalplot.pdf}
+\includegraphics{Figures/7_clinicalplot.pdf}
 
@@
 
+\bibitem[\citeproctext]{ref-bobakCorticalContrastGain1988a}
+Bobak P, Bodis-Wollner I \& Marx MS (1988).
+\href{https://doi.org/10.1113/jphysiol.1988.sp017340}{Cortical contrast
+gain control in human spatial vision.} \emph{The Journal of Physiology}
+\textbf{405,} 421--437.
+
 \bibitem[\citeproctext]{ref-boydMeasuringVolterraKernels1983}
@@
 
+\bibitem[\citeproctext]{ref-ohzawaContrastGainControl1982}
+Ohzawa I, Sclar G \& Freeman RD (1982).
+\href{https://doi.org/10.1038/298266a0}{Contrast gain control in the cat
+visual cortex}. \emph{Nature} \textbf{298,} 266--268.
+
+\bibitem[\citeproctext]{ref-ohzawaContrastGainControl1985}
+Ohzawa I, Sclar G \& Freeman RD (1985).
+\href{https://doi.org/10.1152/jn.1985.54.3.651}{Contrast gain control in
+the cat's visual system}. \emph{Journal of Neurophysiology} \textbf{54,}
+651--667.
+
 \bibitem[\citeproctext]{ref-Pei2014}
@@
 \bibitem[\citeproctext]{ref-reganRapidObjectiveRefraction1973}
-REGAN D (1973). Rapid {Objective Refraction Using Evoked Brain
+Regan D (1973\emph{b}). Rapid {Objective Refraction Using Evoked Brain
 Potentials}. \emph{Investigative Ophthalmology \& Visual Science}
@@
 \bibitem[\citeproctext]{ref-reganEvokedPotentialsSpecific1973}
-Regan D (1973).
+Regan D (1973\emph{a}).
 \href{https://doi.org/10.1016/0042-6989(73)90237-X}{Evoked potentials

### 2025-04-23 review.qmd
Finalizing for pub

@@
 format:
-  latex: default
+  #latex: default
   docx: 
-    omitFigures: false
+    omitFigures: true
   pdf: default
@@
 SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high signal to noise ratio (SNR), and they allow us to analyze interactions between stimulus components using a convenient mathematical framework.
-Here we describe how SSVEPs have been used to study visual contrast over the past 70 years [@Dawson1954].
+Here we describe how SSVEPs have been used to study visual contrast over the past 70 years.
 Because our thinking about SSVEPs is well-described by simple mathematical models, we embed code that illustrates key steps in the modelling and analysis.

### 2025-04-23 review.tex
Finalizing for pub

@@
 mathematical framework. Here we describe how SSVEPs have been used to
-study visual contrast over the past 70 years (Dawson, 1954). Because our
-thinking about SSVEPs is well-described by simple mathematical models,
-we embed code that illustrates key steps in the modelling and analysis.
-This paper can therefore be used both as a review of the use of SSVEP in
-measuring human contrast processing, and as an interactive learning aid.
+study visual contrast over the past 70 years. Because our thinking about
+SSVEPs is well-described by simple mathematical models, we embed code
+that illustrates key steps in the modelling and analysis. This paper can
+therefore be used both as a review of the use of SSVEP in measuring
+human contrast processing, and as an interactive learning aid.
 
@@
 
-\bibitem[\citeproctext]{ref-Dawson1954}
-Dawson GD (1954). \href{https://doi.org/10.1016/0013-4694(54)90007-3}{A
-summation technique for the detection of small evoked potentials}.
-\emph{Electroencephalography and Clinical Neurophysiology} \textbf{6,}
-65--84.
-
 \bibitem[\citeproctext]{ref-De-Vries-Khoe1982}

### 2025-04-23 attentionGratings.py
Cleaning more

@@ -1,104 +1,0 @@
-
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.patches import Circle
-
-def hyper_ratio(params, centerContrast, surroundContrast):
-    Rmax, c50, n, Rzero = params
-    Cc = np.clip(centerContrast, 0, 1)
-    Cs = np.clip(surroundContrast, 0, 1)
-
-    # Validate parameters
-    if any(p <= 0 for p in [Rmax, c50, n]) or Rzero < 0:
-        raise ValueError("Invalid parameters: Rmax, c50, and n must be positive, Rzero must be non-negative")
-
-    response = Rzero + Rmax * Cc**n / (Cc**n + Cs**n + c50**n)
-    return response
-
-def make_grating(image_size, frequency, radius=None, phase_shift=0, orientation=0, contrast=.5, is_annulus=False, inner_radius=None):
-    x = np.linspace(-image_size / 2, image_size / 2, image_size)
-    y = np.linspace(-image_size / 2, image_size / 2, image_size)
-    X, Y = np.meshgrid(x, y)
-
-    theta = np.radians(orientation)
-    X_rot = X * np.cos(theta) + Y * np.sin(theta)
-    grating = np.sin(2 * np.pi * frequency * X_rot + phase_shift)
-
-    if radius is not None:
-        if is_annulus and inner_radius is not None:
-            R = np.sqrt(X**2 + Y**2)
-            mask = (R >= inner_radius) & (R <= radius)
-        else:
-            mask = np.sqrt(X**2 + Y**2) <= radius
-        grating *= mask
-    
-    return grating * contrast
-
-def compute_overlap_area(r_spotlight, r1, r2=None):
-    if r2 is None:
-        return np.pi * r1**2 if r_spotlight >= r1 else np.pi * r_spotlight**2
-    area_outer = np.pi * min(r_spotlight, r2)**2
-    area_inner = np.pi * min(r_spotlight, r1)**2 if r_spotlight > r1 else 0
-    return area_outer - area_inner
-
-
-
-def main():
-    # Simulation parameters
-    image_size = 512
-    frequency = 8
-    contrast = 0.5
-    probe_radius = [100, 200]
-    annulus_width = 50
-    gap = 10
-    spotlight_radius = [95, 400]
-    attentionState = 0.7  # Example modulation by attention state
-
-    # Contrast setup
-    contrasts = np.logspace(-2, -.01, 100)
[…]

### 2025-04-23 attentionGratings2.py

@@ -1,61 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.patches import Circle
-
-def hyper_ratio(params, centerContrast, surroundContrast):
-    Rmax, c50, n, Rzero = params
-    Cc = np.clip(centerContrast, 0, 1)  # Ensure center contrast is in the range [0,1]
-    Cs = np.clip(surroundContrast, 0, 1)  # Ensure surround contrast is in the range [0,1]
-    # Validate parameters
-    if any(p <= 0 for p in [Rmax, c50, n]) or Rzero < 0:
-        raise ValueError("Invalid parameters: Rmax, c50, and n must be positive, Rzero must be non-negative")
-
-    # Calculate the response based on the input parameters
-    response = Rzero + Rmax * Cc**n / (Cc**n + Cs**n + c50**n)
-    return response
-
-def make_grating(image_size, frequency, radius=None, phase_shift=0, orientation=0, contrast=.5, is_annulus=False, inner_radius=None):
-    x = np.linspace(-image_size / 2, image_size / 2, image_size)  # Generate linear space for x-axis
-    y = np.linspace(-image_size / 2, image_size / 2, image_size)  # Generate linear space for y-axis
-    X, Y = np.meshgrid(x, y)  # Create a 2D meshgrid from x and y values
-
-    theta = np.radians(orientation)  # Convert orientation from degrees to radians
-    X_rot = X * np.cos(theta) + Y * np.sin(theta)  # Rotate the grid
-    grating = np.sin(2 * np.pi * frequency * X_rot + phase_shift)  # Generate sinusoidal grating
-
-    if radius is not None:
-        if is_annulus and inner_radius is not None:
-            R = np.sqrt(X**2 + Y**2)  # Calculate radial coordinates
-            mask = (R >= inner_radius) & (R <= radius)  # Create a masked area between the inner and outer radius
-        else:
-            mask = np.sqrt(X**2 + Y**2) <= radius  # Create a circular mask of the specified radius
-        grating *= mask  # Apply the mask to the grating
-
-    return grating * contrast  # Scale the grating by contrast and return it
-def compute_overlap_area(r_spotlight, r1, r2=None):
-    if r2 is None:
-        return np.pi * r1**2 if r_spotlight >= r1 else np.pi * r_spotlight**2  # Return area of full circle if within spotlight radius
-    area_outer = np.pi * min(r_spotlight, r2)**2  # Compute area for the outer radius
[…]
